### 2013-09-07 icml/paper.tex
paper draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ICML 2013 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{article}
\usepackage{import, icml2013}
\input{latex-defs.tex}

\icmltitlerunning{Boosted Mirror Descent}

\begin{document}

\twocolumn[
\icmltitle{Boosted Mirror Descent, Frank-Wolfe, and Herding}

\icmlauthor{Jacob Steinhardt}{jsteinhardt@cs.stanford.edu}
\icmladdress{Stanford University,
             353 Serra Street, Stanford, CA 94305 USA}
\icmlauthor{Jonathan Huggins}{jhuggins@mit.edu}
\icmladdress{Massachusetts Institute of Technology,
             43 Vassar Street, Cambridge, MA 02139 USA}

% You may provide any keywords that you 
% find helpful for describing your paper; these are used to populate 
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{boosting, optimization, mirror descent, frank-wolfe, herding}

\vskip 0.3in
]

\begin{abstract} 
TODO
\end{abstract}

\section{Introduction}
\label{sec:intro}

\section{Boosted Mirror Descent}
\label{sec:algorithm}

Consider the following loss function:
\[ L(u) = \max_{\theta \in \Theta} [\langle \theta, u \rangle - R(\theta)] \]
$R$ is some convex function that we should think of as a measure of complexity 
of $\theta$. Thus, the loss is equal to the maximum amount that $\theta$ can match 
$u$, regularized by the complexity function $R(\theta)$.

Examples:
\begin{itemize}
\item Let $R(\theta) = \langle \theta, u_0 \rangle$. Then $L(u)$ is the 
      \emph{maximum mean discrepancy} between $u$ and $u_0$ 
      relative to $\Theta$.
\item Let $R(\theta) = S^*(\theta)$ where $S^*$ is the Fenchel conjugate of 
      any strongly convex function $S$. Then $L(u) = S(u)$.
\end{itemize}

In this document we consider a family of methods for minimizing $L(u)$, 
which are based on Bach's generalization of the Frank-Wolfe algorithm and can 
be interpreted as boosted mirror descent.

First, let us generalize the setting 
to consider a \emph{two-argument} loss function
\[ L(u,\theta) = h(u) + \langle \theta, u \rangle - R(\theta). \]
In the case $h(u) \equiv 0$, we can get the previous setting by 
considering $\max_{\theta \in \Theta} L(u,\theta)$. We will assume 
throughout this document that $h$ and $R$ are both convex functions.

Boosted mirror descent can be thought of as an algorithm for finding 
``saddle points'' of $L$. It is the following algorithm:
\begin{enumerate}
\item $u_1 = 0$
\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, u_t \rangle - R(\theta)$
\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
\end{enumerate}
Let $\hat{u} = \frac{1}{T} \sum_{t = 1}^{T} u_t$. As long as $h$ is strongly convex, we obtain the 
bound:
\[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/\sqrt{T}). \]
However, it is often the case that $h$ is not strongly convex, whereas $R$ is strongly convex. In this case 
we may wish to use the following slightly different algorithm:
\begin{enumerate}
\item $\theta_1 = 0$
\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$
\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
\end{enumerate}
Once again, if we let $\hat{u} = \frac{1}{T} \sum_{t=1}^T u_t$, then we obtain the following bound, which now 
only requires the convexity of $R$:
\[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/\sqrt{T}). \]

\section{Convergence Proofs}
\label{sec:proofs}
We now prove the convergence results cited in Section~\ref{sec:intro}.
\begin{lemma}
\label{lem:ftl}
Let $l_1,\ldots,l_T$ be functions of $x$ and let $x_1,\ldots,x_T$ 
be chosen via $x_{t+1} \in \arg\min_{x} \frac{1}{t} \sum_{s \leq t} l_s(x)$. 
Then for any $x^*$ we have
\[ \sum_{t=1}^T l_t(x_t) \leq \sum_{t=1}^T l_t(x^*) + \sum_{t=1}^T (l_t(x_t) - l_t(x_{t+1})). \]
\end{lemma}
\begin{proof}
Re-arranging, we want to show that $\sum_{t=1}^T l_t(x_{t+1}) \leq \sum_{t=1}^T l_t(x^*)$ for 
all $x^*$. We will show this by induction. In the base case $T=1$, we want to show that $l_1(x_2) \leq l_1(x^*)$ 
for all $x^*$, which is true because $x_2 \in \arg\min_x l_1(x)$ by construction. For 
the inductive step, suppose that the result holds for $T-1$:
\[ \sum_{t=1}^{T-1} l_t(x_{t+1}) \leq \sum_{t=1}^T l_t(x^*). \]
Now set $x^* = x_{T+1}$ in the above inequality and 
add $l_T(x_{T+1})$ to both sides to get
\begin{align*}
\sum_{t=1}^T l_t(x_{t+1}) &\leq \sum_{t=1}^T l_t(x_{T+1}) \\
 &= \min_{x} \sum_{t=1}^T l_t(x) \\
 &\leq \sum_{t=1}^T l_t(x^*),
\end{align*}
which completes the induction. Here we are using the fact that $x_{T+1}$ 
is the minimizer of $\frac{1}{T} \sum_{t=1}^T l_t(x)$ and hence also of 
$\sum_{t=1}^T l_t(x)$.
\end{proof}
\begin{lemma}
\label{lem:convexity}
Suppose that $f(x)$ is \emph{$\alpha$-strongly convex}: $f(x) \geq f(x_0) + \partial f(x_0)^T(x-x_0) + \frac{\alpha}{2}\|x-x_0\|^2$. 
Let $F(y) := \arg\min_x f(x) + y^Tx$. Then $\|F(y_0)-F(y_1)\| \leq \frac{2}{\alpha} \|y_0-y_1\|$.
\end{lemma}
\begin{proof}
Let $x_0 = F(y_0)$. Then we must have $\partial f(x_0) = -y_0$, hence 
\begin{align*}
f(x)+y_1^Tx &\geq f(x_0) + y_1^Tx_0 - y_0^T(x-x_0) + \frac{\alpha}{2}\|x-x_0\|^2.
\end{align*}
But any minimizer of $f(x) + y_1^Tx$ must at least have a value less than $f(x_0) + y_1^Tx_0$. 
It follows that 
\begin{align*}
         & y_1^Tx_0 \geq y_1^Tx_1 - y_0^T(x_1-x_0) + \frac{\alpha}{2}\|x_1-x_0\|^2 \\
\implies & (y_1-y_0)^T(x_0-x_1) \geq \frac{\alpha}{2}\|x_1-x_0\|^2 \\
\implies & \|y_1-y_0\|\|x_0-x_1\| \geq \frac{\alpha}{2}\|x_1-x_0\|^2 \\
\implies & \|x_1-x_0\| \leq \frac{2}{\alpha} \|y_1-y_0\|,
\end{align*}
as was to be shown.
\end{proof}
\begin{proposition}[Convergence of Method 1]
Consider the updates $\theta_t \in \arg\max_{\theta} \langle \theta, u_t \rangle - R(\theta)$ 
and $u_{t+1} \in \arg\min_u h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$. 
Then for $\hat{u} = \frac{1}{T} \sum_{t=1}^T u_t$, we have
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{T} \sum_{t=1}^T \langle u_t - u_{t+1}, \theta_t \rangle. \]
\end{proposition}
\begin{proof}
Note that $L(u_t, \theta_t) = \max_{\theta} L(u_t, \theta)$ by construction. Also note that, if we invoke 
Lemma~\ref{lem:ftl} with $l_t(u) = L(u, \theta_t)$, then we get the inequality
\[ \sum_{t=1}^T L(u_t, \theta_t) \leq \sum_{t=1}^T L(u^*, \theta_t) + \sum_{t=1}^T \langle u_t - u_{t+1}, \theta_t \rangle. \]
Combining these together, we get the string of inequalities
\begin{align*}
L(\hat{u}, \theta) &= L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) \\
 &\leq \frac{1}{T} \sum_{t=1}^T L(u_t, \theta) \\
 &\leq \frac{1}{T} \sum_{t=1}^T L(u_t, \theta_t) \\
 &\leq \frac{1}{T} \sum_{t=1}^T L(u^*, \theta_t) + \frac{1}{T} \sum_{t=1}^T \langle u_t - u_{t+1}, \theta_t \rangle \\
 &\leq \frac{1}{T} \sum_{t=1}^T \sup_{\theta} L(u^*, \theta) + \frac{1}{T} \sum_{t=1}^T \langle u_t - u_{t+1}, \theta_t \rangle \\
 &= \sup_{\theta} L(u^*, \theta) + \frac{1}{T} \sum_{t=1}^T \langle u_t - u_{t+1}, \theta_t \rangle,
\end{align*}
as was to be shown.
\end{proof}
\begin{corollary}
Suppose that $h$ is $\alpha$-strongly convex and let $R = \sup_{\theta} |\langle \theta, \theta \rangle|$. Then 
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{4R^2\log(T+1)}{\alpha T}. \]
\end{corollary}
\begin{proof}
By Lemma~\ref{lem:convexity}, we have 
\begin{align*}
\|u_t-u_{t+1}\| &\leq \frac{2}{\alpha}\|\frac{1}{t}\sum_{s \leq t}\theta_s - \frac{1}{t+1}\sum_{s \leq t+1}\theta_s\| \\
 &= \frac{2}{\alpha}\|\frac{1}{t(t+1)} \sum_{s \leq t} \theta_s - \frac{1}{t+1} \theta_{t+1}\| \\
 &\leq \frac{2}{\alpha} \left(\frac{1}{t(t+1)} \sum_{s \leq t} \|\theta_s\| + \frac{1}{t+1} \|\theta_{t+1}\|\right) \\
 &\leq \frac{4R}{\alpha (t+1)}.
\end{align*}
It follows from Cauchy-Schwartz that 
\begin{align*}
\sum_{t=1}^T \langle u_t - u_{t+1}, \theta_t \rangle &\leq \sum_{t=1}^T \frac{4R^2}{\alpha (t+1)} \\
 &\leq \frac{4R^2\log(T+1)}{\alpha},
\end{align*}
which completes the proof.
\end{proof}
\begin{proposition}[Convergence of Method 2]
Consider the updates $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$ 
and $\theta_{t+1} \in \arg\max_{\theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$. 
Then for $\hat{u} = \frac{1}{T} \sum_{t=1}^T u_t$, we have 
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{T} \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_{t} \rangle. \]
\end{proposition}
\begin{proof}
If we invoke Lemma~\ref{lem:ftl} with $l_t(\theta) = -L(u_t, \theta)$, then we get the inequality
\[ \sum_{t=1}^T -L(u_t, \theta_t) \leq -\sum_{t=1}^T L(u_t, \theta^*) - \sum_{t=1}^T \langle u_t, \theta_t - \theta_{t+1} \rangle. \]
Re-arranging yields
\[ \sum_{t=1}^T L(u_t, \theta^*) \leq \sum_{t=1}^T L(u_t, \theta_t) + \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_t \rangle. \]
Now, we have the following string of inequalities:
\begin{align*}
L(\hat{u}, \theta) &= L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) \\
 &\leq \frac{1}{T} \sum_{t=1}^T L(u_t, \theta) \\
 &\leq \frac{1}{T} \sum_{t=1}^T L(u_t, \theta_t) + \frac{1}{T} \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_t \rangle \\
 &= \frac{1}{T} \sum_{t=1}^T \inf_{u} L(u, \theta_t) + \frac{1}{T} \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_t \rangle \\
 &\leq \frac{1}{T} \sum_{t=1}^T L(u^*, \theta_t) + \frac{1}{T} \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_t \rangle \\
 &\leq \frac{1}{T} \sum_{t=1}^T \sup_{\theta} L(u^*, \theta) + \frac{1}{T} \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_t \rangle \\
 &= \sup_{\theta} L(u^*, \theta) + \frac{1}{T} \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_t \rangle,
\end{align*}
as was to be shown.
\end{proof}
\begin{corollary}
Suppose that $R$ is $\alpha$-strongly convex and let $R = \sup_{u} |\langle u, u \rangle|$. 
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{4R^2\log(T+1)}{\alpha T}. \]
\end{corollary}
\begin{proof}
The proof is identical to the previous corollary.
\end{proof}

\section{Application to Herding}
\label{sec:herding}
Suppose we are given a family of features $\phi$ together 
with a known value $\bar{\phi}$ for $\bE_{\mu}[\phi]$. We 
would like to construct a distribution that approximately 
matches this distribution, i.e. one for which 
$\bE_{\hat{\mu}}[\phi] \approx \bar{\phi}$. A natural way 
to do this is by using boosted mirror descent to minimize the 
maximum mean discrepancy relative to $\phi$; 
in other words, to minimize
\[ L(\mu) = \max\limits_{i=1}^d |\bE_{\mu}[\phi_i]-\bar{\phi}_i|. \]
If we wish to write this more similarly to the original problem 
formulation, we can write
\[ L(\mu) = \sup_{\|\theta\|_1 \leq 1} \theta^T(\bE_{\mu}[\phi]-\bar{\phi}). \]
One issue is that $\|\theta\|_1$ is not strongly convex, so we relax $\|\theta\|_1$ 
to $\|\theta\|_2$ and write our loss function in the Lagrangian form:
\[ L(\mu, \theta) = \theta^T\bE_{\mu}[\phi] - \theta^T\bar{\phi} - \frac{1}{2}\|\theta\|_2^2. \]
Intriguingly, Method 1 and Method 2 are identical to each other for this choice of $L$ 
(up to a change of index). In this case, $\mu_t = \delta_{x_t}$ for some $x_t$, and 
the updates are the same as the standard herding updates. Let $R = \sup_{x} \|\phi(x)\|_2$; 
then using the Method 2 convergence bound gives us
\begin{align*}
\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2R^2\log(T+1)}{T} \\
 &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{2R^2\log(T+1)}{T} \\
 &= \frac{2R^2\log(T+1)}{T}.
\end{align*}
This is a slightly weaker version of the typical herding bound on MMD.

\section{Herding in Infinite Dimensions}
\label{sec:infinite-case}

\section{Herding and Maximum Entropy}
\label{sec:max-ent}

\section{Conclusion}
\label{sec:conclusion}

%\bibliography{example_paper}
%\bibliographystyle{icml2013}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was
% created by Lise Getoor and Tobias Scheffer, it was slightly modified  
% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz, 
% slightly modified from the 2009 version by Kiri Wagstaff and 
% Sam Roweis's 2008 version, which is slightly modified from 
% Prasad Tadepalli's 2007 version which is a lightly 
% changed version of the previous year's version by Andrew Moore, 
% which was in turn edited from those of Kristian Kersting and 
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2013-09-07 icml/paper.tex
made few small edits and added a couple comments

@@
 Consider the following loss function:
-\[ L(u) = \max_{\theta \in \Theta} [\langle \theta, u \rangle - R(\theta)] \]
-$R$ is some convex function that we should think of as a measure of complexity 
+\[ L(u) = \max_{\theta \in \Theta} \left\{\langle \theta, u \rangle - R(\theta) \right\}, \]
+where $R$ is some convex function that we should think of as a measure of complexity 
 of $\theta$. Thus, the loss is equal to the maximum amount that $\theta$ can match 
@@
 \begin{itemize}
-\item Let $R(\theta) = \langle \theta, u_0 \rangle$. Then $L(u)$ is the 
+\item Let $R(\theta) = \langle \theta, u_0 \rangle$. Then, assuming $-\Theta = \Theta$, $L(u)$ is the 
       \emph{maximum mean discrepancy} between $u$ and $u_0$ 
@@
 considering $\max_{\theta \in \Theta} L(u,\theta)$. We will assume 
-throughout this document that $h$ and $R$ are both convex functions.
+throughout this document that $h$ and $R$ are both convex functions. \NA{JH: Should we comment on the symmetry introduced by this formulation?}
 
@@
 bound:
-\[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/\sqrt{T}). \]
+\[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(\log(T)/T). \]
 However, it is often the case that $h$ is not strongly convex, whereas $R$ is strongly convex. In this case 
@@
 only requires the convexity of $R$:
-\[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/\sqrt{T}). \]
+\[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(\log(T)/T). \]
+\NA{JH: It would be more elegant to argue either here or below that the bounds on the second algorithm follow directly from duality}
 
@@
 would like to construct a distribution that approximately 
-matches this distribution, i.e. one for which 
+matches this distribution, i.e.~one for which 
 $\bE_{\hat{\mu}}[\phi] \approx \bar{\phi}$. A natural way 
@@
 in other words, to minimize
-\[ L(\mu) = \max\limits_{i=1}^d |\bE_{\mu}[\phi_i]-\bar{\phi}_i|. \]
+\[ L(\mu) = \max_{i \in \{1,\dots,d\}} |\bE_{\mu}[\phi_i]-\bar{\phi}_i|. \]
 If we wish to write this more similarly to the original problem 

### 2013-09-09 icml/paper.tex
added max-ent and part of the weighted averaging argument

--- 2013-09-07 icml/paper.tex
@@
 Consider the following loss function:
-\[ L(u) = \max_{\theta \in \Theta} [\langle \theta, u \rangle - R(\theta)] \]
-$R$ is some convex function that we should think of as a measure of complexity 
-of $\theta$. Thus, the loss is equal to the maximum amount that $\theta$ can match 
-$u$, regularized by the complexity function $R(\theta)$.
+\begin{equation}
+L(u) = h(u) + \max_{\theta \in \Theta} [\langle \theta, u \rangle - R(\theta)]
+\end{equation}
+We can think of $h$ as a \emph{primal regularizer} and $R$ as a \emph{dual regularizer}.
 
@@
 \begin{itemize}
-\item Let $R(\theta) = \langle \theta, u_0 \rangle$. Then $L(u)$ is the 
+\item Let $h(u) = 0$ and $R(\theta) = \langle \theta, u_0 \rangle$. Then $L(u)$ is the 
       \emph{maximum mean discrepancy} between $u$ and $u_0$ 
       relative to $\Theta$.
-\item Let $R(\theta) = S^*(\theta)$ where $S^*$ is the Fenchel conjugate of 
+\item Let $h(u) = 0$ and $R(\theta) = S^*(\theta)$ where $S^*$ is the Fenchel conjugate of 
       any strongly convex function $S$. Then $L(u) = S(u)$.
+\item Let $h(u) = \|u\|_1$, $R(u) = \langle \theta, u_0 \rangle + \frac{1}{2} \|\theta\|_2^2$. 
+      Then $L(u) = \|u\|_1 + \frac{1}{2} \|u-u_0\|_2^2$. 
 \end{itemize}
 
-In this document we consider a family of methods for minimizing $L(u)$, 
+In this paper we will consider a family of methods for minimizing $L(u)$, 
 which are based on Bach's generalization of the Frank-Wolfe algorithm and can 
@@
 to consider a \emph{two-argument} loss function
-\[ L(u,\theta) = h(u) + \langle \theta, u \rangle - R(\theta). \]
-In the case $h(u) \equiv 0$, we can get the previous setting by 
-considering $\max_{\theta \in \Theta} L(u,\theta)$. We will assume 
-throughout this document that $h$ and $R$ are both convex functions.
+\begin{equation}
+L(u,\theta) = h(u) + \langle \theta, u \rangle - R(\theta).
+\end{equation}
+We will assume throughout that $h$ and $R$ are both convex functions, 
+and furthermore that $\arg\max_{u} h(u) + \theta^Tu$ can be efficiently 
+computed for all values of $\theta$.
 
-Boosted mirror descent can be thought of as an algorithm for finding 
-``saddle points'' of $L$. It is the following algorithm:
+Mirror descent, together with boosting, yields an algorithm for finding 
+``saddle points'' of $L$. It is the following algorithm (called Algorithm 1).
+For notational convenience, for a sequence of weights $\alpha_1,\alpha_2,\ldots$ 
+let $\hat{u}_t = \frac{\sum_{s=1}^t \alpha_su_s}{\sum_{s=1}^t \alpha_s}$ and let 
+$\hat{\theta}_t = \frac{\sum_{s=1}^t \alpha_s\theta_s}{\sum_{s=1}^t \alpha_s}$.
 \begin{enumerate}
-\item $u_1 = 0$
+\item $u_1 \in \arg\min_u h(u)$
 \item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, u_t \rangle - R(\theta)$
-\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
+\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, u \rangle$
+%\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
 \end{enumerate}
-Let $\hat{u} = \frac{1}{T} \sum_{t = 1}^{T} u_t$. As long as $h$ is strongly convex, we obtain the 
-bound:
-\[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/\sqrt{T}). \]
+As long as $h$ is strongly convex, we obtain the 
+bound (see Proposition~\ref{cor:method-1}):
+\begin{equation}
+\sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/T).
+\end{equation}
+In other words, $\hat{u}_T$ is close to being a global minimum of $L$.
 However, it is often the case that $h$ is not strongly convex, whereas $R$ is strongly convex. In this case 
-we may wish to use the following slightly different algorithm:
+we may wish to use the following slightly different algorithm (called Algorithm 2):
 \begin{enumerate}
-\item $\theta_1 = 0$
+\item $\theta_1 \in \arg\min_{\theta} R(\theta)$
 \item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$
-\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
+%\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
+\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \hat{u}_t \rangle - R(\theta)$
 \end{enumerate}
-Once again, if we let $\hat{u} = \frac{1}{T} \sum_{t=1}^T u_t$, then we obtain the following bound, which now 
-only requires the convexity of $R$:
-\[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/\sqrt{T}). \]
+We then obtain the following bound 
+when $R$ is strongly convex (see Proposition~\ref{cor:method-2}):
+\[ \sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/T). \]
+Algorithm 1 and Algorithm 2 are closely related; indeed, they are dual to each other 
+(performing Algorithm 1 on $L(u,\theta)$ is the same as performing Algorithm 2 on 
+$-L(\theta,u)$).
 
@@
 \label{sec:proofs}
-We now prove the convergence results cited in Section~\ref{sec:intro}.
+We now prove the convergence results cited in Section~\ref{sec:algorithm}. 
+We start with two preliminary lemmas.
 \begin{lemma}
@@
 \end{proof}
-\begin{proposition}[Convergence of Method 1]
+\begin{proposition}[Convergence of Algorithm 1]
+\label{prop:method-1}
 Consider the updates $\theta_t \in \arg\max_{\theta} \langle \theta, u_t \rangle - R(\theta)$ 
-and $u_{t+1} \in \arg\min_u h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$. 
-Then for $\hat{u} = \frac{1}{T} \sum_{t=1}^T u_t$, we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{T} \sum_{t=1}^T \langle u_t - u_{t+1}, \theta_t \rangle. \]
+and $u_{t+1} \in \arg\min_u h(u) + \langle \hat{\theta}_s, u \rangle$. 
+Then we have
+\[ \sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{\sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle}{\sum_{t=1}^T \alpha_t}. \]
 \end{proposition}
@@
 Note that $L(u_t, \theta_t) = \max_{\theta} L(u_t, \theta)$ by construction. Also note that, if we invoke 
-Lemma~\ref{lem:ftl} with $l_t(u) = L(u, \theta_t)$, then we get the inequality
-\[ \sum_{t=1}^T L(u_t, \theta_t) \leq \sum_{t=1}^T L(u^*, \theta_t) + \sum_{t=1}^T \langle u_t - u_{t+1}, \theta_t \rangle. \]
+Lemma~\ref{lem:ftl} with $l_t(u) = \alpha_t L(u, \theta_t)$, then we get the inequality
+\[ \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \leq \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle. \]
 Combining these together, we get the string of inequalities
 \begin{align*}
-L(\hat{u}, \theta) &= L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) \\
- &\leq \frac{1}{T} \sum_{t=1}^T L(u_t, \theta) \\
- &\leq \frac{1}{T} \sum_{t=1}^T L(u_t, \theta_t) \\
- &\leq \frac{1}{T} \sum_{t=1}^T L(u^*, \theta_t) + \frac{1}{T} \sum_{t=1}^T \langle u_t - u_{t+1}, \theta_t \rangle \\
- &\leq \frac{1}{T} \sum_{t=1}^T \sup_{\theta} L(u^*, \theta) + \frac{1}{T} \sum_{t=1}^T \langle u_t - u_{t+1}, \theta_t \rangle \\
- &= \sup_{\theta} L(u^*, \theta) + \frac{1}{T} \sum_{t=1}^T \langle u_t - u_{t+1}, \theta_t \rangle,
+L(\hat{u}_T, \theta) &= L\left(\frac{\sum_{t=1}^T \alpha_tu_t}{\sum_{t=1}^T \alpha_t}, \theta\right) \\
+ &\leq \frac{\sum_{t=1}^T \alpha_t L(u_t, \theta)}{\sum_{t=1}^T \alpha_t} \\
+ &\leq \frac{\sum_{t=1}^T \alpha_t L(u_t, \theta_t)}{\sum_{t=1}^T \alpha_t} \\
+ &\leq \frac{\sum_{t=1}^T \alpha_tL(u^*, \theta_t)}{\sum_{t=1}^T \alpha_t} \\ &\phantom{=} + \frac{\sum_{t=1}^T \alpha_t (L(u_t,\theta_t)-L(u_{t+1},\theta_t))}{\sum_{t=1}^T \alpha_t} \\
+ &= L(u^*, \theta_t) + \frac{\sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle)}{\sum_{t=1}^T \alpha_t} \\ &\phantom{=} + \frac{\sum_{t=1}^{T} \alpha_t(h(u_t)-h(u_{t+1}))}{\sum_{t=1}^T \alpha_t} \\
+ &= L(u^*, \theta_t) + \frac{\sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle}{\sum_{t=1}^T \alpha_t} \\ &\phantom{=} + \frac{\alpha_1 h(u_1) - \alpha_T h(u_{T+1}) + \sum_{t=2}^{T} (\alpha_t-\alpha_{t-1}) h(u_t)}{\sum_{t=1}^T \alpha_t} \\
+ &\leq L(u^*, \theta_t) + \frac{\sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle}{\sum_{t=1}^T \alpha_t} \\
+ &\leq \sup_{\theta} L(u^*, \theta) + \frac{\sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle}{\sum_{t=1}^T \alpha_t},
 \end{align*}
@@
 \begin{corollary}
-Suppose that $h$ is $\alpha$-strongly convex and let $R = \sup_{\theta} |\langle \theta, \theta \rangle|$. Then 
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{4R^2\log(T+1)}{\alpha T}. \]
+\label{cor:method-1}
+Suppose that $h$ is $\gamma$-strongly convex and let $r = \sup_{\theta} |\langle \theta, \theta \rangle|_2$. Then 
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{16r^2}{\gamma T^2}. \]
 \end{corollary}
@@
 \begin{align*}
-\|u_t-u_{t+1}\| &\leq \frac{2}{\alpha}\|\frac{1}{t}\sum_{s \leq t}\theta_s - \frac{1}{t+1}\sum_{s \leq t+1}\theta_s\| \\
- &= \frac{2}{\alpha}\|\frac{1}{t(t+1)} \sum_{s \leq t} \theta_s - \frac{1}{t+1} \theta_{t+1}\| \\
- &\leq \frac{2}{\alpha} \left(\frac{1}{t(t+1)} \sum_{s \leq t} \|\theta_s\| + \frac{1}{t+1} \|\theta_{t+1}\|\right) \\
- &\leq \frac{4R}{\alpha (t+1)}.
+\|u_t-u_{t+1}\| &\leq \frac{2}{\gamma}\|\frac{\sum_{s \leq t}\alpha_s\theta_s}{\sum_{s \leq t} \alpha_s} - \frac{\sum_{s \leq t+1}\alpha_s\theta_s}{\sum_{s \leq t+1} \alpha_s}\| \\
+ &= \frac{2}{\gamma}\|\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\theta_s - \frac{\alpha_{t+1}}{A_{t+1}} \theta_{t+1}\| \\
+ &\leq \frac{2}{\gamma} \left(\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\|\theta_s\| + \frac{\alpha_{t+1}}{A_{t+1}} \|\theta_{t+1}\|\right) \\
+ &\leq \frac{4r\alpha_{t+1}}{\gamma A_{t+1}}.
 \end{align*}
@@
 \begin{align*}
-\sum_{t=1}^T \langle u_t - u_{t+1}, \theta_t \rangle &\leq \sum_{t=1}^T \frac{4R^2}{\alpha (t+1)} \\
- &\leq \frac{4R^2\log(T+1)}{\alpha},
+\frac{1}{A_T} \sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle &\leq \sum_{t=1}^T \frac{4r^2\alpha_{t}\alpha_{t+1}}{\gamma A_{t+1}A_T} \\
+ &\leq \frac{4r^2}{\gamma A_T} \sum_{t=1}^T \frac{\alpha_{t}\alpha_{t+1}}{A_{t+1}}.
 \end{align*}
+In particular, if we let $\alpha_t = t$, then $A_t = \frac{t(t+1)}{2}$ and $\frac{\alpha_{t}\alpha_{t+1}}{A_{t+1}} = 2$. 
+We therefore get
+\begin{equation}
+\frac{1}{A_T} \sum_{t=1}^T \alpha_t\langle u_t-u_{t+1}, \theta_t \rangle \leq \frac{16r^2}{T(T+1)},
+\end{equation}
 which completes the proof.
 \end{proof}
-\begin{proposition}[Convergence of Method 2]
+\begin{proposition}[Convergence of Algorithm 2]
+\label{prop:method-2}
 Consider the updates $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$ 
@@
 \begin{corollary}
+\label{cor:method-2}
 Suppose that $R$ is $\alpha$-strongly convex and let $R = \sup_{u} |\langle u, u \rangle|$. 
@@
 \[ L(\mu, \theta) = \theta^T\bE_{\mu}[\phi] - \theta^T\bar{\phi} - \frac{1}{2}\|\theta\|_2^2. \]
-Intriguingly, Method 1 and Method 2 are identical to each other for this choice of $L$ 
+Intriguingly, Algorithm 1 and Algorithm 2 are identical to each other for this choice of $L$ 
 (up to a change of index). In this case, $\mu_t = \delta_{x_t}$ for some $x_t$, and 
 the updates are the same as the standard herding updates. Let $R = \sup_{x} \|\phi(x)\|_2$; 
-then using the Method 2 convergence bound gives us
+then using the Algorithm 2 convergence bound gives us
 \begin{align*}
@@
 \label{sec:max-ent}
+
+Let us now consider the entropic regularizer $h(\mu) = \bE_{\mu}[\log \mu(x)]$ with $R(\theta) = \bE_{\mu}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.$.
+If we use Method 1, we then end up with 
+\[ \mu_i(x) \propto \exp(\frac{1}{i}(\sum_{j<i} \theta_j)^T\phi(x)), \]
+so $h(\mu) + \bE_{\mu}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the 
+boundary. We can instead consider the following sequence of distributions:
+\begin{enumerate}
+\item $\theta_i \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{\mu_i}[\phi(x)]-\bar{\phi})$
+\item $\mu_{i+1} \propto \exp(\frac{1}{i}(\sum_{j \leq i} \theta_j)^T\phi(x))$
+\end{enumerate}
+The entropic regularizer is not strongly convex but a more general analysis involving 
+Bregman divergence can nevertheless yield a convergence bound in this case. We thus 
+get something that looks like herding but actually gives us an exponential family 
+type distribution.
 

### 2013-09-09 icml/paper.tex
added entropy material and discussion; checked proofs of infinite dimensional issues (none found); added a bunch of comments

--- 2013-09-07 icml/paper.tex
@@
 
+\NA{JHH: I'm thinking that from the beginning we should let $u$ be defined over a different space than $\theta$ and assume a linear map $u \mapsto \eta(u) \in \Theta$.
+Then write $\langle \theta, \eta(u) \rangle$ so the setup will align with the herding case, where $\eta(u) = \bE_{u}[\phi(x)]$.
+}
+
 Consider the following loss function:
@@
 
-In this document we consider a family of methods for minimizing $L(u)$, 
+In this paper we consider a family of methods for minimizing $L(u)$, 
 which are based on Bach's generalization of the Frank-Wolfe algorithm and can 
@@
 considering $\max_{\theta \in \Theta} L(u,\theta)$. We will assume 
-throughout this document that $h$ and $R$ are both convex functions. \NA{JH: Should we comment on the symmetry introduced by this formulation?}
+throughout this document that $h$ and $R$ are both convex functions. \NA{JHH: Should we comment on the symmetry introduced by this formulation?}
 
@@
 \begin{enumerate}
-\item $u_1 = 0$
-\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, u_t \rangle - R(\theta)$
-\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
+\item $u_1 = \argmin_{u} h(u)$
+\item $\theta_{t} \in \argmax_{\theta \in \Theta} \langle \theta, u_t \rangle - R(\theta)$
+\item $u_{t+1} \in \argmin_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
 \end{enumerate}
@@
 \item $\theta_1 = 0$
-\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$
-\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
+\item $u_t \in \argmin_{u} h(u) + \langle \theta_t, u \rangle$
+\item $\theta_{t+1} \in \argmax_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
 \end{enumerate}
@@
 \[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(\log(T)/T). \]
-\NA{JH: It would be more elegant to argue either here or below that the bounds on the second algorithm follow directly from duality}
+\NA{JHH: It would be more elegant to argue either here or below that the bounds on the second algorithm follow directly from duality}
 
@@
 Let $l_1,\ldots,l_T$ be functions of $x$ and let $x_1,\ldots,x_T$ 
-be chosen via $x_{t+1} \in \arg\min_{x} \frac{1}{t} \sum_{s \leq t} l_s(x)$. 
+be chosen via $x_{t+1} \in \argmin_{x} \frac{1}{t} \sum_{s \leq t} l_s(x)$. 
 Then for any $x^*$ we have
@@
 \begin{proof}
+\NA{JHH: Doesn't rely on the spaces $x$'s having any structure nor on any properties of the functions.}
 Re-arranging, we want to show that $\sum_{t=1}^T l_t(x_{t+1}) \leq \sum_{t=1}^T l_t(x^*)$ for 
 all $x^*$. We will show this by induction. In the base case $T=1$, we want to show that $l_1(x_2) \leq l_1(x^*)$ 
-for all $x^*$, which is true because $x_2 \in \arg\min_x l_1(x)$ by construction. For 
+for all $x^*$, which is true because $x_2 \in \argmin_x l_1(x)$ by construction. For 
 the inductive step, suppose that the result holds for $T-1$:
@@
 \end{proof}
-\begin{lemma}
-\label{lem:convexity}
+\begin{lemma} \label{lem:convexity}
 Suppose that $f(x)$ is \emph{$\alpha$-strongly convex}: $f(x) \geq f(x_0) + \partial f(x_0)^T(x-x_0) + \frac{\alpha}{2}\|x-x_0\|^2$. 
-Let $F(y) := \arg\min_x f(x) + y^Tx$. Then $\|F(y_0)-F(y_1)\| \leq \frac{2}{\alpha} \|y_0-y_1\|$.
+Let $F(y) := \argmin_x f(x) + y^Tx$. Then $\|F(y_0)-F(y_1)\| \leq \frac{2}{\alpha} \|y_0-y_1\|$.
 \end{lemma}
 \begin{proof}
+\NA{JHH: Follows from $f$ being strongly convex, the definition of inner product, and Cauchy-Schwartz. We'll just want to be careful about calculating $\alpha$ for our regularizer later on.}
 Let $x_0 = F(y_0)$. Then we must have $\partial f(x_0) = -y_0$, hence 
@@
 \begin{proposition}[Convergence of Method 1]
-Consider the updates $\theta_t \in \arg\max_{\theta} \langle \theta, u_t \rangle - R(\theta)$ 
-and $u_{t+1} \in \arg\min_u h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$. 
+Consider the updates $\theta_t \in \argmax_{\theta} \langle \theta, u_t \rangle - R(\theta)$ 
+and $u_{t+1} \in \argmin_u h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$. 
 Then for $\hat{u} = \frac{1}{T} \sum_{t=1}^T u_t$, we have
@@
 \begin{proof} 
+\NA{JHH: Uses, Lemma~\ref{lem:ftl}, convexity of $h$, taking supremums, and facts that follow from the construction of the algorithm.}
 Note that $L(u_t, \theta_t) = \max_{\theta} L(u_t, \theta)$ by construction. Also note that, if we invoke 
@@
 L(\hat{u}, \theta) &= L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) \\
- &\leq \frac{1}{T} \sum_{t=1}^T L(u_t, \theta) \\
+ &\leq \frac{1}{T} \sum_{t=1}^T L(u_t, \theta) & \\
  &\leq \frac{1}{T} \sum_{t=1}^T L(u_t, \theta_t) \\
@@
 \begin{proof}
+\NA{JHH: Follows from Lemma~\ref{lem:convexity}, triangle inequality, taking supremums, and Cauchy-Schwartz.}
 By Lemma~\ref{lem:convexity}, we have 
@@
 \begin{proposition}[Convergence of Method 2]
-Consider the updates $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$ 
-and $\theta_{t+1} \in \arg\max_{\theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$. 
+Consider the updates $u_t \in \argmin_{u} h(u) + \langle \theta_t, u \rangle$ 
+and $\theta_{t+1} \in \argmax_{\theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$. 
 Then for $\hat{u} = \frac{1}{T} \sum_{t=1}^T u_t$, we have 
@@
 \label{sec:herding}
+
+\NA{JHH: Need to specify the spaces we are working with here.}
 Suppose we are given a family of features $\phi$ together 
@@
 
+Let us now consider the entropic regularizer $h(u) = \tau\bE_{u}[\log u(x)]$ with 
+\[ R(\theta) = \bE_{u}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.\]
+If we use Method 1, we then end up with 
+\[ u_{t}(x) \propto \exp\left\{-\frac{1}{\alpha}\overline{\theta}_{t}^T\phi(x)\right\}, \]
+where $\overline{\theta}_{t} = \frac{1}{t} \sum_{s \le t} \theta_{s}$.
+Hence, $h(u) + \bE_{\mu}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the boundary (when $\tau > 0$).
+We can instead consider the following sequence of distributions:
+\begin{enumerate}
+\item $\theta_t \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{u_t}[\phi(x)]-\bar{\phi})$
+\item $u_{t+1}(x) = Z(\overline{\theta}_{t}, \tau)^{-1} \exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\}$
+\end{enumerate}
+Here $Z(\overline{\theta}_{t}, \tau) = \int\exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\} \dee x$.
+
+The entropic regularizer is not strongly convex but a more general analysis involving 
+Bregman divergence can nevertheless yield a convergence bound in this case. We thus 
+get something that looks like herding but actually gives us an exponential family 
+type distribution. In particular, we have a Gibbs distribution with temperature $\tau$, and 
+in the limit as the temperature goes to zero, we recover herding, since 
+\[ \lim_{\tau \to 0} u_{t+1}(x) = \delta_{x^{*}_{t+1}}, \]
+where $x^{*}_{t+1} = \argmin_{x} \overline{\theta}_{t}^T\phi(x)$. 
+This fact is reminiscent of Welling's original derivation of herding, in which 
+he took the zero temperature limit of the gradient descent algorithm for the optimization problem 
+\[ \argmin_{u} h(u) \quad \text{s.t.}~\bE_{u}[\phi(x)] - \bar\phi = 0. \]
+This optimization is quite similar to the entropic optimization considered here, which 
+we can write as 
+\begin{align*}
+\argmin_{u}& \argmax_{|\theta\|_1 \leq 1} h(u) + \theta^{T}(\bE_{u}[\phi(x)] - \bar \phi)  \\
+&= \argmin_{u} h(u) + \|\bE_{u}[\phi(x)] - \bar\phi\|_{2}. 
+\end{align*}
+In the entropic formulation, the expectation constraint $\bE_{u}[\phi(x)] - \bar\phi = 0$ becomes
+soft, though the two optimizations are equivalent in the zero-temperature limit.
+
+We can view the entropic regularization
+with non-zero temperature as an alternative approximation to herding. If at each step we
+retain a sample $x_{t} \sim u_{t}$, then the sequence $(x_{t})_{t}$ converges to the 
+herding sample path as $\tau \to 0$. So $(x_{t})_{t}$ with non-zero temperature approximate
+herding samples. \NA{JHH: I'm not sure how useful this all is, other than to make contact with 
+the original paper, as I think the Gibbs distribution was exactly what Max was trying to 
+avoid by taking the zero-temperature limit.}
+
 \section{Conclusion}

### 2013-09-10 icml/paper.tex
changed format to unix from dos

--- 2013-09-09 icml/paper.tex
@@
 
+\NA{JHH: I'm thinking that from the beginning we should let $u$ be defined over a different space than $\theta$ and assume a linear map $u \mapsto \eta(u) \in \Theta$.
+Then write $\langle \theta, \eta(u) \rangle$ so the setup will align with the herding case, where $\eta(u) = \bE_{u}[\phi(x)]$.
+}
+
 Consider the following loss function:
 \begin{equation}
-L(u) = h(u) + \max_{\theta \in \Theta} [\langle \theta, u \rangle - R(\theta)]
+L(u) = h(u) + \max_{\theta \in \Theta} \left\{\langle \theta, u \rangle - R(\theta) \right\}.
 \end{equation}
@@
 
+\NA{JHH: Should we comment on the symmetry introduced by this formulation?}
 First, let us generalize the setting 
@@
 $-L(\theta,u)$).
+\NA{JHH: It would be more elegant to argue either here or below that the bounds on the second algorithm follow directly from duality}
 
@@
 \begin{proof}
+\NA{JHH: Follows from $f$ being strongly convex, the definition of inner product, and Cauchy-Schwartz. We'll just want to be careful about calculating $\alpha$ for our regularizer later on.}
 Let $x_0 = F(y_0)$. Then we must have $\partial f(x_0) = -y_0$, hence 
@@
 \begin{proof}
+\NA{JHH: Uses, Lemma~\ref{lem:ftl}, convexity of $h$, taking supremums, and facts that follow from the construction of the algorithm.}
 Note that $L(u_t, \theta_t) = \max_{\theta} L(u_t, \theta)$ by construction. Also note that, if we invoke 
@@
 \label{cor:method-2}
-Suppose that $R$ is $\alpha$-strongly convex and let $R = \sup_{u} |\langle u, u \rangle|$. 
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{4R^2\log(T+1)}{\alpha T}. \]
+Suppose that $R$ is $\gamma$-strongly convex and let $r = \sup_{u} |\langle u, u \rangle|$. Then
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{16r^2}{\gamma T^2}. \]
 \end{corollary}
@@
 (up to a change of index). In this case, $\mu_t = \delta_{x_t}$ for some $x_t$, and 
-the updates are the same as the standard herding updates. Let $R = \sup_{x} \|\phi(x)\|_2$; 
+the updates are the same as the standard herding updates. Let $r = \sup_{x} \|\phi(x)\|_2$; 
 then using the Algorithm 2 convergence bound gives us
@@
 
-Let us now consider the entropic regularizer $h(\mu) = \bE_{\mu}[\log \mu(x)]$ with $R(\theta) = \bE_{\mu}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.$.
+Let us now consider the entropic regularizer $h(u) = \tau\bE_{u}[\log u(x)]$ with 
+\[ R(\theta) = \bE_{u}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.\]
 If we use Method 1, we then end up with 
-\[ \mu_i(x) \propto \exp(\frac{1}{i}(\sum_{j<i} \theta_j)^T\phi(x)), \]
-so $h(\mu) + \bE_{\mu}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the 
-boundary. We can instead consider the following sequence of distributions:
+\[ u_{t}(x) \propto \exp\left\{-\frac{1}{\alpha}\overline{\theta}_{t}^T\phi(x)\right\}, \]
+where $\overline{\theta}_{t} = \frac{1}{t} \sum_{s \le t} \theta_{s}$.
+Hence, $h(u) + \bE_{\mu}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the boundary (when $\tau > 0$).
+We can instead consider the following sequence of distributions:
 \begin{enumerate}
-\item $\theta_i \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{\mu_i}[\phi(x)]-\bar{\phi})$
-\item $\mu_{i+1} \propto \exp(\frac{1}{i}(\sum_{j \leq i} \theta_j)^T\phi(x))$
+\item $\theta_t \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{u_t}[\phi(x)]-\bar{\phi})$
+\item $u_{t+1}(x) = Z(\overline{\theta}_{t}, \tau)^{-1} \exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\}$
 \end{enumerate}
+Here $Z(\overline{\theta}_{t}, \tau) = \int\exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\} \dee x$.
+
 The entropic regularizer is not strongly convex but a more general analysis involving 
@@
 get something that looks like herding but actually gives us an exponential family 
-type distribution.
+type distribution. In particular, we have a Gibbs distribution with temperature $\tau$, and 
+in the limit as the temperature goes to zero, we recover herding, since 
+\[ \lim_{\tau \to 0} u_{t+1}(x) = \delta_{x^{*}_{t+1}}, \]
+where $x^{*}_{t+1} = \argmin_{x} \overline{\theta}_{t}^T\phi(x)$. 
+This fact is reminiscent of Welling's original derivation of herding, in which 
+he took the zero temperature limit of the gradient descent algorithm for the optimization problem 
+\[ \argmin_{u} h(u) \quad \text{s.t.}~\bE_{u}[\phi(x)] - \bar\phi = 0. \]
+This optimization is quite similar to the entropic optimization considered here, which 
+we can write as 
+\begin{align*}
+\argmin_{u}& \argmax_{|\theta\|_1 \leq 1} h(u) + \theta^{T}(\bE_{u}[\phi(x)] - \bar \phi)  \\
+&= \argmin_{u} h(u) + \|\bE_{u}[\phi(x)] - \bar\phi\|_{2}. 
+\end{align*}
+In the entropic formulation, the expectation constraint $\bE_{u}[\phi(x)] - \bar\phi = 0$ becomes
+soft, though the two optimizations are equivalent in the zero-temperature limit.
+
+We can view the entropic regularization
+with non-zero temperature as an alternative approximation to herding. If at each step we
+retain a sample $x_{t} \sim u_{t}$, then the sequence $(x_{t})_{t}$ converges to the 
+herding sample path as $\tau \to 0$. So $(x_{t})_{t}$ with non-zero temperature approximate
+herding samples. \NA{JHH: I'm not sure how useful this all is, other than to make contact with 
+the original paper, as I think the Gibbs distribution was exactly what Max was trying to 
+avoid by taking the zero-temperature limit.}
 

### 2013-09-11 icml/proof.tex
added key lemma for weighted case

\documentclass{article}
\usepackage{import}
\input{latex-defs.tex}

\begin{document}

Throughout this section, assume that $\alpha_1,\ldots,\alpha_T$ is 
a sequence of real numbers and that $A_t = \sum_{s=1}^t \alpha_s$. 
We further require that $A_t \geq 0$ for all $t$.

Also recall that the Bregman divergence is defined by $D_f(x_2 \| x_1) \eqdef f(x_2) - \langle \nabla f(x_1), x_2-x_1 \rangle - f(x_1)$.

A key observation in the following proof is that, if $x_0 = \arg\min f(x) + \langle z, x \rangle$, then $\nabla f(x_0) = -z$.

\NA{JNS: In the following proof, (3) is by the update formula for $x_{t+1}$ together with the preceding observation. 
(4) and (5) are re-arranging the sum. (6) is definition of Bregman divergence. (7) is the observation again. 
(8) is the definition of $x_{T+1}$.}

\begin{lemma}
\label{lem:ftl}
Let $z_1,\ldots,z_T$ be vectors and let $f(x)$ be a strictly convex 
function. Define $\hat{z}_t$ to be $\frac{1}{A_t} \sum_{s=1}^t \alpha_s z_s$.

Let $x_1,\ldots,x_T$ be chosen via $x_{t+1} = \arg\min_{x} f(x) + \langle \hat{z}_t, x\rangle$. 
Then for any $x^*$ we have
\begin{equation}
\frac{1}{A_T} \sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\} \leq f(x^*) + \langle \hat{z}_t, x^* \rangle + \frac{1}{A_T} \sum_{t=1}^T A_t D_{f}(x_t \| x_{t+1}). 
\end{equation}
\end{lemma}
\begin{proof}
We have
\begin{align}
\lefteqn{\sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
 &= \sum_{t=1}^T \{\alpha_t f(x_t) + \langle A_{t-1} \nabla f(x_t) - A_t \nabla f(x_{t+1}), x_t \rangle\} \\
 &= \sum_{t=1}^T \{\alpha_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle\} + A_T(f(x_{T+1})- \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \\
 &= \sum_{t=1}^T \{A_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle - A_t f(x_{t+1})\} + A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \\
 &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \\
 &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) + \langle \hat{z}_T, x_{T+1} \rangle) \\
 &\leq \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x^*) + \langle \hat{z}_T, x^* \rangle). \\
\end{align}
Dividing both sides by $A_T$ completes the proof.
\end{proof}
\end{document}

### 2013-09-12 icml/proof.tex
added remark after proof
small edits to proof of lemma and updated comment to reflect changes

@@
 
-\NA{JNS: In the following proof, (3) is by the update formula for $x_{t+1}$ together with the preceding observation. 
-(4) and (5) are re-arranging the sum. (6) is definition of Bregman divergence. (7) is the observation again. 
+\NA{JNS: In the following proof, (2) is by the update formula for $x_{t+1}$ together with the preceding observation.
+(3) is applying (2).
+(4) and (5) are re-arranging the sum.
+(6) is definition of Bregman divergence.
+(7) is the observation again.
 (8) is the definition of $x_{T+1}$.}
@@
 \begin{proof}
-We have
+Note that 
+\begin{align}
+\alpha_{t}z_{t} = A_{t}\hat z_{t} - A_{t-1}\hat z_{t-1} = - A_{t}\nabla f(x_{t+1}) + A_{t-1} \nabla f(x_{t}) 
+\end{align}
+so we have
 \begin{align}
@@
  &= \sum_{t=1}^T \{\alpha_t f(x_t) + \langle A_{t-1} \nabla f(x_t) - A_t \nabla f(x_{t+1}), x_t \rangle\} \\
- &= \sum_{t=1}^T \{\alpha_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle\} + A_T(f(x_{T+1})- \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \\
- &= \sum_{t=1}^T \{A_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle - A_t f(x_{t+1})\} + A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \\
+ &= \sum_{t=1}^T \{\alpha_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle\} - A_{T}\langle \nabla f(x_{T+1}), x_{T+1} \rangle \\
+ &= \sum_{t=1}^T \{A_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle - A_t f(x_{t+1})\}  \\
+ &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
  &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \\
@@
 \end{proof}
+We also note that $D_f(x_t \| x_{t+1}) = D_{f^*}(\hat{z}_{t+1} \| z_t)$, where $f^*(z) = \sup_x \langle z,x\rangle - f(x)$. 
+This form of the bound will often be more useful to us.
 \end{document}

### 2013-09-12 icml/paper.tex
added note
fixed proofs in section 3

@@
 \NA{JHH: It would be more elegant to argue either here or below that the bounds on the second algorithm follow directly from duality}
+\NA{JNS: The statements aren't actually dual to each other; the first is saying that the primal converges to its 
+    optimum, and the second is saying that the dual converges to its optimum.}
 
@@
 We now prove the convergence results cited in Section~\ref{sec:algorithm}. 
-We start with two preliminary lemmas.
+Throughout this section, assume that $\alpha_1,\ldots,\alpha_T$ is 
+a sequence of real numbers and that $A_t = \sum_{s=1}^t \alpha_s$. 
+We further require that $A_t \geq 0$ for all $t$.
+
+Also recall that the Bregman divergence is defined by 
+$D_f(x_2 \| x_1) \eqdef f(x_2) - \langle \nabla f(x_1), x_2-x_1 \rangle - f(x_1)$.
+
+Our proofs hinge on the following key lemma:
 \begin{lemma}
-\label{lem:ftl}
-Let $l_1,\ldots,l_T$ be functions of $x$ and let $x_1,\ldots,x_T$ 
-be chosen via $x_{t+1} \in \arg\min_{x} \frac{1}{t} \sum_{s \leq t} l_s(x)$. 
+\label{lem:bregman}
+Let $z_1,\ldots,z_T$ be vectors and let $f(x)$ be a strictly convex 
+function. Define $\hat{z}_t$ to be $\frac{1}{A_t} \sum_{s=1}^t \alpha_s z_s$.
+
+Let $x_1,\ldots,x_T$ be chosen via $x_{t+1} = \arg\min_{x} f(x) + \langle \hat{z}_t, x\rangle$. 
 Then for any $x^*$ we have
-\[ \sum_{t=1}^T l_t(x_t) \leq \sum_{t=1}^T l_t(x^*) + \sum_{t=1}^T (l_t(x_t) - l_t(x_{t+1})). \]
+\begin{align*}
+\lefteqn{\frac{1}{A_T} \sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
+\phantom{+} &\leq f(x^*) + \langle \hat{z}_t, x^* \rangle + \frac{1}{A_T} \sum_{t=1}^T A_t D_{f}(x_t \| x_{t+1}). 
+\end{align*}
 \end{lemma}
 \begin{proof}
-\NA{JHH: Doesn't rely on the spaces $x$'s having any structure nor on any properties of the functions.}
-Re-arranging, we want to show that $\sum_{t=1}^T l_t(x_{t+1}) \leq \sum_{t=1}^T l_t(x^*)$ for 
-all $x^*$. We will show this by induction. In the base case $T=1$, we want to show that $l_1(x_2) \leq l_1(x^*)$ 
-for all $x^*$, which is true because $x_2 \in \argmin_x l_1(x)$ by construction. For 
-the inductive step, suppose that the result holds for $T-1$:
-\[ \sum_{t=1}^{T-1} l_t(x_{t+1}) \leq \sum_{t=1}^T l_t(x^*). \]
-Now set $x^* = x_{T+1}$ in the above inequality and 
-add $l_T(x_{T+1})$ to both sides to get
-\begin{align*}
-\sum_{t=1}^T l_t(x_{t+1}) &\leq \sum_{t=1}^T l_t(x_{T+1}) \\
- &= \min_{x} \sum_{t=1}^T l_t(x) \\
- &\leq \sum_{t=1}^T l_t(x^*),
-\end{align*}
-which completes the induction. Here we are using the fact that $x_{T+1}$ 
-is the minimizer of $\frac{1}{T} \sum_{t=1}^T l_t(x)$ and hence also of 
-$\sum_{t=1}^T l_t(x)$.
+First note that, if $x_0 = \arg\min f(x) + \langle z, x \rangle$, 
+then $\nabla f(x_0) = -z$.
+
+\NA{JNS: In the following proof, (2) is by the update formula for $x_{t+1}$ together with the preceding observation.
+(3) is applying (2).
+(4) and (5) are re-arranging the sum.
+(6) is definition of Bregman divergence.
+(7) is the observation again.
+(8) is the definition of $x_{T+1}$.}
+
+Now note that
+\begin{align}
+\alpha_{t}z_{t} = A_{t}\hat z_{t} - A_{t-1}\hat z_{t-1} = - A_{t}\nabla f(x_{t+1}) + A_{t-1} \nabla f(x_{t}) 
+\end{align}
+so we have
+\begin{align}
+\lefteqn{\sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
+ &= \sum_{t=1}^T \{\alpha_t f(x_t) + \langle A_{t-1} \nabla f(x_t) - A_t \nabla f(x_{t+1}), x_t \rangle\} \\
+ &= \sum_{t=1}^T \{\alpha_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle\} - A_{T}\langle \nabla f(x_{T+1}), x_{T+1} \rangle \\
+ &= \sum_{t=1}^T \{A_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle - A_t f(x_{t+1})\}  \\
+ &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
+ &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \\
+ &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) + \langle \hat{z}_T, x_{T+1} \rangle) \\
+ &\leq \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x^*) + \langle \hat{z}_T, x^* \rangle). \\
+\end{align}
+Dividing both sides by $A_T$ completes the proof.
 \end{proof}
+We also note that $D_f(x_t \| x_{t+1}) = D_{f^*}(\hat{z}_{t+1} \| z_t)$, where $f^*(z) = \sup_x \langle z,x\rangle - f(x)$. 
+This form of the bound will often be more useful to us.
 \begin{lemma}
 \label{lem:convexity}
-Suppose that $f(x)$ is \emph{$\alpha$-strongly convex}: $f(x) \geq f(x_0) + \partial f(x_0)^T(x-x_0) + \frac{\alpha}{2}\|x-x_0\|^2$. 
-Let $F(y) := \arg\min_x f(x) + y^Tx$. Then $\|F(y_0)-F(y_1)\| \leq \frac{2}{\alpha} \|y_0-y_1\|$.
+Suppose that $D_f(x' \| x) \geq \frac{1}{2}\|x-x'\|^2$ for some 
+norm $\|\cdot\|$. (In this case we say that $f$ is strongly 
+convex with respect to $\|\cdot\|$.) 
+Then $D_{f^*}(x' \| x) \leq \frac{1}{2}\|x-x'\|_{*}^2$.
 \end{lemma}
-\begin{proof}
-\NA{JHH: Follows from $f$ being strongly convex, the definition of inner product, and Cauchy-Schwartz. We'll just want to be careful about calculating $\alpha$ for our regularizer later on.}
-Let $x_0 = F(y_0)$. Then we must have $\partial f(x_0) = -y_0$, hence 
-\begin{align*}
-f(x)+y_1^Tx &\geq f(x_0) + y_1^Tx_0 - y_0^T(x-x_0) + \frac{\alpha}{2}\|x-x_0\|^2.
-\end{align*}
-But any minimizer of $f(x) + y_1^Tx$ must at least have a value less than $f(x_0) + y_1^Tx_0$. 
-It follows that 
-\begin{align*}
-         & y_1^Tx_0 \geq y_1^Tx_1 - y_0^T(x_1-x_0) + \frac{\alpha}{2}\|x_1-x_0\|^2 \\
-\implies & (y_1-y_0)^T(x_0-x_1) \geq \frac{\alpha}{2}\|x_1-x_0\|^2 \\
-\implies & \|y_1-y_0\|\|x_0-x_1\| \geq \frac{\alpha}{2}\|x_1-x_0\|^2 \\
-\implies & \|x_1-x_0\| \leq \frac{2}{\alpha} \|y_1-y_0\|,
-\end{align*}
-as was to be shown.
-\end{proof}
 \begin{proposition}[Convergence of Algorithm 1]
@@
 Then we have
-\[ \sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{\sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle}{\sum_{t=1}^T \alpha_t}. \]
+\begin{equation}
+\sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_f(x_t \| x_{t+1}).
+\end{equation}
 \end{proposition}
 \begin{proof}
-\NA{JHH: Uses, Lemma~\ref{lem:ftl}, convexity of $h$, taking supremums, and facts that follow from the construction of the algorithm.}
-Note that $L(u_t, \theta_t) = \max_{\theta} L(u_t, \theta)$ by construction. Also note that, if we invoke 
-Lemma~\ref{lem:ftl} with $l_t(u) = \alpha_t L(u, \theta_t)$, then we get the inequality
-\[ \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \leq \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle. \]
+Note that $L(u_t, \theta_t) = \max_{\theta} L(u_t, \theta)$ by construction. 
+Also note that, if we invoke Lemma~\ref{lem:bregman} with $f = h$ and 
+$z_t = \theta_t$, then we get the inequality
+\begin{equation}
+\frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_f(x_t \| x_{t+1}).
+\end{equation}
 Combining these together, we get the string of inequalities
 \begin{align*}
-L(\hat{u}_T, \theta) &= L\left(\frac{\sum_{t=1}^T \alpha_tu_t}{\sum_{t=1}^T \alpha_t}, \theta\right) \\
- &\leq \frac{\sum_{t=1}^T \alpha_t L(u_t, \theta)}{\sum_{t=1}^T \alpha_t} \\
- &\leq \frac{\sum_{t=1}^T \alpha_t L(u_t, \theta_t)}{\sum_{t=1}^T \alpha_t} \\
- &\leq \frac{\sum_{t=1}^T \alpha_tL(u^*, \theta_t)}{\sum_{t=1}^T \alpha_t} \\ &\phantom{=} + \frac{\sum_{t=1}^T \alpha_t (L(u_t,\theta_t)-L(u_{t+1},\theta_t))}{\sum_{t=1}^T \alpha_t} \\
- &= L(u^*, \theta_t) + \frac{\sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle)}{\sum_{t=1}^T \alpha_t} \\ &\phantom{=} + \frac{\sum_{t=1}^{T} \alpha_t(h(u_t)-h(u_{t+1}))}{\sum_{t=1}^T \alpha_t} \\
- &= L(u^*, \theta_t) + \frac{\sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle}{\sum_{t=1}^T \alpha_t} \\ &\phantom{=} + \frac{\alpha_1 h(u_1) - \alpha_T h(u_{T+1}) + \sum_{t=2}^{T} (\alpha_t-\alpha_{t-1}) h(u_t)}{\sum_{t=1}^T \alpha_t} \\
- &\leq L(u^*, \theta_t) + \frac{\sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle}{\sum_{t=1}^T \alpha_t} \\
- &\leq \sup_{\theta} L(u^*, \theta) + \frac{\sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle}{\sum_{t=1}^T \alpha_t},
+L(\hat{u}_T, \theta) &= L\left(\frac{1}{A_T} \sum_{t=1}^T \alpha_tu_t, \theta\right) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_f(u_t \| u_{t+1}) \\
+ &\leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_f(u_t \| u_{t+1}),
 \end{align*}
@@
 \label{cor:method-1}
-Suppose that $h$ is $\gamma$-strongly convex and let $r = \sup_{\theta} |\langle \theta, \theta \rangle|_2$. Then 
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{16r^2}{\gamma T^2}. \]
+Suppose that $h$ is strongly convex with respect to a norm $\|\cdot\|$ 
+and let $r = \sup_{\theta} \|\theta\|_{*}$. Then 
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
+In particular, for $\alpha_t = t$, we have
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T^2}. \]
 \end{corollary}
 \begin{proof}
-\NA{JHH: Follows from Lemma~\ref{lem:convexity}, triangle inequality, taking supremums, and Cauchy-Schwartz.}
 By Lemma~\ref{lem:convexity}, we have 
 \begin{align*}
-\|u_t-u_{t+1}\| &\leq \frac{2}{\gamma}\|\frac{\sum_{s \leq t}\alpha_s\theta_s}{\sum_{s \leq t} \alpha_s} - \frac{\sum_{s \leq t+1}\alpha_s\theta_s}{\sum_{s \leq t+1} \alpha_s}\| \\
- &= \frac{2}{\gamma}\|\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\theta_s - \frac{\alpha_{t+1}}{A_{t+1}} \theta_{t+1}\| \\
- &\leq \frac{2}{\gamma} \left(\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\|\theta_s\| + \frac{\alpha_{t+1}}{A_{t+1}} \|\theta_{t+1}\|\right) \\
- &\leq \frac{4r\alpha_{t+1}}{\gamma A_{t+1}}.
+D_f(x_t \| x_{t+1}) &= D_{f^*}(\hat{\theta}_{t+1} \| \hat{\theta}) \\
+ &\leq \frac{1}{2}\|\theta_{t+1}-\theta\|_{*}^2 \\
+ &= \frac{1}{2}\|\frac{\sum_{s \leq t}\alpha_s\theta_s}{\sum_{s \leq t} \alpha_s} - \frac{\sum_{s \leq t+1}\alpha_s\theta_s}{\sum_{s \leq t+1} \alpha_s}\|_{*}^2 \\
+ &= \frac{1}{2}\|\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\theta_s - \frac{\alpha_{t+1}}{A_{t+1}} \theta_{t+1}\|_{*}^2 \\
+ &\leq \frac{1}{2} \left(\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\|\theta_s\|_{*}^2 + \frac{\alpha_{t+1}}{A_{t+1}} \|\theta_{t+1}\|_{*}^2\right)^2 \\
+ &\leq \frac{2r^2\alpha_{t+1}^2}{A_{t+1}^2}.
 \end{align*}
-It follows from Cauchy-Schwartz that 
+It follows that 
 \begin{align*}
-\frac{1}{A_T} \sum_{t=1}^T \alpha_t \langle u_t - u_{t+1}, \theta_t \rangle &\leq \sum_{t=1}^T \frac{4r^2\alpha_{t}\alpha_{t+1}}{\gamma A_{t+1}A_T} \\
- &\leq \frac{4r^2}{\gamma A_T} \sum_{t=1}^T \frac{\alpha_{t}\alpha_{t+1}}{A_{t+1}}.
+\frac{1}{A_T} \sum_{t=1}^T A_tD_f(x_t \| x_{t+1}) &\leq \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}.
 \end{align*}
-In particular, if we let $\alpha_t = t$, then $A_t = \frac{t(t+1)}{2}$ and $\frac{\alpha_{t}\alpha_{t+1}}{A_{t+1}} = 2$. 
+In particular, if we let $\alpha_t = t$, then $A_t = \frac{t(t+1)}{2}$ and 
+$\frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} = \frac{2(t+1)^2t(t+1)}{(t+1)^2(t+2)^2} = \frac{2t(t+1)}{(t+2)^2} \leq 2$.
 We therefore get
 \begin{equation}
-\frac{1}{A_T} \sum_{t=1}^T \alpha_t\langle u_t-u_{t+1}, \theta_t \rangle \leq \frac{16r^2}{T(T+1)},
+\frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} \leq \frac{8r^2}{T(T+1)},
 \end{equation}
@@
 Consider the updates $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$ 
-and $\theta_{t+1} \in \arg\max_{\theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$. 
-Then for $\hat{u} = \frac{1}{T} \sum_{t=1}^T u_t$, we have 
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{T} \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_{t} \rangle. \]
+and $\theta_{t+1} \in \arg\max_{\theta} \langle \theta, \hat{u}_t \rangle - R(\theta)$. 
+Then we have 
+\begin{equation}
+\sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_{R}(\theta_t \| \theta_{t+1}).
+\end{equation}
 \end{proposition}
 \begin{proof}
-If we invoke Lemma~\ref{lem:ftl} with $l_t(\theta) = -L(u_t, \theta)$, then we get the inequality
-\[ \sum_{t=1}^T -L(u_t, \theta_t) \leq -\sum_{t=1}^T L(u_t, \theta^*) - \sum_{t=1}^T \langle u_t, \theta_t - \theta_{t+1} \rangle. \]
+If we invoke Lemma~\ref{lem:bregman} with $f=R$ and $z_t=-u_t$, then we get 
+the inequality
+\[ \frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta_t) \leq \frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta^*) - \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). \]
 Re-arranging yields
-\[ \sum_{t=1}^T L(u_t, \theta^*) \leq \sum_{t=1}^T L(u_t, \theta_t) + \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_t \rangle. \]
+\[ \frac{1}{A_T }\sum_{t=1}^T \alpha_t L(u_t, \theta^*) \leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). \]
 Now, we have the following string of inequalities:
 \begin{align*}
-L(\hat{u}, \theta) &= L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) \\
- &\leq \frac{1}{T} \sum_{t=1}^T L(u_t, \theta) \\
- &\leq \frac{1}{T} \sum_{t=1}^T L(u_t, \theta_t) + \frac{1}{T} \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_t \rangle \\
- &= \frac{1}{T} \sum_{t=1}^T \inf_{u} L(u, \theta_t) + \frac{1}{T} \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_t \rangle \\
- &\leq \frac{1}{T} \sum_{t=1}^T L(u^*, \theta_t) + \frac{1}{T} \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_t \rangle \\
- &\leq \frac{1}{T} \sum_{t=1}^T \sup_{\theta} L(u^*, \theta) + \frac{1}{T} \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_t \rangle \\
- &= \sup_{\theta} L(u^*, \theta) + \frac{1}{T} \sum_{t=1}^T \langle u_t, \theta_{t+1} - \theta_t \rangle,
+L(\hat{u}, \theta) &= L\left(\frac{1}{A_T} \sum_{t=1}^T \alpha_t u_t, \theta\right) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_R(\theta_t \| \theta_{t+1}) \\
+ &= \frac{1}{A_T} \sum_{t=1}^T \alpha_t \inf_{u} L(u, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
+ &= \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}),
 \end{align*}
@@
 \label{cor:method-2}
-Suppose that $R$ is $\gamma$-strongly convex and let $r = \sup_{u} |\langle u, u \rangle|$. Then
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{16r^2}{\gamma T^2}. \]
+Suppose that $R$ is strongly convex with respect to a norm $\|\cdot\|$ 
+and let $r = \sup_{u} \|u\|_{*}$. Then 
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
+In particular, for $\alpha_t = t$, we have
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T^2}. \]
 \end{corollary}

### 2013-09-12 icml/paper.tex
fixes and corrections, including to final convergence rate result

@@
 
-\NA{JHH: Should we comment on the symmetry introduced by this formulation?}
 First, let us generalize the setting 
@@
 As long as $h$ is strongly convex, we obtain the 
-bound (see Proposition~\ref{cor:method-1}):
+bound (see Corollary~\ref{cor:method-1}):
 \begin{equation}
@@
 We then obtain the following bound 
-when $R$ is strongly convex (see Proposition~\ref{cor:method-2}):
+when $R$ is strongly convex (see Corollary~\ref{cor:method-2}):
 \[ \sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/T). \]
@@
 $-L(\theta,u)$).
-\NA{JHH: It would be more elegant to argue either here or below that the bounds on the second algorithm follow directly from duality}
-\NA{JNS: The statements aren't actually dual to each other; the first is saying that the primal converges to its 
-    optimum, and the second is saying that the dual converges to its optimum.}
 
@@
 Our proofs hinge on the following key lemma:
+%%% MAIN LEMMA %%%
 \begin{lemma}
@@
 \begin{align}
-\alpha_{t}z_{t} = A_{t}\hat z_{t} - A_{t-1}\hat z_{t-1} = - A_{t}\nabla f(x_{t+1}) + A_{t-1} \nabla f(x_{t}) 
+\alpha_{t}z_{t} 
+&= A_{t}\hat z_{t} - A_{t-1}\hat z_{t-1} \\
+&= - A_{t}\nabla f(x_{t+1}) + A_{t-1} \nabla f(x_{t}),
 \end{align}
@@
  &= \sum_{t=1}^T \{\alpha_t f(x_t) + \langle A_{t-1} \nabla f(x_t) - A_t \nabla f(x_{t+1}), x_t \rangle\} \\
- &= \sum_{t=1}^T \{\alpha_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle\} - A_{T}\langle \nabla f(x_{T+1}), x_{T+1} \rangle \\
+ &= \sum_{t=1}^T \{\alpha_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle\} \\
+ & \quad - A_{T}\langle \nabla f(x_{T+1}), x_{T+1} \rangle
+\end{align}
+\begin{align}
  &= \sum_{t=1}^T \{A_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle - A_t f(x_{t+1})\}  \\
  &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
- &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \\
+ &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\}  \\
+ &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
  &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) + \langle \hat{z}_T, x_{T+1} \rangle) \\
- &\leq \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x^*) + \langle \hat{z}_T, x^* \rangle). \\
+ &\leq \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x^*) + \langle \hat{z}_T, x^* \rangle). 
 \end{align}
@@
 \end{lemma}
+
+%%% ALGORITHM 1 ANALYSIS %%%
+% Convergence proposition
 \begin{proposition}[Convergence of Algorithm 1]
@@
 $z_t = \theta_t$, then we get the inequality
-\begin{equation}
-\frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_f(x_t \| x_{t+1}).
-\end{equation}
+\begin{align}
+\frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) 
+&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) \\
+&\quad + \frac{1}{A_T} \sum_{t=1}^T A_tD_f(x_t \| x_{t+1}). \nonumber
+\end{align}
 Combining these together, we get the string of inequalities
@@
 \end{proof}
-\begin{corollary}
-\label{cor:method-1}
+
+% Convergence rate theorem
+\begin{theorem}
+\label{thm:method-1}
 Suppose that $h$ is strongly convex with respect to a norm $\|\cdot\|$ 
@@
 \[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
-In particular, for $\alpha_t = t$, we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T^2}. \]
-\end{corollary}
+\end{theorem}
 \begin{proof}
@@
 \begin{align*}
-D_f(x_t \| x_{t+1}) &= D_{f^*}(\hat{\theta}_{t+1} \| \hat{\theta}) \\
- &\leq \frac{1}{2}\|\theta_{t+1}-\theta\|_{*}^2 \\
+D_f(x_t \| x_{t+1}) &= D_{f^*}(\hat{\theta}_{t+1} \| \hat{\theta}_{t}) \\
+ &\leq \frac{1}{2}\|\hat\theta_{t+1}-\hat\theta_{t}\|_{*}^2 \\
  &= \frac{1}{2}\|\frac{\sum_{s \leq t}\alpha_s\theta_s}{\sum_{s \leq t} \alpha_s} - \frac{\sum_{s \leq t+1}\alpha_s\theta_s}{\sum_{s \leq t+1} \alpha_s}\|_{*}^2 \\
@@
 \end{align*}
-In particular, if we let $\alpha_t = t$, then $A_t = \frac{t(t+1)}{2}$ and 
+\end{proof}
+
+% Convergence rate corollary 
+\begin{corollary} 
+\label{cor:method-1}
+Under the hypotheses of Theorem \ref{thm:method-1}, for $\alpha_{t} = 1$ we have
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2 (\log (T) + 1)}{T}. \]
+and for $\alpha_t = t$ we have
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T}. \]
+\end{corollary}
+\begin{proof}
+If we let $\alpha_t = 1$, then $A_t = t$ and $\frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} = \frac{t}{(t+1)^{2}} \le \frac{1}{t}$.
+We therefore get
+\begin{equation}
+\frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} \leq \frac{2r^2}{T} \sum_{t=1}^{T}\frac{1}{t} \leq \frac{2r^2(\log (T) + 1)}{T+1}.
+\end{equation}
+If we let $\alpha_t = t$, then $A_t = \frac{t(t+1)}{2}$ and 
 $\frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} = \frac{2(t+1)^2t(t+1)}{(t+1)^2(t+2)^2} = \frac{2t(t+1)}{(t+2)^2} \leq 2$.
@@
 \begin{equation}
-\frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} \leq \frac{8r^2}{T(T+1)},
+\frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} \leq \frac{4r^2}{T(T+1)} \sum_{t=1}^{T}2 = \frac{8r^2}{T+1},
 \end{equation}
@@
 \end{proof}
+
+%%% ALGORITHM 2 ANALYSIS %%%
+% Convergence proposition
 \begin{proposition}[Convergence of Algorithm 2]
@@
 the inequality
-\[ \frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta_t) \leq \frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta^*) - \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). \]
+\begin{align}
+\frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta_t) 
+&\leq \frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta^*) \\
+&\quad - \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). \nonumber
+\end{align}
 Re-arranging yields
-\[ \frac{1}{A_T }\sum_{t=1}^T \alpha_t L(u_t, \theta^*) \leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). \]
+\begin{align}
+\frac{1}{A_T }\sum_{t=1}^T \alpha_t L(u_t, \theta^*) 
+&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
+&\quad+ \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). 
+\end{align}
 Now, we have the following string of inequalities:
@@
 \end{proof}
-\begin{corollary}
-\label{cor:method-2}
+
+% Convergence rate theorem
+\begin{theorem}
+\label{thm:method-2}
 Suppose that $R$ is strongly convex with respect to a norm $\|\cdot\|$ 
@@
 \[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
-In particular, for $\alpha_t = t$, we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T^2}. \]
+\end{theorem}
+\begin{corollary}
+\label{cor:method-2}
+Under the hypotheses of Theorem \ref{thm:method-2}, for $\alpha_t = 1$ we have
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2(\log(T) + 1)}{T} \]
+and for $\alpha_t = t$ we have
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T} \]
 \end{corollary}
-\begin{proof}
-The proof is identical to the previous corollary.
+\begin{proof}[Proofs]
+The proofs are identical to those for Algorithm 1.
 \end{proof}
@@
 Suppose we are given a family of features $\phi$ together 
-with a known value $\bar{\phi}$ for $\bE_{\mu}[\phi]$. We 
+with a known value $\bar{\phi}$ for $\bE_{u}[\phi]$. We 
 would like to construct a distribution that approximately 
 matches this distribution, i.e.~one for which 
-$\bE_{\hat{\mu}}[\phi] \approx \bar{\phi}$. A natural way 
+$\bE_{\hat{u}}[\phi] \approx \bar{\phi}$. A natural way 
 to do this is by using boosted mirror descent to minimize the 
@@
 in other words, to minimize
-\[ L(\mu) = \max_{i \in \{1,\dots,d\}} |\bE_{\mu}[\phi_i]-\bar{\phi}_i|. \]
+\[ L(u) = \max_{i \in \{1,\dots,d\}} |\bE_{u}[\phi_i]-\bar{\phi}_i|. \]
 If we wish to write this more similarly to the original problem 
 formulation, we can write
-\[ L(\mu) = \sup_{\|\theta\|_1 \leq 1} \theta^T(\bE_{\mu}[\phi]-\bar{\phi}). \]
+\[ L(u) = \sup_{\|\theta\|_1 \leq 1} \theta^T(\bE_{u}[\phi]-\bar{\phi}). \]
 One issue is that $\|\theta\|_1$ is not strongly convex, so we relax $\|\theta\|_1$ 
 to $\|\theta\|_2$ and write our loss function in the Lagrangian form:
-\[ L(\mu, \theta) = \theta^T\bE_{\mu}[\phi] - \theta^T\bar{\phi} - \frac{1}{2}\|\theta\|_2^2. \]
+\[ L(u, \theta) = \theta^T\bE_{u}[\phi] - \theta^T\bar{\phi} - \frac{1}{2}\|\theta\|_2^2. \]
 Intriguingly, Algorithm 1 and Algorithm 2 are identical to each other for this choice of $L$ 
-(up to a change of index). In this case, $\mu_t = \delta_{x_t}$ for some $x_t$, and 
+(up to a change of index). In this case, $u_t = \delta_{x_t}$ for some $x_t$, and 
 the updates are the same as the standard herding updates. Let $r = \sup_{x} \|\phi(x)\|_2$; 
@@
 \begin{align*}
-\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2R^2\log(T+1)}{T} \\
- &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{2R^2\log(T+1)}{T} \\
- &= \frac{2R^2\log(T+1)}{T}.
+\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{8r^2(\log(T)+1)}{T} \\
+ &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{8R^2(\log(T)+1)}{T} \\
+ &= \frac{8R^2(\log(T)+1)}{T}.
 \end{align*}
@@
 where $\overline{\theta}_{t} = \frac{1}{t} \sum_{s \le t} \theta_{s}$.
-Hence, $h(u) + \bE_{\mu}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the boundary (when $\tau > 0$).
+Hence, $h(u) + \bE_{u}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the boundary (when $\tau > 0$).
 We can instead consider the following sequence of distributions:

### 2013-09-13 icml/paper.tex
small typo fixes

@@
 \begin{equation}
-\sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_f(x_t \| x_{t+1}).
+\sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}).
 \end{equation}
@@
 Also note that, if we invoke Lemma~\ref{lem:bregman} with $f = h$ and 
-$z_t = \theta_t$, then we get the inequality
+$z_t = \theta_t$ then we get the inequality
 \begin{align}
@@
 &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) \\
-&\quad + \frac{1}{A_T} \sum_{t=1}^T A_tD_f(x_t \| x_{t+1}). \nonumber
+&\quad + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}). \nonumber
 \end{align}
@@
  &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_f(u_t \| u_{t+1}) \\
- &\leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_f(u_t \| u_{t+1}),
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_h(u_t \| u_{t+1}) \\
+ &\leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}),
 \end{align*}
@@
 \begin{align*}
-D_f(x_t \| x_{t+1}) &= D_{f^*}(\hat{\theta}_{t+1} \| \hat{\theta}_{t}) \\
+D_h(u_t \| u_{t+1}) &= D_{h^*}(\hat{\theta}_{t+1} \| \hat{\theta}_{t}) \\
  &\leq \frac{1}{2}\|\hat\theta_{t+1}-\hat\theta_{t}\|_{*}^2 \\
@@
 \begin{align*}
-\frac{1}{A_T} \sum_{t=1}^T A_tD_f(x_t \| x_{t+1}) &\leq \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}.
+\frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}) &\leq \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}.
 \end{align*}
@@
 \begin{align*}
-\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{8r^2(\log(T)+1)}{T} \\
- &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{8R^2(\log(T)+1)}{T} \\
- &= \frac{8R^2(\log(T)+1)}{T}.
+\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T} \\
+ &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{8r^2}{T} \\
+ &= \frac{8r^2}{T}.
 \end{align*}
-This is a slightly weaker version of the typical herding bound on MMD.
+This is the same rate as the typical herding bound on MMD.
 

### 2013-09-13 icml/paper.tex
adding sketchy maxent argument

--- 2013-09-12 icml/paper.tex
@@
 (up to a change of index). In this case, $\mu_t = \delta_{x_t}$ for some $x_t$, and 
-the updates are the same as the standard herding updates. Let $r = \sup_{x} \|\phi(x)\|_2$; 
-then using the Algorithm 2 convergence bound gives us
+the updates are (for Algorithm 2):
+\begin{itemize}
+\item $x_t \in \arg\min_{x} \theta_t^T\phi(x)$
+\item $\theta_{t+1} = \frac{1}{A_t} \sum_{s=1}^t \alpha_s \phi(x)$.
+\end{itemize}
+These are the same as the standard herding updates when 
+$\alpha_t = 1$. Note that $R$ is strongly convex with respect to 
+$\|\cdot\|_2$, which is its own dual. Therefore, if we let
+$r = \sup_{x} \|\phi(x)\|_2$, then Corollary~\ref{cor:method-2} gives us
 \begin{align*}
-\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2R^2\log(T+1)}{T} \\
- &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{2R^2\log(T+1)}{T} \\
- &= \frac{2R^2\log(T+1)}{T}.
+\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
+ &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
+ &= \frac{2r^2\log(T+1)}{T}.
 \end{align*}
 This is a slightly weaker version of the typical herding bound on MMD.
+Note that if we use $\alpha_t = t$, which corresponds to using a weighted 
+average, then we obtain a much stronger bound of $\frac{8r^2}{T^2}$ instead.
 
@@
 
-Let us now consider the entropic regularizer $h(u) = \tau\bE_{u}[\log u(x)]$ with 
-\[ R(\theta) = \bE_{u}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.\]
-If we use Method 1, we then end up with 
-\[ u_{t}(x) \propto \exp\left\{-\frac{1}{\alpha}\overline{\theta}_{t}^T\phi(x)\right\}, \]
-where $\overline{\theta}_{t} = \frac{1}{t} \sum_{s \le t} \theta_{s}$.
+Bach et al. have observed that herding appears to be approximately building 
+samples from the maximum entropy distribution for a given collection of moment 
+constraints. In this section we will make this connection more explicit 
+as well as draw other relationships between herding and maximum entropy.
+
+Suppose that instead of setting $h(u) = 0$ we set $h(u)$ to a small multiple of 
+the negative entropy: $h(u) = \eta \bE_{u}[\log u(x)]$. We will still use 
+$R(\theta) = \frac{1}{2}\|\theta\|_2^2$. In this case, Algorithm 2 gives the updates
+\begin{itemize}
+\item $u_t(x) \propto \exp\left(-\frac{1}{\eta}\theta_t^T\phi(x)\right)$
+\item $\theta_{t+1} = \bE_{\hat{u}_t}[\phi(x)]$.
+\end{itemize}
+Using Corollary~\ref{cor:method-2} gives us
+\begin{align*}
+\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
+ &= \inf_{u^*} \left\{-\eta H(u^*) + \frac{1}{2}\|\bE_{u^*}[\phi(x)]-\bar{\phi}\|_2^2\right\} \\ &\phantom{=} + \frac{2r^2\log(T+1)}{T}.
+\end{align*}
+In particular, let $u^*$ be the maximum entropy distribution satisfying $\bE_{u^*}[\phi(x)] = \bar{\phi}$, 
+and let $H^*$ be the corresponding entropy. Then we have 
+\[ -\eta H(\hat{u}) + \|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq -\eta H^* + \frac{2r^2\log(T+1)}{T}. \]
+Since $\|\cdot\|_2^2 \geq 0$, this implies that 
+\begin{equation}
+H(\hat{u}) \geq H^* - \frac{2r^2\log(T+1)}{\eta T},
+\end{equation}
+ and so asymptotically $\hat{u}$ has at least 
+as much entropy as $u^*$.
+
+Now, suppose that our state space has some finite size $S$, so 
+that no distribution has entropy greater than $\log(S)$. Then 
+we conclude that
+\begin{equation}
+\|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq \eta[\log(S)-H^*] + \frac{2r^2\log(T+1)}{T}.
+\end{equation}
+
+If we let $\eta \to 0$ and $\eta T \to \infty$, then we obtain 
+a distribution $\hat{u}$ whose moments are arbitrarily close to 
+the moments of $u^*$, and whose entropy is almost as large as 
+(or larger than) the entropy of $u^*$.
+
+% \begin{lemma}
+% blah
+% \end{lemma}
+% \begin{proof}
+% Let $L(u) = -H(u) + \sup_{\theta} \theta^T(\bE_{u}[\phi(x)]-\bar{\phi})$. 
+% Clearly $L$ is minimized at the maximum entropy distribution $u^*$ satisfying 
+% $\bE_{u}[\phi(x)] = \bar{\phi}$. Let $\theta^*$ be the natural parameters 
+% for $u^*$, then we can lower bound $L(u)$ by $-H(u^*) + (\theta^*)^T(\bE_{u}[\phi(x)]-\bar{\phi})$.
+% 
+% Since $-\lambda H(u) + \|\bE_{u}[\phi(x)]-\bar{\phi}\|_2^2$ is convex, we have
+% $-\lambda H(u) + \|\bE_{u}[\phi(x)]-\bar{\phi}\|_2^2 \geq -\lambda H(u_0) + \|\bE_{u_0}[\phi(x)]-\bar{\phi}\|_2^2 + (\bE_{u_0}[\phi(x)]-\bar{\phi})^T(u-u_0)$.
+% Setting $u = u^*$ yields
+% \[ -\lambda H(u_0) + \frac{1}{2} \|\bE_{u_0}[\phi(x)]-\bar{\phi}\|_2^2 + (\bE_{u_0}[\phi(x)]-\bar{\phi})^T(u-u_0) \leq -\lambda H(u^*). \]
+% 
+% \end{proof}
+
+%% Let us now consider the entropic regularizer $h(u) = \tau\bE_{u}[\log u(x)]$ with 
+%% \[ R(\theta) = \bE_{u}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.\]
+%% If we use Method 1, we then end up with 
+%% \[ u_{t}(x) \propto \exp\left\{-\frac{1}{\alpha}\overline{\theta}_{t}^T\phi(x)\right\}, \]
+%% where $\overline{\theta}_{t} = \frac{1}{t} \sum_{s \le t} \theta_{s}$.
 Hence, $h(u) + \bE_{\mu}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the boundary (when $\tau > 0$).
 We can instead consider the following sequence of distributions:
-\begin{enumerate}
-\item $\theta_t \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{u_t}[\phi(x)]-\bar{\phi})$
-\item $u_{t+1}(x) = Z(\overline{\theta}_{t}, \tau)^{-1} \exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\}$
-\end{enumerate}
-Here $Z(\overline{\theta}_{t}, \tau) = \int\exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\} \dee x$.
-
-The entropic regularizer is not strongly convex but a more general analysis involving 
-Bregman divergence can nevertheless yield a convergence bound in this case. We thus 
-get something that looks like herding but actually gives us an exponential family 
-type distribution. In particular, we have a Gibbs distribution with temperature $\tau$, and 
-in the limit as the temperature goes to zero, we recover herding, since 
-\[ \lim_{\tau \to 0} u_{t+1}(x) = \delta_{x^{*}_{t+1}}, \]
-where $x^{*}_{t+1} = \argmin_{x} \overline{\theta}_{t}^T\phi(x)$. 
-This fact is reminiscent of Welling's original derivation of herding, in which 
-he took the zero temperature limit of the gradient descent algorithm for the optimization problem 
-\[ \argmin_{u} h(u) \quad \text{s.t.}~\bE_{u}[\phi(x)] - \bar\phi = 0. \]
-This optimization is quite similar to the entropic optimization considered here, which 
-we can write as 
-\begin{align*}
-\argmin_{u}& \argmax_{|\theta\|_1 \leq 1} h(u) + \theta^{T}(\bE_{u}[\phi(x)] - \bar \phi)  \\
-&= \argmin_{u} h(u) + \|\bE_{u}[\phi(x)] - \bar\phi\|_{2}. 
-\end{align*}
-In the entropic formulation, the expectation constraint $\bE_{u}[\phi(x)] - \bar\phi = 0$ becomes
-soft, though the two optimizations are equivalent in the zero-temperature limit.
-
-We can view the entropic regularization
-with non-zero temperature as an alternative approximation to herding. If at each step we
-retain a sample $x_{t} \sim u_{t}$, then the sequence $(x_{t})_{t}$ converges to the 
-herding sample path as $\tau \to 0$. So $(x_{t})_{t}$ with non-zero temperature approximate
-herding samples. \NA{JHH: I'm not sure how useful this all is, other than to make contact with 
-the original paper, as I think the Gibbs distribution was exactly what Max was trying to 
-avoid by taking the zero-temperature limit.}
+%% \begin{enumerate}
+%% \item $\theta_t \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{u_t}[\phi(x)]-\bar{\phi})$
+%% \item $u_{t+1}(x) = Z(\overline{\theta}_{t}, \tau)^{-1} \exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\}$
+%% \end{enumerate}
+%% Here $Z(\overline{\theta}_{t}, \tau) = \int\exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\} \dee x$. 
+%% 
+%% The entropic regularizer is not strongly convex but a more general analysis involving 
+%% Bregman divergence can nevertheless yield a convergence bound in this case. We thus 
+%% get something that looks like herding but actually gives us an exponential family 
+%% type distribution. In particular, we have a Gibbs distribution with temperature $\tau$, and 
+%% in the limit as the temperature goes to zero, we recover herding, since 
+%% \[ \lim_{\tau \to 0} u_{t+1}(x) = \delta_{x^{*}_{t+1}}, \]
+%% where $x^{*}_{t+1} = \argmin_{x} \overline{\theta}_{t}^T\phi(x)$. 
+%% This fact is reminiscent of Welling's original derivation of herding, in which 
+%% he took the zero temperature limit of the gradient descent algorithm for the optimization problem 
+%% \[ \argmin_{u} h(u) \quad \text{s.t.}~\bE_{u}[\phi(x)] - \bar\phi = 0. \]
+%% This optimization is quite similar to the entropic optimization considered here, which 
+%% we can write as 
+%% \begin{align*}
+%% \argmin_{u}& \argmax_{|\theta\|_1 \leq 1} h(u) + \theta^{T}(\bE_{u}[\phi(x)] - \bar \phi)  \\
+%% &= \argmin_{u} h(u) + \|\bE_{u}[\phi(x)] - \bar\phi\|_{2}. 
+%% \end{align*}
+%% In the entropic formulation, the expectation constraint $\bE_{u}[\phi(x)] - \bar\phi = 0$ becomes
+%% soft, though the two optimizations are equivalent in the zero-temperature limit. 
+%% 
+%% We can view the entropic regularization
+%% with non-zero temperature as an alternative approximation to herding. If at each step we
+%% retain a sample $x_{t} \sim u_{t}$, then the sequence $(x_{t})_{t}$ converges to the 
+%% herding sample path as $\tau \to 0$. So $(x_{t})_{t}$ with non-zero temperature approximate
+%% herding samples. \NA{JHH: I'm not sure how useful this all is, other than to make contact with 
+%% the original paper, as I think the Gibbs distribution was exactly what Max was trying to 
+%% avoid by taking the zero-temperature limit.}
 

### 2013-09-13 icml/paper.tex
fixed a typo

--- 2013-09-13 icml/paper.tex
@@
 (up to a change of index). In this case, $u_t = \delta_{x_t}$ for some $x_t$, and 
-the updates are the same as the standard herding updates. Let $r = \sup_{x} \|\phi(x)\|_2$; 
-then using the Algorithm 2 convergence bound gives us
+the updates are (for Algorithm 2):
+\begin{itemize}
+\item $x_t \in \arg\min_{x} \theta_t^T\phi(x)$
+\item $\theta_{t+1} = \frac{1}{A_t} \sum_{s=1}^t \alpha_s \phi(x)$.
+\end{itemize}
+These are the same as the standard herding updates when 
+$\alpha_t = 1$. Note that $R$ is strongly convex with respect to 
+$\|\cdot\|_2$, which is its own dual. Therefore, if we let
+$r = \sup_{x} \|\phi(x)\|_2$, then Corollary~\ref{cor:method-2} gives us
 \begin{align*}
-\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T} \\
- &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{8r^2}{T} \\
- &= \frac{8r^2}{T}.
+\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
+ &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
+ &= \frac{2r^2\log(T+1)}{T}.
 \end{align*}
-This is the same rate as the typical herding bound on MMD.
+This is a slightly weaker version of the typical herding bound on MMD.
+Note that if we use $\alpha_t = t$, which corresponds to using a weighted 
+average, then we obtain a stronger bound of $\frac{8r^2}{T}$ instead.
 
@@
 
-Let us now consider the entropic regularizer $h(u) = \tau\bE_{u}[\log u(x)]$ with 
-\[ R(\theta) = \bE_{u}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.\]
-If we use Method 1, we then end up with 
-\[ u_{t}(x) \propto \exp\left\{-\frac{1}{\alpha}\overline{\theta}_{t}^T\phi(x)\right\}, \]
-where $\overline{\theta}_{t} = \frac{1}{t} \sum_{s \le t} \theta_{s}$.
-Hence, $h(u) + \bE_{u}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the boundary (when $\tau > 0$).
+Bach et al. have observed that herding appears to be approximately building 
+samples from the maximum entropy distribution for a given collection of moment 
+constraints. In this section we will make this connection more explicit 
+as well as draw other relationships between herding and maximum entropy.
+
+Suppose that instead of setting $h(u) = 0$ we set $h(u)$ to a small multiple of 
+the negative entropy: $h(u) = \eta \bE_{u}[\log u(x)]$. We will still use 
+$R(\theta) = \frac{1}{2}\|\theta\|_2^2$. In this case, Algorithm 2 gives the updates
+\begin{itemize}
+\item $u_t(x) \propto \exp\left(-\frac{1}{\eta}\theta_t^T\phi(x)\right)$
+\item $\theta_{t+1} = \bE_{\hat{u}_t}[\phi(x)]$.
+\end{itemize}
+Using Corollary~\ref{cor:method-2} gives us
+\begin{align*}
+\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
+ &= \inf_{u^*} \left\{-\eta H(u^*) + \frac{1}{2}\|\bE_{u^*}[\phi(x)]-\bar{\phi}\|_2^2\right\} \\ &\phantom{=} + \frac{2r^2\log(T+1)}{T}.
+\end{align*}
+In particular, let $u^*$ be the maximum entropy distribution satisfying $\bE_{u^*}[\phi(x)] = \bar{\phi}$, 
+and let $H^*$ be the corresponding entropy. Then we have 
+\[ -\eta H(\hat{u}) + \|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq -\eta H^* + \frac{2r^2\log(T+1)}{T}. \]
+Since $\|\cdot\|_2^2 \geq 0$, this implies that 
+\begin{equation}
+H(\hat{u}) \geq H^* - \frac{2r^2\log(T+1)}{\eta T},
+\end{equation}
+ and so asymptotically $\hat{u}$ has at least 
+as much entropy as $u^*$.
+
+Now, suppose that our state space has some finite size $S$, so 
+that no distribution has entropy greater than $\log(S)$. Then 
+we conclude that
+\begin{equation}
+\|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq \eta[\log(S)-H^*] + \frac{2r^2\log(T+1)}{T}.
+\end{equation}
+
+If we let $\eta \to 0$ and $\eta T \to \infty$, then we obtain 
+a distribution $\hat{u}$ whose moments are arbitrarily close to 
+the moments of $u^*$, and whose entropy is almost as large as 
+(or larger than) the entropy of $u^*$.
+
+% \begin{lemma}
+% blah
+% \end{lemma}
+% \begin{proof}
+% Let $L(u) = -H(u) + \sup_{\theta} \theta^T(\bE_{u}[\phi(x)]-\bar{\phi})$. 
+% Clearly $L$ is minimized at the maximum entropy distribution $u^*$ satisfying 
+% $\bE_{u}[\phi(x)] = \bar{\phi}$. Let $\theta^*$ be the natural parameters 
+% for $u^*$, then we can lower bound $L(u)$ by $-H(u^*) + (\theta^*)^T(\bE_{u}[\phi(x)]-\bar{\phi})$.
+% 
+% Since $-\lambda H(u) + \|\bE_{u}[\phi(x)]-\bar{\phi}\|_2^2$ is convex, we have
+% $-\lambda H(u) + \|\bE_{u}[\phi(x)]-\bar{\phi}\|_2^2 \geq -\lambda H(u_0) + \|\bE_{u_0}[\phi(x)]-\bar{\phi}\|_2^2 + (\bE_{u_0}[\phi(x)]-\bar{\phi})^T(u-u_0)$.
+% Setting $u = u^*$ yields
+% \[ -\lambda H(u_0) + \frac{1}{2} \|\bE_{u_0}[\phi(x)]-\bar{\phi}\|_2^2 + (\bE_{u_0}[\phi(x)]-\bar{\phi})^T(u-u_0) \leq -\lambda H(u^*). \]
+% 
+% \end{proof}
+
+%% Let us now consider the entropic regularizer $h(u) = \tau\bE_{u}[\log u(x)]$ with 
+%% \[ R(\theta) = \bE_{u}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.\]
+%% If we use Method 1, we then end up with 
+%% \[ u_{t}(x) \propto \exp\left\{-\frac{1}{\alpha}\overline{\theta}_{t}^T\phi(x)\right\}, \]
+%% where $\overline{\theta}_{t} = \frac{1}{t} \sum_{s \le t} \theta_{s}$.
+Hence, $h(u) + \bE_{\mu}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the boundary (when $\tau > 0$).
 We can instead consider the following sequence of distributions:
-\begin{enumerate}
-\item $\theta_t \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{u_t}[\phi(x)]-\bar{\phi})$
-\item $u_{t+1}(x) = Z(\overline{\theta}_{t}, \tau)^{-1} \exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\}$
-\end{enumerate}
-Here $Z(\overline{\theta}_{t}, \tau) = \int\exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\} \dee x$.
-
-The entropic regularizer is not strongly convex but a more general analysis involving 
-Bregman divergence can nevertheless yield a convergence bound in this case. We thus 
-get something that looks like herding but actually gives us an exponential family 
-type distribution. In particular, we have a Gibbs distribution with temperature $\tau$, and 
-in the limit as the temperature goes to zero, we recover herding, since 
-\[ \lim_{\tau \to 0} u_{t+1}(x) = \delta_{x^{*}_{t+1}}, \]
-where $x^{*}_{t+1} = \argmin_{x} \overline{\theta}_{t}^T\phi(x)$. 
-This fact is reminiscent of Welling's original derivation of herding, in which 
-he took the zero temperature limit of the gradient descent algorithm for the optimization problem 
-\[ \argmin_{u} h(u) \quad \text{s.t.}~\bE_{u}[\phi(x)] - \bar\phi = 0. \]
-This optimization is quite similar to the entropic optimization considered here, which 
-we can write as 
-\begin{align*}
-\argmin_{u}& \argmax_{|\theta\|_1 \leq 1} h(u) + \theta^{T}(\bE_{u}[\phi(x)] - \bar \phi)  \\
-&= \argmin_{u} h(u) + \|\bE_{u}[\phi(x)] - \bar\phi\|_{2}. 
-\end{align*}
-In the entropic formulation, the expectation constraint $\bE_{u}[\phi(x)] - \bar\phi = 0$ becomes
-soft, though the two optimizations are equivalent in the zero-temperature limit.
-
-We can view the entropic regularization
-with non-zero temperature as an alternative approximation to herding. If at each step we
-retain a sample $x_{t} \sim u_{t}$, then the sequence $(x_{t})_{t}$ converges to the 
-herding sample path as $\tau \to 0$. So $(x_{t})_{t}$ with non-zero temperature approximate
-herding samples. \NA{JHH: I'm not sure how useful this all is, other than to make contact with 
-the original paper, as I think the Gibbs distribution was exactly what Max was trying to 
-avoid by taking the zero-temperature limit.}
+%% \begin{enumerate}
+%% \item $\theta_t \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{u_t}[\phi(x)]-\bar{\phi})$
+%% \item $u_{t+1}(x) = Z(\overline{\theta}_{t}, \tau)^{-1} \exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\}$
+%% \end{enumerate}
+%% Here $Z(\overline{\theta}_{t}, \tau) = \int\exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\} \dee x$. 
+%% 
+%% The entropic regularizer is not strongly convex but a more general analysis involving 
+%% Bregman divergence can nevertheless yield a convergence bound in this case. We thus 
+%% get something that looks like herding but actually gives us an exponential family 
+%% type distribution. In particular, we have a Gibbs distribution with temperature $\tau$, and 
+%% in the limit as the temperature goes to zero, we recover herding, since 
+%% \[ \lim_{\tau \to 0} u_{t+1}(x) = \delta_{x^{*}_{t+1}}, \]
+%% where $x^{*}_{t+1} = \argmin_{x} \overline{\theta}_{t}^T\phi(x)$. 
+%% This fact is reminiscent of Welling's original derivation of herding, in which 
+%% he took the zero temperature limit of the gradient descent algorithm for the optimization problem 
+%% \[ \argmin_{u} h(u) \quad \text{s.t.}~\bE_{u}[\phi(x)] - \bar\phi = 0. \]
+%% This optimization is quite similar to the entropic optimization considered here, which 
+%% we can write as 
+%% \begin{align*}
+%% \argmin_{u}& \argmax_{|\theta\|_1 \leq 1} h(u) + \theta^{T}(\bE_{u}[\phi(x)] - \bar \phi)  \\
+%% &= \argmin_{u} h(u) + \|\bE_{u}[\phi(x)] - \bar\phi\|_{2}. 
+%% \end{align*}
+%% In the entropic formulation, the expectation constraint $\bE_{u}[\phi(x)] - \bar\phi = 0$ becomes
+%% soft, though the two optimizations are equivalent in the zero-temperature limit. 
+%% 
+%% We can view the entropic regularization
+%% with non-zero temperature as an alternative approximation to herding. If at each step we
+%% retain a sample $x_{t} \sim u_{t}$, then the sequence $(x_{t})_{t}$ converges to the 
+%% herding sample path as $\tau \to 0$. So $(x_{t})_{t}$ with non-zero temperature approximate
+%% herding samples. \NA{JHH: I'm not sure how useful this all is, other than to make contact with 
+%% the original paper, as I think the Gibbs distribution was exactly what Max was trying to 
+%% avoid by taking the zero-temperature limit.}
 

### 2013-09-24 icml/convex-opt-summary.tex
added convex optimation summary; added some latex defs

\documentclass[reqno,oneside,a4paper]{amsart}
%\usepackage{jhh-misc}
\usepackage{hyperref}
\usepackage{pdflscape}
\usepackage{epsfig}
\usepackage{framed}
\usepackage{natbib}
\usepackage{subfig}
\usepackage{xspace}
\usepackage[normalem]{ulem}

\input{latex-defs.tex}
% inner product command
\newcommand{\ip}[2]{\left\langle #1, #2 \right\rangle}
\def\[#1\]{\begin{align}#1\end{align}}

\begin{document}

\title{Convex Optimization: A Very Short and Incomplete Summary} 
\author{Jonathan Huggins}
\date{\today}

\maketitle

\section{Introduction}

Roughly speaking, we are interested in the following optimization problem:
\[ 
\max_{x} h(x) + f(x),
\]
where $h$ and $f$ are convex, and generally have other properties that aid in fast convergence of optimization algorithms. We assume throughout that all functions are convex. In this document we first outline key concepts. We then state some important lemmas. Finally we give a summary of some important convex optimization algorithms. These notes are based on S\'ebastian Bubeck's excellent notes for his Complexities of Optimization class\footnote{\url{https://blogs.princeton.edu/imabandit/orf523-the-complexities-of-optimization/}} and Francis Bach's ``Duality between subgradient and conditional gradient methods.''\footnote{\url{http://arxiv.org/abs/1211.6302}}

\section{Key Concepts}

\subsection{Conjugates and Gradients}

When the domain of $f$ is Banach space, we can define the convex conjugate of $f$
\[
f^{*}(y) \defined \max_{x} \ip{x}{y} - f(x).
\]
If $f$ is lower semi-continuous, then by the Fenchel-Moreau theorem $f = f^{**}$, i.e.
\[
f(x) = \max_{y} \ip{x}{y} - f^{*}(y).
\]
Furthermore, the gradients of $f^{*}$ and $f$ can be defined as any argument that maximize the previous two equations:
\[
\partial f^{*}(y) &\in \argmax_{x} \ip{x}{y} - f(x) \\
\partial f(x) &\in \argmax_{y} \ip{x}{y} - f^{*}(y).
\]

\subsection{Smoothness, Convexity, Distance} 
We say a function $f$ is $\alpha$-strongly convex if 
\[
f(x)  - f(y) \le \ip{\partial f(x)}{x-y} - \frac{\alpha}{2}\|x - y\|^{2}.
\]
We say a (possibly non-convex) function $g$ is $\beta$-smooth if 
\[
\|\partial g(x) - \partial g(y) \| \le \beta \| x - y\|.
\]

Another useful notion is Bregman divergence: 
\[
D(x,y) \defined f(x) - f(y) - \ip{\partial f(y)}{x - y}.
\]

\NA{JHH: more on Bregman divergence}

\section{Useful Facts}

Here we gather a few useful lemmas.

\blem
If (possibly non-convex) $g$ is $\beta$-smooth, then 
\[
|g(x)  - g(y) - \ip{\partial g(y)}{x-y}| \le \frac{\beta}{2}\|x - y\|^{2}
\]
\elem
\bprf \NA{JHH: prove this} \eprf
\blem
If $f$ is convex and $\beta$-smooth, then 
\[
f(x)  - f(y) \le \ip{\partial f(x)}{x-y} - \frac{1}{2\beta}\|\partial f(x) - \partial f(y)\|^{2}
\]
\elem
\bprf \NA{JHH: prove this} \eprf
\blem A convex function $f$ is $\alpha$-strongly convex $\iff$ $f^{*}$ is $(1/\alpha)$-smooth. \elem
\bprf \NA{JHH: prove this} \eprf

\section{Convex Optimization Algorithms}

The algorithms, along with conditions on the functions and convergence rates, are summarized in the table on the next page. Here we define some of the notation used in specific algorithms. In projected gradient descent $\eta_{t} = 2/\alpha(t+1)$. In Nesterov's accelerated gradient descent,
\[
\lambda_{0} &= 0 \\
\lambda_{t} &= \frac{1 + \sqrt{1 + 4\lambda_{t-1}^{2}}}{2} \\
\gamma_{t}  &= \frac{1 - \lambda_{t}}{\lambda_{t+1}}.
\]
In mirror descent and conditional gradient descent $\rho_{t} = 2/(t + 1)$. Finally, in boosted mirror descent
\[
\alpha_{s} &= s \\
A_{t} &= \sum_{s=1}^{t} \alpha_{s} = t(t+1)/2 \\
\bar x_{t} &= A_{t}^{-1} \sum_{s=1}^{t} \alpha_{s}x_{s} \\
\bar y_{t} &= A_{t}^{-1} \sum_{s=1}^{t} \alpha_{s}y_{s}
\]

\begin{landscape}

\begin{table}[htdp]
\begin{center}
\begin{tabular}{|l|p{4.05cm}|p{4.05cm}|p{4.05cm}|p{4.05cm}|}
\hline & Gradient Descent & Projected Subgrad. Desc. & Conditional Grad. Desc. & Nesterov's Accelerated Grad. Desc.   \\ 
\hline
Objective function  &
$f(x)$ &
$f(x)$ &
$-h^{*}(y) - f^{*}(y)$ &
$f(x)$ \\
\hline
Conditions & 
$f$ is $\beta$-smooth &
$f$ is $\alpha$-s.c., $\|\partial f\| \le L$, and space is compact. & 
$f$ is L.c. and $h$ is $\alpha$-s.c. &
$f$ is $\beta$-smooth \\
\hline
Update equations &
$x_{t+1} = x_{t} - \frac{1}{\beta}\partial f(x_{t})$ &
$y_{t+1} = x_{t} - \eta_{t}\partial f(x_{t})$ and $x_{t+1} = \argmax_{x} \| x - y_{t+1} \|$ &
$x_{t} = \partial h^{*}(-y_{t})$ and $\bar y_{t} = \partial f(x_{t})$ and $y_{t+1} = (1 - \rho_{t})y_{t} + \rho_{t} \bar y_{t}$&
$y_{t+1} = x_{t} - \frac{1}{\beta}\partial f(x_{t})$ and $x_{t+1} = (1- \gamma_{t})y_{t+1} + \gamma_{t}y_{t}$ \\
\hline
Convergence rate &
$\beta R^{2}/t$ &
$L^{2} / \alpha t$ &
$R^{2} / \alpha t$ &
$\beta R^{2}/t^{2}$ \\
\hline
\end{tabular}
\caption{Convex optimization algorithms (gradient descent type). The function $f$ and $g$ are assumed to be convex. Constants are not given in convergence rates. Here $R$ is the radius of the metric space. Shorthands: L.c. = Lipschitz continuous, s.c. = strongly convex. }
\label{table:gd-convex-opt-algs}
\end{center}
\end{table}

\begin{table}[htdp]
\begin{center}
\begin{tabular}{|l|p{5.5cm}|p{5.5cm}|p{5.5cm}|}
\hline & Mirror Descent & ``Primal'' Boosted Mirror Descent & ``Dual'' Boosted Mirror Descent  \\ 
\hline
Objective function & 
$h(x) + f(x)$ & 
$h(x) + f^{*}(x)$ & 
$h(x) + f^{*}(x)$ \\
\hline
Conditions & 
$h$ is smooth \& $\alpha$-s.c. and $f$ is L.c. & 
$h$ is $\alpha$-s.c. and $f^{*}$ is convex& 
$f^{*}$ is $\beta$-smooth and covex \\
\hline
Update equations & 
$y_{t} = \partial f(x_{t})$ and 
$x_{t+1} = \argmin_{x} h(x) - (1 - \rho_{t})\ip{\partial h(x_{t})}{x} + \rho_{t+1}\ip{y_{t}}{x}$ &
$x_{t} = \partial h^{*}(-\bar y_{t})$ and $y_{t+1} = \partial f^{*}(x_{t})$ & 
$x_{t} = \partial h^{*}(-y_{t})$ and $y_{t+1} = \partial f^{*}(\bar x_{t})$ \\
\hline
Convergence rate & 
$R^{2}/\alpha t$ & 
$R^{2}/\alpha t$ & 
$\beta R^{2} / t$ \\
\hline

\end{tabular}
\caption{Convex optimization algorithms (mirror descent type). See Table \ref{table:gd-convex-opt-algs} for conventions. \NA{JHH: should add ISTA and FISTA} }
\end{center}
\label{table:md-convex-opt-algs}
\end{table}

\end{landscape}
\end{document}

### 2013-09-24 icml/paper.tex
added convex optimation summary; added some latex defs
fixed framework

@@
 
-Consider the following loss function:
+Let $U$ and $\Theta$ be convex subsets of two Hilbert spaces $\sH_{U}$ and $\sH_{\Theta}$ and 
+let $i : U \to \Theta$ be a linear map from 
+$U$ to $\Theta$ that preserves the inner product. Consider the following loss function 
+$L : U \to \bR$:
 \begin{equation}
-L(u) = h(u) + \max_{\theta \in \Theta} \left\{\langle \theta, u \rangle - R(\theta) \right\}.
+L(u) = h(u) + \max_{\theta \in \Theta} \left\{\langle \theta, i(u) \rangle - R(\theta) \right\}.
 \end{equation}
@@
 \begin{itemize}
-\item Let $h(u) = 0$ and $R(\theta) = \langle \theta, u_0 \rangle$. Then $L(u)$ is the 
-      \emph{maximum mean discrepancy} between $u$ and $u_0$ 
-      relative to $\Theta$.
-\item Let $h(u) = 0$ and $R(\theta) = S^*(\theta)$ where $S^*$ is the Fenchel conjugate of 
-      any strongly convex function $S$. Then $L(u) = S(u)$.
-\item Let $h(u) = \|u\|_1$, $R(u) = \langle \theta, u_0 \rangle + \frac{1}{2} \|\theta\|_2^2$. 
+\item Let $U$ be the probability simplex over a space $\sX$, let $\phi : \sX \to \bR^d$ be a collection 
+      of statistics, and let $\Theta$ be the unit $L^1$ ball in $\bR^d$. Let $h(u) = 0$, 
+      $\langle \theta, i(u) \rangle = \bE_u[\theta^T\phi(x)]$, and 
+      $R(\theta) = \langle \theta, i(u_0) \rangle$. Then $L(u)$ is the 
+      \emph{maximum mean discrepancy} between $u$ and $u_0$ relative to $\phi$.
+\item Let $h(u) = 0$ and $R(\theta) = S^*(i^{-1}(\theta))$ where $S^*$ is the 
+      Fenchel conjugate of any strongly convex function $S : U \to \bR$. Then $L(u) = S(u)$ 
+      as long as $i^{-1}(\Theta) = \sH_{U}$.
+\item Let $U = \Theta = \bR^d$. Let $h(u) = \|u\|_1$, $R(u) = \langle \theta, u_0 \rangle + \frac{1}{2} \|\theta\|_2^2$. 
       Then $L(u) = \|u\|_1 + \frac{1}{2} \|u-u_0\|_2^2$. 
@@
 \begin{equation}
-L(u,\theta) = h(u) + \langle \theta, u \rangle - R(\theta).
+L(u,\theta) = h(u) + \langle \theta, i(u) \rangle - R(\theta).
 \end{equation}
 We will assume throughout that $h$ and $R$ are both convex functions, 
-and furthermore that $\arg\max_{u} h(u) + \theta^Tu$ can be efficiently 
+and furthermore that $\arg\max_{u} h(u) + \langle \theta, i(u) \rangle$ can be efficiently 
 computed for all values of $\theta$.
@@
 \item $u_1 \in \arg\min_u h(u)$
-\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, u_t \rangle - R(\theta)$
-\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, u \rangle$
+\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, i(u_t) \rangle - R(\theta)$
+\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, i(u) \rangle$
 %\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
@@
 \item $\theta_1 \in \arg\min_{\theta} R(\theta)$
-\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$
+\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, i(u) \rangle$
 %\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
-\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \hat{u}_t \rangle - R(\theta)$
+\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, i(\hat{u}_t) \rangle - R(\theta)$
 \end{enumerate}
@@
 $-L(\theta,u)$).
+
+For convenience, we will abuse notation and use $\langle \theta, u \rangle$ in place of 
+$\langle \theta, i(u) \rangle$ when the map $i$ is clear from context.
 
@@
 \item $x_t \in \arg\min_{x} \theta_t^T\phi(x)$
-\item $\theta_{t+1} = \frac{1}{A_t} \sum_{s=1}^t \alpha_s \phi(x)$.
+\item $\theta_{t+1} = \frac{1}{A_t} \sum_{s=1}^t \alpha_s \phi(x_{s}) - \bar \phi$.
+%\item $\theta_{t+1} = \frac{1}{A_t} \sum_{s=1}^t \alpha_s \phi(x)$.
 \end{itemize}

### 2013-10-01 icml/lp.tex
added L^p argument

\documentclass[reqno,oneside,a4paper]{amsart}
%\usepackage{jhh-misc}
\usepackage{hyperref}
\usepackage{pdflscape}
\usepackage{epsfig}
\usepackage{framed}
\usepackage{natbib}
\usepackage{subfig}
\usepackage{xspace}
\usepackage[normalem]{ulem}

\input{latex-defs.tex}
% inner product command
\newcommand{\ip}[2]{\left\langle #1, #2 \right\rangle}
\def\[#1\]{\begin{align}#1\end{align}}

\begin{document}

\title{Herding convergence for $L^p$ norm} 
\author{Jacob Steinhardt}
\date{\today}

\maketitle

Suppose that we have the loss function $L(u) = \frac{1}{p} \|u-u^*\|_p^p$, where $p > 1$. 
We will show that a modified herding algorithm approaches $u^*$ at a rate of $O(1/T)$.

First, note that $L(u) = \sup_{\theta \in \bR^d} \ip{u-u^*}{\theta} - \frac{1}{q} \|\theta\|_q^q$, 
for $\frac{1}{p} + \frac{1}{q} = 1$ (see Lemma 1).

Now, if we perform herding updates with $R(\theta) = \ip{u^*}{\theta} + \frac{1}{q}\|\theta\|_q^q$, we 
get a convergence rate of
\[ \sum_{t=1}^T \alpha_t D_{R^*}(\hat{u}_{t+1} \| \hat{u}_t). \]
The key part of our argument is to analyze $D_{R^*}(\hat{u}_{t+1} \| \hat{u}_t)$. For convenience, 
let $x = \hat{u}_{t+1}-u^*$, $y = \hat{u}_t-u^*$, and $z = u_{t+1}-u^*$.
First, note that 
\begin{align*}
D_{R^*}(\hat{u}_{t+1} \| \hat{u}_t) &= \frac{1}{p} \sum_{i=1}^d x_i^p-py_i^{p-1}(x_i-y_i)-y_i^p \\
 &= \frac{1}{p} \sum_{i=1}^d (x_i-y_i)\left(\sum_{r=0}^{p-2} y_i^rx_i^{p-r-1}-y_i^{p-1}\right) \\
 &= \frac{1}{p} \sum_{i=1}^d (x_i-y_i)^2 \left(\sum_{r=0}^{p-2} \sum_{s=0}^{p-r-2} y_i^ry_i^sx_i^{p-(r+s)-2}\right) \\
 &= \frac{1}{p} \sum_{i=1}^d (x_i-y_i)^2 \left(\sum_{r+s=0}^{p-2} y_i^{r+s}x_i^{p-(r+s)-2}\right) \\
 &= \frac{1}{p} \sum_{i=1}^d (x_i-y_i)^2 \left(\sum_{t=0}^{p-2} (t+1)y_i^{t}x_i^{p-t-2}\right).
\end{align*}
Now as a side calculation, we have $y_i = \frac{1}{A_t}\sum_{s \leq t} \alpha_s (u_{i,s}-u^*)$ 
and $x_i = \frac{1}{A_{t+1}} \sum_{s \leq t+1} \alpha_s (u_{i,s}-u^*)$. If we simplify we get
$(x_i-y_i)^2 = \frac{\alpha_{t+1}^2}{A_{t+1}^2} (y_i-z_i)^2$. Therefore, we have
\begin{align*}
\lefteqn{\frac{1}{p} \sum_{i=1}^d (x_i-y_i)^2 \left(\sum_{t=0}^{p-2} (t+1)y_i^{t}x_i^{p-t-2}\right)} \\
 &\leq \frac{\alpha_{t+1}^2}{p A_{t+1}^2} \sum_{t=0}^{p-2} \sum_{i=1}^d 2(t+1)(y_i^2+z_i^2)y_i^{t}x_i^{p-t-2}.
\end{align*}
This latter sum has $4\binom{p}{2}$ terms in it, each of the form $\sum_{i=1}^d x_i^ay_i^bz_i^c$ where $a+b+c=p$. 
But we have $x_i^ay_i^bz_i^c \leq \frac{ax_i^p+by_i^p+cz_i^p}{p}$, so $\sum_{i=1}^d x_i^ay_i^bz_i^c \leq \max(\|x\|_p^p, \|y\|_p^p, \|z\|_p^p)$.
Therefore, we obtain the bound
\begin{align*}
\lefteqn{\frac{1}{p} \sum_{i=1}^d (x_i-y_i)^2 \left(\sum_{t=0}^{p-2} (t+1)y_i^{t}x_i^{p-t-2}\right)} \\
 &\leq \frac{4\alpha_{t+1}^2}{p A_{t+1}^2}\binom{p}{2} \sup_{u \in U} \|u-u^*\|_p^p \\
 &= \frac{2\alpha_{t+1}^2(p-1)(2r_p)^p}{A_{t+1}^2}.
\end{align*}
This yields an overall bound of $\frac{8(p-1)(2r_p)^p}{T}$ for the step size $\alpha_t = t$.
\begin{remark}
We can probably shave the $2r_p$ down to just $r_p$ if we're more careful with this argument.
\end{remark}
\begin{remark}
This appears to work for any $p > 1$.
\end{remark}

\section{Proof of Lemma 1}
We want to compute $\sup_{\theta \in \bR^d} \ip{u-u^*}{\theta} - \frac{1}{q} \|\theta\|_q^q$. To do this, 
we will take the derivative with respect to $\theta_i$ and set it to $0$, which yields:
\[ u_i - u_i^* = \theta_i^{q-1}. \]
Therefore, we have
\[ \theta_i = \left(u_i-u_i^*\right)^{\frac{1}{q-1}}. \]
Plugging back in, we get
\begin{align*}
\ip{u-u^*}{\theta} - \frac{1}{q} \|\theta\|_q^q &= \sum_{i=1}^d (u_i-u_i^*)^{\frac{q}{q-1}} \cdot \left(1-\frac{1}{q}\right) \\
 &= \frac{1}{p} \|u-u^*\|_p^p,
\end{align*}
as was to be shown.

\end{document}

### 2013-10-02 icml/convex-opt-summary.tex

@@
 A_{t} &= \sum_{s=1}^{t} \alpha_{s} = t(t+1)/2 \\
-\bar x_{t} &= A_{t}^{-1} \sum_{s=1}^{t} \alpha_{s}x_{s} \\
-\bar y_{t} &= A_{t}^{-1} \sum_{s=1}^{t} \alpha_{s}y_{s}
+\bar u_{t} &= A_{t}^{-1} \sum_{s=1}^{t} \alpha_{s}u_{s} \\
+\bar \theta_{t} &= A_{t}^{-1} \sum_{s=1}^{t} \alpha_{s}\theta_{s}
 \]
@@
 $f(x)$ &
-$-h^{*}(y) - f^{*}(y)$ &
+$h^{*}(y) + f^{*}(y)$ &
 $f(x)$ \\
@@
 $y_{t+1} = x_{t} - \eta_{t}\partial f(x_{t})$ and $x_{t+1} = \argmax_{x} \| x - y_{t+1} \|$ &
-$x_{t} = \partial h^{*}(-y_{t})$ and $\bar y_{t} = \partial f(x_{t})$ and $y_{t+1} = (1 - \rho_{t})y_{t} + \rho_{t} \bar y_{t}$&
+$x_{t} = \partial h^{*}(-\bar y_{t})$ and $y_{t+1} = \partial f(x_{t})$ and $\bar y_{t+1} = (1 - \rho_{t})\bar y_{t} + \rho_{t} y_{t+1}$&
 $y_{t+1} = x_{t} - \frac{1}{\beta}\partial f(x_{t})$ and $x_{t+1} = (1- \gamma_{t})y_{t+1} + \gamma_{t}y_{t}$ \\
@@
 $h(x) + f(x)$ & 
-$h(x) + f^{*}(x)$ & 
-$h(x) + f^{*}(x)$ \\
+$g(x) + R^{*}(x)$ & 
+$g(x) + R^{*}(x)$ \\
 \hline
@@
 $h$ is smooth \& $\alpha$-s.c. and $f$ is L.c. & 
-$h$ is $\alpha$-s.c. and $f^{*}$ is convex& 
-$f^{*}$ is $\beta$-smooth and covex \\
+$g$ is $\alpha$-s.c.& 
+$R^{*}$ is $\beta$-smooth \\
 \hline
@@
 $y_{t} = \partial f(x_{t})$ and 
-$x_{t+1} = \argmin_{x} h(x) - (1 - \rho_{t})\ip{\partial h(x_{t})}{x} + \rho_{t+1}\ip{y_{t}}{x}$ &
-$x_{t} = \partial h^{*}(-\bar y_{t})$ and $y_{t+1} = \partial f^{*}(x_{t})$ & 
-$x_{t} = \partial h^{*}(-y_{t})$ and $y_{t+1} = \partial f^{*}(\bar x_{t})$ \\
+$x_{t+1} = \argmin_{x} h(x) - (1 - \rho_{t+1})\ip{\partial h(x_{t})}{x} + \rho_{t+1}\ip{y_{t}}{x}$ &
+$u_{t} = \partial g^{*}(-\bar \theta_{t})$ and $\theta_{t+1} = \partial R^{*}(u_{t})$ & 
+$u_{t} = \partial g^{*}(-\theta_{t})$ and $\theta_{t+1} = \partial R^{*}(\bar u_{t})$ \\
 \hline
@@
 \end{landscape}
+
+\section{CGD, MD, and Boosted MD}
+
+The Boosted Mirror Descent algorithms are very closely related to Conditional Gradient Descent and Mirror Descent.
+Dual BMD and CGD are related as follows.
+BMD minimizes $g + R^{*}$ while CGD minimizes $h^{*} + f^{*}$.
+Make the identifications $h^{*} = R^{*}$ and $f^{*} = g$.
+Note that $R^{*}$ is $\beta$-smooth $\iff R$ is $(1/\beta)$-strongly convex $\iff h$ is $(1/\beta)$-strongly convex, so the conditions on the functions being optimized are equivalent.
+Then with $\rho_{t} = \alpha_{t+1}/A_{t+1}$, we can rewrite the CGD updates  as
+\[
+x_{t} &= \partial R^{*}(-\bar y_{t})  \\
+y_{t} &= \partial g^{*}(x_{t}) \\
+\bar y_{t+1} &= A_{t+1}^{-1}\sum_{s \le t+1} \alpha_{t} y_{t},
+\]
+which is identical to the dual BMD if we make the substitutions $x_{t} \to -\theta_{t}$ and $y_{t} \to -u_{t}$.
+
+We similarly relate Primal BMD and Mirror Descent.
+BMD minimizes $g + R^{*}$ while CGD minimizes $h + f$.
+Make the identifications $h = g$ and $f = R^{*}$ and note the conditions of $h$ and $g$ are the same.
+Letting $\rho_{t+1} = \alpha_{t}/A_{t}$, we can rewrite the Mirror Descent updates as
+\[
+y_{t} &= \partial R^{*}(x_{t}) \\
+x_{t+1} &= \partial g^{*}((1 - \rho_{t+1})\partial g(x_{t}) - \rho_{t+1}y_{t}) \\
+&= \partial g^{*}(-\bar y_{t}), 
+\]
+where $\bar y_{t}$ is the same as above. Making the substitutions $y_{t} \to \theta_{t}$ and $x_{t} \to u_{t}$ completes the identification of the two methods.
+
 \end{document}

### 2013-10-03 icml/paper.tex
small notational changes and other updates; switched to icml2014 style files

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%% ICML 2013 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass{article}
 
-\documentclass{article}
-\usepackage{import, icml2013}
+\usepackage{hyperref}
+
+\usepackage{icml2014}
+
 \input{latex-defs.tex}
@@
 
+\icmlauthor{Jonathan Huggins \NA{add footnote on authorship}}{jhuggins@mit.edu}
+\icmladdress{MIT CSAIL,
+             77 Massachusetts Avenue, Cambridge, MA 02139 USA}
 \icmlauthor{Jacob Steinhardt}{jsteinhardt@cs.stanford.edu}
@@
              353 Serra Street, Stanford, CA 94305 USA}
-\icmlauthor{Jonathan Huggins}{jhuggins@mit.edu}
-\icmladdress{Massachusetts Institute of Technology,
-             43 Vassar Street, Cambridge, MA 02139 USA}
 
@@
 
-\NA{JHH: I'm thinking that from the beginning we should let $u$ be defined over a different space than $\theta$ and assume a linear map $u \mapsto \eta(u) \in \Theta$.
-Then write $\langle \theta, \eta(u) \rangle$ so the setup will align with the herding case, where $\eta(u) = \bE_{u}[\phi(x)]$.
-}
-
-Let $U$ and $\Theta$ be convex subsets of two Hilbert spaces $\sH_{U}$ and $\sH_{\Theta}$ and 
-let $i : U \to \Theta$ be a linear map from 
+Let $U$ and $\Theta$ be convex subsets of two Banach spaces $\sB_{U}$ and $\sB_{\Theta}$ and 
+let $\imath : U \to \Theta$ be a linear map from 
 $U$ to $\Theta$ that preserves the inner product. Consider the following loss function 
@@
 \begin{equation}
-L(u) = h(u) + \max_{\theta \in \Theta} \left\{\langle \theta, i(u) \rangle - R(\theta) \right\}.
+L(u) = h(u) + \max_{\theta \in \Theta} \left\{\langle \theta, \imath(u) \rangle - R(\theta) \right\}.
 \end{equation}
@@
       of statistics, and let $\Theta$ be the unit $L^1$ ball in $\bR^d$. Let $h(u) = 0$, 
-      $\langle \theta, i(u) \rangle = \bE_u[\theta^T\phi(x)]$, and 
-      $R(\theta) = \langle \theta, i(u_0) \rangle$. Then $L(u)$ is the 
+      $\langle \theta, \imath(u) \rangle = \bE_u[\theta^T\phi(x)]$, and 
+      $R(\theta) = \langle \theta, \imath(u_0) \rangle$. Then $L(u)$ is the 
       \emph{maximum mean discrepancy} between $u$ and $u_0$ relative to $\phi$.
@@
       Fenchel conjugate of any strongly convex function $S : U \to \bR$. Then $L(u) = S(u)$ 
-      as long as $i^{-1}(\Theta) = \sH_{U}$.
+      as long as $i^{-1}(\Theta) = \sB_{U}$.
 \item Let $U = \Theta = \bR^d$. Let $h(u) = \|u\|_1$, $R(u) = \langle \theta, u_0 \rangle + \frac{1}{2} \|\theta\|_2^2$. 
@@
 \begin{equation}
-L(u,\theta) = h(u) + \langle \theta, i(u) \rangle - R(\theta).
+L(u,\theta) = h(u) + \langle \theta, \imath(u) \rangle - R(\theta).
 \end{equation}
 We will assume throughout that $h$ and $R$ are both convex functions, 
-and furthermore that $\arg\max_{u} h(u) + \langle \theta, i(u) \rangle$ can be efficiently 
+and furthermore that $\arg\max_{u} h(u) + \langle \theta, \imath(u) \rangle$ can be efficiently 
 computed for all values of $\theta$.
@@
 \item $u_1 \in \arg\min_u h(u)$
-\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, i(u_t) \rangle - R(\theta)$
-\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, i(u) \rangle$
+\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, \imath(u_t) \rangle - R(\theta)$
+\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, \imath(u) \rangle$
 %\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
@@
 \item $\theta_1 \in \arg\min_{\theta} R(\theta)$
-\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, i(u) \rangle$
+\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, \imath(u) \rangle$
 %\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
-\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, i(\hat{u}_t) \rangle - R(\theta)$
+\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \imath(\hat{u}_t) \rangle - R(\theta)$
 \end{enumerate}
@@
 For convenience, we will abuse notation and use $\langle \theta, u \rangle$ in place of 
-$\langle \theta, i(u) \rangle$ when the map $i$ is clear from context.
+$\langle \theta, \imath(u) \rangle$ when the map $i$ is clear from context.
 
@@
 
-Bach et al. have observed that herding appears to be approximately building 
+Bach et al.~have observed that herding appears to be approximately building 
 samples from the maximum entropy distribution for a given collection of moment 
@@
 \item $u_t(x) \propto \exp\left(-\frac{1}{\eta}\theta_t^T\phi(x)\right)$
-\item $\theta_{t+1} = \bE_{\hat{u}_t}[\phi(x)]$.
+\item $\theta_{t+1} = \bE_{\hat{u}_t}[\phi(x)] - \bar\phi$.
 \end{itemize}
-Using Corollary~\ref{cor:method-2} gives us
+Corollary~\ref{cor:method-2} implies that for all $u^{*}$,
 \begin{align*}
-\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
- &= \inf_{u^*} \left\{-\eta H(u^*) + \frac{1}{2}\|\bE_{u^*}[\phi(x)]-\bar{\phi}\|_2^2\right\} \\ &\phantom{=} + \frac{2r^2\log(T+1)}{T}.
+\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) &\leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
+ &= -\eta H(u^*) + \frac{1}{2}\|\bE_{u^*}[\phi(x)]-\bar{\phi}\|_2^2 \\ &\phantom{=} + \frac{8r^2}{T}.
 \end{align*}
@@
 and let $H^*$ be the corresponding entropy. Then we have 
-\[ -\eta H(\hat{u}) + \|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq -\eta H^* + \frac{2r^2\log(T+1)}{T}. \]
+\[ -\eta H(\hat{u}) + \|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq -\eta H^* + \frac{8r^2}{T}. \]
 Since $\|\cdot\|_2^2 \geq 0$, this implies that 
 \begin{equation}
-H(\hat{u}) \geq H^* - \frac{2r^2\log(T+1)}{\eta T},
+H(\hat{u}) \geq H^* - \frac{8r^2}{\eta T},
 \end{equation}
@@
 \begin{equation}
-\|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq \eta[\log(S)-H^*] + \frac{2r^2\log(T+1)}{T}.
+\|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq \eta[\log(S)-H^*] + \frac{8r^2}{T}.
 \end{equation}
@@
 
-%% Let us now consider the entropic regularizer $h(u) = \tau\bE_{u}[\log u(x)]$ with 
-%% \[ R(\theta) = \bE_{u}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.\]
-%% If we use Method 1, we then end up with 
-%% \[ u_{t}(x) \propto \exp\left\{-\frac{1}{\alpha}\overline{\theta}_{t}^T\phi(x)\right\}, \]
-%% where $\overline{\theta}_{t} = \frac{1}{t} \sum_{s \le t} \theta_{s}$.
+%\subsection{Old stuff}
+%
+% Let us now consider the entropic regularizer $h(u) = \tau\bE_{u}[\log u(x)]$ with 
+% \[ R(\theta) = \bE_{u}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.\]
+% If we use Method 1, we then end up with 
+% \[ u_{t}(x) \propto \exp\left\{-\frac{1}{\alpha}\overline{\theta}_{t}^T\phi(x)\right\}, \]
+% where $\overline{\theta}_{t} = \frac{1}{t} \sum_{s \le t} \theta_{s}$.
 Hence, $h(u) + \bE_{\mu}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the boundary (when $\tau > 0$).
 We can instead consider the following sequence of distributions:
-%% \begin{enumerate}
-%% \item $\theta_t \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{u_t}[\phi(x)]-\bar{\phi})$
-%% \item $u_{t+1}(x) = Z(\overline{\theta}_{t}, \tau)^{-1} \exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\}$
-%% \end{enumerate}
-%% Here $Z(\overline{\theta}_{t}, \tau) = \int\exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\} \dee x$. 
-%% 
-%% The entropic regularizer is not strongly convex but a more general analysis involving 
-%% Bregman divergence can nevertheless yield a convergence bound in this case. We thus 
-%% get something that looks like herding but actually gives us an exponential family 
-%% type distribution. In particular, we have a Gibbs distribution with temperature $\tau$, and 
-%% in the limit as the temperature goes to zero, we recover herding, since 
-%% \[ \lim_{\tau \to 0} u_{t+1}(x) = \delta_{x^{*}_{t+1}}, \]
-%% where $x^{*}_{t+1} = \argmin_{x} \overline{\theta}_{t}^T\phi(x)$. 
-%% This fact is reminiscent of Welling's original derivation of herding, in which 
-%% he took the zero temperature limit of the gradient descent algorithm for the optimization problem 
-%% \[ \argmin_{u} h(u) \quad \text{s.t.}~\bE_{u}[\phi(x)] - \bar\phi = 0. \]
-%% This optimization is quite similar to the entropic optimization considered here, which 
-%% we can write as 
-%% \begin{align*}
-%% \argmin_{u}& \argmax_{|\theta\|_1 \leq 1} h(u) + \theta^{T}(\bE_{u}[\phi(x)] - \bar \phi)  \\
-%% &= \argmin_{u} h(u) + \|\bE_{u}[\phi(x)] - \bar\phi\|_{2}. 
-%% \end{align*}
-%% In the entropic formulation, the expectation constraint $\bE_{u}[\phi(x)] - \bar\phi = 0$ becomes
-%% soft, though the two optimizations are equivalent in the zero-temperature limit. 
-%% 
-%% We can view the entropic regularization
-%% with non-zero temperature as an alternative approximation to herding. If at each step we
-%% retain a sample $x_{t} \sim u_{t}$, then the sequence $(x_{t})_{t}$ converges to the 
-%% herding sample path as $\tau \to 0$. So $(x_{t})_{t}$ with non-zero temperature approximate
-%% herding samples. \NA{JHH: I'm not sure how useful this all is, other than to make contact with 
-%% the original paper, as I think the Gibbs distribution was exactly what Max was trying to 
-%% avoid by taking the zero-temperature limit.}
+% \begin{enumerate}
+% \item $\theta_t \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{u_t}[\phi(x)]-\bar{\phi})$
+% \item $u_{t+1}(x) = Z(\overline{\theta}_{t}, \tau)^{-1} \exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\}$
+% \end{enumerate}
+% Here $Z(\overline{\theta}_{t}, \tau) = \int\exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\} \dee x$. 
+% 
+% The entropic regularizer is not strongly convex but a more general analysis involving 
+% Bregman divergence can nevertheless yield a convergence bound in this case. We thus 
+% get something that looks like herding but actually gives us an exponential family 
+% type distribution. In particular, we have a Gibbs distribution with temperature $\tau$, and 
+% in the limit as the temperature goes to zero, we recover herding, since 
+% \[ \lim_{\tau \to 0} u_{t+1}(x) = \delta_{x^{*}_{t+1}}, \]
+% where $x^{*}_{t+1} = \argmin_{x} \overline{\theta}_{t}^T\phi(x)$. 
+% This fact is reminiscent of Welling's original derivation of herding, in which 
+% he took the zero temperature limit of the gradient descent algorithm for the optimization problem 
+% \[ \argmin_{u} h(u) \quad \text{s.t.}~\bE_{u}[\phi(x)] - \bar\phi = 0. \]
+% This optimization is quite similar to the entropic optimization considered here, which 
+% we can write as 
+% \begin{align*}
+% \argmin_{u}& \argmax_{|\theta\|_1 \leq 1} h(u) + \theta^{T}(\bE_{u}[\phi(x)] - \bar \phi)  \\
+% &= \argmin_{u} h(u) + \|\bE_{u}[\phi(x)] - \bar\phi\|_{2}. 
+% \end{align*}
+% In the entropic formulation, the expectation constraint $\bE_{u}[\phi(x)] - \bar\phi = 0$ becomes
+% soft, though the two optimizations are equivalent in the zero-temperature limit. 
+% 
+% We can view the entropic regularization
+% with non-zero temperature as an alternative approximation to herding. If at each step we
+% retain a sample $x_{t} \sim u_{t}$, then the sequence $(x_{t})_{t}$ converges to the 
+% herding sample path as $\tau \to 0$. So $(x_{t})_{t}$ with non-zero temperature approximate
+% herding samples. \NA{JHH: I'm not sure how useful this all is, other than to make contact with 
+% the original paper, as I think the Gibbs distribution was exactly what Max was trying to 
+% avoid by taking the zero-temperature limit.}
 

### 2013-10-03 icml/herding-connection.tex
split into subfiles

--- 2013-10-03 icml/paper.tex
@@
-\documentclass{article}
-
-\usepackage{hyperref}
-
-\usepackage{icml2014}
-
-\input{latex-defs.tex}
-
-\icmltitlerunning{Boosted Mirror Descent}
-
+\documentclass[paper.tex]{subfiles}
 \begin{document}
-
-\twocolumn[
-\icmltitle{Boosted Mirror Descent, Frank-Wolfe, and Herding}
-
-\icmlauthor{Jonathan Huggins \NA{add footnote on authorship}}{jhuggins@mit.edu}
-\icmladdress{MIT CSAIL,
-             77 Massachusetts Avenue, Cambridge, MA 02139 USA}
-\icmlauthor{Jacob Steinhardt}{jsteinhardt@cs.stanford.edu}
-\icmladdress{Stanford University,
-             353 Serra Street, Stanford, CA 94305 USA}
-
-% You may provide any keywords that you 
-% find helpful for describing your paper; these are used to populate 
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{boosting, optimization, mirror descent, frank-wolfe, herding}
-
-\vskip 0.3in
-]
-
-\begin{abstract} 
-TODO
-\end{abstract}
-
-\section{Introduction}
-\label{sec:intro}
-
-\section{Boosted Mirror Descent}
-\label{sec:algorithm}
-
-Let $U$ and $\Theta$ be convex subsets of two Banach spaces $\sB_{U}$ and $\sB_{\Theta}$ and 
-let $\imath : U \to \Theta$ be a linear map from 
-$U$ to $\Theta$ that preserves the inner product. Consider the following loss function 
-$L : U \to \bR$:
-\begin{equation}
-L(u) = h(u) + \max_{\theta \in \Theta} \left\{\langle \theta, \imath(u) \rangle - R(\theta) \right\}.
-\end{equation}
-We can think of $h$ as a \emph{primal regularizer} and $R$ as a \emph{dual regularizer}.
-
-Examples:
-\begin{itemize}
-\item Let $U$ be the probability simplex over a space $\sX$, let $\phi : \sX \to \bR^d$ be a collection 
-      of statistics, and let $\Theta$ be the unit $L^1$ ball in $\bR^d$. Let $h(u) = 0$, 
-      $\langle \theta, \imath(u) \rangle = \bE_u[\theta^T\phi(x)]$, and 
-      $R(\theta) = \langle \theta, \imath(u_0) \rangle$. Then $L(u)$ is the 
-      \emph{maximum mean discrepancy} between $u$ and $u_0$ relative to $\phi$.
-\item Let $h(u) = 0$ and $R(\theta) = S^*(i^{-1}(\theta))$ where $S^*$ is the 
-      Fenchel conjugate of any strongly convex function $S : U \to \bR$. Then $L(u) = S(u)$ 
-      as long as $i^{-1}(\Theta) = \sB_{U}$.
-\item Let $U = \Theta = \bR^d$. Let $h(u) = \|u\|_1$, $R(u) = \langle \theta, u_0 \rangle + \frac{1}{2} \|\theta\|_2^2$. 
-      Then $L(u) = \|u\|_1 + \frac{1}{2} \|u-u_0\|_2^2$. 
-\end{itemize}
-
-In this paper we will consider a family of methods for minimizing $L(u)$, 
-which are based on Bach's generalization of the Frank-Wolfe algorithm and can 
-be interpreted as boosted mirror descent.
-
-First, let us generalize the setting 
-to consider a \emph{two-argument} loss function
-\begin{equation}
-L(u,\theta) = h(u) + \langle \theta, \imath(u) \rangle - R(\theta).
-\end{equation}
-We will assume throughout that $h$ and $R$ are both convex functions, 
-and furthermore that $\arg\max_{u} h(u) + \langle \theta, \imath(u) \rangle$ can be efficiently 
-computed for all values of $\theta$.
-
-Mirror descent, together with boosting, yields an algorithm for finding 
-``saddle points'' of $L$. It is the following algorithm (called Algorithm 1).
-For notational convenience, for a sequence of weights $\alpha_1,\alpha_2,\ldots$ 
-let $\hat{u}_t = \frac{\sum_{s=1}^t \alpha_su_s}{\sum_{s=1}^t \alpha_s}$ and let 
-$\hat{\theta}_t = \frac{\sum_{s=1}^t \alpha_s\theta_s}{\sum_{s=1}^t \alpha_s}$.
-\begin{enumerate}
-\item $u_1 \in \arg\min_u h(u)$
-\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, \imath(u_t) \rangle - R(\theta)$
-\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, \imath(u) \rangle$
-%\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
-\end{enumerate}
-As long as $h$ is strongly convex, we obtain the 
-bound (see Corollary~\ref{cor:method-1}):
-\begin{equation}
-\sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/T).
-\end{equation}
-In other words, $\hat{u}_T$ is close to being a global minimum of $L$.
-However, it is often the case that $h$ is not strongly convex, whereas $R$ is strongly convex. In this case 
-we may wish to use the following slightly different algorithm (called Algorithm 2):
-\begin{enumerate}
-\item $\theta_1 \in \arg\min_{\theta} R(\theta)$
-\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, \imath(u) \rangle$
-%\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
-\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \imath(\hat{u}_t) \rangle - R(\theta)$
-\end{enumerate}
-We then obtain the following bound 
-when $R$ is strongly convex (see Corollary~\ref{cor:method-2}):
-\[ \sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/T). \]
-Algorithm 1 and Algorithm 2 are closely related; indeed, they are dual to each other 
-(performing Algorithm 1 on $L(u,\theta)$ is the same as performing Algorithm 2 on 
-$-L(\theta,u)$).
-
-For convenience, we will abuse notation and use $\langle \theta, u \rangle$ in place of 
-$\langle \theta, \imath(u) \rangle$ when the map $i$ is clear from context.
-
-\section{Convergence Proofs}
-\label{sec:proofs}
-We now prove the convergence results cited in Section~\ref{sec:algorithm}. 
-Throughout this section, assume that $\alpha_1,\ldots,\alpha_T$ is 
-a sequence of real numbers and that $A_t = \sum_{s=1}^t \alpha_s$. 
-We further require that $A_t \geq 0$ for all $t$.
-
-Also recall that the Bregman divergence is defined by 
-$D_f(x_2 \| x_1) \eqdef f(x_2) - \langle \nabla f(x_1), x_2-x_1 \rangle - f(x_1)$.
-
-Our proofs hinge on the following key lemma:
-%%% MAIN LEMMA %%%
-\begin{lemma}
-\label{lem:bregman}
-Let $z_1,\ldots,z_T$ be vectors and let $f(x)$ be a strictly convex 
-function. Define $\hat{z}_t$ to be $\frac{1}{A_t} \sum_{s=1}^t \alpha_s z_s$.
-
-Let $x_1,\ldots,x_T$ be chosen via $x_{t+1} = \arg\min_{x} f(x) + \langle \hat{z}_t, x\rangle$. 
-Then for any $x^*$ we have
-\begin{align*}
-\lefteqn{\frac{1}{A_T} \sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
-\phantom{+} &\leq f(x^*) + \langle \hat{z}_t, x^* \rangle + \frac{1}{A_T} \sum_{t=1}^T A_t D_{f}(x_t \| x_{t+1}). 
-\end{align*}
-\end{lemma}
-\begin{proof}
-First note that, if $x_0 = \arg\min f(x) + \langle z, x \rangle$, 
-then $\nabla f(x_0) = -z$.
-
-\NA{JNS: In the following proof, (2) is by the update formula for $x_{t+1}$ together with the preceding observation.
-(3) is applying (2).
-(4) and (5) are re-arranging the sum.
-(6) is definition of Bregman divergence.
-(7) is the observation again.
-(8) is the definition of $x_{T+1}$.}
-
-Now note that
-\begin{align}
-\alpha_{t}z_{t} 
-&= A_{t}\hat z_{t} - A_{t-1}\hat z_{t-1} \\
-&= - A_{t}\nabla f(x_{t+1}) + A_{t-1} \nabla f(x_{t}),
-\end{align}
-so we have
-\begin{align}
-\lefteqn{\sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
- &= \sum_{t=1}^T \{\alpha_t f(x_t) + \langle A_{t-1} \nabla f(x_t) - A_t \nabla f(x_{t+1}), x_t \rangle\} \\
- &= \sum_{t=1}^T \{\alpha_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle\} \\
- & \quad - A_{T}\langle \nabla f(x_{T+1}), x_{T+1} \rangle
-\end{align}
-\begin{align}
- &= \sum_{t=1}^T \{A_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle - A_t f(x_{t+1})\}  \\
- &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
- &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\}  \\
- &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
- &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) + \langle \hat{z}_T, x_{T+1} \rangle) \\
- &\leq \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x^*) + \langle \hat{z}_T, x^* \rangle). 
-\end{align}
-Dividing both sides by $A_T$ completes the proof.
-\end{proof}
-We also note that $D_f(x_t \| x_{t+1}) = D_{f^*}(\hat{z}_{t+1} \| z_t)$, where $f^*(z) = \sup_x \langle z,x\rangle - f(x)$. 
-This form of the bound will often be more useful to us.
-\begin{lemma}
-\label{lem:convexity}
-Suppose that $D_f(x' \| x) \geq \frac{1}{2}\|x-x'\|^2$ for some 
-norm $\|\cdot\|$. (In this case we say that $f$ is strongly 
-convex with respect to $\|\cdot\|$.) 
-Then $D_{f^*}(x' \| x) \leq \frac{1}{2}\|x-x'\|_{*}^2$.
-\end{lemma}
-
-%%% ALGORITHM 1 ANALYSIS %%%
-% Convergence proposition
-\begin{proposition}[Convergence of Algorithm 1]
-\label{prop:method-1}
-Consider the updates $\theta_t \in \arg\max_{\theta} \langle \theta, u_t \rangle - R(\theta)$ 
-and $u_{t+1} \in \arg\min_u h(u) + \langle \hat{\theta}_s, u \rangle$. 
-Then we have
-\begin{equation}
-\sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}).
-\end{equation}
-\end{proposition}
-\begin{proof}
-Note that $L(u_t, \theta_t) = \max_{\theta} L(u_t, \theta)$ by construction. 
-Also note that, if we invoke Lemma~\ref{lem:bregman} with $f = h$ and 
-$z_t = \theta_t$ then we get the inequality
-\begin{align}
-\frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) 
-&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) \\
-&\quad + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}). \nonumber
-\end{align}
-Combining these together, we get the string of inequalities
-\begin{align*}
-L(\hat{u}_T, \theta) &= L\left(\frac{1}{A_T} \sum_{t=1}^T \alpha_tu_t, \theta\right) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_h(u_t \| u_{t+1}) \\
- &\leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}),
-\end{align*}
-as was to be shown.
-\end{proof}
-
-% Convergence rate theorem
-\begin{theorem}
-\label{thm:method-1}
-Suppose that $h$ is strongly convex with respect to a norm $\|\cdot\|$ 
-and let $r = \sup_{\theta} \|\theta\|_{*}$. Then 
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
-\end{theorem}
-\begin{proof}
-By Lemma~\ref{lem:convexity}, we have 
-\begin{align*}
-D_h(u_t \| u_{t+1}) &= D_{h^*}(\hat{\theta}_{t+1} \| \hat{\theta}_{t}) \\
- &\leq \frac{1}{2}\|\hat\theta_{t+1}-\hat\theta_{t}\|_{*}^2 \\
- &= \frac{1}{2}\|\frac{\sum_{s \leq t}\alpha_s\theta_s}{\sum_{s \leq t} \alpha_s} - \frac{\sum_{s \leq t+1}\alpha_s\theta_s}{\sum_{s \leq t+1} \alpha_s}\|_{*}^2 \\
- &= \frac{1}{2}\|\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\theta_s - \frac{\alpha_{t+1}}{A_{t+1}} \theta_{t+1}\|_{*}^2 \\
- &\leq \frac{1}{2} \left(\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\|\theta_s\|_{*}^2 + \frac{\alpha_{t+1}}{A_{t+1}} \|\theta_{t+1}\|_{*}^2\right)^2 \\
- &\leq \frac{2r^2\alpha_{t+1}^2}{A_{t+1}^2}.
-\end{align*}
-It follows that 
-\begin{align*}
-\frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}) &\leq \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}.
-\end{align*}
-\end{proof}
-
-% Convergence rate corollary 
-\begin{corollary} 
-\label{cor:method-1}
-Under the hypotheses of Theorem \ref{thm:method-1}, for $\alpha_{t} = 1$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2 (\log (T) + 1)}{T}. \]
-and for $\alpha_t = t$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T}. \]
-\end{corollary}
-\begin{proof}
-If we let $\alpha_t = 1$, then $A_t = t$ and $\frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} = \frac{t}{(t+1)^{2}} \le \frac{1}{t}$.
-We therefore get
-\begin{equation}
-\frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} \leq \frac{2r^2}{T} \sum_{t=1}^{T}\frac{1}{t} \leq \frac{2r^2(\log (T) + 1)}{T+1}.
-\end{equation}
-If we let $\alpha_t = t$, then $A_t = \frac{t(t+1)}{2}$ and 
-$\frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} = \frac{2(t+1)^2t(t+1)}{(t+1)^2(t+2)^2} = \frac{2t(t+1)}{(t+2)^2} \leq 2$.
-We therefore get
-\begin{equation}
-\frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} \leq \frac{4r^2}{T(T+1)} \sum_{t=1}^{T}2 = \frac{8r^2}{T+1},
-\end{equation}
-which completes the proof.
-\end{proof}
-
-%%% ALGORITHM 2 ANALYSIS %%%
-% Convergence proposition
-\begin{proposition}[Convergence of Algorithm 2]
-\label{prop:method-2}
-Consider the updates $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$ 
-and $\theta_{t+1} \in \arg\max_{\theta} \langle \theta, \hat{u}_t \rangle - R(\theta)$. 
-Then we have 
-\begin{equation}
-\sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_{R}(\theta_t \| \theta_{t+1}).
-\end{equation}
-\end{proposition}
-\begin{proof}
-If we invoke Lemma~\ref{lem:bregman} with $f=R$ and $z_t=-u_t$, then we get 
-the inequality
-\begin{align}
-\frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta_t) 
-&\leq \frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta^*) \\
-&\quad - \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). \nonumber
-\end{align}
-Re-arranging yields
-\begin{align}
-\frac{1}{A_T }\sum_{t=1}^T \alpha_t L(u_t, \theta^*) 
-&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
-&\quad+ \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). 
-\end{align}
-Now, we have the following string of inequalities:
-\begin{align*}
-L(\hat{u}, \theta) &= L\left(\frac{1}{A_T} \sum_{t=1}^T \alpha_t u_t, \theta\right) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_R(\theta_t \| \theta_{t+1}) \\
- &= \frac{1}{A_T} \sum_{t=1}^T \alpha_t \inf_{u} L(u, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
- &= \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}),
-\end{align*}
-as was to be shown.
-\end{proof}
-
-% Convergence rate theorem
-\begin{theorem}
-\label{thm:method-2}
-Suppose that $R$ is strongly convex with respect to a norm $\|\cdot\|$ 
-and let $r = \sup_{u} \|u\|_{*}$. Then 
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
-\end{theorem}
-\begin{corollary}
-\label{cor:method-2}
-Under the hypotheses of Theorem \ref{thm:method-2}, for $\alpha_t = 1$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2(\log(T) + 1)}{T} \]
-and for $\alpha_t = t$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T} \]
-\end{corollary}
-\begin{proof}[Proofs]
-The proofs are identical to those for Algorithm 1.
-\end{proof}
-
 \section{Application to Herding}
@@
 
-\section{Herding in Infinite Dimensions}
+\subsection{Herding in Infinite Dimensions}
 \label{sec:infinite-case}
 
-\section{Herding and Maximum Entropy}
+\subsection{Herding and Maximum Entropy}
 \label{sec:max-ent}
@@
 
-% \begin{lemma}
-% blah
-% \end{lemma}
-% \begin{proof}
-% Let $L(u) = -H(u) + \sup_{\theta} \theta^T(\bE_{u}[\phi(x)]-\bar{\phi})$. 
-% Clearly $L$ is minimized at the maximum entropy distribution $u^*$ satisfying 
-% $\bE_{u}[\phi(x)] = \bar{\phi}$. Let $\theta^*$ be the natural parameters 
-% for $u^*$, then we can lower bound $L(u)$ by $-H(u^*) + (\theta^*)^T(\bE_{u}[\phi(x)]-\bar{\phi})$.
-% 
-% Since $-\lambda H(u) + \|\bE_{u}[\phi(x)]-\bar{\phi}\|_2^2$ is convex, we have
-% $-\lambda H(u) + \|\bE_{u}[\phi(x)]-\bar{\phi}\|_2^2 \geq -\lambda H(u_0) + \|\bE_{u_0}[\phi(x)]-\bar{\phi}\|_2^2 + (\bE_{u_0}[\phi(x)]-\bar{\phi})^T(u-u_0)$.
-% Setting $u = u^*$ yields
-% \[ -\lambda H(u_0) + \frac{1}{2} \|\bE_{u_0}[\phi(x)]-\bar{\phi}\|_2^2 + (\bE_{u_0}[\phi(x)]-\bar{\phi})^T(u-u_0) \leq -\lambda H(u^*). \]
-% 
-% \end{proof}
-
-%\subsection{Old stuff}
-%
-% Let us now consider the entropic regularizer $h(u) = \tau\bE_{u}[\log u(x)]$ with 
-% \[ R(\theta) = \bE_{u}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.\]
-% If we use Method 1, we then end up with 
-% \[ u_{t}(x) \propto \exp\left\{-\frac{1}{\alpha}\overline{\theta}_{t}^T\phi(x)\right\}, \]
-% where $\overline{\theta}_{t} = \frac{1}{t} \sum_{s \le t} \theta_{s}$.
-Hence, $h(u) + \bE_{\mu}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the boundary (when $\tau > 0$).
-We can instead consider the following sequence of distributions:
-% \begin{enumerate}
-% \item $\theta_t \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{u_t}[\phi(x)]-\bar{\phi})$
-% \item $u_{t+1}(x) = Z(\overline{\theta}_{t}, \tau)^{-1} \exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\}$
-% \end{enumerate}
-% Here $Z(\overline{\theta}_{t}, \tau) = \int\exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\} \dee x$. 
-% 
-% The entropic regularizer is not strongly convex but a more general analysis involving 
-% Bregman divergence can nevertheless yield a convergence bound in this case. We thus 
-% get something that looks like herding but actually gives us an exponential family 
-% type distribution. In particular, we have a Gibbs distribution with temperature $\tau$, and 
-% in the limit as the temperature goes to zero, we recover herding, since 
-% \[ \lim_{\tau \to 0} u_{t+1}(x) = \delta_{x^{*}_{t+1}}, \]
-% where $x^{*}_{t+1} = \argmin_{x} \overline{\theta}_{t}^T\phi(x)$. 
-% This fact is reminiscent of Welling's original derivation of herding, in which 
-% he took the zero temperature limit of the gradient descent algorithm for the optimization problem 
-% \[ \argmin_{u} h(u) \quad \text{s.t.}~\bE_{u}[\phi(x)] - \bar\phi = 0. \]
-% This optimization is quite similar to the entropic optimization considered here, which 
-% we can write as 
-% \begin{align*}
-% \argmin_{u}& \argmax_{|\theta\|_1 \leq 1} h(u) + \theta^{T}(\bE_{u}[\phi(x)] - \bar \phi)  \\
-% &= \argmin_{u} h(u) + \|\bE_{u}[\phi(x)] - \bar\phi\|_{2}. 
-% \end{align*}
-% In the entropic formulation, the expectation constraint $\bE_{u}[\phi(x)] - \bar\phi = 0$ becomes
-% soft, though the two optimizations are equivalent in the zero-temperature limit. 
-% 
-% We can view the entropic regularization
-% with non-zero temperature as an alternative approximation to herding. If at each step we
-% retain a sample $x_{t} \sim u_{t}$, then the sequence $(x_{t})_{t}$ converges to the 
-% herding sample path as $\tau \to 0$. So $(x_{t})_{t}$ with non-zero temperature approximate
-% herding samples. \NA{JHH: I'm not sure how useful this all is, other than to make contact with 
-% the original paper, as I think the Gibbs distribution was exactly what Max was trying to 
-% avoid by taking the zero-temperature limit.}
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-%\bibliography{example_paper}
-%\bibliographystyle{icml2013}
-
 \end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was
-% created by Lise Getoor and Tobias Scheffer, it was slightly modified  
-% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz, 
-% slightly modified from the 2009 version by Kiri Wagstaff and 
-% Sam Roweis's 2008 version, which is slightly modified from 
-% Prasad Tadepalli's 2007 version which is a lightly 
-% changed version of the previous year's version by Andrew Moore, 
-% which was in turn edited from those of Kristian Kersting and 
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2013-10-03 icml/introduction.tex
split into subfiles

--- 2013-10-03 icml/paper.tex
@@
-\documentclass{article}
-
-\usepackage{hyperref}
-
-\usepackage{icml2014}
-
-\input{latex-defs.tex}
-
-\icmltitlerunning{Boosted Mirror Descent}
-
+\documentclass[paper.tex]{subfiles}
 \begin{document}
-
-\twocolumn[
-\icmltitle{Boosted Mirror Descent, Frank-Wolfe, and Herding}
-
-\icmlauthor{Jonathan Huggins \NA{add footnote on authorship}}{jhuggins@mit.edu}
-\icmladdress{MIT CSAIL,
-             77 Massachusetts Avenue, Cambridge, MA 02139 USA}
-\icmlauthor{Jacob Steinhardt}{jsteinhardt@cs.stanford.edu}
-\icmladdress{Stanford University,
-             353 Serra Street, Stanford, CA 94305 USA}
-
-% You may provide any keywords that you 
-% find helpful for describing your paper; these are used to populate 
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{boosting, optimization, mirror descent, frank-wolfe, herding}
-
-\vskip 0.3in
-]
-
-\begin{abstract} 
-TODO
-\end{abstract}
-
-\section{Introduction}
-\label{sec:intro}
-
 \section{Boosted Mirror Descent}
@@
 
-\section{Convergence Proofs}
-\label{sec:proofs}
-We now prove the convergence results cited in Section~\ref{sec:algorithm}. 
-Throughout this section, assume that $\alpha_1,\ldots,\alpha_T$ is 
-a sequence of real numbers and that $A_t = \sum_{s=1}^t \alpha_s$. 
-We further require that $A_t \geq 0$ for all $t$.
-
-Also recall that the Bregman divergence is defined by 
-$D_f(x_2 \| x_1) \eqdef f(x_2) - \langle \nabla f(x_1), x_2-x_1 \rangle - f(x_1)$.
-
-Our proofs hinge on the following key lemma:
-%%% MAIN LEMMA %%%
-\begin{lemma}
-\label{lem:bregman}
-Let $z_1,\ldots,z_T$ be vectors and let $f(x)$ be a strictly convex 
-function. Define $\hat{z}_t$ to be $\frac{1}{A_t} \sum_{s=1}^t \alpha_s z_s$.
-
-Let $x_1,\ldots,x_T$ be chosen via $x_{t+1} = \arg\min_{x} f(x) + \langle \hat{z}_t, x\rangle$. 
-Then for any $x^*$ we have
-\begin{align*}
-\lefteqn{\frac{1}{A_T} \sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
-\phantom{+} &\leq f(x^*) + \langle \hat{z}_t, x^* \rangle + \frac{1}{A_T} \sum_{t=1}^T A_t D_{f}(x_t \| x_{t+1}). 
-\end{align*}
-\end{lemma}
-\begin{proof}
-First note that, if $x_0 = \arg\min f(x) + \langle z, x \rangle$, 
-then $\nabla f(x_0) = -z$.
-
-\NA{JNS: In the following proof, (2) is by the update formula for $x_{t+1}$ together with the preceding observation.
-(3) is applying (2).
-(4) and (5) are re-arranging the sum.
-(6) is definition of Bregman divergence.
-(7) is the observation again.
-(8) is the definition of $x_{T+1}$.}
-
-Now note that
-\begin{align}
-\alpha_{t}z_{t} 
-&= A_{t}\hat z_{t} - A_{t-1}\hat z_{t-1} \\
-&= - A_{t}\nabla f(x_{t+1}) + A_{t-1} \nabla f(x_{t}),
-\end{align}
-so we have
-\begin{align}
-\lefteqn{\sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
- &= \sum_{t=1}^T \{\alpha_t f(x_t) + \langle A_{t-1} \nabla f(x_t) - A_t \nabla f(x_{t+1}), x_t \rangle\} \\
- &= \sum_{t=1}^T \{\alpha_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle\} \\
- & \quad - A_{T}\langle \nabla f(x_{T+1}), x_{T+1} \rangle
-\end{align}
-\begin{align}
- &= \sum_{t=1}^T \{A_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle - A_t f(x_{t+1})\}  \\
- &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
- &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\}  \\
- &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
- &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) + \langle \hat{z}_T, x_{T+1} \rangle) \\
- &\leq \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x^*) + \langle \hat{z}_T, x^* \rangle). 
-\end{align}
-Dividing both sides by $A_T$ completes the proof.
-\end{proof}
-We also note that $D_f(x_t \| x_{t+1}) = D_{f^*}(\hat{z}_{t+1} \| z_t)$, where $f^*(z) = \sup_x \langle z,x\rangle - f(x)$. 
-This form of the bound will often be more useful to us.
-\begin{lemma}
-\label{lem:convexity}
-Suppose that $D_f(x' \| x) \geq \frac{1}{2}\|x-x'\|^2$ for some 
-norm $\|\cdot\|$. (In this case we say that $f$ is strongly 
-convex with respect to $\|\cdot\|$.) 
-Then $D_{f^*}(x' \| x) \leq \frac{1}{2}\|x-x'\|_{*}^2$.
-\end{lemma}
-
-%%% ALGORITHM 1 ANALYSIS %%%
-% Convergence proposition
-\begin{proposition}[Convergence of Algorithm 1]
-\label{prop:method-1}
-Consider the updates $\theta_t \in \arg\max_{\theta} \langle \theta, u_t \rangle - R(\theta)$ 
-and $u_{t+1} \in \arg\min_u h(u) + \langle \hat{\theta}_s, u \rangle$. 
-Then we have
-\begin{equation}
-\sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}).
-\end{equation}
-\end{proposition}
-\begin{proof}
-Note that $L(u_t, \theta_t) = \max_{\theta} L(u_t, \theta)$ by construction. 
-Also note that, if we invoke Lemma~\ref{lem:bregman} with $f = h$ and 
-$z_t = \theta_t$ then we get the inequality
-\begin{align}
-\frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) 
-&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) \\
-&\quad + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}). \nonumber
-\end{align}
-Combining these together, we get the string of inequalities
-\begin{align*}
-L(\hat{u}_T, \theta) &= L\left(\frac{1}{A_T} \sum_{t=1}^T \alpha_tu_t, \theta\right) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_h(u_t \| u_{t+1}) \\
- &\leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}),
-\end{align*}
-as was to be shown.
-\end{proof}
-
-% Convergence rate theorem
-\begin{theorem}
-\label{thm:method-1}
-Suppose that $h$ is strongly convex with respect to a norm $\|\cdot\|$ 
-and let $r = \sup_{\theta} \|\theta\|_{*}$. Then 
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
-\end{theorem}
-\begin{proof}
-By Lemma~\ref{lem:convexity}, we have 
-\begin{align*}
-D_h(u_t \| u_{t+1}) &= D_{h^*}(\hat{\theta}_{t+1} \| \hat{\theta}_{t}) \\
- &\leq \frac{1}{2}\|\hat\theta_{t+1}-\hat\theta_{t}\|_{*}^2 \\
- &= \frac{1}{2}\|\frac{\sum_{s \leq t}\alpha_s\theta_s}{\sum_{s \leq t} \alpha_s} - \frac{\sum_{s \leq t+1}\alpha_s\theta_s}{\sum_{s \leq t+1} \alpha_s}\|_{*}^2 \\
- &= \frac{1}{2}\|\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\theta_s - \frac{\alpha_{t+1}}{A_{t+1}} \theta_{t+1}\|_{*}^2 \\
- &\leq \frac{1}{2} \left(\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\|\theta_s\|_{*}^2 + \frac{\alpha_{t+1}}{A_{t+1}} \|\theta_{t+1}\|_{*}^2\right)^2 \\
- &\leq \frac{2r^2\alpha_{t+1}^2}{A_{t+1}^2}.
-\end{align*}
-It follows that 
-\begin{align*}
-\frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}) &\leq \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}.
-\end{align*}
-\end{proof}
-
-% Convergence rate corollary 
-\begin{corollary} 
-\label{cor:method-1}
-Under the hypotheses of Theorem \ref{thm:method-1}, for $\alpha_{t} = 1$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2 (\log (T) + 1)}{T}. \]
-and for $\alpha_t = t$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T}. \]
-\end{corollary}
-\begin{proof}
-If we let $\alpha_t = 1$, then $A_t = t$ and $\frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} = \frac{t}{(t+1)^{2}} \le \frac{1}{t}$.
-We therefore get
-\begin{equation}
-\frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} \leq \frac{2r^2}{T} \sum_{t=1}^{T}\frac{1}{t} \leq \frac{2r^2(\log (T) + 1)}{T+1}.
-\end{equation}
-If we let $\alpha_t = t$, then $A_t = \frac{t(t+1)}{2}$ and 
-$\frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} = \frac{2(t+1)^2t(t+1)}{(t+1)^2(t+2)^2} = \frac{2t(t+1)}{(t+2)^2} \leq 2$.
-We therefore get
-\begin{equation}
-\frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} \leq \frac{4r^2}{T(T+1)} \sum_{t=1}^{T}2 = \frac{8r^2}{T+1},
-\end{equation}
-which completes the proof.
-\end{proof}
-
-%%% ALGORITHM 2 ANALYSIS %%%
-% Convergence proposition
-\begin{proposition}[Convergence of Algorithm 2]
-\label{prop:method-2}
-Consider the updates $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$ 
-and $\theta_{t+1} \in \arg\max_{\theta} \langle \theta, \hat{u}_t \rangle - R(\theta)$. 
-Then we have 
-\begin{equation}
-\sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_{R}(\theta_t \| \theta_{t+1}).
-\end{equation}
-\end{proposition}
-\begin{proof}
-If we invoke Lemma~\ref{lem:bregman} with $f=R$ and $z_t=-u_t$, then we get 
-the inequality
-\begin{align}
-\frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta_t) 
-&\leq \frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta^*) \\
-&\quad - \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). \nonumber
-\end{align}
-Re-arranging yields
-\begin{align}
-\frac{1}{A_T }\sum_{t=1}^T \alpha_t L(u_t, \theta^*) 
-&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
-&\quad+ \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). 
-\end{align}
-Now, we have the following string of inequalities:
-\begin{align*}
-L(\hat{u}, \theta) &= L\left(\frac{1}{A_T} \sum_{t=1}^T \alpha_t u_t, \theta\right) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_R(\theta_t \| \theta_{t+1}) \\
- &= \frac{1}{A_T} \sum_{t=1}^T \alpha_t \inf_{u} L(u, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
- &= \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}),
-\end{align*}
-as was to be shown.
-\end{proof}
-
-% Convergence rate theorem
-\begin{theorem}
-\label{thm:method-2}
-Suppose that $R$ is strongly convex with respect to a norm $\|\cdot\|$ 
-and let $r = \sup_{u} \|u\|_{*}$. Then 
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
-\end{theorem}
-\begin{corollary}
-\label{cor:method-2}
-Under the hypotheses of Theorem \ref{thm:method-2}, for $\alpha_t = 1$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2(\log(T) + 1)}{T} \]
-and for $\alpha_t = t$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T} \]
-\end{corollary}
-\begin{proof}[Proofs]
-The proofs are identical to those for Algorithm 1.
-\end{proof}
-
-\section{Application to Herding}
-\label{sec:herding}
-
-\NA{JHH: Need to specify the spaces we are working with here.}
-Suppose we are given a family of features $\phi$ together 
-with a known value $\bar{\phi}$ for $\bE_{u}[\phi]$. We 
-would like to construct a distribution that approximately 
-matches this distribution, i.e.~one for which 
-$\bE_{\hat{u}}[\phi] \approx \bar{\phi}$. A natural way 
-to do this is by using boosted mirror descent to minimize the 
-maximum mean discrepancy relative to $\phi$; 
-in other words, to minimize
-\[ L(u) = \max_{i \in \{1,\dots,d\}} |\bE_{u}[\phi_i]-\bar{\phi}_i|. \]
-If we wish to write this more similarly to the original problem 
-formulation, we can write
-\[ L(u) = \sup_{\|\theta\|_1 \leq 1} \theta^T(\bE_{u}[\phi]-\bar{\phi}). \]
-One issue is that $\|\theta\|_1$ is not strongly convex, so we relax $\|\theta\|_1$ 
-to $\|\theta\|_2$ and write our loss function in the Lagrangian form:
-\[ L(u, \theta) = \theta^T\bE_{u}[\phi] - \theta^T\bar{\phi} - \frac{1}{2}\|\theta\|_2^2. \]
-Intriguingly, Algorithm 1 and Algorithm 2 are identical to each other for this choice of $L$ 
-(up to a change of index). In this case, $u_t = \delta_{x_t}$ for some $x_t$, and 
-the updates are (for Algorithm 2):
-\begin{itemize}
-\item $x_t \in \arg\min_{x} \theta_t^T\phi(x)$
-\item $\theta_{t+1} = \frac{1}{A_t} \sum_{s=1}^t \alpha_s \phi(x_{s}) - \bar \phi$.
-%\item $\theta_{t+1} = \frac{1}{A_t} \sum_{s=1}^t \alpha_s \phi(x)$.
-\end{itemize}
-These are the same as the standard herding updates when 
-$\alpha_t = 1$. Note that $R$ is strongly convex with respect to 
-$\|\cdot\|_2$, which is its own dual. Therefore, if we let
-$r = \sup_{x} \|\phi(x)\|_2$, then Corollary~\ref{cor:method-2} gives us
-\begin{align*}
-\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
- &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
- &= \frac{2r^2\log(T+1)}{T}.
-\end{align*}
-This is a slightly weaker version of the typical herding bound on MMD.
-Note that if we use $\alpha_t = t$, which corresponds to using a weighted 
-average, then we obtain a stronger bound of $\frac{8r^2}{T}$ instead.
-
-\section{Herding in Infinite Dimensions}
-\label{sec:infinite-case}
-
-\section{Herding and Maximum Entropy}
-\label{sec:max-ent}
-
-Bach et al.~have observed that herding appears to be approximately building 
-samples from the maximum entropy distribution for a given collection of moment 
-constraints. In this section we will make this connection more explicit 
-as well as draw other relationships between herding and maximum entropy.
-
-Suppose that instead of setting $h(u) = 0$ we set $h(u)$ to a small multiple of 
-the negative entropy: $h(u) = \eta \bE_{u}[\log u(x)]$. We will still use 
-$R(\theta) = \frac{1}{2}\|\theta\|_2^2$. In this case, Algorithm 2 gives the updates
-\begin{itemize}
-\item $u_t(x) \propto \exp\left(-\frac{1}{\eta}\theta_t^T\phi(x)\right)$
-\item $\theta_{t+1} = \bE_{\hat{u}_t}[\phi(x)] - \bar\phi$.
-\end{itemize}
-Corollary~\ref{cor:method-2} implies that for all $u^{*}$,
-\begin{align*}
-\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) &\leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
- &= -\eta H(u^*) + \frac{1}{2}\|\bE_{u^*}[\phi(x)]-\bar{\phi}\|_2^2 \\ &\phantom{=} + \frac{8r^2}{T}.
-\end{align*}
-In particular, let $u^*$ be the maximum entropy distribution satisfying $\bE_{u^*}[\phi(x)] = \bar{\phi}$, 
-and let $H^*$ be the corresponding entropy. Then we have 
-\[ -\eta H(\hat{u}) + \|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq -\eta H^* + \frac{8r^2}{T}. \]
-Since $\|\cdot\|_2^2 \geq 0$, this implies that 
-\begin{equation}
-H(\hat{u}) \geq H^* - \frac{8r^2}{\eta T},
-\end{equation}
-and so asymptotically $\hat{u}$ has at least 
-as much entropy as $u^*$.
-
-Now, suppose that our state space has some finite size $S$, so 
-that no distribution has entropy greater than $\log(S)$. Then 
-we conclude that
-\begin{equation}
-\|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq \eta[\log(S)-H^*] + \frac{8r^2}{T}.
-\end{equation}
-
-If we let $\eta \to 0$ and $\eta T \to \infty$, then we obtain 
-a distribution $\hat{u}$ whose moments are arbitrarily close to 
-the moments of $u^*$, and whose entropy is almost as large as 
-(or larger than) the entropy of $u^*$.
-
-% \begin{lemma}
-% blah
-% \end{lemma}
-% \begin{proof}
-% Let $L(u) = -H(u) + \sup_{\theta} \theta^T(\bE_{u}[\phi(x)]-\bar{\phi})$. 
-% Clearly $L$ is minimized at the maximum entropy distribution $u^*$ satisfying 
-% $\bE_{u}[\phi(x)] = \bar{\phi}$. Let $\theta^*$ be the natural parameters 
-% for $u^*$, then we can lower bound $L(u)$ by $-H(u^*) + (\theta^*)^T(\bE_{u}[\phi(x)]-\bar{\phi})$.
-% 
-% Since $-\lambda H(u) + \|\bE_{u}[\phi(x)]-\bar{\phi}\|_2^2$ is convex, we have
-% $-\lambda H(u) + \|\bE_{u}[\phi(x)]-\bar{\phi}\|_2^2 \geq -\lambda H(u_0) + \|\bE_{u_0}[\phi(x)]-\bar{\phi}\|_2^2 + (\bE_{u_0}[\phi(x)]-\bar{\phi})^T(u-u_0)$.
-% Setting $u = u^*$ yields
-% \[ -\lambda H(u_0) + \frac{1}{2} \|\bE_{u_0}[\phi(x)]-\bar{\phi}\|_2^2 + (\bE_{u_0}[\phi(x)]-\bar{\phi})^T(u-u_0) \leq -\lambda H(u^*). \]
-% 
-% \end{proof}
-
-%\subsection{Old stuff}
-%
-% Let us now consider the entropic regularizer $h(u) = \tau\bE_{u}[\log u(x)]$ with 
-% \[ R(\theta) = \bE_{u}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.\]
-% If we use Method 1, we then end up with 
-% \[ u_{t}(x) \propto \exp\left\{-\frac{1}{\alpha}\overline{\theta}_{t}^T\phi(x)\right\}, \]
-% where $\overline{\theta}_{t} = \frac{1}{t} \sum_{s \le t} \theta_{s}$.
-Hence, $h(u) + \bE_{\mu}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the boundary (when $\tau > 0$).
-We can instead consider the following sequence of distributions:
-% \begin{enumerate}
-% \item $\theta_t \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{u_t}[\phi(x)]-\bar{\phi})$
-% \item $u_{t+1}(x) = Z(\overline{\theta}_{t}, \tau)^{-1} \exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\}$
-% \end{enumerate}
-% Here $Z(\overline{\theta}_{t}, \tau) = \int\exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\} \dee x$. 
-% 
-% The entropic regularizer is not strongly convex but a more general analysis involving 
-% Bregman divergence can nevertheless yield a convergence bound in this case. We thus 
-% get something that looks like herding but actually gives us an exponential family 
-% type distribution. In particular, we have a Gibbs distribution with temperature $\tau$, and 
-% in the limit as the temperature goes to zero, we recover herding, since 
-% \[ \lim_{\tau \to 0} u_{t+1}(x) = \delta_{x^{*}_{t+1}}, \]
-% where $x^{*}_{t+1} = \argmin_{x} \overline{\theta}_{t}^T\phi(x)$. 
-% This fact is reminiscent of Welling's original derivation of herding, in which 
-% he took the zero temperature limit of the gradient descent algorithm for the optimization problem 
-% \[ \argmin_{u} h(u) \quad \text{s.t.}~\bE_{u}[\phi(x)] - \bar\phi = 0. \]
-% This optimization is quite similar to the entropic optimization considered here, which 
-% we can write as 
-% \begin{align*}
-% \argmin_{u}& \argmax_{|\theta\|_1 \leq 1} h(u) + \theta^{T}(\bE_{u}[\phi(x)] - \bar \phi)  \\
-% &= \argmin_{u} h(u) + \|\bE_{u}[\phi(x)] - \bar\phi\|_{2}. 
-% \end{align*}
-% In the entropic formulation, the expectation constraint $\bE_{u}[\phi(x)] - \bar\phi = 0$ becomes
-% soft, though the two optimizations are equivalent in the zero-temperature limit. 
-% 
-% We can view the entropic regularization
-% with non-zero temperature as an alternative approximation to herding. If at each step we
-% retain a sample $x_{t} \sim u_{t}$, then the sequence $(x_{t})_{t}$ converges to the 
-% herding sample path as $\tau \to 0$. So $(x_{t})_{t}$ with non-zero temperature approximate
-% herding samples. \NA{JHH: I'm not sure how useful this all is, other than to make contact with 
-% the original paper, as I think the Gibbs distribution was exactly what Max was trying to 
-% avoid by taking the zero-temperature limit.}
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-%\bibliography{example_paper}
-%\bibliographystyle{icml2013}
-
 \end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was
-% created by Lise Getoor and Tobias Scheffer, it was slightly modified  
-% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz, 
-% slightly modified from the 2009 version by Kiri Wagstaff and 
-% Sam Roweis's 2008 version, which is slightly modified from 
-% Prasad Tadepalli's 2007 version which is a lightly 
-% changed version of the previous year's version by Andrew Moore, 
-% which was in turn edited from those of Kristian Kersting and 
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2013-10-03 icml/paper.tex
split into subfiles

--- 2013-09-13 icml/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%% ICML 2013 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \documentclass{article}
-\usepackage{import, icml2013}
+\usepackage{import, subfiles}
+\usepackage{hyperref, icml2014}
 \input{latex-defs.tex}
@@
 
-\section{Introduction}
-\label{sec:intro}
-
-\section{Boosted Mirror Descent}
-\label{sec:algorithm}
-
-\NA{JHH: I'm thinking that from the beginning we should let $u$ be defined over a different space than $\theta$ and assume a linear map $u \mapsto \eta(u) \in \Theta$.
-Then write $\langle \theta, \eta(u) \rangle$ so the setup will align with the herding case, where $\eta(u) = \bE_{u}[\phi(x)]$.
-}
-
-Consider the following loss function:
-\begin{equation}
-L(u) = h(u) + \max_{\theta \in \Theta} \left\{\langle \theta, u \rangle - R(\theta) \right\}.
-\end{equation}
-We can think of $h$ as a \emph{primal regularizer} and $R$ as a \emph{dual regularizer}.
-
-Examples:
-\begin{itemize}
-\item Let $h(u) = 0$ and $R(\theta) = \langle \theta, u_0 \rangle$. Then $L(u)$ is the 
-      \emph{maximum mean discrepancy} between $u$ and $u_0$ 
-      relative to $\Theta$.
-\item Let $h(u) = 0$ and $R(\theta) = S^*(\theta)$ where $S^*$ is the Fenchel conjugate of 
-      any strongly convex function $S$. Then $L(u) = S(u)$.
-\item Let $h(u) = \|u\|_1$, $R(u) = \langle \theta, u_0 \rangle + \frac{1}{2} \|\theta\|_2^2$. 
-      Then $L(u) = \|u\|_1 + \frac{1}{2} \|u-u_0\|_2^2$. 
-\end{itemize}
-
-In this paper we will consider a family of methods for minimizing $L(u)$, 
-which are based on Bach's generalization of the Frank-Wolfe algorithm and can 
-be interpreted as boosted mirror descent.
-
-First, let us generalize the setting 
-to consider a \emph{two-argument} loss function
-\begin{equation}
-L(u,\theta) = h(u) + \langle \theta, u \rangle - R(\theta).
-\end{equation}
-We will assume throughout that $h$ and $R$ are both convex functions, 
-and furthermore that $\arg\max_{u} h(u) + \theta^Tu$ can be efficiently 
-computed for all values of $\theta$.
-
-Mirror descent, together with boosting, yields an algorithm for finding 
-``saddle points'' of $L$. It is the following algorithm (called Algorithm 1).
-For notational convenience, for a sequence of weights $\alpha_1,\alpha_2,\ldots$ 
-let $\hat{u}_t = \frac{\sum_{s=1}^t \alpha_su_s}{\sum_{s=1}^t \alpha_s}$ and let 
-$\hat{\theta}_t = \frac{\sum_{s=1}^t \alpha_s\theta_s}{\sum_{s=1}^t \alpha_s}$.
-\begin{enumerate}
-\item $u_1 \in \arg\min_u h(u)$
-\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, u_t \rangle - R(\theta)$
-\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, u \rangle$
-%\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
-\end{enumerate}
-As long as $h$ is strongly convex, we obtain the 
-bound (see Corollary~\ref{cor:method-1}):
-\begin{equation}
-\sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/T).
-\end{equation}
-In other words, $\hat{u}_T$ is close to being a global minimum of $L$.
-However, it is often the case that $h$ is not strongly convex, whereas $R$ is strongly convex. In this case 
-we may wish to use the following slightly different algorithm (called Algorithm 2):
-\begin{enumerate}
-\item $\theta_1 \in \arg\min_{\theta} R(\theta)$
-\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$
-%\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
-\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \hat{u}_t \rangle - R(\theta)$
-\end{enumerate}
-We then obtain the following bound 
-when $R$ is strongly convex (see Corollary~\ref{cor:method-2}):
-\[ \sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/T). \]
-Algorithm 1 and Algorithm 2 are closely related; indeed, they are dual to each other 
-(performing Algorithm 1 on $L(u,\theta)$ is the same as performing Algorithm 2 on 
-$-L(\theta,u)$).
-
-\section{Convergence Proofs}
-\label{sec:proofs}
-We now prove the convergence results cited in Section~\ref{sec:algorithm}. 
-Throughout this section, assume that $\alpha_1,\ldots,\alpha_T$ is 
-a sequence of real numbers and that $A_t = \sum_{s=1}^t \alpha_s$. 
-We further require that $A_t \geq 0$ for all $t$.
-
-Also recall that the Bregman divergence is defined by 
-$D_f(x_2 \| x_1) \eqdef f(x_2) - \langle \nabla f(x_1), x_2-x_1 \rangle - f(x_1)$.
-
-Our proofs hinge on the following key lemma:
-%%% MAIN LEMMA %%%
-\begin{lemma}
-\label{lem:bregman}
-Let $z_1,\ldots,z_T$ be vectors and let $f(x)$ be a strictly convex 
-function. Define $\hat{z}_t$ to be $\frac{1}{A_t} \sum_{s=1}^t \alpha_s z_s$.
-
-Let $x_1,\ldots,x_T$ be chosen via $x_{t+1} = \arg\min_{x} f(x) + \langle \hat{z}_t, x\rangle$. 
-Then for any $x^*$ we have
-\begin{align*}
-\lefteqn{\frac{1}{A_T} \sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
-\phantom{+} &\leq f(x^*) + \langle \hat{z}_t, x^* \rangle + \frac{1}{A_T} \sum_{t=1}^T A_t D_{f}(x_t \| x_{t+1}). 
-\end{align*}
-\end{lemma}
-\begin{proof}
-First note that, if $x_0 = \arg\min f(x) + \langle z, x \rangle$, 
-then $\nabla f(x_0) = -z$.
-
-\NA{JNS: In the following proof, (2) is by the update formula for $x_{t+1}$ together with the preceding observation.
-(3) is applying (2).
-(4) and (5) are re-arranging the sum.
-(6) is definition of Bregman divergence.
-(7) is the observation again.
-(8) is the definition of $x_{T+1}$.}
-
-Now note that
-\begin{align}
-\alpha_{t}z_{t} 
-&= A_{t}\hat z_{t} - A_{t-1}\hat z_{t-1} \\
-&= - A_{t}\nabla f(x_{t+1}) + A_{t-1} \nabla f(x_{t}),
-\end{align}
-so we have
-\begin{align}
-\lefteqn{\sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
- &= \sum_{t=1}^T \{\alpha_t f(x_t) + \langle A_{t-1} \nabla f(x_t) - A_t \nabla f(x_{t+1}), x_t \rangle\} \\
- &= \sum_{t=1}^T \{\alpha_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle\} \\
- & \quad - A_{T}\langle \nabla f(x_{T+1}), x_{T+1} \rangle
-\end{align}
-\begin{align}
- &= \sum_{t=1}^T \{A_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle - A_t f(x_{t+1})\}  \\
- &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
- &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\}  \\
- &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
- &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) + \langle \hat{z}_T, x_{T+1} \rangle) \\
- &\leq \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x^*) + \langle \hat{z}_T, x^* \rangle). 
-\end{align}
-Dividing both sides by $A_T$ completes the proof.
-\end{proof}
-We also note that $D_f(x_t \| x_{t+1}) = D_{f^*}(\hat{z}_{t+1} \| z_t)$, where $f^*(z) = \sup_x \langle z,x\rangle - f(x)$. 
-This form of the bound will often be more useful to us.
-\begin{lemma}
-\label{lem:convexity}
-Suppose that $D_f(x' \| x) \geq \frac{1}{2}\|x-x'\|^2$ for some 
-norm $\|\cdot\|$. (In this case we say that $f$ is strongly 
-convex with respect to $\|\cdot\|$.) 
-Then $D_{f^*}(x' \| x) \leq \frac{1}{2}\|x-x'\|_{*}^2$.
-\end{lemma}
-
-%%% ALGORITHM 1 ANALYSIS %%%
-% Convergence proposition
-\begin{proposition}[Convergence of Algorithm 1]
-\label{prop:method-1}
-Consider the updates $\theta_t \in \arg\max_{\theta} \langle \theta, u_t \rangle - R(\theta)$ 
-and $u_{t+1} \in \arg\min_u h(u) + \langle \hat{\theta}_s, u \rangle$. 
-Then we have
-\begin{equation}
-\sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}).
-\end{equation}
-\end{proposition}
-\begin{proof}
-Note that $L(u_t, \theta_t) = \max_{\theta} L(u_t, \theta)$ by construction. 
-Also note that, if we invoke Lemma~\ref{lem:bregman} with $f = h$ and 
-$z_t = \theta_t$ then we get the inequality
-\begin{align}
-\frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) 
-&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) \\
-&\quad + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}). \nonumber
-\end{align}
-Combining these together, we get the string of inequalities
-\begin{align*}
-L(\hat{u}_T, \theta) &= L\left(\frac{1}{A_T} \sum_{t=1}^T \alpha_tu_t, \theta\right) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_h(u_t \| u_{t+1}) \\
- &\leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}),
-\end{align*}
-as was to be shown.
-\end{proof}
-
-% Convergence rate theorem
-\begin{theorem}
-\label{thm:method-1}
-Suppose that $h$ is strongly convex with respect to a norm $\|\cdot\|$ 
-and let $r = \sup_{\theta} \|\theta\|_{*}$. Then 
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
-\end{theorem}
-\begin{proof}
-By Lemma~\ref{lem:convexity}, we have 
-\begin{align*}
-D_h(u_t \| u_{t+1}) &= D_{h^*}(\hat{\theta}_{t+1} \| \hat{\theta}_{t}) \\
- &\leq \frac{1}{2}\|\hat\theta_{t+1}-\hat\theta_{t}\|_{*}^2 \\
- &= \frac{1}{2}\|\frac{\sum_{s \leq t}\alpha_s\theta_s}{\sum_{s \leq t} \alpha_s} - \frac{\sum_{s \leq t+1}\alpha_s\theta_s}{\sum_{s \leq t+1} \alpha_s}\|_{*}^2 \\
- &= \frac{1}{2}\|\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\theta_s - \frac{\alpha_{t+1}}{A_{t+1}} \theta_{t+1}\|_{*}^2 \\
- &\leq \frac{1}{2} \left(\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\|\theta_s\|_{*}^2 + \frac{\alpha_{t+1}}{A_{t+1}} \|\theta_{t+1}\|_{*}^2\right)^2 \\
- &\leq \frac{2r^2\alpha_{t+1}^2}{A_{t+1}^2}.
-\end{align*}
-It follows that 
-\begin{align*}
-\frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}) &\leq \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}.
-\end{align*}
-\end{proof}
-
-% Convergence rate corollary 
-\begin{corollary} 
-\label{cor:method-1}
-Under the hypotheses of Theorem \ref{thm:method-1}, for $\alpha_{t} = 1$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2 (\log (T) + 1)}{T}. \]
-and for $\alpha_t = t$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T}. \]
-\end{corollary}
-\begin{proof}
-If we let $\alpha_t = 1$, then $A_t = t$ and $\frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} = \frac{t}{(t+1)^{2}} \le \frac{1}{t}$.
-We therefore get
-\begin{equation}
-\frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} \leq \frac{2r^2}{T} \sum_{t=1}^{T}\frac{1}{t} \leq \frac{2r^2(\log (T) + 1)}{T+1}.
-\end{equation}
-If we let $\alpha_t = t$, then $A_t = \frac{t(t+1)}{2}$ and 
-$\frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} = \frac{2(t+1)^2t(t+1)}{(t+1)^2(t+2)^2} = \frac{2t(t+1)}{(t+2)^2} \leq 2$.
-We therefore get
-\begin{equation}
-\frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} \leq \frac{4r^2}{T(T+1)} \sum_{t=1}^{T}2 = \frac{8r^2}{T+1},
-\end{equation}
-which completes the proof.
-\end{proof}
-
-%%% ALGORITHM 2 ANALYSIS %%%
-% Convergence proposition
-\begin{proposition}[Convergence of Algorithm 2]
-\label{prop:method-2}
-Consider the updates $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$ 
-and $\theta_{t+1} \in \arg\max_{\theta} \langle \theta, \hat{u}_t \rangle - R(\theta)$. 
-Then we have 
-\begin{equation}
-\sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_{R}(\theta_t \| \theta_{t+1}).
-\end{equation}
-\end{proposition}
-\begin{proof}
-If we invoke Lemma~\ref{lem:bregman} with $f=R$ and $z_t=-u_t$, then we get 
-the inequality
-\begin{align}
-\frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta_t) 
-&\leq \frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta^*) \\
-&\quad - \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). \nonumber
-\end{align}
-Re-arranging yields
-\begin{align}
-\frac{1}{A_T }\sum_{t=1}^T \alpha_t L(u_t, \theta^*) 
-&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
-&\quad+ \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). 
-\end{align}
-Now, we have the following string of inequalities:
-\begin{align*}
-L(\hat{u}, \theta) &= L\left(\frac{1}{A_T} \sum_{t=1}^T \alpha_t u_t, \theta\right) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_R(\theta_t \| \theta_{t+1}) \\
- &= \frac{1}{A_T} \sum_{t=1}^T \alpha_t \inf_{u} L(u, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
- &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
- &= \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}),
-\end{align*}
-as was to be shown.
-\end{proof}
-
-% Convergence rate theorem
-\begin{theorem}
-\label{thm:method-2}
-Suppose that $R$ is strongly convex with respect to a norm $\|\cdot\|$ 
-and let $r = \sup_{u} \|u\|_{*}$. Then 
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
-\end{theorem}
-\begin{corollary}
-\label{cor:method-2}
-Under the hypotheses of Theorem \ref{thm:method-2}, for $\alpha_t = 1$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2(\log(T) + 1)}{T} \]
-and for $\alpha_t = t$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T} \]
-\end{corollary}
-\begin{proof}[Proofs]
-The proofs are identical to those for Algorithm 1.
-\end{proof}
-
-\section{Application to Herding}
-\label{sec:herding}
-
-\NA{JHH: Need to specify the spaces we are working with here.}
-Suppose we are given a family of features $\phi$ together 
-with a known value $\bar{\phi}$ for $\bE_{u}[\phi]$. We 
-would like to construct a distribution that approximately 
-matches this distribution, i.e.~one for which 
-$\bE_{\hat{u}}[\phi] \approx \bar{\phi}$. A natural way 
-to do this is by using boosted mirror descent to minimize the 
-maximum mean discrepancy relative to $\phi$; 
-in other words, to minimize
-\[ L(u) = \max_{i \in \{1,\dots,d\}} |\bE_{u}[\phi_i]-\bar{\phi}_i|. \]
-If we wish to write this more similarly to the original problem 
-formulation, we can write
-\[ L(u) = \sup_{\|\theta\|_1 \leq 1} \theta^T(\bE_{u}[\phi]-\bar{\phi}). \]
-One issue is that $\|\theta\|_1$ is not strongly convex, so we relax $\|\theta\|_1$ 
-to $\|\theta\|_2$ and write our loss function in the Lagrangian form:
-\[ L(u, \theta) = \theta^T\bE_{u}[\phi] - \theta^T\bar{\phi} - \frac{1}{2}\|\theta\|_2^2. \]
-Intriguingly, Algorithm 1 and Algorithm 2 are identical to each other for this choice of $L$ 
-(up to a change of index). In this case, $u_t = \delta_{x_t}$ for some $x_t$, and 
-the updates are (for Algorithm 2):
-\begin{itemize}
-\item $x_t \in \arg\min_{x} \theta_t^T\phi(x)$
-\item $\theta_{t+1} = \frac{1}{A_t} \sum_{s=1}^t \alpha_s \phi(x)$.
-\end{itemize}
-These are the same as the standard herding updates when 
-$\alpha_t = 1$. Note that $R$ is strongly convex with respect to 
-$\|\cdot\|_2$, which is its own dual. Therefore, if we let
-$r = \sup_{x} \|\phi(x)\|_2$, then Corollary~\ref{cor:method-2} gives us
-\begin{align*}
-\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
- &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
- &= \frac{2r^2\log(T+1)}{T}.
-\end{align*}
-This is a slightly weaker version of the typical herding bound on MMD.
-Note that if we use $\alpha_t = t$, which corresponds to using a weighted 
-average, then we obtain a stronger bound of $\frac{8r^2}{T}$ instead.
-
-\section{Herding in Infinite Dimensions}
-\label{sec:infinite-case}
-
-\section{Herding and Maximum Entropy}
-\label{sec:max-ent}
-
-Bach et al. have observed that herding appears to be approximately building 
-samples from the maximum entropy distribution for a given collection of moment 
-constraints. In this section we will make this connection more explicit 
-as well as draw other relationships between herding and maximum entropy.
-
-Suppose that instead of setting $h(u) = 0$ we set $h(u)$ to a small multiple of 
-the negative entropy: $h(u) = \eta \bE_{u}[\log u(x)]$. We will still use 
-$R(\theta) = \frac{1}{2}\|\theta\|_2^2$. In this case, Algorithm 2 gives the updates
-\begin{itemize}
-\item $u_t(x) \propto \exp\left(-\frac{1}{\eta}\theta_t^T\phi(x)\right)$
-\item $\theta_{t+1} = \bE_{\hat{u}_t}[\phi(x)]$.
-\end{itemize}
-Using Corollary~\ref{cor:method-2} gives us
-\begin{align*}
-\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
- &= \inf_{u^*} \left\{-\eta H(u^*) + \frac{1}{2}\|\bE_{u^*}[\phi(x)]-\bar{\phi}\|_2^2\right\} \\ &\phantom{=} + \frac{2r^2\log(T+1)}{T}.
-\end{align*}
-In particular, let $u^*$ be the maximum entropy distribution satisfying $\bE_{u^*}[\phi(x)] = \bar{\phi}$, 
-and let $H^*$ be the corresponding entropy. Then we have 
-\[ -\eta H(\hat{u}) + \|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq -\eta H^* + \frac{2r^2\log(T+1)}{T}. \]
-Since $\|\cdot\|_2^2 \geq 0$, this implies that 
-\begin{equation}
-H(\hat{u}) \geq H^* - \frac{2r^2\log(T+1)}{\eta T},
-\end{equation}
- and so asymptotically $\hat{u}$ has at least 
-as much entropy as $u^*$.
-
-Now, suppose that our state space has some finite size $S$, so 
-that no distribution has entropy greater than $\log(S)$. Then 
-we conclude that
-\begin{equation}
-\|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq \eta[\log(S)-H^*] + \frac{2r^2\log(T+1)}{T}.
-\end{equation}
-
-If we let $\eta \to 0$ and $\eta T \to \infty$, then we obtain 
-a distribution $\hat{u}$ whose moments are arbitrarily close to 
-the moments of $u^*$, and whose entropy is almost as large as 
-(or larger than) the entropy of $u^*$.
-
-% \begin{lemma}
-% blah
-% \end{lemma}
-% \begin{proof}
-% Let $L(u) = -H(u) + \sup_{\theta} \theta^T(\bE_{u}[\phi(x)]-\bar{\phi})$. 
-% Clearly $L$ is minimized at the maximum entropy distribution $u^*$ satisfying 
-% $\bE_{u}[\phi(x)] = \bar{\phi}$. Let $\theta^*$ be the natural parameters 
-% for $u^*$, then we can lower bound $L(u)$ by $-H(u^*) + (\theta^*)^T(\bE_{u}[\phi(x)]-\bar{\phi})$.
-% 
-% Since $-\lambda H(u) + \|\bE_{u}[\phi(x)]-\bar{\phi}\|_2^2$ is convex, we have
-% $-\lambda H(u) + \|\bE_{u}[\phi(x)]-\bar{\phi}\|_2^2 \geq -\lambda H(u_0) + \|\bE_{u_0}[\phi(x)]-\bar{\phi}\|_2^2 + (\bE_{u_0}[\phi(x)]-\bar{\phi})^T(u-u_0)$.
-% Setting $u = u^*$ yields
-% \[ -\lambda H(u_0) + \frac{1}{2} \|\bE_{u_0}[\phi(x)]-\bar{\phi}\|_2^2 + (\bE_{u_0}[\phi(x)]-\bar{\phi})^T(u-u_0) \leq -\lambda H(u^*). \]
-% 
-% \end{proof}
-
-%% Let us now consider the entropic regularizer $h(u) = \tau\bE_{u}[\log u(x)]$ with 
-%% \[ R(\theta) = \bE_{u}[\theta^T\bar{\phi}] + \left\{ \begin{array}{llc} \infty & : & \|\theta\|_1 > 1 \\ 0 & : & \|\theta\|_1 \leq 1 \end{array} \right.\]
-%% If we use Method 1, we then end up with 
-%% \[ u_{t}(x) \propto \exp\left\{-\frac{1}{\alpha}\overline{\theta}_{t}^T\phi(x)\right\}, \]
-%% where $\overline{\theta}_{t} = \frac{1}{t} \sum_{s \le t} \theta_{s}$.
-Hence, $h(u) + \bE_{\mu}[\theta^T\phi(x)] - R(\theta)$ is no longer maximized at the boundary (when $\tau > 0$).
-We can instead consider the following sequence of distributions:
-%% \begin{enumerate}
-%% \item $\theta_t \in \argmax_{\|\theta\|_1 \leq 1} \theta^T (\bE_{u_t}[\phi(x)]-\bar{\phi})$
-%% \item $u_{t+1}(x) = Z(\overline{\theta}_{t}, \tau)^{-1} \exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\}$
-%% \end{enumerate}
-%% Here $Z(\overline{\theta}_{t}, \tau) = \int\exp\left\{-\frac{1}{\tau}\overline{\theta}_{t}^T\phi(x)\right\} \dee x$. 
-%% 
-%% The entropic regularizer is not strongly convex but a more general analysis involving 
-%% Bregman divergence can nevertheless yield a convergence bound in this case. We thus 
-%% get something that looks like herding but actually gives us an exponential family 
-%% type distribution. In particular, we have a Gibbs distribution with temperature $\tau$, and 
-%% in the limit as the temperature goes to zero, we recover herding, since 
-%% \[ \lim_{\tau \to 0} u_{t+1}(x) = \delta_{x^{*}_{t+1}}, \]
-%% where $x^{*}_{t+1} = \argmin_{x} \overline{\theta}_{t}^T\phi(x)$. 
-%% This fact is reminiscent of Welling's original derivation of herding, in which 
-%% he took the zero temperature limit of the gradient descent algorithm for the optimization problem 
-%% \[ \argmin_{u} h(u) \quad \text{s.t.}~\bE_{u}[\phi(x)] - \bar\phi = 0. \]
-%% This optimization is quite similar to the entropic optimization considered here, which 
-%% we can write as 
-%% \begin{align*}
-%% \argmin_{u}& \argmax_{|\theta\|_1 \leq 1} h(u) + \theta^{T}(\bE_{u}[\phi(x)] - \bar \phi)  \\
-%% &= \argmin_{u} h(u) + \|\bE_{u}[\phi(x)] - \bar\phi\|_{2}. 
-%% \end{align*}
-%% In the entropic formulation, the expectation constraint $\bE_{u}[\phi(x)] - \bar\phi = 0$ becomes
-%% soft, though the two optimizations are equivalent in the zero-temperature limit. 
-%% 
-%% We can view the entropic regularization
-%% with non-zero temperature as an alternative approximation to herding. If at each step we
-%% retain a sample $x_{t} \sim u_{t}$, then the sequence $(x_{t})_{t}$ converges to the 
-%% herding sample path as $\tau \to 0$. So $(x_{t})_{t}$ with non-zero temperature approximate
-%% herding samples. \NA{JHH: I'm not sure how useful this all is, other than to make contact with 
-%% the original paper, as I think the Gibbs distribution was exactly what Max was trying to 
-%% avoid by taking the zero-temperature limit.}
-
-\section{Conclusion}
-\label{sec:conclusion}
+\subfile{introduction.tex}
+\subfile{herding-connection.tex}
+\subfile{convergence-proofs.tex}
+%\subfile{lower-bounds.tex}
+%\subfile{chen-proofs.tex}
+%\subfile{conclusion.tex}
 

### 2013-10-04 icml/introduction.tex
first draft of introduction

@@
 \begin{document}
+
+\section{Introduction} 
+\label{sec:intro}
+
+\NA{JHH: Need to add in all these citations}
+
+The herding algorithm, which was first presented in \citet{Welling:2009a}, has received increasing attention since its introduction.
+Herding was originally introduced as a deterministic method for performing joint learning and inference in Markov random field (MRFs).
+This derivation was based on performing fixed step size gradient descent on the zero temperature limit the maximum entropy optimization problem for an MRF.
+Herding for MRFs was further explored in \citet{Welling:2009a,Gelfand:2010,Bornn:2013}.
+
+Let $\sX$ be an observation space, $\phi : \sX \to \sB_{\Theta}$ a feature map to an inner produce space, and $\bar\phi = \bE_{x \sim p}[\phi(x)]$ the empirical moments of some (possible empirical) distribution.
+Then the herding algorithm is given by the follow simple procedure:
+\[
+x_{t} &\in \argmin_{x \in \sX} \ip{\theta_{t}}{\phi(x)} \\
+\theta_{t} &= \theta_{t-1}  + \bar \phi - \phi(x_{t}).
+\]
+
+An alternative interpretation of herding \citep{Chen:2010a,Huszar:2012,Bach:2012a} is as a moment matching algorithm.
+In particular, it generates a set of pseudo-samples $\{ x_{t} \}$  which match the moments $\bar\phi$.
+That is, the error $\sE_{T}^{2} \eqdef \| \bar\phi - T^{-1}\sum_{t=1}^{T} \phi(x_{t})\|^{2} \to 0$ as $T \to \infty$.
+If $\sB_{\Theta}$ is a reproducing kernel Hilbert space generated by $\phi$, $\sE_{T}^{2}$ is equivalent to the maximum mean discrepancy (MMD) between $p$ and the empirical distribution $\hat p_{T} \eqdef T^{-1}\sum_{t=1}^{T} \delta_{x_{t}}$ \citep{Huszar:2012}:
+\(
+\sE_{T}^{2} &= \text{MMD}^{2}(p, \hat p_{T})  \\
+&\eqdef \sup_{\theta \in \Theta} \left|\bE_{p}\ip{\theta}{\phi(x)} - \bE_{\hat p_{T}}\ip{\theta}{\phi(x)} \right|,
+\)
+where $\Theta = \{ \theta \in \sB_{\Theta} \ | \  \|\theta\| = 1 \}$.
+
+If we are interested in generating samples from $p$ (or some distribution with the same moments as $p$), herding provides an attractive alternative to standard i.i.d.
+sampling for two reasons.
+First, it is deterministic, so we do not have to worry about getting ``unlucky'' and generating a set of samples that poorly match $p$.
+Second, in the case where $\sB_{\theta}$ is finite dimensional, the error $\sE_{T}$ of the herding estimator converges to zero at a rate of $O(1/T^{2})$ whereas the convergence rate of i.i.d.~sampling is $O(1/T)$ \citep{Chen:2010a}.
+The infinite dimensional case, however, is not as well understood.
+\citet{Bach:2012a} recently showed that the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} does not apply in the infinite dimensional case.
+In addition, \citet{Bach:2012a} introduced an interpretation of herding as a special case of conditional gradient descent (\algname{CGD}) minimizing the square norm.
+They showed that \algname{CGD} with line search produces faster convergence rates: $O(1/T)$ in the infinite dimensional case and exponential in the finite dimensional valse.
+However, they found that empirically that line search had poor approximation properties since led to distributions far from the maximum entropy distribution.
+
+In this paper, we make the following contributions:
+\begin{itemize}
+\item In Section \ref{sec:algorithm} we introduce a novel pair of convex algorithms: primal and dual boosted mirror descent (\algname{PBMD} and \algname{DBMD}).
+We show how they relate to \algname{CGD} and mirror descent (\algname{MD}).
+\item We then use \algname{DBMD} in Section \ref{sec:herding} to define a class of generalized herding algorithms.
+Connections with the learning interpretation are made by showing how to obtain herding as the limit of an entropically normalized case of \algname{BMD}.
+\item Section \ref{sec:proofs} presents convergence rates for $\algname{PBMD}$ and $\algname{DBMD}$.
+As special cases, we recover a $O(\log T/ T)$ convergence rate for herding and a $O(1/T)$ rate for a herding-like algorithm that uses unequally weighted samples.
+Both of these results apply to the infinite dimensional case.
+\item In Section \ref{sec:lower-bounds} we prove matching lower bounds  showing that ``sparse'' herding-like methods cannot converge faster than $O(1/T)$ in the the infinite dimensional case and that herding converges at the slower $O(\log T / T)$ rate.
+\item Finally, in section \ref{sec:chen} we generalize the $O(1/T^{2})$ proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \algname{BMD}. 
+\end{itemize}
+
 \section{Boosted Mirror Descent}

### 2013-10-04 icml/introduction.tex
intro improved; log(t)/t writeup added

--- 2013-10-03 icml/introduction.tex
@@
 \begin{document}
+
+\section{Introduction} 
+\label{sec:intro}
+
+The herding algorithm of \citet{Welling:2009a} has been significantly analyzed and extended since its introduction.
+Herding was originally conceived of as a deterministic method for performing joint learning and inference in Markov random field (MRFs).
+The ``learning'' interpretation of herding was derived as a fixed step size gradient descent algorithm on the zero temperature limit of a the maximum entropy optimization problem for the MRF.
+Herding for MRFs was further explored in \citet{Welling:2009a,Gelfand:2010,Bornn:2013}.
+
+In order to define herding, let $\sX$ be an observation space, $\phi : \sX \to \sB_{\Theta}$ a feature map to an inner product space, and $\bar\phi = \bE_{x \sim p}[\phi(x)]$ the empirical moments of some (possible empirical) distribution.
+Then the herding algorithm is given by the follow simple procedure to generate pseudosamples from $\sX$:
+\[
+x_{t} &\in \argmin_{x \in \sX} \ip{\theta_{t}}{\phi(x)} \\
+\theta_{t} &= \theta_{t-1}  + \bar \phi - \phi(x_{t}).
+\]
+
+An alternative interpretation of herding \citep{Chen:2010a,Huszar:2012,Bach:2012a} is as a moment matching algorithm.
+In particular, it generates a set of pseudosamples $\{ x_{t} \}$  which match the moments $\bar\phi$.
+That is, the squared error $\sE_{T}^{2} \eqdef \| \bar\phi - T^{-1}\sum_{t=1}^{T} \phi(x_{t})\|^{2} \to 0$ as $T \to \infty$.
+If $\sB_{\Theta}$ is a reproducing kernel Hilbert space (RKHS) with feature map $\phi$, $\sE_{T}^{2}$ is equivalent to the maximum mean discrepancy (MMD) between $p$ and the empirical distribution $\hat p_{T} \eqdef T^{-1}\sum_{t=1}^{T} \delta_{x_{t}}$ \citep{Huszar:2012}:
+\(
+\sE_{T}^{2} &= \text{MMD}^{2}(p, \hat p_{T})  \\
+&\eqdef \sup_{\theta \in \Theta} \left|\bE_{p}\ip{\theta}{\phi(x)} - \bE_{\hat p_{T}}\ip{\theta}{\phi(x)} \right|,
+\)
+where $\Theta = \{ \theta \in \sB_{\Theta} \ | \  \|\theta\| = 1 \}$.
+
+If we are interested in generating samples from $p$ (or some distribution with the same moments as $p$), herding provides an attractive alternative to standard i.i.d.
+sampling for two reasons.
+First, it is deterministic, so we do not have to worry about getting ``unlucky'' and generating a set of samples that poorly represent  $p$.
+Second, in the case where $\sB_{\theta}$ is finite dimensional, the squared error of the herding estimator converges to zero at a rate of $O(1/T^{2})$ whereas the convergence rate of i.i.d.~sampling is $O(1/T)$ \citep{Chen:2010a}.
+The infinite dimensional case, however, is not as well understood.
+\citet{Bach:2012a} recently showed that the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} does not apply in the infinite dimensional case.
+In addition, \citet{Bach:2012a} introduced an interpretation of herding as a special case of conditional gradient descent (\algname{CGD}) minimizing the square norm of an RKHS.
+They showed that \algname{CGD} with line search leads to faster convergence rates: $O(1/T)$ in the infinite dimensional case and exponential in the finite dimensional case.
+However, they found  empirically that line search had poor approximation properties since led to distributions far from the maximum entropy distribution.
+
+In this paper, we use a novel gradient descent algorithm to investigate herding and its generalizations:
+\begin{itemize}
+\item In Section \ref{sec:algorithm} we introduce a novel pair of convex optimization algorithms, which we call primal and dual boosted mirror descent (\algname{PBMD} and \algname{DBMD}).
+We show how \algname{PBMD} and \algname{DBMD} relate to \algname{CGD} and mirror descent (\algname{MD}).
+\item We then use \algname{DBMD} in Section \ref{sec:herding} to define a class of generalized herding algorithms.
+Connections with the ``learning'' interpretation are made by showing how to obtain herding as the limit of an entropically normalized case of \algname{BMD}.
+\item Section \ref{sec:proofs} presents convergence rates for $\algname{PBMD}$ and $\algname{DBMD}$.
+As special cases, we recover a $O(\log T/ T)$ convergence rate for herding and a $O(1/T)$ rate for a herding-like algorithm that uses unequally weighted samples.
+Both of these results apply to the infinite dimensional case.
+\item In Section \ref{sec:lower-bounds} we prove matching lower bounds to the upper bounds given in Section \ref{sec:proofs}, showing that any ``sparse'' herding-like method cannot converge faster than $O(1/T)$ in the the infinite dimensional case and that herding converges at the slower $O(\log T / T)$ rate.
+\item Finally, in section \ref{sec:chen} we generalize the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \algname{BMD}. 
+\end{itemize}
+
 \section{Boosted Mirror Descent}

### 2013-10-04 log-t-over-t-example.tex
intro improved; log(t)/t writeup added

\documentclass[reqno,oneside,a4paper]{amsart}
\usepackage{jhh-misc}
\usepackage{hyperref}
\usepackage{pdflscape}
\usepackage{epsfig}
\usepackage{framed}
\usepackage{natbib}
\usepackage{subfig}
\usepackage{xspace}
\usepackage[normalem]{ulem}

% inner product command
\newcommand{\ip}[2]{\left\langle #1, #2 \right\rangle}
\def\[#1\]{\begin{align}#1\end{align}}

\begin{document}

\title[$1/T$ counterexample?]{Counterexample to $1/T$ convergence of herding in infinite dimensions?} 
\author{Jonathan Huggins}
\date{\today}

\maketitle

We build off the same counterexample that shows $O(1/T)$ sparse convergence in infinite dimensions. The state space is $\naturals$ and the features are of the form $\phi_{n} = \delta_{n}$. Let the target distribution be $p(n) = \frac{c}{n^{1+\eps}}$, where $c$ is the normalizing constant, which is approximately $\eps$ as $\eps \to 0$. Let $k(n)$ be the number of herding samples equal to $n$, let $\bar n = \max \{ n \given k(n) \ge 1 \}$ be the largest $n$ that has been sampled, and let $t = \sum_{n=1}^{\bar n} k(n)$ be the total number of samples. Then 
\begin{align*}
\theta_{t} = \left(\frac{k(1)}{t} - c, \frac{k(2)}{t} - \frac{c}{2^{1+\eps}}, \dots, \frac{k(n)}{t} - \frac{c}{n^{1+\eps}}, \dots, \frac{k(\bar n)}{t} - \frac{c}{\bar n^{1+\eps}}, - \frac{c}{(\bar n + 1)^{1+\eps}},\dots\right)\
\end{align*}
The $k(n)$ are chosen by the herding algorithm such that 
\[
k(n) &= \min\left\{ k~\bigg|~\frac{k}{t} - \frac{c}{n^{1+\eps}} > -\frac{c}{(\bar n + 1)^{1+\eps}} \right\} \\
&= \left\lceil ct\left(\frac{1}{n^{1+\eps}} - \frac{1}{\bar n^{1+\eps}}\right)\right\rceil. \label{eq:k}
\]

The goal is to find $\bar n$ for a fixed $t$. If the convergence rate is $O(\log t / t)$, then herding should have only put mass on the first $O(t / \log t)$ elements since the error will then be $O((\log t / t)^{1 - 2\eps})$. Hence, $\bar n$ should be roughly $O(t / \log t)$.

Now, let $m(k) = \max\{ n \given k(n) = k \}$ be the maximum $n$ with $k$ samples and let $\bar k = k(1)$ be the maximum number of samples for any $n$. In this case we have the relationship
\[
t = \sum_{k=1}^{\bar k} m(k).
\]
We can approximate $m(k)$ by solving for $n$ in \eqref{eq:k}, replacing the ceiling with $+ 1$, which should be a tight upper bound:
\[
k &\approx  ct\left(\frac{1}{m(k)^{1+\eps}} - \frac{1}{\bar n^{1+\eps}}\right)  + 1\\
\implies m(k) &\approx \left(\frac{c t \bar n^{1+\eps}}{(k-1) \bar n^{1+\eps} + c t}\right)^{1/(1 + \eps)} \\
\implies m(k) &\approx \frac{(c t)^{1/(1 + \eps)} \bar n}{(k-1)^{1/(1 + \eps)} \bar n + (c t)^{1/(1 + \eps)}}.
\]
Note also that $\bar k \approx ct\left(1 - \frac{1}{\bar n^{1+\eps}}\right) + 1$.

To first order, we have $c \approx \eps$. Making this substitution, we want to solve
\[
t &=  \sum_{k=1}^{\bar k} m(k) \\
&= \sum_{k=1}^{\bar k} \frac{(\eps t)^{1/(1 + \eps)} \bar n}{(k-1)^{1/(1 + \eps)} \bar n + (\eps t)^{1/(1 + \eps)}} \\
&\approx \int_{1}^{\bar k} \frac{(\eps t)^{1/(1 + \eps)} \bar n}{(x-1)^{1/(1 + \eps)} \bar n + (\eps t)^{1/(1 + \eps)}}\,dx
\]

\end{document}

### 2013-10-04 icml/herding-connection.tex
2/3 of the way done with herding section

@@
 
-\NA{JHH: Need to specify the spaces we are working with here.}
-Suppose we are given a family of features $\phi$ together 
-with a known value $\bar{\phi}$ for $\bE_{u}[\phi]$. We 
-would like to construct a distribution that approximately 
-matches this distribution, i.e.~one for which 
-$\bE_{\hat{u}}[\phi] \approx \bar{\phi}$. A natural way 
-to do this is by using boosted mirror descent to minimize the 
-maximum mean discrepancy relative to $\phi$; 
-in other words, to minimize
-\[ L(u) = \max_{i \in \{1,\dots,d\}} |\bE_{u}[\phi_i]-\bar{\phi}_i|. \]
-If we wish to write this more similarly to the original problem 
-formulation, we can write
-\[ L(u) = \sup_{\|\theta\|_1 \leq 1} \theta^T(\bE_{u}[\phi]-\bar{\phi}). \]
-One issue is that $\|\theta\|_1$ is not strongly convex, so we relax $\|\theta\|_1$ 
-to $\|\theta\|_2$ and write our loss function in the Lagrangian form:
-\[ L(u, \theta) = \theta^T\bE_{u}[\phi] - \theta^T\bar{\phi} - \frac{1}{2}\|\theta\|_2^2. \]
-Intriguingly, Algorithm 1 and Algorithm 2 are identical to each other for this choice of $L$ 
-(up to a change of index). In this case, $u_t = \delta_{x_t}$ for some $x_t$, and 
-the updates are (for Algorithm 2):
+The standard setup for herding is as follows: we are given a 
+space $\sX$ and a feature mapping $\phi : \sX \to \sH$ for some 
+reproducing kernel Hilbert space $\sH$ \cite{RKHS}. 
+We define $\Delta_{\sX}$ to be the simplex 
+of probability distributions over $\sX$ and the marginal polytope 
+$\sM \eqdef \{\bE_{\mu}[\phi(x)] \mid \mu \in \Delta_{\sX}\}$.
+
+The goal of herding is to construct samples $x_1,\ldots,x_T \in \sX$ 
+such that $\|\frac{1}{T}\sum_{t=1}^T \phi(x_t)-\bar{\phi}\|_{\sH}^2$ 
+is minimized for some given $\bar{\phi} \in \sM$.
+
+The herding algorithm uses the following updates\footnote{The original 
+algorithm includes an additional $\bar{\phi}$ term in the third bullet 
+point and indexes $x_t$ as $x_{t+1}$.} [\cite{chen}, equations (1) and (2)]:
 \begin{itemize}
-\item $x_t \in \arg\min_{x} \theta_t^T\phi(x)$
-\item $\theta_{t+1} = \frac{1}{A_t} \sum_{s=1}^t \alpha_s \phi(x_{s}) - \bar \phi$.
-%\item $\theta_{t+1} = \frac{1}{A_t} \sum_{s=1}^t \alpha_s \phi(x)$.
+\item $w_1 = \bar{\phi}$
+\item $x_t \in \arg\max_{x \in \sX} \ip{w_t}{\phi(x)}$
+\item $w_{t+1} = \sum_{s=1}^t [\bar{\phi}-\phi(x_s)]$
 \end{itemize}
-These are the same as the standard herding updates when 
-$\alpha_t = 1$. Note that $R$ is strongly convex with respect to 
-$\|\cdot\|_2$, which is its own dual. Therefore, if we let
-$r = \sup_{x} \|\phi(x)\|_2$, then Corollary~\ref{cor:method-2} gives us
+If we define $\theta_t = -w_t/t$, then these updates are equivalent to
+\begin{itemize}
+\item $\theta_1 = -\bar{\phi}$
+\item $\mu_t \in \arg\min_{\mu \in \Delta_{\sX}} \ip{\theta_t}{\bE_{\mu}[\phi(x)]}$
+\item $\theta_{t+1} = \frac{1}{t} \sum_{s=1}^t [\bE_{\mu_s}[\phi(x)] - \bar{\phi}]$
+\end{itemize}
+Note that we have replaced the point $x_t$ with a distribution $\mu_t$. 
+This does not change the algorithm as $\bE_{\mu}[\theta_t^T\phi(x)]$ will 
+always be minimized at an extreme point of $\Delta_{\sX}$, which corresponds 
+to a point mass.
+
+The above updates correspond to Algorithm 2 for $U = \sX$, $\Theta = \sH$, 
+$\alpha_t = 1$, $h \equiv 0$, 
+$R(\theta) = \ip{\bar{\phi}}{\theta} + \frac{1}{2}\|\theta\|_{\sH}^2$, 
+and $\imath(u)$ the canonical extension of $\phi$ from $\sX$ to $\Delta_{\sX}$.
+To see this, note that 
+\begin{align}
+\lefteqn{\arg\min_{\theta \in \sH} \ip{\imath(u)}{\theta} - R(\theta)} \\
+ &= \arg\min_{\theta \in \sH} \ip{\bE_{u}[\phi(x)]-\bar{\phi}}{\theta} - \frac{1}{2}\|\theta\|_{\sH}^2 \\
+ &= \bE_u[\phi(x)] - \bar{\phi}.
+\end{align}
+For $u = 0$, this gives us the first bullet point. For $u = \hat{u}_t$, 
+this gives us the third bullet point. Finally, the second bullet point 
+is already in the form given in Algorithm 2.
+
+Note that $R$ is convex with respect to $\|\cdot\|_2$ and that 
+$R^*(u) = \frac{1}{2}\|\bar{\phi}-\bE_{u}[\phi(x)]\|_2^2$. We can 
+therefore apply Corollary~\ref{thm:method-2} to obtain the bound
 \begin{align*}
-\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
- &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
- &= \frac{2r^2\log(T+1)}{T}.
+\lefteqn{\frac{1}{2}\|\bar{\phi}-\frac{1}{T}\sum_{t=1}^T \phi(x_t)\|_2^2} \\
+ &\leq \inf_{u^*} \frac{1}{2}\|\bar{\phi}-\bE_{u^*}[\phi(x)]\|_2^2 + \frac{2r^2\log(T+1)}{T} \\
+ &= \frac{2r^2\log(T+1)}{T},
 \end{align*}
+where $r \eqdef \sup_{m \in \sM} \|m\|_2$.
+
 This is a slightly weaker version of the typical herding bound on MMD.
@@
 
+The convergence bound for herding implicitly assumes that 
+$\sM$ is bounded in the $l^2$ norm. Assuming that $\sH$ is 
+finite-dimensional, this will always be true. However, in 
+the infinite-dimensional case this may no longer be true. 
+For instance, consider $\sX = \{1,\ldots,N\}$ and let 
+$\phi : \sX \to l^2(\bZ_{\geq 0})$ be
+defined by $\phi(n)_i = \frac{1}{\sqrt{n+i}}$. Then 
+$\|\phi(n)\|_2 = \infty$ for each $n \in \sX$, and so herding 
+yields no convergence guarantees. Indeed, the algorithm 
+does not even necessarily make sense: since $\theta$ is 
+no longer guaranteed to have finite $l^2$ norm, the 
+inner product $\ip{\imath(u)}{\theta}$ may not be defined.
+
+Note, however, that in this case
+$\|\phi(n)\|_3 = \sqrt[3]{\sum_{i=0}^{\infty} (n+i)^{-3/2}} < \infty$.
+(In fact, the maximum is equal to
+$\|\phi(1)\|_3 = \sqrt[3]{\zeta(3/2)} < 1.378$.) This motivates 
+a generalization of herding in which we minimize 
+$\|\bE_{u}[\phi(x)]-\bar{\phi}\|_3^3$ instead of 
+$\|\bE_{u}[\phi(x)]-\bar{\phi}\|_2^2$.
+
+More generally, for any 
+$p \geq 2$ and $q$ such that $\frac{1}{p} + \frac{1}{q} = 1$, we can take 
+$h \equiv 0$, $R(\theta) = \ip{\bar{\phi}}{\theta} + \frac{1}{q}\|\theta\|_q^q$. 
+Then $R^*(u) = \frac{1}{p}\|\bE_{u}[\phi(x)]-\bar{\phi}\|_p^p$. Define the 
+``correction function'' $C_q$ by
+\begin{equation}
+C_q(x)_i \eqdef \sign(x_i) \cdot |x_i|^{\frac{1}{q-1}}.
+\end{equation}
+We obtain the following \emph{generalized herding} updates:
+\begin{itemize}
+\item $\theta_{1} = C_q(-\bar{\phi})$
+\item $x_{t} \in \arg\min_{x \in \sX} \ip{\phi(x)}{\theta_t}$
+\item $\theta_{t+1} = C_q\left(\frac{1}{t} \sum_{s=1}^t [\phi(x_s)-\bar{\phi}]\right)$
+\end{itemize}
+
 \subsection{Herding and Maximum Entropy}
@@
 samples from the maximum entropy distribution for a given collection of moment 
-constraints. In this section we will make this connection more explicit 
-as well as draw other relationships between herding and maximum entropy.
+constraints. In this section we present a herding-style algorithm that more 
+explicitly yields a maximum-entropy distribution. Our hope is that comparing 
+the two algorithms may help to make the connection between herding and maximum 
+entropy more explicit.
 
-Suppose that instead of setting $h(u) = 0$ we set $h(u)$ to a small multiple of 
-the negative entropy: $h(u) = \eta \bE_{u}[\log u(x)]$. We will still use 
-$R(\theta) = \frac{1}{2}\|\theta\|_2^2$. In this case, Algorithm 2 gives the updates
+The key idea is to use the primal regularizer $h(u) = -\eta H(u)$, where 
+$H(u)$ is the entropy of $u$ and $\eta \in \bR_{>0}$. We will still use 
+$R(\theta) = \ip{\bar{\phi}}{\theta} + \frac{1}{2} \|\theta\|_{\sH}^2$.
+In this case, Algorithm 2 gives the updates
 \begin{itemize}
-\item $u_t(x) \propto \exp\left(-\frac{1}{\eta}\theta_t^T\phi(x)\right)$
+\item $\theta_1 = -\bar{\phi}$
+\item $u_t(x) \propto \exp\left(-\frac{1}{\eta}\ip{\theta_t}{\phi(x)}\right)$
 \item $\theta_{t+1} = \bE_{\hat{u}_t}[\phi(x)] - \bar\phi$.
 \end{itemize}
-Corollary~\ref{cor:method-2} implies that for all $u^{*}$,
+Corollary~\ref{cor:method-2} implies that, for 
+$\hat{u}_T = \frac{1}{T} \sum_{t=1}^T u_t$, we have
 \begin{align*}
-\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) &\leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
- &= -\eta H(u^*) + \frac{1}{2}\|\bE_{u^*}[\phi(x)]-\bar{\phi}\|_2^2 \\ &\phantom{=} + \frac{8r^2}{T}.
+\lefteqn{-\eta H(\hat{u}_T) + \|\bar{\phi}-\bE_{\hat{u}_T}[\phi(x)]\|_{\sH}^2} \\
+ &\leq \inf_{u} -\eta H(u) + \|\bar{\phi}-\bE_{u}[\phi(x)]\|_{\sH}^2 + \frac{2r^2\log(T+1)}{T} \\
+ &\leq -\eta H_{\max}(\bar{\phi}) + \frac{2r^2\log(T+1)}{T},
 \end{align*}
-In particular, let $u^*$ be the maximum entropy distribution satisfying $\bE_{u^*}[\phi(x)] = \bar{\phi}$, 
-and let $H^*$ be the corresponding entropy. Then we have 
-\[ -\eta H(\hat{u}) + \|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq -\eta H^* + \frac{8r^2}{T}. \]
-Since $\|\cdot\|_2^2 \geq 0$, this implies that 
-\begin{equation}
-H(\hat{u}) \geq H^* - \frac{8r^2}{\eta T},
-\end{equation}
-and so asymptotically $\hat{u}$ has at least 
-as much entropy as $u^*$.
-
-Now, suppose that our state space has some finite size $S$, so 
-that no distribution has entropy greater than $\log(S)$. Then 
-we conclude that
-\begin{equation}
-\|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq \eta[\log(S)-H^*] + \frac{8r^2}{T}.
-\end{equation}
-
-If we let $\eta \to 0$ and $\eta T \to \infty$, then we obtain 
-a distribution $\hat{u}$ whose moments are arbitrarily close to 
-the moments of $u^*$, and whose entropy is almost as large as 
-(or larger than) the entropy of $u^*$.
+where $H_{\max}(\bar{\phi})$ is the maximum entropy of any distribution whose 
+moments are $\bar{\phi}$.
 

### 2013-10-04 icml/lower-bounds.tex

\documentclass[paper.tex]{subfiles}

\begin{document}
\section{Lower Bounds on Convergence Rates}
\label{sec:lower-bounds}

It is an open question whether it is possible to achieve the fast $O(1/T^{2})$ convergence of finite-dimensional herding in the infinite-dimensional case. We now show that this is not in general possible. In particular, the upper bounds in the previous section for herding and herding-like algorithms have matching lower bounds for at least some combinations of distributions and feature functions.

Let $\sX$ be the set of vectors $(x_{1},x_{2},\dots) \in [0,1]^{\infty}$ such that $\sum_{n=1}^{\infty} x_{n} = 1$ and a finite number of the $x_{n}$ are non-zero. Define the infinite feature vector $\phi$ to be the imbedding of $\sX$ into $\ell^{1}$. Let the target distribution $p(x)$ be chosen such that $\bar\phi_{n} = \alpha/n^{1+\eps}$, where the normalization constant $\alpha = 1/\zeta(1 + \eps)$ is the reciprocal of the Riemann zeta function. Note that as $\eps \to 0$, $\alpha \approx \eps$. We will consider the standard $\ell^{2}$ herding loss function $L(\theta) = \|\theta - \bar \phi\|_2^2$.

Define a {\em sparse update method} to be any method that produces an estimator of $\bar \phi$ at iteration $t$ of the form 
\[
\theta_{t} = \sum_{s =1}^{t}w_{t} \phi(x_{t}).
\]
In other words, the $t$-th estimator will have at most $t$ non-zero entries, and so $\|x-x^{*}\|_2^2 \ge \alpha^{2}\sum_{n > t} n^{-2-2\eps} = O(\alpha^{2}n^{-1-2\eps})$. Therefore, any sparse update method must have convergence rate no faster than $n^{-1-2\eps}$, for every $\eps > 0$.

Sparse update methods can be viewed as charactering ``herding-like algorithms,'' including the set of generalized herding algorithms based on \algname{DBMD} and \algname{CGD} with line search. Those sparse methods are of interest because they provide (weighted) ``samples'' matching the moments of the distribution of interest. Since they only operate on the extreme points of the simplex, the optimization problems required to use them are more tractable.

\end{document}

### 2013-10-04 icml/introduction.tex
split of bmd section. added shortcuts for algorithm names

@@
 \(
-\sE_{T}^{2} &= \text{MMD}^{2}(p, \hat p_{T})  \\
+\sE_{T}^{2} &= \MMD^{2}(p, \hat p_{T})  \\
 &\eqdef \sup_{\theta \in \Theta} \left|\bE_{p}\ip{\theta}{\phi(x)} - \bE_{\hat p_{T}}\ip{\theta}{\phi(x)} \right|,
@@
 \citet{Bach:2012a} recently showed that the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} does not apply in the infinite dimensional case.
-In addition, \citet{Bach:2012a} introduced an interpretation of herding as a special case of conditional gradient descent (\algname{CGD}) minimizing the square norm of an RKHS.
-They showed that \algname{CGD} with line search leads to faster convergence rates: $O(1/T)$ in the infinite dimensional case and exponential in the finite dimensional case.
+In addition, \citet{Bach:2012a} introduced an interpretation of herding as a special case of conditional gradient descent (\cgd) minimizing the square norm of an RKHS.
+They showed that \cgd with line search leads to faster convergence rates: $O(1/T)$ in the infinite dimensional case and exponential in the finite dimensional case.
 However, they found  empirically that line search had poor approximation properties since led to distributions far from the maximum entropy distribution.
@@
 \begin{itemize}
-\item In Section \ref{sec:algorithm} we introduce a novel pair of convex optimization algorithms, which we call primal and dual boosted mirror descent (\algname{PBMD} and \algname{DBMD}).
-We show how \algname{PBMD} and \algname{DBMD} relate to \algname{CGD} and mirror descent (\algname{MD}).
-\item We then use \algname{DBMD} in Section \ref{sec:herding} to define a class of generalized herding algorithms.
-Connections with the ``learning'' interpretation are made by showing how to obtain herding as the limit of an entropically normalized case of \algname{BMD}.
-\item Section \ref{sec:proofs} presents convergence rates for $\algname{PBMD}$ and $\algname{DBMD}$.
+\item In Section \ref{sec:algorithm} we introduce a novel pair of convex optimization algorithms, which we call primal and dual boosted mirror descent (\primal and \dual).
+We show how \primal and \dual relate to \cgd and mirror descent (\md).
+\item We then use \dual in Section \ref{sec:herding} to define a class of generalized herding algorithms.
+Connections with the ``learning'' interpretation are made by showing how to obtain herding as the limit of an entropically normalized case of \dual.
+\item Section \ref{sec:proofs} presents convergence rates for \primal and \dual.
 As special cases, we recover a $O(\log T/ T)$ convergence rate for herding and a $O(1/T)$ rate for a herding-like algorithm that uses unequally weighted samples.
@@
 \item In Section \ref{sec:lower-bounds} we prove matching lower bounds to the upper bounds given in Section \ref{sec:proofs}, showing that any ``sparse'' herding-like method cannot converge faster than $O(1/T)$ in the the infinite dimensional case and that herding converges at the slower $O(\log T / T)$ rate.
-\item Finally, in section \ref{sec:chen} we generalize the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \algname{BMD}. 
+\item Finally, in section \ref{sec:chen} we generalize the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \bmd. 
 \end{itemize}
 
-\section{Boosted Mirror Descent}
-\label{sec:algorithm}
-
-Let $U$ and $\Theta$ be convex subsets of two Banach spaces $\sB_{U}$ and $\sB_{\Theta}$ and 
-let $\imath : U \to \Theta$ be a linear map from 
-$U$ to $\Theta$ that preserves the inner product. Consider the following loss function 
-$L : U \to \bR$:
-\begin{equation}
-L(u) = h(u) + \max_{\theta \in \Theta} \left\{\langle \theta, \imath(u) \rangle - R(\theta) \right\}.
-\end{equation}
-We can think of $h$ as a \emph{primal regularizer} and $R$ as a \emph{dual regularizer}.
-
-Examples:
-\begin{itemize}
-\item Let $U$ be the probability simplex over a space $\sX$, let $\phi : \sX \to \bR^d$ be a collection 
-      of statistics, and let $\Theta$ be the unit $L^1$ ball in $\bR^d$. Let $h(u) = 0$, 
-      $\langle \theta, \imath(u) \rangle = \bE_u[\theta^T\phi(x)]$, and 
-      $R(\theta) = \langle \theta, \imath(u_0) \rangle$. Then $L(u)$ is the 
-      \emph{maximum mean discrepancy} between $u$ and $u_0$ relative to $\phi$.
-\item Let $h(u) = 0$ and $R(\theta) = S^*(i^{-1}(\theta))$ where $S^*$ is the 
-      Fenchel conjugate of any strongly convex function $S : U \to \bR$. Then $L(u) = S(u)$ 
-      as long as $i^{-1}(\Theta) = \sB_{U}$.
-\item Let $U = \Theta = \bR^d$. Let $h(u) = \|u\|_1$, $R(u) = \langle \theta, u_0 \rangle + \frac{1}{2} \|\theta\|_2^2$. 
-      Then $L(u) = \|u\|_1 + \frac{1}{2} \|u-u_0\|_2^2$. 
-\end{itemize}
-
-In this paper we will consider a family of methods for minimizing $L(u)$, 
-which are based on Bach's generalization of the Frank-Wolfe algorithm and can 
-be interpreted as boosted mirror descent.
-
-First, let us generalize the setting 
-to consider a \emph{two-argument} loss function
-\begin{equation}
-L(u,\theta) = h(u) + \langle \theta, \imath(u) \rangle - R(\theta).
-\end{equation}
-We will assume throughout that $h$ and $R$ are both convex functions, 
-and furthermore that $\arg\max_{u} h(u) + \langle \theta, \imath(u) \rangle$ can be efficiently 
-computed for all values of $\theta$.
-
-Mirror descent, together with boosting, yields an algorithm for finding 
-``saddle points'' of $L$. It is the following algorithm (called Algorithm 1).
-For notational convenience, for a sequence of weights $\alpha_1,\alpha_2,\ldots$ 
-let $\hat{u}_t = \frac{\sum_{s=1}^t \alpha_su_s}{\sum_{s=1}^t \alpha_s}$ and let 
-$\hat{\theta}_t = \frac{\sum_{s=1}^t \alpha_s\theta_s}{\sum_{s=1}^t \alpha_s}$.
-\begin{enumerate}
-\item $u_1 \in \arg\min_u h(u)$
-\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, \imath(u_t) \rangle - R(\theta)$
-\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, \imath(u) \rangle$
-%\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
-\end{enumerate}
-As long as $h$ is strongly convex, we obtain the 
-bound (see Corollary~\ref{cor:method-1}):
-\begin{equation}
-\sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/T).
-\end{equation}
-In other words, $\hat{u}_T$ is close to being a global minimum of $L$.
-However, it is often the case that $h$ is not strongly convex, whereas $R$ is strongly convex. In this case 
-we may wish to use the following slightly different algorithm (called Algorithm 2):
-\begin{enumerate}
-\item $\theta_1 \in \arg\min_{\theta} R(\theta)$
-\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, \imath(u) \rangle$
-%\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
-\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \imath(\hat{u}_t) \rangle - R(\theta)$
-\end{enumerate}
-We then obtain the following bound 
-when $R$ is strongly convex (see Corollary~\ref{cor:method-2}):
-\[ \sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/T). \]
-Algorithm 1 and Algorithm 2 are closely related; indeed, they are dual to each other 
-(performing Algorithm 1 on $L(u,\theta)$ is the same as performing Algorithm 2 on 
-$-L(\theta,u)$).
-
-For convenience, we will abuse notation and use $\langle \theta, u \rangle$ in place of 
-$\langle \theta, \imath(u) \rangle$ when the map $i$ is clear from context.
-
 \end{document}

### 2013-10-04 icml/lower-bounds.tex
split of bmd section. added shortcuts for algorithm names

@@
 
-Sparse update methods can be viewed as charactering ``herding-like algorithms,'' including the set of generalized herding algorithms based on \algname{DBMD} and \algname{CGD} with line search.
+Sparse update methods can be viewed as charactering ``herding-like algorithms,'' including the set of generalized herding algorithms based on \dual and \cgd with line search.
 Those sparse methods are of interest because they provide (weighted) ``samples'' matching the moments of the distribution of interest.

### 2013-10-04 icml/herding-connection.tex
finished proofs for herding connection, still need somewhat better exposition

@@
 \end{itemize}
+By Proposition~\ref{prop:method-2}, we know that
+\begin{align}
+\frac{1}{p}\|\bar{\phi}-\frac{1}{T}\sum_{t=1}^T \phi(x_t)\|_p^p \leq \frac{1}{A_T} \sum_{t=1}^T A_tD_{R^*}(\hat{u}_{t+1}\|\hat{u}_t).
+\end{align}
+We can no longer use Corollary~\ref{cor:method-2} since $\|\cdot\|_p^p$ is not strongly convex. 
+Instead, we use the following two inequalities (proved in Appendix~\ref{apx:inequalities}):
+\begin{lemma}
+Let $x, y \in \bR_{\geq 0}$. Then:
+\begin{equation}
+x^p-py^{p-1}x+(p-1)y^p \leq \binom{p}{2}(x-y)^2\max(x,y)^{p-2}
+\end{equation}
+and
+\begin{equation}
+(y-z)^2\max(x,y)^{p-2} \leq 4\max(x,y,z)^p
+\end{equation}
+\end{lemma}
+Using these, we obtain the bound
+\begin{align}
+D_{R*}(\hat{u}_{t+1} \| \hat{u}_t) &= \frac{1}{p} \sum_{i} \hat{u}_{t+1,i}^p - p\hat{u}_{t+1,i}\hat{u}_{t,i}^{p-1} + (p-1)\hat{u}_{t,i}^{p} \\
+ &\leq \frac{1}{p}\binom{p}{2} \sum_i (\hat{u}_{t+1,i}-\hat{u}_{t,i})^2\max(\hat{u}_{t+1,i},\hat{u}_{t,i})^{p-2} \\
+ &= \frac{1}{p}\binom{p}{2}\frac{\alpha_{t+1}^2}{A_{t+1}^2} \sum_i (u_{t+1,i}-\hat{u}_{t})^2\max(\hat{u}_{t+1,i},\hat{u}_{t,i})^{p-2} \\
+ &\leq 2(p-1)\frac{\alpha_{t+1}^2}{A_{t+1}^2} \sum_i \max(\hat{u}_{t,i},\hat{u}_{t+1,i},u_{t+1,i})^p \\
+ &\leq 2(p-1)\frac{\alpha_{t+1}^2}{A_{t+1}^2} \sum_i \hat{u}_{t,i}^p+\hat{u}_{t+1,i}^p+u_{t+1,i}^p \\
+ &\leq 6(p-1)\frac{\alpha_{t+1}^2}{A_{t+1}^2} r_p^p
+\end{align}
+Plugging back in yields a convergence rate of $\frac{6(p-1)r_p^p(\log(T)+1)}{T}$ for $\alpha_t = 1$ 
+and $\frac{24(p-1)r_p^p}{T}$ for $\alpha_t = t$. We therefore obtain a convergence rate of $O(T^{-1/p})$ 
+in the $l^p$ norm, even if $\sM$ is not bounded in the $l^2$ norm.
 

### 2013-10-04 icml/boosted-mirror-descent.tex
split of bmd section. added shortcuts for algorithm names
tightened lower bound argument; started supplementary material file

--- 2013-10-03 icml/introduction.tex
@@
 \begin{document}
+
 \section{Boosted Mirror Descent}
@@
 
-Let $U$ and $\Theta$ be convex subsets of two Banach spaces $\sB_{U}$ and $\sB_{\Theta}$ and 
-let $\imath : U \to \Theta$ be a linear map from 
-$U$ to $\Theta$ that preserves the inner product. Consider the following loss function 
-$L : U \to \bR$:
-\begin{equation}
-L(u) = h(u) + \max_{\theta \in \Theta} \left\{\langle \theta, \imath(u) \rangle - R(\theta) \right\}.
-\end{equation}
+Our results primarily rely on a pair of novel convex optimization algorithms we call primal and dual boosted mirror descent.
+
+\subsection{Convex Optimization}
+
+Before introducing the boosted mirror descent (\bmd) algorithms, we recall some basic definitions related convex optimization.
+Let $f$ be a real function whose domain is a Banach space.
+The {\em convex conjugate} $f^{*}$ of $f$ is defined as
+\[
+f^{*}(y) \eqdef \max_{x} \ip{x}{y} - f(x).
+\]
+If $f$ is lower semi-continuous, then by the Fenchel-Moreau theorem $f = f^{**}$, i.e.
+\[
+f(x) = \max_{y} \ip{x}{y} - f^{*}(y).
+\]
+Throughout the paper, we will assume without further notice that functions are lower semi-continuous when required.
+The gradients of $f^{*}$ and $f$ can be defined as any argument that maximizes the previous two equations:
+\[
+\partial f^{*}(y) &\in \argmax_{x} \ip{x}{y} - f(x) \\
+\partial f(x) &\in \argmax_{y} \ip{x}{y} - f^{*}(y).
+\]
+
+In order to prove fast convergence, we will require additional assumptions beyond convexity. We say a function $f$ is {\em $\alpha$-strongly convex} if 
+\[
+f(x)  - f(y) \le \ip{\partial f(x)}{x-y} - \frac{\alpha}{2}\|x - y\|^{2}.
+\]
+and that a (possibly non-convex) function $g$ is $\beta$-smooth if 
+\[
+\|\partial g(x) - \partial g(y) \| \le \beta \| x - y\|.
+\]
+Strong convexity and smoothness are dual to each in the sense that a convex function $f$ is $\alpha$-strongly convex if and only if $f^{*}$ is $(1/\alpha)$-smooth.
+
+\subsection{The Boosted Mirror Descent Algorithms}
+
+We can now define the \bmd  algorithms.
+Let $U$ and $\Theta$ be convex subsets of the Banach spaces $\sB_{U}$ and $\sB_{\Theta}$, respectively, and let $\imath : U \to \Theta$ be a linear imbedding from $U$ to $\Theta$ that preserves the inner product.
+Consider the two-argument loss function $L : U \times \Theta \to \bR$
+\[
+L(u,\theta) = h(u) + \ip{\theta}{\imath(u)} - R(\theta)
+\]
+for which we wish to find arguments that obtain $L_{*} = \min_{u}\max_{\theta} L(u, \theta)$. Taking only max, we have the one-argument loss function:
+\[
+L(u)
+&= h(u) + \max_{\theta \in \Theta} \left\{\ip{\theta}{\imath(u)} - R(\theta) \right\} \\
+&= h(u) + R^{*}(\imath(u)).
+\]
 We can think of $h$ as a \emph{primal regularizer} and $R$ as a \emph{dual regularizer}.
-
-Examples:
-\begin{itemize}
+The pair of regularization functions $h$ and $R$ provides flexible optimization framework, as seen from the following examples:
+\begin{enumerate}
 \item Let $U$ be the probability simplex over a space $\sX$, let $\phi : \sX \to \bR^d$ be a collection 
@@
 \item Let $h(u) = 0$ and $R(\theta) = S^*(i^{-1}(\theta))$ where $S^*$ is the 
-      Fenchel conjugate of any strongly convex function $S : U \to \bR$. Then $L(u) = S(u)$ 
+      convex conjugate of any strongly convex function $S : U \to \bR$. Then $L(u) = S(u)$ 
       as long as $i^{-1}(\Theta) = \sB_{U}$.
-\item Let $U = \Theta = \bR^d$. Let $h(u) = \|u\|_1$, $R(u) = \langle \theta, u_0 \rangle + \frac{1}{2} \|\theta\|_2^2$. 
+\item Let $U = \Theta = \bR^d$ and $\imath$ the identity map. Let $h(u) = \|u\|_1$, $R(u) = \langle \theta, u_0 \rangle + \frac{1}{2} \|\theta\|_2^2$. 
       Then $L(u) = \|u\|_1 + \frac{1}{2} \|u-u_0\|_2^2$. 
-\end{itemize}
+\end{enumerate}
 
-In this paper we will consider a family of methods for minimizing $L(u)$, 
-which are based on Bach's generalization of the Frank-Wolfe algorithm and can 
-be interpreted as boosted mirror descent.
-
-First, let us generalize the setting 
-to consider a \emph{two-argument} loss function
-\begin{equation}
-L(u,\theta) = h(u) + \langle \theta, \imath(u) \rangle - R(\theta).
-\end{equation}
-We will assume throughout that $h$ and $R$ are both convex functions, 
-and furthermore that $\arg\max_{u} h(u) + \langle \theta, \imath(u) \rangle$ can be efficiently 
+Going forward, for convenience we will abuse notation and use $\ip{\theta}{u}$ in place of 
+$\ip{\theta}{\imath(u)}$ when the map $\imath$ is clear from context. We will assume 
+throughout that $h$ and $R$ are both convex functions, and furthermore that 
+$\arg\max_{u} h(u) + \ip{\theta}{u}$ can be efficiently 
 computed for all values of $\theta$.
 
+%In this paper we will consider a family of methods for minimizing $L(u)$, 
+%which are based on Bach's generalization of the Frank-Wolfe algorithm and can 
+%be interpreted as boosted mirror descent.
+
 Mirror descent, together with boosting, yields an algorithm for finding 
-``saddle points'' of $L$. It is the following algorithm (called Algorithm 1).
+``saddle points'' of $L$. 
 For notational convenience, for a sequence of weights $\alpha_1,\alpha_2,\ldots$ 
 let $\hat{u}_t = \frac{\sum_{s=1}^t \alpha_su_s}{\sum_{s=1}^t \alpha_s}$ and let 
-$\hat{\theta}_t = \frac{\sum_{s=1}^t \alpha_s\theta_s}{\sum_{s=1}^t \alpha_s}$.
+$\hat{\theta}_t = \frac{\sum_{s=1}^t \alpha_s\theta_s}{\sum_{s=1}^t \alpha_s}$. Then 
+the {\em primal boosted mirror descent} (\primal) algorithm is:
 \begin{enumerate}
 \item $u_1 \in \arg\min_u h(u)$
-\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, \imath(u_t) \rangle - R(\theta)$
-\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, \imath(u) \rangle$
-%\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
+\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, u_t \rangle - R(\theta) = \partial R^{*}(u_{t})$
+\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, u \rangle = \partial h^{*}(-\hat\theta_{t})$
 \end{enumerate}
@@
 \begin{equation}
-\sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/T).
+\sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq L_{*} + O(1/T).
 \end{equation}
-In other words, $\hat{u}_T$ is close to being a global minimum of $L$.
-However, it is often the case that $h$ is not strongly convex, whereas $R$ is strongly convex. In this case 
-we may wish to use the following slightly different algorithm (called Algorithm 2):
+In other words, $\hat{u}_T$ is close to being a global minimum of $L(u)$.
+However, it is often the case that $h$ is not strongly convex, whereas $R$ is strongly convex. As we shall see, this will be the case for our application to herding.
+In this case we may wish to use the following slightly different algorithm, which we call {\em dual boosted mirror descent} (\dual):
 \begin{enumerate}
 \item $\theta_1 \in \arg\min_{\theta} R(\theta)$
-\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, \imath(u) \rangle$
-%\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
-\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \imath(\hat{u}_t) \rangle - R(\theta)$
+\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle = \partial h^{*}(-\theta_{t})$
+\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \hat{u}_t \rangle - R(\theta) = \partial R^{*}(\hat u_{t})$
 \end{enumerate}
@@
 when $R$ is strongly convex (see Corollary~\ref{cor:method-2}):
-\[ \sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/T). \]
-Algorithm 1 and Algorithm 2 are closely related; indeed, they are dual to each other 
-(performing Algorithm 1 on $L(u,\theta)$ is the same as performing Algorithm 2 on 
-$-L(\theta,u)$).
+\[ \sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq L_{*} + O(1/T). \]
+As suggested by the name \primal and \dual are closely related; indeed, they are dual to each other in the sense that 
+performing \primal on $L(u,\theta)$ is the same as performing \dual on 
+$-L(\theta,u)$.
 
-For convenience, we will abuse notation and use $\langle \theta, u \rangle$ in place of 
-$\langle \theta, \imath(u) \rangle$ when the map $i$ is clear from context.
+\subsection{Connections to Other Gradient Descent Algorithms}
+
+The \bmd algorithms are very closely related to \cgd and \md.
 

### 2013-10-04 icml/lower-bounds.tex
tightened lower bound argument; started supplementary material file

@@
 
-It is an open question whether it is possible to achieve the fast $O(1/T^{2})$ convergence of finite-dimensional herding in the infinite-dimensional case.
+It has been an open question whether it is possible to achieve the fast $O(1/T^{2})$ convergence of finite-dimensional herding in the infinite-dimensional case.
 We now show that this is not in general possible.
@@
 Define the infinite feature vector $\phi$ to be the imbedding of $\sX$ into $\ell^{1}$.
-Let the target distribution $p(x)$ be chosen such that $\bar\phi_{n} = \alpha/n^{1+\eps}$, where the normalization constant $\alpha = 1/\zeta(1 + \eps)$ is the reciprocal of the Riemann zeta function.
-Note that as $\eps \to 0$, $\alpha \approx \eps$.
+Let the target distribution $p(x)$ be chosen such that $\bar\phi_{n} = c_{\alpha}/\alpha(n)$, where $\alpha(n)$ is any superlinear function (that is, $\alpha \in \omega(n)$) and $c_{\alpha}^{-1} = \sum_{n=1}^{\infty} \alpha(n) < \infty$.
 We will consider the standard $\ell^{2}$ herding loss function $L(\theta) = \|\theta - \bar \phi\|_2^2$.
@@
 \]
-In other words, the $t$-th estimator will have at most $t$ non-zero entries, and so $\|x-x^{*}\|_2^2 \ge \alpha^{2}\sum_{n > t} n^{-2-2\eps} = O(\alpha^{2}n^{-1-2\eps})$.
-Therefore, any sparse update method must have convergence rate no faster than $n^{-1-2\eps}$, for every $\eps > 0$.
+In other words, the $t$-th estimator will have at most $t$ non-zero entries, and so 
+\[
+\|x-x^{*}\|_2^2 \ge c_{\alpha}^{2}\sum_{n > t} 1/\alpha(n)^{2} = \Theta(1/\beta(t)) \label{eq:convergence-lb}
+\]
+for some $\beta \in \omega(t)$.
+But for any $\beta \in \omega(t)$ we can choose some $\alpha \in \omega(n)$ to obtain \eqref{eq:convergence-lb}.
+Therefore, any sparse update method must have convergence rate no faster than $\Theta(1/\beta(t))$ for all superlinear $\beta$, and hence no faster than $\Theta(1/t)$.
 
@@
 
+\NA{JH: move definition and discussion of sparse update method to beginning}
+
 \end{document}

### 2013-10-04 icml/supplementary-material.tex
tightened lower bound argument; started supplementary material file

--- 2013-10-02 icml/convex-opt-summary.tex
@@
 \usepackage{hyperref}
-\usepackage{pdflscape}
 \usepackage{epsfig}
@@
 \input{latex-defs.tex}
-% inner product command
-\newcommand{\ip}[2]{\left\langle #1, #2 \right\rangle}
-\def\[#1\]{\begin{align}#1\end{align}}
 
@@
 
-\title{Convex Optimization: A Very Short and Incomplete Summary} 
-\author{Jonathan Huggins}
-\date{\today}
+\title{Supplementary Material} 
+\author{}
+\date{}
 
@@
 
-\section{Introduction}
-
-Roughly speaking, we are interested in the following optimization problem:
-\[ 
-\max_{x} h(x) + f(x),
-\]
-where $h$ and $f$ are convex, and generally have other properties that aid in fast convergence of optimization algorithms.
-We assume throughout that all functions are convex.
-In this document we first outline key concepts.
-We then state some important lemmas.
-Finally we give a summary of some important convex optimization algorithms.
-These notes are based on S\'ebastian Bubeck's excellent notes for his Complexities of Optimization class\footnote{\url{https://blogs.princeton.edu/imabandit/orf523-the-complexities-of-optimization/}} and Francis Bach's ``Duality between subgradient and conditional gradient methods.''\footnote{\url{http://arxiv.org/abs/1211.6302}}
-
-\section{Key Concepts}
-
-\subsection{Conjugates and Gradients}
-
-When the domain of $f$ is Banach space, we can define the convex conjugate of $f$
-\[
-f^{*}(y) \defined \max_{x} \ip{x}{y} - f(x).
-\]
-If $f$ is lower semi-continuous, then by the Fenchel-Moreau theorem $f = f^{**}$, i.e.
-\[
-f(x) = \max_{y} \ip{x}{y} - f^{*}(y).
-\]
-Furthermore, the gradients of $f^{*}$ and $f$ can be defined as any argument that maximize the previous two equations:
-\[
-\partial f^{*}(y) &\in \argmax_{x} \ip{x}{y} - f(x) \\
-\partial f(x) &\in \argmax_{y} \ip{x}{y} - f^{*}(y).
-\]
-
-\subsection{Smoothness, Convexity, Distance} 
-We say a function $f$ is $\alpha$-strongly convex if 
-\[
-f(x)  - f(y) \le \ip{\partial f(x)}{x-y} - \frac{\alpha}{2}\|x - y\|^{2}.
-\]
-We say a (possibly non-convex) function $g$ is $\beta$-smooth if 
-\[
-\|\partial g(x) - \partial g(y) \| \le \beta \| x - y\|.
-\]
-
-Another useful notion is Bregman divergence: 
-\[
-D(x,y) \defined f(x) - f(y) - \ip{\partial f(y)}{x - y}.
-\]
-
-\NA{JHH: more on Bregman divergence}
-
-\section{Useful Facts}
-
-Here we gather a few useful lemmas.
-
-\blem
-If (possibly non-convex) $g$ is $\beta$-smooth, then 
-\[
-|g(x)  - g(y) - \ip{\partial g(y)}{x-y}| \le \frac{\beta}{2}\|x - y\|^{2}
-\]
-\elem
-\bprf \NA{JHH: prove this} \eprf
-\blem
-If $f$ is convex and $\beta$-smooth, then 
-\[
-f(x)  - f(y) \le \ip{\partial f(x)}{x-y} - \frac{1}{2\beta}\|\partial f(x) - \partial f(y)\|^{2}
-\]
-\elem
-\bprf \NA{JHH: prove this} \eprf
-\blem A convex function $f$ is $\alpha$-strongly convex $\iff$ $f^{*}$ is $(1/\alpha)$-smooth. \elem
-\bprf \NA{JHH: prove this} \eprf
-
-\section{Convex Optimization Algorithms}
-
-The algorithms, along with conditions on the functions and convergence rates, are summarized in the table on the next page.
-Here we define some of the notation used in specific algorithms.
-In projected gradient descent $\eta_{t} = 2/\alpha(t+1)$.
-In Nesterov's accelerated gradient descent,
-\[
-\lambda_{0} &= 0 \\
-\lambda_{t} &= \frac{1 + \sqrt{1 + 4\lambda_{t-1}^{2}}}{2} \\
-\gamma_{t}  &= \frac{1 - \lambda_{t}}{\lambda_{t+1}}.
-\]
-In mirror descent and conditional gradient descent $\rho_{t} = 2/(t + 1)$. Finally, in boosted mirror descent
-\[
-\alpha_{s} &= s \\
-A_{t} &= \sum_{s=1}^{t} \alpha_{s} = t(t+1)/2 \\
-\bar u_{t} &= A_{t}^{-1} \sum_{s=1}^{t} \alpha_{s}u_{s} \\
-\bar \theta_{t} &= A_{t}^{-1} \sum_{s=1}^{t} \alpha_{s}\theta_{s}
-\]
-
-\begin{landscape}
-
-\begin{table}[htdp]
-\begin{center}
-\begin{tabular}{|l|p{4.05cm}|p{4.05cm}|p{4.05cm}|p{4.05cm}|}
-\hline & Gradient Descent & Projected Subgrad. Desc. & Conditional Grad. Desc. & Nesterov's Accelerated Grad. Desc.   \\ 
-\hline
-Objective function  &
-$f(x)$ &
-$f(x)$ &
-$h^{*}(y) + f^{*}(y)$ &
-$f(x)$ \\
-\hline
-Conditions & 
-$f$ is $\beta$-smooth &
-$f$ is $\alpha$-s.c., $\|\partial f\| \le L$, and space is compact. & 
-$f$ is L.c. and $h$ is $\alpha$-s.c. &
-$f$ is $\beta$-smooth \\
-\hline
-Update equations &
-$x_{t+1} = x_{t} - \frac{1}{\beta}\partial f(x_{t})$ &
-$y_{t+1} = x_{t} - \eta_{t}\partial f(x_{t})$ and $x_{t+1} = \argmax_{x} \| x - y_{t+1} \|$ &
-$x_{t} = \partial h^{*}(-\bar y_{t})$ and $y_{t+1} = \partial f(x_{t})$ and $\bar y_{t+1} = (1 - \rho_{t})\bar y_{t} + \rho_{t} y_{t+1}$&
-$y_{t+1} = x_{t} - \frac{1}{\beta}\partial f(x_{t})$ and $x_{t+1} = (1- \gamma_{t})y_{t+1} + \gamma_{t}y_{t}$ \\
-\hline
-Convergence rate &
-$\beta R^{2}/t$ &
-$L^{2} / \alpha t$ &
-$R^{2} / \alpha t$ &
-$\beta R^{2}/t^{2}$ \\
-\hline
-\end{tabular}
-\caption{Convex optimization algorithms (gradient descent type).
-The function $f$ and $g$ are assumed to be convex.
-Constants are not given in convergence rates.
-Here $R$ is the radius of the metric space.
-Shorthands: L.c.
-= Lipschitz continuous, s.c.
-= strongly convex.
-}
-\label{table:gd-convex-opt-algs}
-\end{center}
-\end{table}
-
-\begin{table}[htdp]
-\begin{center}
-\begin{tabular}{|l|p{5.5cm}|p{5.5cm}|p{5.5cm}|}
-\hline & Mirror Descent & ``Primal'' Boosted Mirror Descent & ``Dual'' Boosted Mirror Descent  \\ 
-\hline
-Objective function & 
-$h(x) + f(x)$ & 
-$g(x) + R^{*}(x)$ & 
-$g(x) + R^{*}(x)$ \\
-\hline
-Conditions & 
-$h$ is smooth \& $\alpha$-s.c. and $f$ is L.c. & 
-$g$ is $\alpha$-s.c.& 
-$R^{*}$ is $\beta$-smooth \\
-\hline
-Update equations & 
-$y_{t} = \partial f(x_{t})$ and 
-$x_{t+1} = \argmin_{x} h(x) - (1 - \rho_{t+1})\ip{\partial h(x_{t})}{x} + \rho_{t+1}\ip{y_{t}}{x}$ &
-$u_{t} = \partial g^{*}(-\bar \theta_{t})$ and $\theta_{t+1} = \partial R^{*}(u_{t})$ & 
-$u_{t} = \partial g^{*}(-\theta_{t})$ and $\theta_{t+1} = \partial R^{*}(\bar u_{t})$ \\
-\hline
-Convergence rate & 
-$R^{2}/\alpha t$ & 
-$R^{2}/\alpha t$ & 
-$\beta R^{2} / t$ \\
-\hline
-
-\end{tabular}
-\caption{Convex optimization algorithms (mirror descent type). See Table \ref{table:gd-convex-opt-algs} for conventions. \NA{JHH: should add ISTA and FISTA} }
-\end{center}
-\label{table:md-convex-opt-algs}
-\end{table}
-
-\end{landscape}
+\appendix
 
@@
 
-The Boosted Mirror Descent algorithms are very closely related to Conditional Gradient Descent and Mirror Descent.
-Dual BMD and CGD are related as follows.
-BMD minimizes $g + R^{*}$ while CGD minimizes $h^{*} + f^{*}$.
+The \bmd algorithms are very closely related to \cgd and \md.
+\dual and \cgd are related as follows.
+\dual minimizes $g + R^{*}$ while \cgd minimizes $h^{*} + f^{*}$.
 Make the identifications $h^{*} = R^{*}$ and $f^{*} = g$.
 Note that $R^{*}$ is $\beta$-smooth $\iff R$ is $(1/\beta)$-strongly convex $\iff h$ is $(1/\beta)$-strongly convex, so the conditions on the functions being optimized are equivalent.
-Then with $\rho_{t} = \alpha_{t+1}/A_{t+1}$, we can rewrite the CGD updates  as
+Then with $\rho_{t} = \alpha_{t+1}/A_{t+1}$, we can rewrite the \cgd updates  s
 \[
@@
 \]
-which is identical to the dual BMD if we make the substitutions $x_{t} \to -\theta_{t}$ and $y_{t} \to -u_{t}$.
+which is identical to the \dual if we make the substitutions $x_{t} \to -\theta_{t}$ and $y_{t} \to -u_{t}$.
 
-We similarly relate Primal BMD and Mirror Descent.
-BMD minimizes $g + R^{*}$ while CGD minimizes $h + f$.
+We similarly relate \primal and \md.
+\primal minimizes $g + R^{*}$ while \md minimizes $h + f$.
 Make the identifications $h = g$ and $f = R^{*}$ and note the conditions of $h$ and $g$ are the same.
-Letting $\rho_{t+1} = \alpha_{t}/A_{t}$, we can rewrite the Mirror Descent updates as
+Letting $\rho_{t+1} = \alpha_{t}/A_{t}$, we can rewrite the \md updates as
 \[

### 2013-10-04 icml/introduction.tex
updated intro

@@
 
-The herding algorithm of \citet{Welling:2009a} has been significantly analyzed and extended since its introduction.
-Herding was originally conceived of as a deterministic method for performing joint learning and inference in Markov random field (MRFs).
-The ``learning'' interpretation of herding was derived as a fixed step size gradient descent algorithm on the zero temperature limit of a the maximum entropy optimization problem for the MRF.
-Herding for MRFs was further explored in \citet{Welling:2009a,Gelfand:2010,Bornn:2013}.
+Many algorithms, such as SMC \cite{smc} and MCMC \cite{mcmc}, approximate a 
+distribution by samples. Most such algorithms require access to a full posterior and 
+may be prone to local optima. In contrast, the herding algorithm introduced in 
+\cite{Welling:2009a} generates samples whose moments are guaranteed to converge 
+to any given collection of feasible moments. Moreover, \emph{only} the moments 
+are required; a full posterior is not needed.
 
-In order to define herding, let $\sX$ be an observation space, $\phi : \sX \to \sB_{\Theta}$ a feature map to an inner product space, and $\bar\phi = \bE_{x \sim p}[\phi(x)]$ the empirical moments of some (possible empirical) distribution.
-Then the herding algorithm is given by the follow simple procedure to generate pseudosamples from $\sX$:
+Since its introduction, herding has been shown to have many desirable properties.
+For instance, Chen et al showed that the moments converge at a rate of $O(1/T)$ for 
+$T$ samples, which is even faster than the $O(1/\sqrt{T})$ that would be obtained 
+by sampling the exact posterior. Furthermore, Bach et al show that herding is 
+equivalent to conditional gradient, and Husz\'{a}r and Duvenaud show that 
+herding is Bayesian quadrature. Herding was originally conceived of as a method 
+for inference in MRF, and this line of work has been further explored in 
+\citet{Welling:2009a,Gelfand:2010,Bornn:2013}.
+
+In order to define herding, let $\sX$ be an observation space, $\phi : \sX \to \sB_{\Theta}$ a 
+feature map to an inner product space, and $\bar\phi = \bE_{x \sim p}[\phi(x)]$ the empirical 
+moments of some distribution. Then the herding algorithm is the following simple procedure to 
+generate pseudosamples from $\sX$:
 \[
@@
 \]
-
-An alternative interpretation of herding \citep{Chen:2010a,Huszar:2012,Bach:2012a} is as a moment matching algorithm.
+In this paper we generalize the herding algorithm to the following set of updates, which 
+require only the convexity of $h$ and $R$:
+\[
+u_t &\in \argmin_{u} h(u) + \ip{\theta_t}{u} \\
+\theta_{t+1} &\in \argmax_{\theta} \ip{\theta}{\hat{u}_t} - R(\theta),
+\]
+where $\hat{u}_t$ is a certain weighted average of $u_1,\ldots,u_t$. Our 
+algorithm is described in more detail in Section~\ref{sec:algorithm}. This 
+generalization has several advantages:
+\begin{itemize}
+\item It allows us to cast herding in the framework of mirror descent, 
+      which lets us better understand its convergence properties.
+\item It allows us to obtain generalizations of herding for infinite 
+      dimensional spaces that may apply even if standard herding does not.
+\item It makes the known connections between herding, subgradient, and 
+      conditional gradient methods more clear. 
+\end{itemize}
+In particular, while it was known that herding was a form of conditional 
+gradient \cite{TODO} and hence (via the relationship in \cite{TODO}) also 
+a form of subgradient descent, our framework exposes all three algorithms 
+as special cases. A full table of the relationships for various choices of 
+$h$ and $R$ is given in Figure~\ref{fig:table}.
+% -- Doesn't seem relevant enough to justify the amount of technical detail
+%An alternative interpretation of herding \citep{Chen:2010a,Huszar:2012,Bach:2012a} is as a moment matching algorithm.
 In particular, it generates a set of pseudosamples $\{ x_{t} \}$  which match the moments $\bar\phi$.
@@
 If $\sB_{\Theta}$ is a reproducing kernel Hilbert space (RKHS) with feature map $\phi$, $\sE_{T}^{2}$ is equivalent to the maximum mean discrepancy (MMD) between $p$ and the empirical distribution $\hat p_{T} \eqdef T^{-1}\sum_{t=1}^{T} \delta_{x_{t}}$ \citep{Huszar:2012}:
-\(
-\sE_{T}^{2} &= \MMD^{2}(p, \hat p_{T})  \\
-&\eqdef \sup_{\theta \in \Theta} \left|\bE_{p}\ip{\theta}{\phi(x)} - \bE_{\hat p_{T}}\ip{\theta}{\phi(x)} \right|,
-\)
-where $\Theta = \{ \theta \in \sB_{\Theta} \ | \  \|\theta\| = 1 \}$.
+%\(
+%\sE_{T}^{2} &= \MMD^{2}(p, \hat p_{T})  \\
+%&\eqdef \sup_{\theta \in \Theta} \left|\bE_{p}\ip{\theta}{\phi(x)} - \bE_{\hat p_{T}}\ip{\theta}{\phi(x)} \right|,
+%\)
+%where $\Theta = \{ \theta \in \sB_{\Theta} \ | \  \|\theta\| = 1 \}$.
 
-If we are interested in generating samples from $p$ (or some distribution with the same moments as $p$), herding provides an attractive alternative to standard i.i.d.
+%If we are interested in generating samples from $p$ (or some distribution with the same moments as $p$), herding provides an attractive alternative to standard i.i.d.
 sampling for two reasons.
@@
 
-In this paper, we use a novel gradient descent algorithm to investigate herding and its generalizations:
+In summary, we introduce a novel gradient descent algorithm to investigate herding and its generalizations:
 \begin{itemize}
-\item In Section \ref{sec:algorithm} we introduce a novel pair of convex optimization algorithms, which we call primal and dual boosted mirror descent (\primal and \dual).
+\item In Section \ref{sec:algorithm} we introduce a pair of convex optimization algorithms, which we call primal and dual boosted mirror descent (\primal and \dual).
 We show how \primal and \dual relate to \cgd and mirror descent (\md).
 \item We then use \dual in Section \ref{sec:herding} to define a class of generalized herding algorithms.
-Connections with the ``learning'' interpretation are made by showing how to obtain herding as the limit of an entropically normalized case of \dual.
+Connections with the ``learning'' interpretation of herding are made by showing how to obtain herding as the limit of an entropically normalized case of \dual.
 \item Section \ref{sec:proofs} presents convergence rates for \primal and \dual.
-As special cases, we recover a $O(\log T/ T)$ convergence rate for herding and a $O(1/T)$ rate for a herding-like algorithm that uses unequally weighted samples.
+As special cases, we recover a $O(\log T/ T)$ convergence rate for classical herding and a $O(1/T)$ rate for weighted herding.
 Both of these results apply to the infinite dimensional case.

### 2013-10-04 icml/boosted-mirror-descent.tex
draft of connections subsection

@@
 \end{enumerate}
-As long as $h$ is strongly convex, we obtain the 
+As long as $h$ is strongly convex, for the proper choice of $\alpha_{t}$ we obtain the 
 bound (see Corollary~\ref{cor:method-1}):
@@
 \[ \sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq L_{*} + O(1/T). \]
-As suggested by the name \primal and \dual are closely related; indeed, they are dual to each other in the sense that 
-performing \primal on $L(u,\theta)$ is the same as performing \dual on 
-$-L(\theta,u)$.
 
@@
 
-The \bmd algorithms are very closely related to \cgd and \md.
+As suggested by their names, the \bmd algorithms are closely related: they are dual to each other in the sense that 
+performing \primal on $L(u,\theta)$ is the same as performing \dual on $-L(\theta,u)$. This relationship is obvious by inspection of the update equations:
+\(
+\text{Primal} && \text{Dual} \\
+\theta_{t} &= \partial R^{*}(u_{t}) & \theta_{t+1} &= \partial R^{*}(\hat u_{t}) \\
+u_{t+1} &= \partial h^{*}(-\hat\theta_{t}) & u_t &= \partial h^{*}(-\theta_{t})
+\)
+This duality is analogous to that between mirror descent and conjugate gradient descent \citep{Bach:2012b}.
+\md minimizes $g(x) + f(x)$ and requires that $g$ be $\alpha$-strongly convex while \cgd minimizes $g^{*}(x) + f^{*}(x)$ and requires that $g$ be $\alpha$-strongly convex.
+The update equations are
+\(
+y_{t} &= \partial f(x_{t})  \\
+\bar y_{t} &=  (1 - \rho_{t+1})\partial g(x_{t}) - \rho_{t+1}y_{t}  \\
+x_{t+1} &=  \partial g^{*}(\bar y_{t}) 
+\)
+for \md and
+\(
+ y_{t+1} &= \partial f(x_{t}) \\
+\bar y_{t+1} &= (1 - \rho_{t})\bar y_{t} + \rho_{t} y_{t+1}  \\
+x_{t} &= \partial g^{*}(-\bar y_{t})
+\)
+for \cgd. 
+In fact, the \bmd algorithms are very closely related to \cgd and \md: \cgd can be obtained from \dual and \md obtained from \primal.
+The relationships are given in Table~\ref{tab:connections}.
+The table also shows how generalized herding and $q$-herding --- which are discussed in the next section --- are obtained from \bmd.
 
+\renewcommand{\arraystretch}{1.5}
+\begin{center}
+\begin{table*}[t]
+\caption{Formulation of herding and convex optimization algorithms in terms of \bmd. Full details relating \md and \cgd to \bmd are given in the Supplementary Material.}
+\newcolumntype{L}{>{\raggedright\arraybackslash}X}
+\begin{tabularx}{\textwidth}{L|lllLL}
+Algorithm 			& $h(u)$ 	& $R(\theta)$ & $\alpha_{t}$ & \bmd Algorithm & Notes 
+\\ \hline
+generalized herding & 0			& $\ip{\bar\phi}{\theta} + \frac{1}{2} \| \theta \|_{\sH}^{2}$ 	& unconstrained 	& \dual &
+\\ 
+$q$-herding			& 0			& $\ip{\bar\phi}{\theta} + \frac{1}{q} \| \theta \|_{q}^{q}$ 	& unconstrained 	& \dual & $q \ge 2$
+\\ 
+\md					& $g(x)$	& $f^{*}(y)$				& $\rho_{t}/\prod_{s=1}^{t} (1 - \rho_{s})$				& \primal & 
+\\ 
+\cgd				& $f^{*}(y)$& $g(x)$					& $\rho_{t}/\prod_{s=1}^{t} (1 - \rho_{s})$				& \dual &
+\\
+\end{tabularx}
+\label{tab:connections}
+\end{table*}
+\end{center}
 \end{document}

### 2013-10-04 icml/introduction.tex
draft of connections subsection

@@
 Since its introduction, herding has been shown to have many desirable properties.
-For instance, Chen et al showed that the moments converge at a rate of $O(1/T)$ for 
-$T$ samples, which is even faster than the $O(1/\sqrt{T})$ that would be obtained 
+For instance, Chen et al showed that the moments converge at a rate of $O(1/T^{2})$ for 
+$T$ samples, which is even faster than the $O(1/T)$ that would be obtained 
 by sampling the exact posterior. Furthermore, Bach et al show that herding is 
-equivalent to conditional gradient, and Husz\'{a}r and Duvenaud show that 
+equivalent to conditional gradient, and \citet{Huszar:2012} show that 
 herding is Bayesian quadrature. Herding was originally conceived of as a method 
 for inference in MRF, and this line of work has been further explored in 
-\citet{Welling:2009a,Gelfand:2010,Bornn:2013}.
+\citet{Welling:2009a}, \citet{Gelfand:2010} and \citet{Bornn:2013}.
 
@@
 In particular, while it was known that herding was a form of conditional 
-gradient \cite{TODO} and hence (via the relationship in \cite{TODO}) also 
+gradient \cite{Bach:2012a} and hence (via the relationship in \cite{Bach:2012b}) also 
 a form of subgradient descent, our framework exposes all three algorithms 
 as special cases. A full table of the relationships for various choices of 
-$h$ and $R$ is given in Figure~\ref{fig:table}.
+$h$ and $R$ is given in Table~\ref{tab:connections}.
 % -- Doesn't seem relevant enough to justify the amount of technical detail

### 2013-10-04 icml/herding-connection.tex
added a bit more detail to q-herding

@@
 We define $\Delta_{\sX}$ to be the simplex 
-of probability distributions over $\sX$ and the marginal polytope 
+of probability distributions over $\sX$ and define the marginal polytope 
 $\sM \eqdef \{\bE_{\mu}[\phi(x)] \mid \mu \in \Delta_{\sX}\}$.
@@
 
-Note that $R$ is convex with respect to $\|\cdot\|_2$ and that 
-$R^*(u) = \frac{1}{2}\|\bar{\phi}-\bE_{u}[\phi(x)]\|_2^2$. We can 
+Note that $R$ is strongly convex with respect to $\|\cdot\|_{\sH}$ and that 
+$R^*(u) = \frac{1}{2}\|\bar{\phi}-\bE_{u}[\phi(x)]\|_{\sH}^2$. We can 
 therefore apply Corollary~\ref{thm:method-2} to obtain the bound
@@
 The convergence bound for herding implicitly assumes that 
-$\sM$ is bounded in the $l^2$ norm. Assuming that $\sH$ is 
-finite-dimensional, this will always be true. However, in 
+$\sM$ is bounded in norm. Assuming that $\sX$ is compact 
+and that $\sH$ is finite-dimensional, this will always be true. However, in 
 the infinite-dimensional case this may no longer be true. 
 For instance, consider $\sX = \{1,\ldots,N\}$ and let 
-$\phi : \sX \to l^2(\bZ_{\geq 0})$ be
+$\phi : \sX \to l^3(\bZ_{\geq 0})$ be
 defined by $\phi(n)_i = \frac{1}{\sqrt{n+i}}$. Then 
@@
 $h \equiv 0$, $R(\theta) = \ip{\bar{\phi}}{\theta} + \frac{1}{q}\|\theta\|_q^q$. 
-Then $R^*(u) = \frac{1}{p}\|\bE_{u}[\phi(x)]-\bar{\phi}\|_p^p$. Define the 
-``correction function'' $C_q$ by
+Then $R^*(u) = \frac{1}{p}\|\bE_{u}[\phi(x)]-\bar{\phi}\|_p^p$. And 
+$\left(\frac{\partial R^*}{\partial u}\right)_i = \sign(u_i) \cdot |u_i|^{p-1}$. 
+Therefore, define the ``correction function'' $C_p$ by
 \begin{equation}
-C_q(x)_i \eqdef \sign(x_i) \cdot |x_i|^{\frac{1}{q-1}}.
+C_p(u)_i \eqdef \sign(u_i) \cdot |u_i|^{p-1}.
 \end{equation}
@@
 \begin{itemize}
-\item $\theta_{1} = C_q(-\bar{\phi})$
+\item $\theta_{1} = C_p(-\bar{\phi})$
 \item $x_{t} \in \arg\min_{x \in \sX} \ip{\phi(x)}{\theta_t}$
-\item $\theta_{t+1} = C_q\left(\frac{1}{t} \sum_{s=1}^t [\phi(x_s)-\bar{\phi}]\right)$
+\item $\theta_{t+1} = C_p\left(\frac{1}{t} \sum_{s=1}^t [\phi(x_s)-\bar{\phi}]\right)$
 \end{itemize}
+Note that $\|C_p(u)\|_q^q = \sum_{i} |u_i|^{q(p-1)} = \sum_{i} |u_i|^p = \|u\|_p^p$. 
+Therefore, $C_p(u) \in l^q(\bZ_{\geq 0})$ whenever $u \in l^p(\bZ_{\geq 0})$. By 
+H\"{o}lder's inequality, it follows that $\ip{u}{\theta_t}$ exists for all 
+$u \in l^p(\bZ_{\geq 0})$ and hence the $q$-herding updates are well-defined.
+
 By Proposition~\ref{prop:method-2}, we know that

### 2013-10-04 icml/herding-connection.tex
explicit construction of alpha from beta in lower bound proof

--- 2013-10-04 icml/herding-connection.tex
@@
 algorithm includes an additional $\bar{\phi}$ term in the third bullet 
-point and indexes $x_t$ as $x_{t+1}$.} [\cite{chen}, equations (1) and (2)]:
+point and indexes $x_t$ as $x_{t+1}$.} [\cite{Chen:2010}, equations (1) and (2)]:
 \begin{itemize}
@@
 
-The above updates correspond to Algorithm 2 for $U = \sX$, $\Theta = \sH$, 
+The above updates correspond to \dual for $U = \sX$, $\Theta = \sH$, 
 $\alpha_t = 1$, $h \equiv 0$, 
@@
 this gives us the third bullet point. Finally, the second bullet point 
-is already in the form given in Algorithm 2.
+is already in the form given in \dual.
 
@@
 \end{itemize}
+By Proposition~\ref{prop:method-2}, we know that
+\begin{align}
+\frac{1}{p}\|\bar{\phi}-\frac{1}{T}\sum_{t=1}^T \phi(x_t)\|_p^p \leq \frac{1}{A_T} \sum_{t=1}^T A_tD_{R^*}(\hat{u}_{t+1}\|\hat{u}_t).
+\end{align}
+We can no longer use Corollary~\ref{cor:method-2} since $\|\cdot\|_p^p$ is not strongly convex. 
+Instead, we use the following two inequalities (proofs are given in the Supplementary Material):
+\begin{lemma}
+Let $x, y \in \bR_{\geq 0}$. Then:
+\begin{equation*}
+x^p-py^{p-1}x+(p-1)y^p \leq \binom{p}{2}(x-y)^2\max(x,y)^{p-2}
+\end{equation*}
+and
+\begin{equation*}
+(y-z)^2\max(x,y)^{p-2} \leq 4\max(x,y,z)^p
+\end{equation*}
+\end{lemma}
+Using these, we obtain the bound
+\(
+\lefteqn{D_{R*}(\hat{u}_{t+1} \| \hat{u}_t)} \\
+ &= \frac{1}{p} \sum_{i} \hat{u}_{t+1,i}^p - p\hat{u}_{t+1,i}\hat{u}_{t,i}^{p-1} + (p-1)\hat{u}_{t,i}^{p} \\
+ &\leq \frac{1}{p}\binom{p}{2} \sum_i (\hat{u}_{t+1,i}-\hat{u}_{t,i})^2\max(\hat{u}_{t+1,i},\hat{u}_{t,i})^{p-2} \\
+ &= \frac{1}{p}\binom{p}{2}\frac{\alpha_{t+1}^2}{A_{t+1}^2} \sum_i (u_{t+1,i}-\hat{u}_{t})^2\max(\hat{u}_{t+1,i},\hat{u}_{t,i})^{p-2} \\
+ &\leq 2(p-1)\frac{\alpha_{t+1}^2}{A_{t+1}^2} \sum_i \max(\hat{u}_{t,i},\hat{u}_{t+1,i},u_{t+1,i})^p \\
+ &\leq 2(p-1)\frac{\alpha_{t+1}^2}{A_{t+1}^2} \sum_i \hat{u}_{t,i}^p+\hat{u}_{t+1,i}^p+u_{t+1,i}^p \\
+ &\leq 6(p-1)\frac{\alpha_{t+1}^2}{A_{t+1}^2} r_p^p
+\)
+Plugging back in yields a convergence rate of $\frac{6(p-1)r_p^p(\log(T)+1)}{T}$ for $\alpha_t = 1$ 
+and $\frac{24(p-1)r_p^p}{T}$ for $\alpha_t = t$. We therefore obtain a convergence rate of $O(T^{-1/p})$ 
+in the $l^p$ norm, even if $\sM$ is not bounded in the $l^2$ norm.
 
@@
 
-Bach et al.~have observed that herding appears to be approximately building 
+\citet{Welling:2009a} and \citet{Bach:2012b} have observed that herding appears to be approximately building 
 samples from the maximum entropy distribution for a given collection of moment 
@@
 $R(\theta) = \ip{\bar{\phi}}{\theta} + \frac{1}{2} \|\theta\|_{\sH}^2$.
-In this case, Algorithm 2 gives the updates
+In this case, the \dual updates are
 \begin{itemize}
@@
 
+Note that as $\eta \to 0$, $u_{t}$ concentrates around its mode, converging to delta function, which is the herding sample.
+
 \end{document}

### 2013-10-04 icml/lower-bounds.tex
explicit construction of alpha from beta in lower bound proof

@@
 
-Let $\sX$ be the set of vectors $(x_{1},x_{2},\dots) \in [0,1]^{\infty}$ such that $\sum_{n=1}^{\infty} x_{n} = 1$ and a finite number of the $x_{n}$ are non-zero.
-Define the infinite feature vector $\phi$ to be the imbedding of $\sX$ into $\ell^{1}$.
-Let the target distribution $p(x)$ be chosen such that $\bar\phi_{n} = c_{\alpha}/\alpha(n)$, where $\alpha(n)$ is any superlinear function (that is, $\alpha \in \omega(n)$) and $c_{\alpha}^{-1} = \sum_{n=1}^{\infty} \alpha(n) < \infty$.
-We will consider the standard $\ell^{2}$ herding loss function $L(\theta) = \|\theta - \bar \phi\|_2^2$.
-
-Define a {\em sparse update method} to be any method that produces an estimator of $\bar \phi$ at iteration $t$ of the form 
+For the matching $O(1/T)$ lower bound we analyze sparse update methods.
+Formally, define a {\em sparse update method} to be any method that produces an estimator of $\bar \phi$ at iteration $t$ of the form
 \[
@@
 \]
-In other words, the $t$-th estimator will have at most $t$ non-zero entries, and so 
+Sparse update methods can be viewed as characterizing ``herding-like'' algorithms, including the generalized herding and \cgd with line search.
+These sparse methods -- which only operate on the extreme points of the simplex -- are of interest because they provide (weighted) ``samples'' matching the moments of the distribution of interest.
+
+The bound setup is as follows.
+Let $\sX$ be the set of vectors $(x_{1},x_{2},\dots) \in [0,1]^{\infty}$ such that $\sum_{n=1}^{\infty} x_{n} = 1$ and a finite number of the $x_{n}$ are non-zero.
+Define the infinite feature vector $\phi$ to be the imbedding of $\sX$ into $\ell^{1}$.
+Let the target distribution $p(x)$ be chosen such that $\bar\phi_{n} = c/\alpha(n)$, where $\alpha(n)$ is any superlinear function and $c^{-1} = \sum_{n=1}^{\infty} \alpha(n) < \infty$.
+Without loss of generality we can assume that $\alpha$ is monotonically increasing.
+We will consider the standard $\ell^{2}$ herding loss function $L(\theta) = \|\theta - \bar \phi\|_2^2$.
+
+Since we are only considering sparse update methods, the $t$-th estimator will have at most $t$ non-zero entries, and so 
 \[
-\|x-x^{*}\|_2^2 \ge c_{\alpha}^{2}\sum_{n > t} 1/\alpha(n)^{2} = \Theta(1/\beta(t)) \label{eq:convergence-lb}
+\|x-x^{*}\|_2^2 \ge c^{2}\sum_{n > t} 1/\alpha(n)^{2} = \Theta(1/\beta(t)) \label{eq:convergence-lb}
 \]
-for some $\beta \in \omega(t)$.
-But for any $\beta \in \omega(t)$ we can choose some $\alpha \in \omega(n)$ to obtain \eqref{eq:convergence-lb}.
+for some (monotonically increasing) superlinear function $\beta$.
+But for any such $\beta$ we can choose some $\alpha$ to obtain \eqref{eq:convergence-lb}.
+In particular, let $1/\gamma(t) = \frac{\dee}{\dee t}\frac{1}{\beta(t)}$, where $\gamma(t) \in \omega(t^{2})$.
+Then we can choose $\alpha = \sqrt \gamma \in \omega(t)$ to satisfy \eqref{eq:convergence-lb}.
+
 Therefore, any sparse update method must have convergence rate no faster than $\Theta(1/\beta(t))$ for all superlinear $\beta$, and hence no faster than $\Theta(1/t)$.
 
-Sparse update methods can be viewed as charactering ``herding-like algorithms,'' including the set of generalized herding algorithms based on \dual and \cgd with line search.
-Those sparse methods are of interest because they provide (weighted) ``samples'' matching the moments of the distribution of interest.
-Since they only operate on the extreme points of the simplex, the optimization problems required to use them are more tractable.
-
-\NA{JH: move definition and discussion of sparse update method to beginning}
-
 \end{document}

### 2013-10-04 icml/boosted-mirror-descent.tex
small corrections, expositional changes, and improved citations

@@
 Strong convexity and smoothness are dual to each in the sense that a convex function $f$ is $\alpha$-strongly convex if and only if $f^{*}$ is $(1/\alpha)$-smooth.
+\NA{JH: should double check that we actually need smoothness}
 
@@
 \]
-for which we wish to find arguments that obtain $L_{*} = \min_{u}\max_{\theta} L(u, \theta)$. Taking only max, we have the one-argument loss function:
+for which we wish to find arguments that obtain $L_{*} = \min_{u}\max_{\theta} L(u, \theta)$. Taking only the maximum, we have the one-argument loss function
 \[
@@
 We can think of $h$ as a \emph{primal regularizer} and $R$ as a \emph{dual regularizer}.
-The pair of regularization functions $h$ and $R$ provides flexible optimization framework, as seen from the following examples:
+The pair of regularization functions $h$ and $R$ define a flexible class of functions of optimize, as seen from the following examples:
 \begin{enumerate}
@@
 Mirror descent, together with boosting, yields an algorithm for finding 
-``saddle points'' of $L$. 
+``saddle points'' of $L(u, \theta)$. 
 For notational convenience, for a sequence of weights $\alpha_1,\alpha_2,\ldots$ 
-let $\hat{u}_t = \frac{\sum_{s=1}^t \alpha_su_s}{\sum_{s=1}^t \alpha_s}$ and let 
-$\hat{\theta}_t = \frac{\sum_{s=1}^t \alpha_s\theta_s}{\sum_{s=1}^t \alpha_s}$. Then 
+let 
+\(
+\hat{u}_t = \frac{\sum_{s=1}^t \alpha_su_s}{\sum_{s=1}^t \alpha_s} 
+\quad \text{and} \quad 
+\hat{\theta}_t = \frac{\sum_{s=1}^t \alpha_s\theta_s}{\sum_{s=1}^t \alpha_s}.
+\)
+Then 
 the {\em primal boosted mirror descent} (\primal) algorithm is:
@@
 In other words, $\hat{u}_T$ is close to being a global minimum of $L(u)$.
-However, it is often the case that $h$ is not strongly convex, whereas $R$ is strongly convex. As we shall see, this will be the case for our application to herding.
+However, it is often the case that $h$ is not strongly convex but $R$ is strongly convex. As we shall see, this will be the case for our application to herding.
 In this case we may wish to use the following slightly different algorithm, which we call {\em dual boosted mirror descent} (\dual):
@@
 \md minimizes $g(x) + f(x)$ and requires that $g$ be $\alpha$-strongly convex while \cgd minimizes $g^{*}(x) + f^{*}(x)$ and requires that $g$ be $\alpha$-strongly convex.
-The update equations are
+The update equations for \md are
 \(
@@
 \)
-for \md and
+and for \cgd are
 \(
@@
 \bar y_{t+1} &= (1 - \rho_{t})\bar y_{t} + \rho_{t} y_{t+1}  \\
-x_{t} &= \partial g^{*}(-\bar y_{t})
+x_{t} &= \partial g^{*}(-\bar y_{t}).
 \)
-for \cgd. 
 In fact, the \bmd algorithms are very closely related to \cgd and \md: \cgd can be obtained from \dual and \md obtained from \primal.
@@
 \begin{table*}[t]
-\caption{Formulation of herding and convex optimization algorithms in terms of \bmd. Full details relating \md and \cgd to \bmd are given in the Supplementary Material.}
+\caption{Formulation of herding and convex optimization algorithms in terms of \bmd. Full details relating \md and \cgd to \bmd are given in the Supplementary Material. See also \citet{Bach:2012b}.}
 \newcolumntype{L}{>{\raggedright\arraybackslash}X}
@@
 \\ \hline
+herding 			& 0			& $\ip{\bar\phi}{\theta} + \frac{1}{2} \| \theta \|_{\sH}^{2}$ 	& 1				 	& \dual &
+\\
 generalized herding & 0			& $\ip{\bar\phi}{\theta} + \frac{1}{2} \| \theta \|_{\sH}^{2}$ 	& unconstrained 	& \dual &
 \\ 
-$q$-herding			& 0			& $\ip{\bar\phi}{\theta} + \frac{1}{q} \| \theta \|_{q}^{q}$ 	& unconstrained 	& \dual & $q \ge 2$
+$q$-herding			& 0			& $\ip{\bar\phi}{\theta} + \frac{1}{q} \| \theta \|_{q}^{q}$ 	& unconstrained 	& \dual & $1 \le q \le 2$
 \\ 

### 2013-10-04 icml/herding-connection.tex
small corrections, expositional changes, and improved citations

@@
 
-The standard setup for herding is as follows: we are given a 
+We now derive herding as a special case of \dual. The standard setup for herding is as follows: we are given a 
 space $\sX$ and a feature mapping $\phi : \sX \to \sH$ for some 
@@
 such that $\|\frac{1}{T}\sum_{t=1}^T \phi(x_t)-\bar{\phi}\|_{\sH}^2$ 
-is minimized for some given $\bar{\phi} \in \sM$.
+is minimized for some given $\bar{\phi} \in \sM$. One can typically
+think of $\bar \phi$ as being the expected value of $\phi$ under some 
+distribution $p \in \Delta_{\sX}$.
 
@@
 algorithm includes an additional $\bar{\phi}$ term in the third bullet 
-point and indexes $x_t$ as $x_{t+1}$.} [\cite{Chen:2010}, equations (1) and (2)]:
+point and indexes $x_t$ as $x_{t+1}$.} \citep[][equations (1) and (2)]{Chen:2010a}:
 \begin{itemize}

### 2013-10-04 icml/introduction.tex
small corrections, expositional changes, and improved citations

@@
 
-Many algorithms, such as SMC \cite{smc} and MCMC \cite{mcmc}, approximate a 
+Many algorithms, such as SMC \citep{smc} and MCMC \citep{mcmc}, approximate a 
 distribution by samples. Most such algorithms require access to a full posterior and 
 may be prone to local optima. In contrast, the herding algorithm introduced in 
-\cite{Welling:2009a} generates samples whose moments are guaranteed to converge 
+\citet{Welling:2009a} generates samples whose moments are guaranteed to converge 
 to any given collection of feasible moments. Moreover, \emph{only} the moments 
@@
 $T$ samples, which is even faster than the $O(1/T)$ that would be obtained 
-by sampling the exact posterior. Furthermore, Bach et al show that herding is 
+by sampling independently from the exact posterior. Furthermore, \citet{Bach:2012a} show that herding is 
 equivalent to conditional gradient, and \citet{Huszar:2012} show that 
-herding is Bayesian quadrature. Herding was originally conceived of as a method 
-for inference in MRF, and this line of work has been further explored in 
+herding is equivalent to Bayesian quadrature. Herding was originally conceived of as a method 
+for inference in Markov random fields, and this line of work has been further explored in 
 \citet{Welling:2009a}, \citet{Gelfand:2010} and \citet{Bornn:2013}.
@@
 In particular, while it was known that herding was a form of conditional 
-gradient \cite{Bach:2012a} and hence (via the relationship in \cite{Bach:2012b}) also 
+gradient \cite{Bach:2012a} and hence (via the relationship in \citet{Bach:2012b}) also 
 a form of subgradient descent, our framework exposes all three algorithms 

### 2013-10-04 icml/boosted-mirror-descent.tex
more small fixes; tried to improve l.b. section

@@
 \]
-and that a (possibly non-convex) function $g$ is $\beta$-smooth if 
-\[
-\|\partial g(x) - \partial g(y) \| \le \beta \| x - y\|.
-\]
-Strong convexity and smoothness are dual to each in the sense that a convex function $f$ is $\alpha$-strongly convex if and only if $f^{*}$ is $(1/\alpha)$-smooth.
+%and that a (possibly non-convex) function $g$ is $\beta$-smooth if 
+%\[
+%\|\partial g(x) - \partial g(y) \| \le \beta \| x - y\|.
+%\]
+%Strong convexity and smoothness are dual to each in the sense that a convex function $f$ is $\alpha$-strongly convex if and only if $f^{*}$ is $(1/\alpha)$-smooth.
 \NA{JH: should double check that we actually need smoothness}
@@
 The relationships are given in Table~\ref{tab:connections}.
-The table also shows how generalized herding and $q$-herding --- which are discussed in the next section --- are obtained from \bmd.
+The table also shows how $\alpha_{t}$-weighted herding and $q$-herding --- which are discussed in the next section --- are obtained from \bmd.
 
@@
 \newcolumntype{L}{>{\raggedright\arraybackslash}X}
-\begin{tabularx}{\textwidth}{L|lllLL}
+\begin{tabularx}{\textwidth}{L|lllcL}
 Algorithm 			& $h(u)$ 	& $R(\theta)$ & $\alpha_{t}$ & \bmd Algorithm & Notes 
@@
 \\
-generalized herding & 0			& $\ip{\bar\phi}{\theta} + \frac{1}{2} \| \theta \|_{\sH}^{2}$ 	& unconstrained 	& \dual &
+$\alpha_{t}$-weighted herding 	& 0	& $\ip{\bar\phi}{\theta} + \frac{1}{2} \| \theta \|_{\sH}^{2}$ 	& unconstrained 	& \dual & $\sum_{s =1}^{t} \alpha_{t} > 0~\forall t$
 \\ 

### 2013-10-04 icml/conclusion.tex
more small fixes; tried to improve l.b. section

--- 2013-10-04 icml/boosted-mirror-descent.tex
@@
 
-\section{Boosted Mirror Descent}
-\label{sec:algorithm}
+\section{Conclusion} 
+\label{sec:conclusion}
 
-Our results primarily rely on a pair of novel convex optimization algorithms we call primal and dual boosted mirror descent.
+\NA{Write me!}
 
-\subsection{Convex Optimization}
-
-Before introducing the boosted mirror descent (\bmd) algorithms, we recall some basic definitions related convex optimization.
-Let $f$ be a real function whose domain is a Banach space.
-The {\em convex conjugate} $f^{*}$ of $f$ is defined as
-\[
-f^{*}(y) \eqdef \max_{x} \ip{x}{y} - f(x).
-\]
-If $f$ is lower semi-continuous, then by the Fenchel-Moreau theorem $f = f^{**}$, i.e.
-\[
-f(x) = \max_{y} \ip{x}{y} - f^{*}(y).
-\]
-Throughout the paper, we will assume without further notice that functions are lower semi-continuous when required.
-The gradients of $f^{*}$ and $f$ can be defined as any argument that maximizes the previous two equations:
-\[
-\partial f^{*}(y) &\in \argmax_{x} \ip{x}{y} - f(x) \\
-\partial f(x) &\in \argmax_{y} \ip{x}{y} - f^{*}(y).
-\]
-
-In order to prove fast convergence, we will require additional assumptions beyond convexity. We say a function $f$ is {\em $\alpha$-strongly convex} if 
-\[
-f(x)  - f(y) \le \ip{\partial f(x)}{x-y} - \frac{\alpha}{2}\|x - y\|^{2}.
-\]
-and that a (possibly non-convex) function $g$ is $\beta$-smooth if 
-\[
-\|\partial g(x) - \partial g(y) \| \le \beta \| x - y\|.
-\]
-Strong convexity and smoothness are dual to each in the sense that a convex function $f$ is $\alpha$-strongly convex if and only if $f^{*}$ is $(1/\alpha)$-smooth.
-
-\subsection{The Boosted Mirror Descent Algorithms}
-
-We can now define the \bmd  algorithms.
-Let $U$ and $\Theta$ be convex subsets of the Banach spaces $\sB_{U}$ and $\sB_{\Theta}$, respectively, and let $\imath : U \to \Theta$ be a linear imbedding from $U$ to $\Theta$ that preserves the inner product.
-Consider the two-argument loss function $L : U \times \Theta \to \bR$
-\[
-L(u,\theta) = h(u) + \ip{\theta}{\imath(u)} - R(\theta)
-\]
-for which we wish to find arguments that obtain $L_{*} = \min_{u}\max_{\theta} L(u, \theta)$. Taking only max, we have the one-argument loss function:
-\[
-L(u)
-&= h(u) + \max_{\theta \in \Theta} \left\{\ip{\theta}{\imath(u)} - R(\theta) \right\} \\
-&= h(u) + R^{*}(\imath(u)).
-\]
-We can think of $h$ as a \emph{primal regularizer} and $R$ as a \emph{dual regularizer}.
-The pair of regularization functions $h$ and $R$ provides flexible optimization framework, as seen from the following examples:
-\begin{enumerate}
-\item Let $U$ be the probability simplex over a space $\sX$, let $\phi : \sX \to \bR^d$ be a collection 
-      of statistics, and let $\Theta$ be the unit $L^1$ ball in $\bR^d$. Let $h(u) = 0$, 
-      $\langle \theta, \imath(u) \rangle = \bE_u[\theta^T\phi(x)]$, and 
-      $R(\theta) = \langle \theta, \imath(u_0) \rangle$. Then $L(u)$ is the 
-      \emph{maximum mean discrepancy} between $u$ and $u_0$ relative to $\phi$.
-\item Let $h(u) = 0$ and $R(\theta) = S^*(i^{-1}(\theta))$ where $S^*$ is the 
-      convex conjugate of any strongly convex function $S : U \to \bR$. Then $L(u) = S(u)$ 
-      as long as $i^{-1}(\Theta) = \sB_{U}$.
-\item Let $U = \Theta = \bR^d$ and $\imath$ the identity map. Let $h(u) = \|u\|_1$, $R(u) = \langle \theta, u_0 \rangle + \frac{1}{2} \|\theta\|_2^2$. 
-      Then $L(u) = \|u\|_1 + \frac{1}{2} \|u-u_0\|_2^2$. 
-\end{enumerate}
-
-Going forward, for convenience we will abuse notation and use $\ip{\theta}{u}$ in place of 
-$\ip{\theta}{\imath(u)}$ when the map $\imath$ is clear from context. We will assume 
-throughout that $h$ and $R$ are both convex functions, and furthermore that 
-$\arg\max_{u} h(u) + \ip{\theta}{u}$ can be efficiently 
-computed for all values of $\theta$.
-
-%In this paper we will consider a family of methods for minimizing $L(u)$, 
-%which are based on Bach's generalization of the Frank-Wolfe algorithm and can 
-%be interpreted as boosted mirror descent.
-
-Mirror descent, together with boosting, yields an algorithm for finding 
-``saddle points'' of $L$. 
-For notational convenience, for a sequence of weights $\alpha_1,\alpha_2,\ldots$ 
-let $\hat{u}_t = \frac{\sum_{s=1}^t \alpha_su_s}{\sum_{s=1}^t \alpha_s}$ and let 
-$\hat{\theta}_t = \frac{\sum_{s=1}^t \alpha_s\theta_s}{\sum_{s=1}^t \alpha_s}$. Then 
-the {\em primal boosted mirror descent} (\primal) algorithm is:
-\begin{enumerate}
-\item $u_1 \in \arg\min_u h(u)$
-\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, u_t \rangle - R(\theta) = \partial R^{*}(u_{t})$
-\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, u \rangle = \partial h^{*}(-\hat\theta_{t})$
-\end{enumerate}
-As long as $h$ is strongly convex, for the proper choice of $\alpha_{t}$ we obtain the 
-bound (see Corollary~\ref{cor:method-1}):
-\begin{equation}
-\sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq L_{*} + O(1/T).
-\end{equation}
-In other words, $\hat{u}_T$ is close to being a global minimum of $L(u)$.
-However, it is often the case that $h$ is not strongly convex, whereas $R$ is strongly convex. As we shall see, this will be the case for our application to herding.
-In this case we may wish to use the following slightly different algorithm, which we call {\em dual boosted mirror descent} (\dual):
-\begin{enumerate}
-\item $\theta_1 \in \arg\min_{\theta} R(\theta)$
-\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle = \partial h^{*}(-\theta_{t})$
-\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \hat{u}_t \rangle - R(\theta) = \partial R^{*}(\hat u_{t})$
-\end{enumerate}
-We then obtain the following bound 
-when $R$ is strongly convex (see Corollary~\ref{cor:method-2}):
-\[ \sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq L_{*} + O(1/T). \]
-
-\subsection{Connections to Other Gradient Descent Algorithms}
-
-As suggested by their names, the \bmd algorithms are closely related: they are dual to each other in the sense that 
-performing \primal on $L(u,\theta)$ is the same as performing \dual on $-L(\theta,u)$. This relationship is obvious by inspection of the update equations:
-\(
-\text{Primal} && \text{Dual} \\
-\theta_{t} &= \partial R^{*}(u_{t}) & \theta_{t+1} &= \partial R^{*}(\hat u_{t}) \\
-u_{t+1} &= \partial h^{*}(-\hat\theta_{t}) & u_t &= \partial h^{*}(-\theta_{t})
-\)
-This duality is analogous to that between mirror descent and conjugate gradient descent \citep{Bach:2012b}.
-\md minimizes $g(x) + f(x)$ and requires that $g$ be $\alpha$-strongly convex while \cgd minimizes $g^{*}(x) + f^{*}(x)$ and requires that $g$ be $\alpha$-strongly convex.
-The update equations are
-\(
-y_{t} &= \partial f(x_{t})  \\
-\bar y_{t} &=  (1 - \rho_{t+1})\partial g(x_{t}) - \rho_{t+1}y_{t}  \\
-x_{t+1} &=  \partial g^{*}(\bar y_{t}) 
-\)
-for \md and
-\(
- y_{t+1} &= \partial f(x_{t}) \\
-\bar y_{t+1} &= (1 - \rho_{t})\bar y_{t} + \rho_{t} y_{t+1}  \\
-x_{t} &= \partial g^{*}(-\bar y_{t})
-\)
-for \cgd. 
-In fact, the \bmd algorithms are very closely related to \cgd and \md: \cgd can be obtained from \dual and \md obtained from \primal.
-The relationships are given in Table~\ref{tab:connections}.
-The table also shows how generalized herding and $q$-herding --- which are discussed in the next section --- are obtained from \bmd.
-
-\renewcommand{\arraystretch}{1.5}
-\begin{center}
-\begin{table*}[t]
-\caption{Formulation of herding and convex optimization algorithms in terms of \bmd. Full details relating \md and \cgd to \bmd are given in the Supplementary Material.}
-\newcolumntype{L}{>{\raggedright\arraybackslash}X}
-\begin{tabularx}{\textwidth}{L|lllLL}
-Algorithm 			& $h(u)$ 	& $R(\theta)$ & $\alpha_{t}$ & \bmd Algorithm & Notes 
-\\ \hline
-generalized herding & 0			& $\ip{\bar\phi}{\theta} + \frac{1}{2} \| \theta \|_{\sH}^{2}$ 	& unconstrained 	& \dual &
-\\ 
-$q$-herding			& 0			& $\ip{\bar\phi}{\theta} + \frac{1}{q} \| \theta \|_{q}^{q}$ 	& unconstrained 	& \dual & $q \ge 2$
-\\ 
-\md					& $g(x)$	& $f^{*}(y)$				& $\rho_{t}/\prod_{s=1}^{t} (1 - \rho_{s})$				& \primal & 
-\\ 
-\cgd				& $f^{*}(y)$& $g(x)$					& $\rho_{t}/\prod_{s=1}^{t} (1 - \rho_{s})$				& \dual &
-\\
-\end{tabularx}
-\label{tab:connections}
-\end{table*}
-\end{center}
 \end{document}

### 2013-10-04 icml/herding-connection.tex
more small fixes; tried to improve l.b. section

@@
 space $\sX$ and a feature mapping $\phi : \sX \to \sH$ for some 
-reproducing kernel Hilbert space $\sH$ \cite{RKHS}. 
+reproducing kernel Hilbert space $\sH$ \cite{Hofmann:2008}. 
 We define $\Delta_{\sX}$ to be the simplex 
@@
 \end{align*}
-where $r \eqdef \sup_{m \in \sM} \|m\|_2$.
+where $r \eqdef \sup_{m \in \sM} \|m\|_2$. This is a slightly weaker 
+version of the typical herding bound on MMD.
 
-This is a slightly weaker version of the typical herding bound on MMD.
-Note that if we use $\alpha_t = t$, which corresponds to using a weighted 
-average, then we obtain a stronger bound of $\frac{8r^2}{T}$ instead.
+If we do not restrict $\alpha_{t}$ to be 1, we obtain a class of algorithms
+we call {\em $\alpha_{t}$-weighted herding}. For the $t$-weighted herding 
+algorithm with we obtain a stronger bound of $\frac{8r^2}{T}$ instead.
 
@@
 \end{equation}
-We obtain the following \emph{generalized herding} updates:
+We obtain the following \emph{$q$-herding} updates:
 \begin{itemize}
@@
 $u \in l^p(\bZ_{\geq 0})$ and hence the $q$-herding updates are well-defined.
+
+\NA{JH: make this into a theorem?}
 
@@
 
-Note that as $\eta \to 0$, $u_{t}$ concentrates around its mode, converging to delta function, which is the herding sample.
+Note that as $\eta \to 0$, $u_{t}$ concentrates around its mode, converging to a delta function at the herding sample location.
+It is thus possible that, under the proper regularity conditions, it can be shown using entropically regularized \bmd that herding converges to a distribution that is close to the maximum entropy distribution.
+However, we leave this line of investigation for future work.
 

### 2013-10-04 icml/introduction.tex
more small fixes; tried to improve l.b. section

@@
 distribution by samples. Most such algorithms require access to a full posterior and 
-may be prone to local optima. In contrast, the herding algorithm introduced in 
+may be prone to becoming stuck local modes. In contrast, the herding algorithm introduced in 
 \citet{Welling:2009a} generates samples whose moments are guaranteed to converge 
@@
 In particular, while it was known that herding was a form of conditional 
-gradient \cite{Bach:2012a} and hence (via the relationship in \citet{Bach:2012b}) also 
+gradient descent \cite{Bach:2012a} and hence (via the relationship in \citet{Bach:2012b}) also 
 a form of subgradient descent, our framework exposes all three algorithms 
-as special cases. A full table of the relationships for various choices of 
+as special cases. A full table outlining the relationships for various choices of 
 $h$ and $R$ is given in Table~\ref{tab:connections}.
@@
 
-In summary, we introduce a novel gradient descent algorithm to investigate herding and its generalizations:
+In summary, we introduce a novel gradient descent algorithm to investigate herding and its generalizations, and make the following contributions:
 \begin{itemize}
@@
 We show how \primal and \dual relate to \cgd and mirror descent (\md).
-\item We then use \dual in Section \ref{sec:herding} to define a class of generalized herding algorithms.
+\item We then use \dual in Section \ref{sec:herding} to define two classes of generalized herding algorithms.
 Connections with the ``learning'' interpretation of herding are made by showing how to obtain herding as the limit of an entropically normalized case of \dual.
@@
 Both of these results apply to the infinite dimensional case.
-\item In Section \ref{sec:lower-bounds} we prove matching lower bounds to the upper bounds given in Section \ref{sec:proofs}, showing that any ``sparse'' herding-like method cannot converge faster than $O(1/T)$ in the the infinite dimensional case and that herding converges at the slower $O(\log T / T)$ rate.
+\item In Section \ref{sec:lower-bounds} we prove a matching lower bound to the upper bound given in Section \ref{sec:proofs}, showing that any ``sparse'' herding-like method cannot converge faster than $O(1/T)$ in the the infinite dimensional case.% and that herding converges at the slower $O(\log T / T)$ rate.
 \item Finally, in section \ref{sec:chen} we generalize the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \bmd. 

### 2013-10-04 icml/lower-bounds.tex
more small fixes; tried to improve l.b. section

@@
 \begin{document}
-\section{Lower Bounds on Convergence Rates}
+\section{A General Lower Bound on Convergence Rates}
 \label{sec:lower-bounds}
 
-It has been an open question whether it is possible to achieve the fast $O(1/T^{2})$ convergence of finite-dimensional herding in the infinite-dimensional case.
-We now show that this is not in general possible.
-In particular, the upper bounds in the previous section for herding and herding-like algorithms have matching lower bounds for at least some combinations of distributions and feature functions.
-
-For the matching $O(1/T)$ lower bound we analyze sparse update methods.
-Formally, define a {\em sparse update method} to be any method that produces an estimator of $\bar \phi$ at iteration $t$ of the form
+Although it is possible to achieve the fast $O(1/T^{2})$ convergence rate for finite-dimensional herding, \citet{Bach:2012a} note that without further assumptions herding has only a $O(1/T)$ convergence rate in the infinite-dimensional case.
+We now provide an example showing rate is tight for {\em any} herding-like algorithm.
+By ``herding-like,'' we mean algorithms that approximate the distribution with at most one additional sample at each step.
+We formalize this idea by defining a {\em sparse update method} to be any method that produces an estimator of $\bar \phi$ at iteration $t$ of the form
 \[
@@
 \]
-Sparse update methods can be viewed as characterizing ``herding-like'' algorithms, including the generalized herding and \cgd with line search.
+Sparse update methods include $\alpha_{t}$-weighted herding and \cgd with line search.
 These sparse methods -- which only operate on the extreme points of the simplex -- are of interest because they provide (weighted) ``samples'' matching the moments of the distribution of interest.
 
-The bound setup is as follows.
+Our example is as follows.
 Let $\sX$ be the set of vectors $(x_{1},x_{2},\dots) \in [0,1]^{\infty}$ such that $\sum_{n=1}^{\infty} x_{n} = 1$ and a finite number of the $x_{n}$ are non-zero.

### 2013-10-04 icml/paper.tex
first draft of introduction
intro improved; log(t)/t writeup added
added bib filea
split of bmd section. added shortcuts for algorithm names
draft of connections subsection
added smc and mcmc to bibliography
more small fixes; tried to improve l.b. section

@@
 \usepackage{hyperref, icml2014}
+\usepackage{tabularx}
 \input{latex-defs.tex}
@@
 % the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{boosting, optimization, mirror descent, frank-wolfe, herding}
+\icmlkeywords{boosting, optimization, mirror descent, frank-wolfe, conjugate gradient descent, herding}
 
@@
 \begin{abstract} 
-TODO
+We introduce a unifying framework for understanding herding and 
+related algorithms. The framework is based on pair of optimization 
+algorithms we call primal and dual boosted mirror descent.
+We define two classes of generalized herding algorithms, 
+including one for infinite dimensional spaces
+that may apply even when herding does not. A number 
+of convergence bounds that apply to herding and generalized
+herding are obtained. Along the way we provide
+a more cohesive picture of the connections between herding, 
+conditional gradient descent and related methods, and maximum entropy. 
 \end{abstract}
@@
 \subfile{introduction.tex}
+\subfile{boosted-mirror-descent.tex}
 \subfile{herding-connection.tex}
 \subfile{convergence-proofs.tex}
-%\subfile{lower-bounds.tex}
+\subfile{lower-bounds.tex}
 %\subfile{chen-proofs.tex}
-%\subfile{conclusion.tex}
+\subfile{conclusion.tex}
 
-%\bibliography{example_paper}
-%\bibliographystyle{icml2013}
+\bibliography{herding.bib}
+\bibliographystyle{icml2014}
 
 \end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was
-% created by Lise Getoor and Tobias Scheffer, it was slightly modified  
-% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz, 
-% slightly modified from the 2009 version by Kiri Wagstaff and 
-% Sam Roweis's 2008 version, which is slightly modified from 
-% Prasad Tadepalli's 2007 version which is a lightly 
-% changed version of the previous year's version by Andrew Moore, 
-% which was in turn edited from those of Kristian Kersting and 
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2013-10-04 icml/herding-connection.tex
lots of minor fixes

@@
 
-The above updates correspond to \dual for $U = \sX$, $\Theta = \sH$, 
-$\alpha_t = 1$, $h \equiv 0$, 
-$R(\theta) = \ip{\bar{\phi}}{\theta} + \frac{1}{2}\|\theta\|_{\sH}^2$, 
-and $\imath(u)$ the canonical extension of $\phi$ from $\sX$ to $\Delta_{\sX}$.
+The above updates correspond to \dual for $U = \Delta_{\sX}$, $\Theta = \sH$, 
+$\ip{u}{\theta} = \ip{\bE_u[\phi(x)]}{\theta}_{\sH}$, $\alpha_t = 1$, $h \equiv 0$, and
+$R(\theta) = \ip{\bar{\phi}}{\theta}_{\sH} + \frac{1}{2}\|\theta\|_{\sH}^2$.
 To see this, note that 
 \begin{align}
-\lefteqn{\arg\min_{\theta \in \sH} \ip{\imath(u)}{\theta} - R(\theta)} \\
- &= \arg\min_{\theta \in \sH} \ip{\bE_{u}[\phi(x)]-\bar{\phi}}{\theta} - \frac{1}{2}\|\theta\|_{\sH}^2 \\
+\lefteqn{\arg\min_{\theta \in \sH} \ip{u}{\theta} - R(\theta)} \\
+ &= \arg\min_{\theta \in \sH} \ip{\bE_{u}[\phi(x)]-\bar{\phi}}{\theta}_{\sH} - \frac{1}{2}\|\theta\|_{\sH}^2 \\
  &= \bE_u[\phi(x)] - \bar{\phi}.
@@
 If we do not restrict $\alpha_{t}$ to be 1, we obtain a class of algorithms
-we call {\em $\alpha_{t}$-weighted herding}. For the $t$-weighted herding 
-algorithm with we obtain a stronger bound of $\frac{8r^2}{T}$ instead.
+we call {\em $\alpha_{t}$-weighted herding}. For the $\alpha_t = t$ we obtain 
+a stronger bound of $\frac{8r^2}{T}$.
 
-\subsection{Herding in Infinite Dimensions}
+\subsection{$q$-herding}
 \label{sec:infinite-case}

### 2013-10-04 icml/boosted-mirror-descent.tex

@@
 
-Our results primarily rely on a pair of novel convex optimization algorithms we call primal and dual boosted mirror descent.
+Our results primarily rely on a pair of novel convex optimization algorithms we call primal and dual boosted mirror descent. They are inspired by the decomposition used in \cite{Bach:2012b}.
 
@@
 
-Before introducing the boosted mirror descent (\bmd) algorithms, we recall some basic definitions related convex optimization.
+Before introducing the boosted mirror descent (\bmd) algorithms, we recall some basic definitions related to convex optimization.
 Let $f$ be a real function whose domain is a Banach space.
@@
 \]
-If $f$ is lower semi-continuous, then by the Fenchel-Moreau theorem $f = f^{**}$, i.e.
+If $f$ is lower semi-continuous and convex, then by the 
+Fenchel-Moreau theorem $f = f^{**}$, i.e.
 \[
@@
 Throughout the paper, we will assume without further notice that functions are lower semi-continuous when required.
-The gradients of $f^{*}$ and $f$ can be defined as any argument that maximizes the previous two equations:
+The (sub)gradients of $f^{*}$ and $f$ can be defined as any argument that maximizes the previous two equations:
 \[
@@
 
-In order to prove fast convergence, we will require additional assumptions beyond convexity. We say a function $f$ is {\em $\alpha$-strongly convex} if 
+In order to prove fast convergence, we will require additional assumptions beyond convexity. 
+Define the \emph{Bregman divergence} of a convex function to be
+\[ D_f(x \| y) = f(x)-\ip{\partial f(y), y-x} - f(y), \]
+which is a measure of ``how convex'' $f$ is between $x$ and $y$. 
+We say that a function $f$ is \emph{strongly convex} with respect 
+to a norm $\|\cdot\|$ if
 \[
-f(x)  - f(y) \le \ip{\partial f(x)}{x-y} - \frac{\alpha}{2}\|x - y\|^{2}.
+D_f(x \| y) \geq \frac{1}{2}\|x-y\|^2
 \]
-%and that a (possibly non-convex) function $g$ is $\beta$-smooth if 
-%\[
-%\|\partial g(x) - \partial g(y) \| \le \beta \| x - y\|.
-%\]
-%Strong convexity and smoothness are dual to each in the sense that a convex function $f$ is $\alpha$-strongly convex if and only if $f^{*}$ is $(1/\alpha)$-smooth.
-\NA{JH: should double check that we actually need smoothness}
+A related notion is \emph{strong smoothness}: a convex function $f$ 
+is \emph{strongly smooth} with respect to $\|\cdot\|$ if
+\[ D_f(x \| y) \leq \frac{1}{2}\|x-y\|^2 \]
+An important result is that $f$ is strongly convex with respect to $\|\cdot\|$ 
+if and only if its Fenchel conjugate $f^*$ is strongly smooth with respect 
+to the dual norm $\|\cdot\|_*$.
 
@@
 
-We can now define the \bmd  algorithms.
-Let $U$ and $\Theta$ be convex subsets of the Banach spaces $\sB_{U}$ and $\sB_{\Theta}$, respectively, and let $\imath : U \to \Theta$ be a linear imbedding from $U$ to $\Theta$ that preserves the inner product.
+We can now define the \bmd  algorithms. Let $U$ and $\Theta$ be convex subsets 
+of the Banach spaces $\sB_{U}$ and $\sB_{\Theta}$, respectively, and let 
+$\ip{\cdot}{\cdot}$ be a bilinear map on $\sB_U \times \sB_{\Theta}$; typically 
+$\ip{\cdot}{\cdot}$ will correspond to an inner product but we will use the 
+bracket notation for convenience even when this is not the case.
 Consider the two-argument loss function $L : U \times \Theta \to \bR$ 
+defined as
 \[
-L(u,\theta) = h(u) + \ip{\theta}{\imath(u)} - R(\theta)
+L(u,\theta) = h(u) + \ip{u}{\theta} - R(\theta)
 \]
@@
 L(u)
-&= h(u) + \max_{\theta \in \Theta} \left\{\ip{\theta}{\imath(u)} - R(\theta) \right\} \\
-&= h(u) + R^{*}(\imath(u)).
+&= h(u) + \max_{\theta \in \Theta} \left\{\ip{u}{\theta} - R(\theta) \right\} \\
+&= h(u) + R^{*}(u).
 \]
 We can think of $h$ as a \emph{primal regularizer} and $R$ as a \emph{dual regularizer}.
-The pair of regularization functions $h$ and $R$ define a flexible class of functions of optimize, as seen from the following examples:
+The pair of regularization functions $h$ and $R$ define a flexible class of functions to optimize, as seen from the following examples:
 \begin{enumerate}
@@
       of statistics, and let $\Theta$ be the unit $L^1$ ball in $\bR^d$. Let $h(u) = 0$, 
-      $\langle \theta, \imath(u) \rangle = \bE_u[\theta^T\phi(x)]$, and 
-      $R(\theta) = \langle \theta, \imath(u_0) \rangle$. Then $L(u)$ is the 
+      $\langle u, \theta \rangle = \bE_u[\theta^T\phi(x)]$, and 
+      $R(\theta) = \langle u_0, \theta \rangle$. Then $L(u)$ is the 
       \emph{maximum mean discrepancy} between $u$ and $u_0$ relative to $\phi$.
@@
       as long as $i^{-1}(\Theta) = \sB_{U}$.
-\item Let $U = \Theta = \bR^d$ and $\imath$ the identity map. Let $h(u) = \|u\|_1$, $R(u) = \langle \theta, u_0 \rangle + \frac{1}{2} \|\theta\|_2^2$. 
+\item Let $U = \Theta = \bR^d$. Let $h(u) = \|u\|_1$, $R(u) = \langle u_0, \theta \rangle + \frac{1}{2} \|\theta\|_2^2$. 
       Then $L(u) = \|u\|_1 + \frac{1}{2} \|u-u_0\|_2^2$. 
@@
 
-Going forward, for convenience we will abuse notation and use $\ip{\theta}{u}$ in place of 
-$\ip{\theta}{\imath(u)}$ when the map $\imath$ is clear from context. We will assume 
+We will assume 
 throughout that $h$ and $R$ are both convex functions, and furthermore that 
-$\arg\max_{u} h(u) + \ip{\theta}{u}$ can be efficiently 
+$\arg\max_{u} h(u) + \ip{u}{\theta}$ can be efficiently 
 computed for all values of $\theta$.
@@
 
-Mirror descent, together with boosting, yields an algorithm for finding 
+Mirror descent, combined with boosting, yields an algorithm for finding 
 ``saddle points'' of $L(u, \theta)$. 
@@
 As suggested by their names, the \bmd algorithms are closely related: they are dual to each other in the sense that 
-performing \primal on $L(u,\theta)$ is the same as performing \dual on $-L(\theta,u)$. This relationship is obvious by inspection of the update equations:
+performing \primal on $L(u,\theta)$ is the same as performing \dual on $L'(\theta, u) \eqdef -L(u,\theta)$. This relationship is obvious by inspection of the update equations:
 \(
@@
 This duality is analogous to that between mirror descent and conjugate gradient descent \citep{Bach:2012b}.
-\md minimizes $g(x) + f(x)$ and requires that $g$ be $\alpha$-strongly convex while \cgd minimizes $g^{*}(x) + f^{*}(x)$ and requires that $g$ be $\alpha$-strongly convex.
+\md minimizes $g(x) + f(x)$ and requires that $g$ be $\alpha$-strongly convex while \cgd minimizes $f^{*}(x) + g^{*}(x)$ and requires that $g$ be $\alpha$-strongly convex.
 The update equations for \md are

### 2013-10-04 icml/introduction.tex

@@
 distribution by samples. Most such algorithms require access to a full posterior and 
-may be prone to becoming stuck local modes. In contrast, the herding algorithm introduced in 
+may be prone to becoming stuck in local modes. In contrast, the herding algorithm introduced in 
 \citet{Welling:2009a} generates samples whose moments are guaranteed to converge 
@@
 Since its introduction, herding has been shown to have many desirable properties.
-For instance, Chen et al showed that the moments converge at a rate of $O(1/T^{2})$ for 
-$T$ samples, which is even faster than the $O(1/T)$ that would be obtained 
+For instance, \citet{Chen:2010a} showed that the moments converge at a rate of $O(1/T^{2})$ for 
+$T$ samples, which is even faster than the $O(1/T)$ rate that would be obtained 
 by sampling independently from the exact posterior. Furthermore, \citet{Bach:2012a} show that herding is 
@@
 for inference in Markov random fields, and this line of work has been further explored in 
-\citet{Welling:2009a}, \citet{Gelfand:2010} and \citet{Bornn:2013}.
+\citet{Welling:2009a}, \citet{Gelfand:2010}, and \citet{Bornn:2013}.
 
@@
 \begin{itemize}
-\item It allows us to cast herding in the framework of mirror descent, 
+\item It allows us to cast herding in the framework of mirror descent \cite{Beck:2003}, 
       which lets us better understand its convergence properties.
 \item It allows us to obtain generalizations of herding for infinite 
-      dimensional spaces that may apply even if standard herding does not.
+      dimensional spaces that apply in cases where standard herding does not.
 \item It makes the known connections between herding, subgradient, and 
@@
 $h$ and $R$ is given in Table~\ref{tab:connections}.
-% -- Doesn't seem relevant enough to justify the amount of technical detail
-%An alternative interpretation of herding \citep{Chen:2010a,Huszar:2012,Bach:2012a} is as a moment matching algorithm.
-In particular, it generates a set of pseudosamples $\{ x_{t} \}$  which match the moments $\bar\phi$.
-That is, the squared error $\sE_{T}^{2} \eqdef \| \bar\phi - T^{-1}\sum_{t=1}^{T} \phi(x_{t})\|^{2} \to 0$ as $T \to \infty$.
-If $\sB_{\Theta}$ is a reproducing kernel Hilbert space (RKHS) with feature map $\phi$, $\sE_{T}^{2}$ is equivalent to the maximum mean discrepancy (MMD) between $p$ and the empirical distribution $\hat p_{T} \eqdef T^{-1}\sum_{t=1}^{T} \delta_{x_{t}}$ \citep{Huszar:2012}:
-%\(
-%\sE_{T}^{2} &= \MMD^{2}(p, \hat p_{T})  \\
-%&\eqdef \sup_{\theta \in \Theta} \left|\bE_{p}\ip{\theta}{\phi(x)} - \bE_{\hat p_{T}}\ip{\theta}{\phi(x)} \right|,
-%\)
-%where $\Theta = \{ \theta \in \sB_{\Theta} \ | \  \|\theta\| = 1 \}$.
-
-%If we are interested in generating samples from $p$ (or some distribution with the same moments as $p$), herding provides an attractive alternative to standard i.i.d.
-sampling for two reasons.
-First, it is deterministic, so we do not have to worry about getting ``unlucky'' and generating a set of samples that poorly represent  $p$.
-Second, in the case where $\sB_{\theta}$ is finite dimensional, the squared error of the herding estimator converges to zero at a rate of $O(1/T^{2})$ whereas the convergence rate of i.i.d.~sampling is $O(1/T)$ \citep{Chen:2010a}.
-The infinite dimensional case, however, is not as well understood.
-\citet{Bach:2012a} recently showed that the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} does not apply in the infinite dimensional case.
-In addition, \citet{Bach:2012a} introduced an interpretation of herding as a special case of conditional gradient descent (\cgd) minimizing the square norm of an RKHS.
-They showed that \cgd with line search leads to faster convergence rates: $O(1/T)$ in the infinite dimensional case and exponential in the finite dimensional case.
-However, they found  empirically that line search had poor approximation properties since led to distributions far from the maximum entropy distribution.
 
@@
 \item In Section \ref{sec:algorithm} we introduce a pair of convex optimization algorithms, which we call primal and dual boosted mirror descent (\primal and \dual).
-We show how \primal and \dual relate to \cgd and mirror descent (\md).
+We show how \primal and \dual relate to conditional gradient (\cgd) and mirror descent (\md).
 \item We then use \dual in Section \ref{sec:herding} to define two classes of generalized herding algorithms.

### 2013-10-04 icml/lower-bounds.tex

@@
 Although it is possible to achieve the fast $O(1/T^{2})$ convergence rate for finite-dimensional herding, \citet{Bach:2012a} note that without further assumptions herding has only a $O(1/T)$ convergence rate in the infinite-dimensional case.
-We now provide an example showing rate is tight for {\em any} herding-like algorithm.
+We now provide an example showing this bound is tight for {\em any} herding-like algorithm.
 By ``herding-like,'' we mean algorithms that approximate the distribution with at most one additional sample at each step.
@@
 \[
-\theta_{t} = \sum_{s =1}^{t}w_{t} \phi(x_{t}).
+\theta_{t} = \sum_{s =1}^{t}w_{t} \phi(x_{t}),
 \]
-Sparse update methods include $\alpha_{t}$-weighted herding and \cgd with line search.
+where $w_{t} \in \bR$.
+Sparse update methods include herding, $\alpha_{t}$-weighted herding, \cgd, and  \cgd with line search.
 These sparse methods -- which only operate on the extreme points of the simplex -- are of interest because they provide (weighted) ``samples'' matching the moments of the distribution of interest.
 
-Our example is as follows.
-Let $\sX$ be the set of vectors $(x_{1},x_{2},\dots) \in [0,1]^{\infty}$ such that $\sum_{n=1}^{\infty} x_{n} = 1$ and a finite number of the $x_{n}$ are non-zero.
-Define the infinite feature vector $\phi$ to be the imbedding of $\sX$ into $\ell^{1}$.
-Let the target distribution $p(x)$ be chosen such that $\bar\phi_{n} = c/\alpha(n)$, where $\alpha(n)$ is any superlinear function and $c^{-1} = \sum_{n=1}^{\infty} \alpha(n) < \infty$.
+Our example is constructed as follows.
+Let the sample space be $\sX = \{ 1, 2, 3, \dots\}$.
+Define the infinite feature vector $\phi$ to be $\phi_{n}(x) = \bI(x = n)$, so $\sM$ is the unit $\ell^{1}$ ball.
+For target distribution $p \in \Delta_{\sX}$, we therefore have  that the mean vector is the vector of probabilities: $\bar\phi_{n} = p(n)$.
+Let $p(x) = c/\alpha(n)$, where $\alpha(n)$ is any superlinear function (i.e.~$\alpha \in \omega(n)$) and $c^{-1} = \sum_{n=1}^{\infty} \alpha(n) < \infty$.
 Without loss of generality we can assume that $\alpha$ is monotonically increasing.
@@
 \[
-\|x-x^{*}\|_2^2 \ge c^{2}\sum_{n > t} 1/\alpha(n)^{2} = \Theta(1/\beta(t)) \label{eq:convergence-lb}
+\|\theta_{t}- \bar \phi\|_2^2 \ge c^{2}\sum_{n > t} 1/\alpha(n)^{2} = \Theta(1/\beta(t)) \label{eq:convergence-lb}
 \]
@@
 Then we can choose $\alpha = \sqrt \gamma \in \omega(t)$ to satisfy \eqref{eq:convergence-lb}.
-
 Therefore, any sparse update method must have convergence rate no faster than $\Theta(1/\beta(t))$ for all superlinear $\beta$, and hence no faster than $\Theta(1/t)$.

### 2013-10-04 icml/supplementary-material.tex
draft of connections subsection

@@
 \dual minimizes $g + R^{*}$ while \cgd minimizes $h^{*} + f^{*}$.
-Make the identifications $h^{*} = R^{*}$ and $f^{*} = g$.
-Note that $R^{*}$ is $\beta$-smooth $\iff R$ is $(1/\beta)$-strongly convex $\iff h$ is $(1/\beta)$-strongly convex, so the conditions on the functions being optimized are equivalent.
-Then with $\rho_{t} = \alpha_{t+1}/A_{t+1}$, we can rewrite the \cgd updates  s
+Make the identifications $h^{*} = R^{*}$ and $f^{*} = g$ and note the conditions on the functions being optimized are equivalent.
+Then with $\rho_{t} = \alpha_{t+1}/A_{t+1}$, we can rewrite the \cgd updates as
 \[
@@
 y_{t} &= \partial R^{*}(x_{t}) \\
-x_{t+1} &= \partial g^{*}((1 - \rho_{t+1})\partial g(x_{t}) - \rho_{t+1}y_{t}) \\
-&= \partial g^{*}(-\bar y_{t}), 
+\bar y_{t} &= \partial g^{*}((1 - \rho_{t+1})\partial g(x_{t}) - \rho_{t+1}y_{t}) \\
+x_{t+1} &= \partial g^{*}(-\bar y_{t}), 
 \]
@@
 
+Finally, note that if we write $\alpha_{t} = \rho_{t}/\prod_{s=1}^{t} (1 - \rho_{s})$, we have $A_{t} = \prod_{s=1}^{t} (1 - \rho_{s})$ and hence $\alpha_{t} / A_{t} = \rho_{t}$.
+Hence, up to a change in index, we have a bijection between the $\rho_{t}$ and the $\alpha_{t}$.
+
+\section{Proof Lemma used in $q$-Herding Convergence Proof}
+
+Two simple equalities were used in the proof of the convergence rate of $q$-herding.
+We now provide proofs.
+Using the fact that $x^{n} - y^{n} = (x - y)\sum_{i=0}^{n-1}x^{i-1}y^{n-i-1}$, for the first inequality we have
+\(
+x^{p} - py^{p-1}x + (p-1)y^{p}
+&= x^{p} - y^{p} - py^{p-1}x + py^{p} \\
+&= (x - y)\sum_{r=0}^{p-2} (y^{r}x^{p-r-1} - y^{p-1}) \\
+&= (x - y)^{2}\sum_{r=0}^{p-2}\sum_{s=0}^{p-r-2}x^{p-r-s-2}y^{r+s} \\
+&\le (x - y)^{2}\sum_{r=0}^{p-2}\sum_{s=0}^{p-r-2}\max(x, y)^{p-2} \\
+&= {p \choose 2}(x - y)^{2}\max(x, y)^{p-2}.
+\)
+For the second inequality we have
+\(
+(y-z)^{2}\max(x,y)^{p-2} 
+&\le 4\max(y,z)^{2}\max(x,y)^{p-2} \\
+&\le 4\max(x,y,z)^{p}.
+\)
+
 \end{document}

### 2013-10-04 icml/chen-proofs.tex
added chen proofs

\documentclass[paper.tex]{subfiles}

\begin{document}
\section{An Upper Bound in Infinite Dimensions}
\label{sec:chen-proofs}

An intriguing aspect of classical herding is that it is known to have a convergence rate of 
$O(1/T^2)$ under suitable conditions, as was shown in \citet{Chen:2010b}. In this section, 
we will need to assume that $h(x)$ is the indicator function for $U$ (i.e. $0$ on $U$ and 
$\infty$ outside of it) and that $R(\theta) = \ip{u_0}{\theta} + R_0(\theta)$, where 
$R_0$ exhibits \emph{quadratic scaling}: $R_0(s\theta) = s^2R_0(\theta)$. Note that 
this implies that $R_0^*(u-u_0) = R^*(u)$.
For simplicity 
we will focus on the case $\alpha_t = 1$. Then we have 
the following bound:
\begin{theorem}
TODO
\end{theorem}
\begin{proof}
The key quantity we will analyze is $(t-1)^2R(\theta_t)$. Note that we have
\begin{align*}
\lefteqn{(t-1)^2R(\theta_t) - t^2R(\theta_{t+1})} \\
 &= (t-1)^2R^*(\hat{u}_{t-1})-t^2R^*(\hat{u}_t) \\
 &= (t-1)^2R_0^*(\hat{u}_{t-1})-t^2R_0^*(\hat{u}_t-z) \\
 &= (t-1)^2\left[R^*(\hat{u}_{t-1})-R_0^*((t/(t-1))(\hat{u}_t-z))\right] \\
 &= (t-1)^2\left[R^*(\hat{u}_{t-1})-R^*\left(\frac{t\hat{u}_t-z}{t-1}\right)\right] \\
 &= (t-1)^2\left[\ip{\frac{\partial R^*}{\partial \hat{u}_{t-1}}}{\hat{u}_{t-1}-\frac{t\hat{u}_t-z}{t-1}}-D_{R^*}\left(\frac{t\hat{u}_t-z}{t-1} \middle\| \hat{u}_{t-1}\right)\right] \\
 &\geq (t-1)^2\left[\ip{\theta_t}{-\frac{1}{t-1}(u_t-z)}-\frac{1}{2}\left\|\frac{u_t-z}{t-1}\right\|_*^2\right] \\
 &= -(t-1)\ip{\theta_t}{u_t-z}-\frac{1}{2}\|u_t-z\|_*^2 \\
 &= -(t-1)\left[\inf_{u \in U} \ip{\theta_t}{u} - \ip{\theta_t}{z}\right]-\frac{1}{2}\|u_t-z\|_*^2 \\
 &\geq -(t-1)\left[\sup_{\theta} \inf_{u \in U} \ip{\theta}{u-z}\right]-2r_*^2.
\end{align*}

\end{proof}

\end{document}

### 2013-10-04 icml/conclusion.tex
first pass at a conclusion

@@
 
-\NA{Write me!}
+In this paper we have introduced the primal and dual boosted mirror descent algorithms in order to investigate the properties of herding and a number of novel classes of generalized herding algorithms.
+$\alpha_{t}$-weighted herding is very similar to standard herding, but the pseudosamples are not required to have equal weights.
+We also defined $q$-herding.
+The $q$-herding algorithm minimizes the distance between the sample estimator and the target moments as measured by the $\ell^{p}$ norm (with $p$ the conjugate of $q$, $p \ge 2$).
+For $p = 2$, we get herding.
+But in infinite dimensional space, $q$-herding with $p > 2$ may some times be applied when standard herding is inapplicable.
+
+Using \bmd we also place the convergence results of \citet{Chen:2010} and \citet{Bach:2010} into a single framework.
+Furthermore, we were able to extend those convergence results to certain classes of herding-like algorithms including $\alpha_{t}$-weighted herding, $q$-herding, and \NA{chen} herding.
+
+Our results thus extend the applicability and flexibility of deterministic algorithms for generating pseudosamples.
+While not universally applicable, these deterministic algorithms offer an increasingly attractive alternative to random sampling for certain types of problems and model classes because the can offer faster convergence rates, convergence guarantees without ``with high probability'' conditions, and simple implementations.
 

### 2013-10-04 icml/paper.tex
saving
first pass at a conclusion

@@
 
-\icmltitlerunning{Boosted Mirror Descent}
+\icmltitlerunning{Herding and Boosted Mirror Descent}
 
@@
 \twocolumn[
-\icmltitle{Boosted Mirror Descent, Frank-Wolfe, and Herding}
+\icmltitle{Herding and its Generalizations via Boosted Mirror Descent}
 
@@
 related algorithms. The framework is based on pair of optimization 
-algorithms we call primal and dual boosted mirror descent.
+algorithms that we call primal and dual boosted mirror descent.
 We define two classes of generalized herding algorithms, 
@@
 
-\bibliography{herding.bib}
+\bibliography{herding}
 \bibliographystyle{icml2014}

### 2013-10-04 icml/chen-proofs.tex
done with chen proof

@@
 \begin{theorem}
-TODO
+\label{thm:chen}
+Suppose that $R$ is strongly convex with respect to the norm $\|\cdot\|$ and 
+let $r_* \eqdef \sup_{u \in U} \|u\|_*$. Also suppose that $h \equiv 0$ and 
+that $R_0$ exhibits quadratic scaling as above. Define 
+\[ \gamma^* \eqdef \inf_{\theta \in \Theta} \sup_{u \in U} \frac{\ip{-\theta}{u-u_0}}{\sqrt{R(\theta)}}. \]
+Then $R(\theta_{t+1}) \leq \frac{2r_*^2+(2r_*^2/\gamma_*)^2}{t^2}$ as long 
+as $\gamma^* > 0$.
 \end{theorem}
-\begin{proof}
+This corresponds to Proposition 1 of \citet{Chen:2010b} for $R(\theta) = \frac{1}{2}\|\theta\|_2^2$. Note that if there is an $l^2$ ball of radius $\epsilon$ 
+around $u_0$ that is contained in $u$ then $\gamma^* \geq \frac{\epsilon}{r_*}$ 
+in this case, which is how Chen's result follows from ours.
+\begin{proof}[Proof of Theorem~\ref{thm:chen}]
 The key quantity we will analyze is $(t-1)^2R(\theta_t)$. Note that we have
@@
  &= -(t-1)\left[\inf_{u \in U} \ip{\theta_t}{u} - \ip{\theta_t}{z}\right]-\frac{1}{2}\|u_t-z\|_*^2 \\
- &\geq -(t-1)\left[\sup_{\theta} \inf_{u \in U} \ip{\theta}{u-z}\right]-2r_*^2.
+ &\geq (t-1)\sqrt{R(\theta_t)}\left[\inf_{\theta} \sup_{u \in U} \ip{-\theta}{u-u_0}\right]-2r_*^2 \\
+ &= \sqrt{(t-1)^2R(\theta_t)}\gamma^*-2r_*^2.
 \end{align*}
-
+There are now two possibilities for each $t$:
+\begin{enumerate}
+\item $t^2R(\theta_{t+1}) < (t-1)^2R(\theta_t)$, or
+\item $(t-1)^2R(\theta_t) \leq \frac{4r_*^4}{(\gamma^*)^2}$.
+\end{enumerate}
+In particular, consider the value of $t$ for which 
+$t^2R(\theta_{t+1})$ is maximized. Then by assumption 
+case $1$ cannot hold and so $(t-1)^2R(\theta_t) \leq \frac{4r_*^4}{(\gamma^*)^2}$.
+But then by applying the above sequence of inequalities 
+we also have 
+\begin{align*}
+\lefteqn{t^2R(\theta_{t+1}) - (t-1)^2R(\theta_t)} \\
+ &\leq 2r_*^2 - \sqrt{(t-1)^2R(\theta_t)}\gamma^* \\
+ &\leq 2r_*^2,
+\end{align*}
+hence $t^2R(\theta_{t+1}) \leq 2r_*^2 + \frac{4r_*^4}{(\gamma_*)^2}$. 
+Since $t^2R(\theta_{t+1})$ was maximal by assumption, we have
+$R(\theta_{t+1}) \leq \frac{2r_*^2+(2r_*/\gamma_*)^2}{t^2}$, which 
+completes the proof.
 \end{proof}

### 2013-10-04 icml/introduction.tex
put some proof stuff into supp mat

@@
 \item In Section \ref{sec:lower-bounds} we prove a matching lower bound to the upper bound given in Section \ref{sec:proofs}, showing that any ``sparse'' herding-like method cannot converge faster than $O(1/T)$ in the the infinite dimensional case.% and that herding converges at the slower $O(\log T / T)$ rate.
-\item Finally, in section \ref{sec:chen} we generalize the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \bmd. 
+\item Finally, in section \ref{sec:chen-proofs} we generalize the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \bmd. 
 \end{itemize}

### 2013-10-04 icml/supplementary-material.tex
put some proof stuff into supp mat

@@
 
+\section{Proof of Key Lemma}
+
+We now prove the key lemma, which we restate here, used in the convergence proofs of the \bmd algorithms.  
+\begin{lemma}
+\label{lem:bregman}
+Let $z_1,\ldots,z_T$ be vectors and let $f(x)$ be a strictly convex 
+function. Define $\hat{z}_t$ to be $\frac{1}{A_t} \sum_{s=1}^t \alpha_s z_s$.
+
+Let $x_1,\ldots,x_T$ be chosen via $x_{t+1} = \arg\min_{x} f(x) + \langle \hat{z}_t, x\rangle$. 
+Then for any $x^*$ we have
+\begin{align*}
+\lefteqn{\frac{1}{A_T} \sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
+\phantom{+} &\leq f(x^*) + \langle \hat{z}_t, x^* \rangle + \frac{1}{A_T} \sum_{t=1}^T A_t D_{f}(x_t \| x_{t+1}). 
+\end{align*}
+\end{lemma}
+\begin{proof}
+First note that, if $x_0 = \arg\min f(x) + \langle z, x \rangle$, 
+then $\nabla f(x_0) = -z$.
+
+%\NA{JNS: In the following proof, (2) is by the update formula for $x_{t+1}$ together with the preceding observation.
+(3) is applying (2).
+(4) and (5) are re-arranging the sum.
+(6) is definition of Bregman divergence.
+(7) is the observation again.
+(8) is the definition of $x_{T+1}$.}
+
+Now note that
+\begin{align}
+\alpha_{t}z_{t} 
+&= A_{t}\hat z_{t} - A_{t-1}\hat z_{t-1} \\
+&= - A_{t}\nabla f(x_{t+1}) + A_{t-1} \nabla f(x_{t}),
+\end{align}
+so we have
+\begin{align}
+\lefteqn{\sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
+ &= \sum_{t=1}^T \{\alpha_t f(x_t) + \langle A_{t-1} \nabla f(x_t) - A_t \nabla f(x_{t+1}), x_t \rangle\} \\
+ &= \sum_{t=1}^T \{\alpha_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle\} \\
+ & \quad - A_{T}\langle \nabla f(x_{T+1}), x_{T+1} \rangle
+\end{align}
+\begin{align}
+ &= \sum_{t=1}^T \{A_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle - A_t f(x_{t+1})\}  \\
+ &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
+ &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\}  \\
+ &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
+ &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) + \langle \hat{z}_T, x_{T+1} \rangle) \\
+ &\leq \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x^*) + \langle \hat{z}_T, x^* \rangle). 
+\end{align}
+Dividing both sides by $A_T$ completes the proof.
+\end{proof}
+
 \end{document}

### 2013-10-04 icml/chen-proofs.tex
version without chen

@@
 An intriguing aspect of classical herding is that it is known to have a convergence rate of 
-$O(1/T^2)$ under suitable conditions, as was shown in \citet{Chen:2010b}. In this section, 
+$O(1/T^2)$ under suitable conditions, as was shown in \citet{Chen:2010a}. In this section, 
 we will need to assume that $h(x)$ is the indicator function for $U$ (i.e. $0$ on $U$ and 
@@
 \end{theorem}
-This corresponds to Proposition 1 of \citet{Chen:2010b} for $R(\theta) = \frac{1}{2}\|\theta\|_2^2$. Note that if there is an $l^2$ ball of radius $\epsilon$ 
+This corresponds to Proposition 1 of \citet{Chen:2010b} for $R(\theta) = \frac{1}{2}\|\theta\|_2^2$. Note that if there is an $\ell^2$ ball of radius $\epsilon$ 
 around $u_0$ that is contained in $u$ then $\gamma^* \geq \frac{\epsilon}{r_*}$ 

### 2013-10-04 icml/conclusion.tex
version without chen

--- 2013-10-04 icml/conclusion.tex
@@
 
-\NA{Write me!}
+In this paper we have introduced the primal and dual boosted mirror descent algorithms in order to investigate the properties of herding and a number of novel classes of generalized herding algorithms.
+In particular, $\alpha_{t}$-weighted herding is very similar to standard herding, but the pseudosamples are not required to have equal weights.
+We also defined the $q$-herding algorithm, which minimizes the distance between the sample estimator and the target moments as measured by the $\ell^{p}$ norm (with $p$ the conjugate of $q$, $p \ge 2$).
+For $p = 2$, we get herding.
+But in infinite dimensional space, $q$-herding with $p > 2$ may some times be applied when standard herding is inapplicable.
+We have also presented a number of convergence bounds for various classes of herding-like algorithms.
+We have also shown the limitations of herding-like algorithms in infinite dimensional spaces.
+
+%Using \bmd we also place the convergence results of \citet{Chen:2010a} and \citet{Bach:2012a} into a single framework.
+Furthermore, we were able to extend those convergence results to certain classes of herding-like algorithms including $\alpha_{t}$-weighted herding and $q$-herding%, and \NA{chen} herding.
+
+Our results thus extend the applicability and flexibility of deterministic algorithms for generating pseudosamples.
+While not universally applicable, these deterministic algorithms offer an increasingly attractive alternative to random sampling for certain types of problems and model classes because the can offer faster convergence rates, avoid local optima, come with convergence guarantees without the ``with high probability'' condition, and are often simple to implement.
 

### 2013-10-04 icml/conclusion.tex
changed conclusion

@@
 But in infinite dimensional space, $q$-herding with $p > 2$ may some times be applied when standard herding is inapplicable.
-We have also presented a number of convergence bounds for various classes of herding-like algorithms.
-We have also shown the limitations of herding-like algorithms in infinite dimensional spaces.
+We hope that our results will provide a useful consolidation of the many known phenomena related to herding, as well as
+form the basis for more new pseudo-sampling algorithms such as $q$-herding.
+
+%We have also presented a number of convergence bounds for various classes of herding-like algorithms. We have also shown the limitations of herding-like algorithms in infinite dimensional spaces.
 
@@
 
-Our results thus extend the applicability and flexibility of deterministic algorithms for generating pseudosamples.
+%Our results thus extend the applicability and flexibility of deterministic algorithms for generating pseudosamples.
 While not universally applicable, these deterministic algorithms offer an increasingly attractive alternative to random sampling for certain types of problems and model classes because the can offer faster convergence rates, avoid local optima, come with convergence guarantees without the ``with high probability'' condition, and are often simple to implement.

### 2013-10-04 icml/introduction.tex
version without chen
final fixes

--- 2013-10-04 icml/introduction.tex
@@
 Many algorithms, such as SMC \citep{smc} and MCMC \citep{mcmc}, approximate a 
-distribution by samples. Most such algorithms require access to a full posterior and 
+distribution with samples. Most such algorithms require access to a full posterior and 
 may be prone to becoming stuck in local modes. In contrast, the herding algorithm introduced in 
@@
 \]
-In this paper we generalize the herding algorithm to the following set of updates, which 
+In this paper we generalize the herding algorithm to minimize $h(h) + R^{*}(u)$ via the following set of updates, which 
 require only the convexity of $h$ and $R$:
@@
 \item In Section \ref{sec:lower-bounds} we prove a matching lower bound to the upper bound given in Section \ref{sec:proofs}, showing that any ``sparse'' herding-like method cannot converge faster than $O(1/T)$ in the the infinite dimensional case.% and that herding converges at the slower $O(\log T / T)$ rate.
-\item Finally, in section \ref{sec:chen} we generalize the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \bmd. 
+%\item Finally, in section \ref{sec:chen-proofs} we generalize the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \bmd. 
 \end{itemize}

### 2013-10-04 icml/paper.tex
final fixes

@@
 \begin{abstract} 
+Much has been learned about herding since it was first proposed.
 We introduce a unifying framework for understanding herding and 
@@
 a more cohesive picture of the connections between herding, 
-conditional gradient descent and related methods, and maximum entropy. 
+conditional gradient descent, and maximum entropy. 
 \end{abstract}

### 2013-10-25 nips/paper.tex
added nips workshop paper

\documentclass{article} % For LaTeX2e
\usepackage{import, subfiles}
\usepackage{tabularx}
\usepackage{nips13submit_e,times}
\input{latex-defs.tex}

\title{A Greedy Framework for First-Order Optimization}
\author{
Jacob Steinhardt\thanks{Both authors contributed equally to this work.} \\
Department of Computer Science\\
Stanford University\\
Stanford, CA 94305 \\
\texttt{jsteinhardt@cs.stanford.edu} \\
\And
Jonathan Huggins\footnote[1]{}\\
Department of EECS \\
Massachusetts Institute of Technology \\
Cambridge, MA 02139 \\
\texttt{jhuggins@mit.edu}
}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy

\begin{document}
\maketitle

Recent work has shown many connections between conditional gradient 
and other first-order optimization methods, such as herding \cite{Bach:2012a} 
and subgradient descent \cite{Bach:2012b}. By considering a type of 
\emph{proximal conditional method}, which we call boosted mirror 
descent (\bmd), we are able to unify all of these algorithms into 
a single framework, which can be interpreted as taking successive 
arg-mins of a sequence of surrogate functions. Using a standard 
online learning analysis based on Bregman divergences, we are 
able to demonstrate an $O(1/T)$ convergence rate for all algorithms 
in this class.

More concretely, suppose that we are given a function 
$L : U \times \Theta \to \bR$ defined by
\[ L(u, \theta) = h(u) + \ip{u}{\theta} - R(\theta) \]
and wish to find the \emph{saddle point} 
\[ L_* \eqdef \min_{u} \max_{\theta} L(u, \theta). \]
We should think of $u$ as the primal variables and $\theta$ 
as the dual variables; we will assume throughout that 
$h$ and $R$ are both convex. We will also abuse notation 
and define $L(u) \eqdef \max_{\theta} L(u, \theta)$; we 
can equivalently write $L(u)$ as 
\[ L(u) = h(u) + R^*(u), \]
where $R^*$ is the Fenchel conjugate of $R$. Note that 
$L(u)$ is a convex function. Moreover, since 
$R \leftrightarrow R^*$ is a one-to-one mapping, for 
\emph{any} convex function $L$ and \emph{any} decomposition 
of $L$ into convex functions $h$ and $R^*$, we get a 
corresponding two-argument function $L(u, \theta)$.

Given the function $L(u, \theta)$, we define the following 
optimization procedure, which will generate a sequence of 
points $(u_1,\theta_1), (u_2, \theta_2), \ldots$ converging 
to a saddle point of $L$. First, take a sequence of weights
$\alpha_1, \alpha_2, \ldots$, and for notational convenience 
define
\(
\hat{u}_t = \frac{\sum_{s=1}^t \alpha_su_s}{\sum_{s=1}^t \alpha_s} 
\quad \text{and} \quad 
\hat{\theta}_t = \frac{\sum_{s=1}^t \alpha_s\theta_s}{\sum_{s=1}^t \alpha_s}.
\)
Then 
the {\em primal boosted mirror descent} (\primal) algorithm is:
\begin{enumerate}
\item $u_1 \in \arg\min_u h(u)$
\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, u_t \rangle - R(\theta) = \partial R^{*}(u_{t})$
\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, u \rangle = \partial h^{*}(-\hat\theta_{t})$
\end{enumerate}
As long as $h$ is strongly convex, for the proper choice of $\alpha_{t}$ we 
obtain the bound (see Corollary~\ref{cor:method-1}):
\begin{equation}
\sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq L_{*} + O(1/T).
\end{equation}

As an example, suppose that we are given a $\gamma$-strongly convex function 
$f$: that is, $f(x) = \frac{\gamma}{2}\|x\|_2^2 + f_0(x)$, where $f_0$ is convex. 
Then we let $h(x) = \frac{\gamma}{2}\|x\|_2^2$, $R^*(x) = f_0(x)$, and obtain 
the updates:
\begin{enumerate}
\item $u_1 = 0$
\item $\theta_t = \partial f_0(u_t)$
\item $u_{t+1} = -\frac{1}{\gamma}\hat{\theta}_t = -\frac{\sum_{s=1}^t \alpha_s\partial f_0(u_s)}{\gamma \sum_{s=1}^t \alpha_s}$
\end{enumerate}
We therefore obtain a variant on subgradient descent where $u_{t+1}$ 
is a weighted average of the first $t$ subgradients (times a step size 
$1/\gamma$). Note that these are the subgradients of $f_0$, which are related 
to the subgradients of $f$ by $\partial f_0(x) = \partial f(x) - \gamma x$.

We can also concern the dual form of our mirror descent algorithm 
({\em dual boosted mirror descent}, or \dual):
\begin{enumerate}
\item $\theta_1 \in \arg\min_{\theta} R(\theta)$
\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle = \partial h^{*}(-\theta_{t})$
\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \hat{u}_t \rangle - R(\theta) = \partial R^{*}(\hat u_{t})$
\end{enumerate}
Convergence now hinges upon strong convexity of $R$ rather than 
$h$, and we obtain the same $1/T$ convergence 
rate (see Corollary~\ref{cor:method-2}):
\[ \sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq L_{*} + O(1/T). \]
An important special case is $h(u) = \left\{ \begin{array}{ccl} 0 & : & u \in K \\ \infty & : & u \not\in K \end{array} \right.$, where $K$ is some 
compact set. Also let $R^* = f$, where $f$ is an arbitrary strongly convex 
function. Then we are minimizing $f$ over the compact set $K$, and we obtain 
the updates from conditional gradient:
\begin{enumerate}
\item $\theta_1 = \partial f(0)$
\item $u_t \in \arg\min_{u \in K} \langle \theta_t, u \rangle$
\item $\theta_{t+1} = \partial f(\hat{u}_t)$
\end{enumerate}
Our notation is slightly different from previous presentations in 
that we use linear weights ($\alpha_t$) instead of geometric weights 
(often denoted $\rho_t$, as in \cite{Bach:2012b}). However, under the 
mapping $\alpha_t = \rho_t/\prod_{s=1}^t (1-\rho_s)$, we obtain an 
algorithm equivalent to the usual formulation.

Our framework is intriguing in that it involves a purely greedy minimization 
of surrogate loss functions (alternating between the primal and dual), and yet 
is powerful enough to capture both subgradient descent and conditional gradient 
descent, as well as a host of other first-order methods, including the low-rank 
SDP solver introduced by Arora, Hazan, and Kale \cite{aroraSDP}. Briefly, the 
AHK algorithm seeks to maximize 
$\sum_{j=1}^m \frac{1}{2}(\Tr(A_j^TX)-b_j)^2$ subject to the 
constraints $X \succeq 0$ and $\Tr(X) \leq \rho$.\footnote{This is actually a 
variant of their algorithm, which we present for ease of exposition.} We 
can then define 
\[ h(X) = \left\{ \begin{array}{ccl} 0 & : & X \succeq 0\mathrm{ \ and \ }\Tr(X) \leq \rho \\ \infty & : & \mathrm{else} \end{array} \right.\]
and
\[ R^*(X) = \sum_{j=1}^m \frac{1}{2}(\Tr(A_j^TX)-b_j)^2. \]
Note that this decomposition is actually a special case of the conditional 
gradient decomposition above, and so we obtain the updates 
\[ X_{t+1} \in \operatorname{argmin}_{X \succeq 0, Tr(X) \leq \rho} \sum_{j=1}^m \left[\Tr(A_j^T\hat{X}_t)-b_j\right]\Tr(A_j^TX), \]
whose solution turns out to be $\rho vv^T$, where $v$ is the top singular 
vector of the matrix
$-\sum_{j=1}^m \left[\Tr(A_j^T\hat{X}_t)-b_j\right]A_j$. This example serves both 
to illustrate the flexibility of our framework and to highlight the interesting 
fact that the Arora-Hazan-Kale SDP algorithm is a special case of conditional 
gradient (to get the original formulation in \cite{aroraSDP}, we need to replace 
the function $\frac{1}{2}x^2$ with $x_{+}\log x_{+}$, where $x_{+} = \max(x,0)$).

We end by stating our formal convergence results. For the primal algorithm 
(\primal) we have:
\begin{theorem}
\label{thm:method-1}
Suppose that $h$ is strongly convex with respect to a norm $\|\cdot\|$ 
and let $r = \sup_{\theta} \|\theta\|_{*}$. Then 
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
\end{theorem}
\begin{corollary} 
\label{cor:method-1}
Under the hypotheses of Theorem \ref{thm:method-1}, for $\alpha_{t} = 1$ we have
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2 (\log (T) + 1)}{T}. \]
and for $\alpha_t = t$ we have
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T}. \]
\end{corollary}
Similarly, for the dual algorithm (\dual) we have:
\begin{theorem}
\label{thm:method-2}
Suppose that $R$ is strongly convex with respect to a norm $\|\cdot\|$ 
and let $r = \sup_{u} \|u\|_{*}$. Then 
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
\end{theorem}
\begin{corollary}
\label{cor:method-2}
Under the hypotheses of Theorem \ref{thm:method-2}, for $\alpha_t = 1$ we have
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2(\log(T) + 1)}{T} \]
and for $\alpha_t = t$ we have
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T} \]
\end{corollary}
Thus, a step size of $\alpha_t = t$ yields the claimed $O(1/T)$ convergence rate.

\bibliography{herding}
\bibliographystyle{plain}

\end{document}

### 2013-10-25 nips/paper.tex
added q-herding

@@
 
+\paragraph{Introduction.} 
 Recent work has shown many connections between conditional gradient 
@@
 
+\paragraph{Setup.} 
 More concretely, suppose that we are given a function 
@@
 
+\paragraph{Primal algorithm.}
 Given the function $L(u, \theta)$, we define the following 
@@
 
+\paragraph{Dual algorithm.}
 We can also concern the dual form of our mirror descent algorithm 
@@
 
+\paragraph{Discussion.}
 Our framework is intriguing in that it involves a purely greedy minimization 
@@
 
+\paragraph{$q$-herding.}
+In addition to unifying several existing methods, our framework allows us to 
+extend herding to a an algorithm that we call \emph{$q$-herding}. Herding is 
+an algorithm for constructing pseudosamples that match a specified collection 
+of moments from a distribution; it was originally introduced by Welling 
+\cite{Welling:2009a} and was shown to be a special case of conditional gradient 
+by Bach et al. \cite{Bach:2012a}. It can be cast as trying to minimize 
+$\|\bE_{\mu}[\phi(x)]-\bar{\phi}\|_2^2$ for $\mu$ in the probability simplex over 
+$\sX$, for a given $\phi : \sX \to \bR^d$ and $\bar{\phi} \in \bR^d$. As shown in 
+\cite{Bach:2012a}, the herding updates are equivalent to \dual with 
+$h(\mu) \equiv 0$ and $R(\theta) = \theta^T\bar{\phi} + \frac{1}{2}\|\theta\|_2^2$.
+The implicit assumption in the herding algorithm is that $\|\phi(x)\|_2$ is 
+bounded. We are able to construct a more general algorithm that only requires 
+$\|\phi(x)\|_p$ to be bounded for some $p \geq 2$. This \emph{$q$-herding} 
+algorithm works by taking 
+$R(\theta) = \theta^T\bar{\phi} + \frac{1}{q}\|\theta\|_q^q$, where 
+$\frac{1}{p} + \frac{1}{q} = 1$. In this case our convergence results imply that 
+$\|\bE_{\mu}[\phi(x)]-\bar{\phi}\|_p^p$ decays at a rate of $O(1/T)$.
+
+\paragraph{Convergence results.}
 We end by stating our formal convergence results. For the primal algorithm 

### 2013-11-12 nips/paper.tex
added convergence proofs and further lit discussion to nips workshop paper

@@
 
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
+\newcommand{\fix}{\marginpar{\textcolor{NAColor}{FIX}}}
+\newcommand{\new}{\marginpar{\textcolor{NAColor}{NEW}}}
+\allowdisplaybreaks
 
@@
 \paragraph{Dual algorithm.}
-We can also concern the dual form of our mirror descent algorithm 
+We can also consider the dual form of our mirror descent algorithm 
 ({\em dual boosted mirror descent}, or \dual):
 \begin{enumerate}
-\item $\theta_1 \in \arg\min_{\theta} R(\theta)$
+\item $\theta_1 \in \arg\min_{\theta} R(\theta)$a
 \item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle = \partial h^{*}(-\theta_{t})$
@@
 is powerful enough to capture both subgradient descent and conditional gradient 
-descent, as well as a host of other first-order methods, including the low-rank 
-SDP solver introduced by Arora, Hazan, and Kale \cite{aroraSDP}. Briefly, the 
-AHK algorithm seeks to maximize 
+descent, as well as a host of other first-order methods.
+
+An example of a first-order method captured by our framework is 
+the low-rank SDP solver introduced by Arora, Hazan, and Kale \cite{aroraSDP}. 
+Briefly, the AHK algorithm seeks to maximize 
 $\sum_{j=1}^m \frac{1}{2}(\Tr(A_j^TX)-b_j)^2$ subject to the 
@@
 
+\NA{JHH: expand? other papers for inclusion?} \new Our algorithms and 
+analyses complement those in \cite{Mairal:2013}. Our primal and dual 
+algorithms exploit strong convexity and smoothness, respectively,\footnote{
+In the dual algorithm, the strong convexity of $R$ implies that 
+$R^{*}$ is strongly smooth.} whereas the algorithms given by \cite{Mairal:2013} 
+are based on (strongly) convex surrogate functions. One direction for 
+future investigation would use surrogate functions to generalize 
+our primal algorithm in the spirit of the generalized Frank-Wolfe
+scheme \cite[Algorithm 3]{Mairal:2013}.
+
 \paragraph{$q$-herding.}
 In addition to unifying several existing methods, our framework allows us to 
-extend herding to a an algorithm that we call \emph{$q$-herding}. Herding is 
+extend herding to an algorithm we call \emph{$q$-herding}. Herding is 
 an algorithm for constructing pseudosamples that match a specified collection 
@@
 \cite{Welling:2009a} and was shown to be a special case of conditional gradient 
-by Bach et al. \cite{Bach:2012a}. It can be cast as trying to minimize 
-$\|\bE_{\mu}[\phi(x)]-\bar{\phi}\|_2^2$ for $\mu$ in the probability simplex over 
-$\sX$, for a given $\phi : \sX \to \bR^d$ and $\bar{\phi} \in \bR^d$. As shown in 
+by Bach et al.~\cite{Bach:2012a}. Let $\sM_{\sX}$ be the probability simplex over 
+$\sX$. Herding can be cast as trying to minimize 
+$\|\bE_{\mu}[\phi(x)]-\bar{\phi}\|_2^2$ over $\mu \in \sM_{\sX}$, for a
+given $\phi : \sX \to \bR^d$ and $\bar{\phi} \in \bR^d$. As shown in 
 \cite{Bach:2012a}, the herding updates are equivalent to \dual with 
 $h(\mu) \equiv 0$ and $R(\theta) = \theta^T\bar{\phi} + \frac{1}{2}\|\theta\|_2^2$.
+
 The implicit assumption in the herding algorithm is that $\|\phi(x)\|_2$ is 
@@
 
+\NA{JHH: needs work} \new The solutions generated by conditional gradient have 
+an attractive sparsity property: the solution after the $t$-th iteration is 
+a convex combination of $t$ extreme points of the optimized space. In the 
+case of herding, the extreme points of $\sM_{\sX}$ are delta
+functions and each delta function $\delta_{x}$ represents the element $x$ 
+being selected as a pseudosample. In fact, conjugate gradient (and hence herding
+and $q$-herding) is asymptotically optimal amongst algorithms with sparse 
+solutions \cite{Jaggi:2013}. For further discussion of the sparsity properties 
+of conjugate gradient, see Jaggi~\cite{Jaggi:2013}.
+
 \paragraph{Convergence results.}
-We end by stating our formal convergence results. For the primal algorithm 
+We end by stating our formal convergence results. Proofs are given in the Appendix. For the primal algorithm 
 (\primal) we have:
@@
 
+\newpage
+
+\section*{Appendix: Convergence Proofs} 
+\label{sec:proofs}
+
+We now prove the convergence results given above. Throughout this section, assume that $\alpha_1,\ldots,\alpha_T$ is 
+a sequence of real numbers and that $A_t = \sum_{s=1}^t \alpha_s$. 
+We further require that $A_t > 0$ for all $t$.
+
+Also recall that the Bregman divergence is defined by 
+$D_f(x_2 \| x_1) \eqdef f(x_2) - \langle \partial f(x_1), x_2-x_1 \rangle - f(x_1)$.
+
+Our proofs hinge on the following key lemma:
+
+%%% MAIN LEMMA %%%
+\begin{lemma}
+\label{lem:bregman}
+Let $z_1,\ldots,z_T$ be vectors and let $f(x)$ be a strictly convex 
+function. Define $\hat{z}_t$ to be $\frac{1}{A_t} \sum_{s=1}^t \alpha_s z_s$.
+Let $x_1,\ldots,x_T$ be chosen via $x_{t+1} = \arg\min_{x} f(x) + \langle \hat{z}_t, x\rangle$. 
+Then for any $x^*$ we have
+\begin{align*}
+\lefteqn{\frac{1}{A_T} \sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
+\phantom{+} &\leq f(x^*) + \langle \hat{z}_t, x^* \rangle + \frac{1}{A_T} \sum_{t=1}^T A_t D_{f}(x_t \| x_{t+1}). 
+\end{align*}
+\end{lemma}
+
+%%% START PROOF %%%
+\begin{proof}
+First note that, if $x_0 = \arg\min f(x) + \langle z, x \rangle$, 
+then $\partial f(x_0) = -z$.
+
+Now note that
+\begin{align}
+\alpha_{t}z_{t} 
+&= A_{t}\hat z_{t} - A_{t-1}\hat z_{t-1} \\
+&= - A_{t}\partial f(x_{t+1}) + A_{t-1} \partial f(x_{t}),
+\end{align}
+so we have
+\begin{align} 
+\lefteqn{\sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
+ &= \sum_{t=1}^T \{\alpha_t f(x_t) + \langle A_{t-1} \partial f(x_t) - A_t \partial f(x_{t+1}), x_t \rangle\} \\
+ &= \sum_{t=1}^T \{\alpha_t f(x_t) - \langle A_{t} \partial f(x_{t+1}), x_t-x_{t+1} \rangle\} \\
+ & \quad - A_{T}\langle \partial f(x_{T+1}), x_{T+1} \rangle \\
+ &= \sum_{t=1}^T \{A_t f(x_t) - \langle A_{t} \partial f(x_{t+1}), x_t-x_{t+1} \rangle - A_t f(x_{t+1})\}  \\
+ &\quad+ A_T(f(x_{T+1}) - \langle \partial f(x_{T+1}), x_{T+1} \rangle) \nonumber \\ 
+ &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\}  \\ 
+ &\quad+ A_T(f(x_{T+1}) - \langle \partial f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
+ &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) + \langle \hat{z}_T, x_{T+1} \rangle) \\
+ &\leq \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x^*) + \langle \hat{z}_T, x^* \rangle). 
+\end{align}
+Dividing both sides by $A_T$ completes the proof.
+\end{proof}
+%%% END PROOF %%%
+
+We also note that $D_f(x_t \| x_{t+1}) = D_{f^*}(\hat{z}_{t+1} \| z_t)$, where $f^*(z) = \sup_x \langle z,x\rangle - f(x)$. 
+This form of the bound will often be more useful to us.
+\begin{lemma}
+\label{lem:convexity}
+Suppose that $D_f(x' \| x) \geq \frac{1}{2}\|x-x'\|^2$ for some 
+norm $\|\cdot\|$. (In this case we say that $f$ is strongly 
+convex with respect to $\|\cdot\|$.) 
+Then $D_{f^*}(x' \| x) \leq \frac{1}{2}\|x-x'\|_{*}^2$.
+\end{lemma}
+
+%%% ALGORITHM 1 ANALYSIS %%%
+We require a final supporting proposition before proving Theorem \ref{thm:method-1}.
+
+% Convergence proposition
+\begin{proposition}[Convergence of \primal]
+\label{prop:method-1}
+Consider the updates $\theta_t \in \arg\max_{\theta} \langle \theta, u_t \rangle - R(\theta)$ 
+and $u_{t+1} \in \arg\min_u h(u) + \langle \hat{\theta}_s, u \rangle$. 
+Then we have
+\begin{equation}
+\sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}).
+\end{equation}
+\end{proposition}
+\begin{proof}
+Note that $L(u_t, \theta_t) = \max_{\theta} L(u_t, \theta)$ by construction. 
+Also note that, if we invoke Lemma~\ref{lem:bregman} with $f = h$ and 
+$z_t = \theta_t$ then we get the inequality
+\begin{align}
+\frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) 
+&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) \\
+&\quad + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}). \nonumber
+\end{align}
+Combining these together, we get the string of inequalities
+\begin{align*} 
+L(\hat{u}_T, \theta) &= L\left(\frac{1}{A_T} \sum_{t=1}^T \alpha_tu_t, \theta\right) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_h(u_t \| u_{t+1}) \\
+ &\leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}),
+\end{align*}
+as was to be shown.
+\end{proof}
+
+% Convergence rate theorem
+%\begin{theorem}
+%\label{thm:method-1}
+%Suppose that $h$ is strongly convex with respect to a norm $\|\cdot\|$ 
+%and let $r = \sup_{\theta} \|\theta\|_{*}$. Then 
+%\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
+%\end{theorem}
+\begin{proof}[Proof of Theorem \ref{thm:method-1}]
+By Lemma~\ref{lem:convexity}, we have 
+\begin{align*} 
+D_h(u_t \| u_{t+1}) &= D_{h^*}(\hat{\theta}_{t+1} \| \hat{\theta}_{t}) \\
+ &\leq \frac{1}{2}\|\hat\theta_{t+1}-\hat\theta_{t}\|_{*}^2 \\
+ &= \frac{1}{2}\|\frac{\sum_{s \leq t}\alpha_s\theta_s}{\sum_{s \leq t} \alpha_s} - \frac{\sum_{s \leq t+1}\alpha_s\theta_s}{\sum_{s \leq t+1} \alpha_s}\|_{*}^2 \\
+ &= \frac{1}{2}\|\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\theta_s - \frac{\alpha_{t+1}}{A_{t+1}} \theta_{t+1}\|_{*}^2 \\
+ &\leq \frac{1}{2} \left(\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\|\theta_s\|_{*}^2 + \frac{\alpha_{t+1}}{A_{t+1}} \|\theta_{t+1}\|_{*}^2\right)^2 \\
+ &\leq \frac{2r^2\alpha_{t+1}^2}{A_{t+1}^2}.
+\end{align*}
+It follows that 
+\begin{align*}
+\frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}) &\leq \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}.
+\end{align*}
+\end{proof}
+
+% Convergence rate corollary 
+%\begin{corollary} 
+%\label{cor:method-1}
+%Under the hypotheses of Theorem \ref{thm:method-1}, for $\alpha_{t} = 1$ we have
+%\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2 (\log (T) + 1)}{T}. \]
+%and for $\alpha_t = t$ we have
+%\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T}. \]
+%\end{corollary}
+\begin{proof}[Proof of Corollary \ref{cor:method-1}]
+If we let $\alpha_t = 1$, then $A_t = t$ and $\frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} = \frac{t}{(t+1)^{2}} \le \frac{1}{t}$.
+We therefore get
+\begin{equation}
+\frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} \leq \frac{2r^2}{T} \sum_{t=1}^{T}\frac{1}{t} \leq \frac{2r^2(\log (T) + 1)}{T+1}.
+\end{equation}
+If we let $\alpha_t = t$, then $A_t = \frac{t(t+1)}{2}$ and 
+$\frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} = \frac{2(t+1)^2t(t+1)}{(t+1)^2(t+2)^2} = \frac{2t(t+1)}{(t+2)^2} \leq 2$.
+We therefore get
+\begin{equation}
+\frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2} \leq \frac{4r^2}{T(T+1)} \sum_{t=1}^{T}2 = \frac{8r^2}{T+1},
+\end{equation}
+which completes the proof.
+\end{proof}
+
+%%% ALGORITHM 2 ANALYSIS %%%
+The proof of Theorem \ref{thm:method-2} requires an analagous supporting proposition to that of Theorem \ref{thm:method-1}.
+% Convergence proposition
+\begin{proposition}[Convergence of \dual]
+\label{prop:method-2}
+Consider the updates $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$ 
+and $\theta_{t+1} \in \arg\max_{\theta} \langle \theta, \hat{u}_t \rangle - R(\theta)$. 
+Then we have 
+\begin{equation}
+\sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_{R}(\theta_t \| \theta_{t+1}).
+\end{equation}
+\end{proposition}
+\begin{proof}
+If we invoke Lemma~\ref{lem:bregman} with $f=R$ and $z_t=-u_t$, then we get 
+the inequality
+\begin{align}
+\frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta_t) 
+&\leq \frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta^*) \\
+&\quad - \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). \nonumber
+\end{align}
+Re-arranging yields
+\begin{align}
+\frac{1}{A_T }\sum_{t=1}^T \alpha_t L(u_t, \theta^*) 
+&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
+&\quad+ \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). 
+\end{align}
+Now, we have the following string of inequalities:
+\begin{align*} 
+L(\hat{u}, \theta) &= L\left(\frac{1}{A_T} \sum_{t=1}^T \alpha_t u_t, \theta\right) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_R(\theta_t \| \theta_{t+1}) \\
+ &= \frac{1}{A_T} \sum_{t=1}^T \alpha_t \inf_{u} L(u, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
+ &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}) \\
+ &= \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}),
+\end{align*}
+as was to be shown.
+\end{proof}
+
+% Convergence rate theorem
+\begin{theorem}
+\label{thm:method-2}
+Suppose that $R$ is strongly convex with respect to a norm $\|\cdot\|$ 
+and let $r = \sup_{u} \|u\|_{*}$. Then 
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
+\end{theorem}
+\begin{corollary}
+\label{cor:method-2}
+Under the hypotheses of Theorem \ref{thm:method-2}, for $\alpha_t = 1$ we have
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2(\log(T) + 1)}{T} \]
+and for $\alpha_t = t$ we have
+\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T} \]
+\end{corollary}
+\begin{proof}[Proofs]
+The proofs are identical to those for \primal.
+\end{proof}
+
 \end{document}

### 2013-12-02 nips-poster/nips-poster.tex
created skeleton for NIPS poster

--- 2013-10-25 nips/paper.tex
@@
-\documentclass{article} % For LaTeX2e
-\usepackage{import, subfiles}
+\documentclass[final]{beamer}
+\mode<presentation>
+{
+ %\usetheme{Berlin}
+ % \usetheme{Aachen}
+%  \usetheme{Oldi6}
+%  \usetheme{I6td}
+ \usetheme{I6dv}
+%  \usetheme{I6pd}
+%  \usetheme{I6pd2}
+}
+
+\usepackage{bm}
+
+\usepackage{times}
+\usepackage{amsmath,amssymb,amsthm}
+
+\usepackage[english]{babel}
+\usepackage[latin1]{inputenc}
+
+\usepackage{graphicx}
 \usepackage{tabularx}
-\usepackage{nips13submit_e,times}
+\usepackage{natbib}
+
 \input{latex-defs.tex}
 
-\title{A Greedy Framework for First-Order Optimization}
-\author{
-Jacob Steinhardt\thanks{Both authors contributed equally to this work.} \\
-Department of Computer Science\\
-Stanford University\\
-Stanford, CA 94305 \\
-\texttt{jsteinhardt@cs.stanford.edu} \\
-\And
-Jonathan Huggins\footnote[1]{}\\
-Department of EECS \\
-Massachusetts Institute of Technology \\
-Cambridge, MA 02139 \\
-\texttt{jhuggins@mit.edu}
-}
+\usepackage[orientation=landscape,size=custom,width=129,height=110,scale=1.4]{beamerposter}
 
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
+% Display a grid to help align images
+%\beamertemplategridbackground[1cm]
 
-\nipsfinalcopy
+\title%[] %(optional, use only with long paper titles)
+{A Greedy Framework for First-Order Optimization}
+
+%\subtitle {Include Only If Paper Has a Subtitle}
+
+\author%[Author, Another] % (optional, use only with lots of authors)
+{Jonathan Huggins$^{*1}$, Jacob Steinhardt$^{*2}$}
+
+\institute[] % (optional, but mostly needed)
+{$^{*}$Both authors contributed equally to this work. $^1$Massachusetts Institute of Technology. $^{2}$Stanford University.}
+
+\date{December 2013}
 
 \begin{document}
-\maketitle
 
+\begin{frame}{} 
+  	%\titlepage  
+%\vfill
+    \begin{columns}
+\begin{column}{0.48\linewidth}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%% COLUMN 1		 						%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% BEGIN INTRODUCTION BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{block}{\large Introduction}
 Recent work has shown many connections between conditional gradient 
@@
 in this class. 
+\end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% END INTRODUCTION BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% BEGIN ALGORITHM BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{block}{\large The Boosted Mirror Descent Algorithm}
 
@@
 
+\begin{columns}[t]
+%%% PRIMAL ALGORITHM SUB-BLOCK %%%
+\begin{column}{0.45\linewidth}
+\begin{block}{Primal Algorithm}
 Given the function $L(u, \theta)$, we define the following 
@@
 to the subgradients of $f$ by $\partial f_0(x) = \partial f(x) - \gamma x$.
+\end{block}
+\end{column}
 
-We can also concern the dual form of our mirror descent algorithm 
+%%% DUAL ALGORITHM SUB-BLOCK %%%
+\begin{column}{0.45\linewidth}
+\begin{block}{Dual Algorithm}
+We can also consider the dual form of our mirror descent algorithm 
 ({\em dual boosted mirror descent}, or \dual):
 \begin{enumerate}
-\item $\theta_1 \in \arg\min_{\theta} R(\theta)$
+\item $\theta_1 \in \arg\min_{\theta} R(\theta)$a
 \item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle = \partial h^{*}(-\theta_{t})$
@@
 algorithm equivalent to the usual formulation.
+\end{block}
 
-Our framework is intriguing in that it involves a purely greedy minimization 
-of surrogate loss functions (alternating between the primal and dual), and yet 
-is powerful enough to capture both subgradient descent and conditional gradient 
-descent, as well as a host of other first-order methods, including the low-rank 
-SDP solver introduced by Arora, Hazan, and Kale \cite{aroraSDP}. Briefly, the 
-AHK algorithm seeks to maximize 
-$\sum_{j=1}^m \frac{1}{2}(\Tr(A_j^TX)-b_j)^2$ subject to the 
-constraints $X \succeq 0$ and $\Tr(X) \leq \rho$.\footnote{This is actually a 
-variant of their algorithm, which we present for ease of exposition.} We 
-can then define 
-\[ h(X) = \left\{ \begin{array}{ccl} 0 & : & X \succeq 0\mathrm{ \ and \ }\Tr(X) \leq \rho \\ \infty & : & \mathrm{else} \end{array} \right.\]
-and
-\[ R^*(X) = \sum_{j=1}^m \frac{1}{2}(\Tr(A_j^TX)-b_j)^2. \]
-Note that this decomposition is actually a special case of the conditional 
-gradient decomposition above, and so we obtain the updates 
-\[ X_{t+1} \in \operatorname{argmin}_{X \succeq 0, Tr(X) \leq \rho} \sum_{j=1}^m \left[\Tr(A_j^T\hat{X}_t)-b_j\right]\Tr(A_j^TX), \]
-whose solution turns out to be $\rho vv^T$, where $v$ is the top singular 
-vector of the matrix
-$-\sum_{j=1}^m \left[\Tr(A_j^T\hat{X}_t)-b_j\right]A_j$. This example serves both 
-to illustrate the flexibility of our framework and to highlight the interesting 
-fact that the Arora-Hazan-Kale SDP algorithm is a special case of conditional 
-gradient (to get the original formulation in \cite{aroraSDP}, we need to replace 
-the function $\frac{1}{2}x^2$ with $x_{+}\log x_{+}$, where $x_{+} = \max(x,0)$).
+\end{column}
+\end{columns}
+\end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% END ALGORITHM BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{column}
 
-We end by stating our formal convergence results. For the primal algorithm 
-(\primal) we have:
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%% COLUMN 2		 						%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{column}{0.48\linewidth}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% BEGIN CONVERGENCE BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{block}{\large Convergence of Boosted Mirror Descent}
+\begin{columns}[t]
+%%% PRIMAL CONVERGENCE SUB-BLOCK %%%
+\begin{column}{0.45\linewidth}
+\begin{block}{Primal Algorithm}
 \begin{theorem}
@@
 \end{corollary}
-Similarly, for the dual algorithm (\dual) we have:
+\end{block}
+\end{column}
+
+%%% DUAL CONVERGENCE SUB-BLOCK %%%
+\begin{column}{0.45\linewidth}
+\begin{block}{Dual Convergence}
 \begin{theorem}
@@
 \end{corollary}
-Thus, a step size of $\alpha_t = t$ yields the claimed $O(1/T)$ convergence rate.
+\end{block}
+\end{column}
+\end{columns}
+\end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% END CONVERGENCE BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% BEGIN AHK APPLICATION BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{block}{\large Application: Solving SDPs}
+
+\end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% END AHK APPLICATION BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% BEGIN Q-HERDING APPLICATION BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{block}{\large Application: $q$-herding}
+
+\end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% END Q-HERDING APPLICATION BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% BEGIN ACKNOWLEDGMENTS BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{block}{\large Acknowledgements}
+We thank Cameron Freer for helpful conversations that led to this work.
+JHH was supported by the U.S.
+Government under FA9550-11-C-0028 and awarded by the Department of Defense, Air Force Office of Scientific Research, National Defense Science and Engineering Graduate (NDSEG) Fellowship, 32 CFR 168a.
+\end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% END ACKNOWLEDGMENTS BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% BEGIN BIBLIOGRAPHY BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{block}{\small References}
+\begin{footnotesize}
+\bibliographystyle{plainnat}
 \bibliography{herding}
-\bibliographystyle{plain}
+\end{footnotesize}
+\end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% END BIBLIOGRAPHY BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+\end{column}
+
+\end{columns}
+\end{frame}
 \end{document}

### 2013-12-02 nips/paper.tex
revised paper, updated related work

@@
 Then 
-the {\em primal boosted mirror descent} (\primal) algorithm is:
+the {\em primal boosted mirror descent} (\primal) algorithm is the following iterative procedure:
 \begin{enumerate}
@@
 \begin{enumerate}
-\item $\theta_1 \in \arg\min_{\theta} R(\theta)$a
+\item $\theta_1 \in \arg\min_{\theta} R(\theta)$
 \item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle = \partial h^{*}(-\theta_{t})$
@@
 function. Then we are minimizing $f$ over the compact set $K$, and we obtain 
-the updates from conditional gradient:
+the following updates, which are equivalent to (standard) conditional gradient 
+or Frank-Wolfe:
 \begin{enumerate}
@@
 
+\paragraph{Optimality.}
+The solutions generated by conditional gradient have 
+an attractive sparsity property: the solution after the $t$-th iteration is 
+a convex combination of $t$ extreme points of the optimized space. In 
+\cite{Jaggi:2013} it is shown that conditional gradient is asymptotically 
+optimal amongst algorithms with sparse solutions. This 
+suggests that the $1/T$ convergence rate obtained in this paper cannot be 
+improved without further assumptions. For further discussion of the properties 
+of conditional gradient, see Jaggi~\cite{Jaggi:2013}.
+
+\paragraph{Primal vs. dual algorithms.}
+It is worth taking a moment to contrast the primal and dual algorithms. Note 
+that \emph{both} algorithms have a primal convergence guarantee (i.e. both 
+guarantee that $\hat{u}_T$ is close to optimal). The sense in which the latter 
+algorithm is a ``dual'' algorithm is that it hinges on strong convexity of 
+the dual, as opposed to strong convexity of the primal. If we care about dual 
+convergence instead of primal convergence, the only change we need to make 
+is to take $\hat{\theta}_T$ at the end of the algorithm instead of $\hat{u}_T$. 
+(This can be seeing by defining the loss function $L_2(\theta, u) \eqdef -L(u, \theta)$, 
+which inverts the role of $u$ and $\theta$ but yields the same updates as above.)
+
 \paragraph{Discussion.}
@@
 of surrogate loss functions (alternating between the primal and dual), and yet 
-is powerful enough to capture both subgradient descent and conditional gradient 
-descent, as well as a host of other first-order methods.
+is powerful enough to capture both (a variant of) subgradient descent and 
+conditional gradient descent, as well as a host of other first-order methods.
 
-An example of a first-order method captured by our framework is 
+An example of a more complex first-order method captured by our framework is 
 the low-rank SDP solver introduced by Arora, Hazan, and Kale \cite{aroraSDP}. 
-Briefly, the AHK algorithm seeks to maximize 
+Briefly, the AHK algorithm seeks to minimize 
 $\sum_{j=1}^m \frac{1}{2}(\Tr(A_j^TX)-b_j)^2$ subject to the 
@@
 the function $\frac{1}{2}x^2$ with $x_{+}\log x_{+}$, where $x_{+} = \max(x,0)$). 
+It is worth noting that the analysis in \cite{aroraSDP} appears to be tighter in 
+their special case than the analysis we give here. We plan to investigate this 
+phenomenon in future work.
 
-\NA{JHH: expand? other papers for inclusion?} \new Our algorithms and 
-analyses complement those in \cite{Mairal:2013}. Our primal and dual 
-algorithms exploit strong convexity and smoothness, respectively,\footnote{
-In the dual algorithm, the strong convexity of $R$ implies that 
-$R^{*}$ is strongly smooth.} whereas the algorithms given by \cite{Mairal:2013} 
-are based on (strongly) convex surrogate functions. One direction for 
-future investigation would use surrogate functions to generalize 
-our primal algorithm in the spirit of the generalized Frank-Wolfe
-scheme \cite[Algorithm 3]{Mairal:2013}.
+\paragraph{Related work.}
+A generalized Frank-Wolfe algorithm has also been proposed in \cite{Mairal:2013}, 
+which considers first-order convex surrogate functions more generally. Algorithm 3 
+of \cite{Mairal:2013} turns out to be equivalent to a line-search formulation of the 
+algorithm we propose in this paper. However, the perspective is quite different. 
+While both algorithms share the intuition of trying to construct a ``proximal Frank-Wolfe'' 
+algorithm, the general framework of \cite{Mairal:2013} studies convex \emph{upper bounds}, 
+rather than the convex \emph{lower bounds} of this paper (although for the 
+``proximal gradient surrogates'' that they suggest using, their Frank-Wolfe algorithm 
+does indeed deal with lower bounds). Another difference is that their analysis focuses 
+on the Lipschitz properties of the objective function, whereas we use the more general 
+notion of strong convexity with respect to an arbitrary norm. We believe that this 
+geometric perspective is useful for tailoring the proximal function to specific applications, 
+as demonstrated in our discussion of $q$-herding below. Finally, the saddle point 
+perspective on Frank-Wolfe is not present in \cite{Mairal:2013}, although they 
+do discuss the general idea of saddle-point methods in relation to convex surrogates.
 
@@
 $\frac{1}{p} + \frac{1}{q} = 1$. In this case our convergence results imply that 
-$\|\bE_{\mu}[\phi(x)]-\bar{\phi}\|_p^p$ decays at a rate of $O(1/T)$.
-
-\NA{JHH: needs work} \new The solutions generated by conditional gradient have 
-an attractive sparsity property: the solution after the $t$-th iteration is 
-a convex combination of $t$ extreme points of the optimized space. In the 
-case of herding, the extreme points of $\sM_{\sX}$ are delta
-functions and each delta function $\delta_{x}$ represents the element $x$ 
-being selected as a pseudosample. In fact, conjugate gradient (and hence herding
-and $q$-herding) is asymptotically optimal amongst algorithms with sparse 
-solutions \cite{Jaggi:2013}. For further discussion of the sparsity properties 
-of conjugate gradient, see Jaggi~\cite{Jaggi:2013}.
+$\|\bE_{\mu}[\phi(x)]-\bar{\phi}\|_p^p$ decays at a rate of $O(1/T)$ (proof to 
+appear in an extended version of this paper).
 
@@
 Suppose that $h$ is strongly convex with respect to a norm $\|\cdot\|$ 
-and let $r = \sup_{\theta} \|\theta\|_{*}$. Then 
+and let $r = \sup_{\theta} \|\theta\|_{*}$. Then, for any $u^*$,
 \[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
@@
 Suppose that $R$ is strongly convex with respect to a norm $\|\cdot\|$ 
-and let $r = \sup_{u} \|u\|_{*}$. Then 
+and let $r = \sup_{u} \|u\|_{*}$. Then, for any $u^*$,
 \[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
@@
 Thus, a step size of $\alpha_t = t$ yields the claimed $O(1/T)$ convergence rate.
+
+\paragraph{Acknowledgements.}
+Thanks to Simon Lacoste-Julien, Martin Jaggi, Percy Liang, Arun Chaganty, and Sida Wang 
+for helpful discussions. Thanks also to the anonymous reviewer for suggesting improvements 
+to the paper. This work was supported by the Hertz Foundation and by the NDSEG graduate 
+fellowship.
 
@@
 \begin{align*}
-\lefteqn{\frac{1}{A_T} \sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
-\phantom{+} &\leq f(x^*) + \langle \hat{z}_t, x^* \rangle + \frac{1}{A_T} \sum_{t=1}^T A_t D_{f}(x_t \| x_{t+1}). 
+\frac{1}{A_T} \sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\} \leq f(x^*) + \langle \hat{z}_t, x^* \rangle + \frac{1}{A_T} \sum_{t=1}^T A_t D_{f}(x_t \| x_{t+1}). 
+% \lefteqn{\frac{1}{A_T} \sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
+% \phantom{+} &\leq f(x^*) + \langle \hat{z}_t, x^* \rangle + \frac{1}{A_T} \sum_{t=1}^T A_t D_{f}(x_t \| x_{t+1}). 
 \end{align*}
@@
 We also note that $D_f(x_t \| x_{t+1}) = D_{f^*}(\hat{z}_{t+1} \| z_t)$, where $f^*(z) = \sup_x \langle z,x\rangle - f(x)$. 
-This form of the bound will often be more useful to us.
+This form of the bound will often be more useful to us. Another useful property of Bregman divergences is the following lemma 
+which relates strong convexity of $f$ to strong smoothness of $f^*$:
 \begin{lemma}
@@
 Suppose that $D_f(x' \| x) \geq \frac{1}{2}\|x-x'\|^2$ for some 
-norm $\|\cdot\|$. (In this case we say that $f$ is strongly 
+norm $\|\cdot\|$ and for all $x, x'$. (In this case we say that $f$ is strongly 
 convex with respect to $\|\cdot\|$.) 
-Then $D_{f^*}(x' \| x) \leq \frac{1}{2}\|x-x'\|_{*}^2$.
+Then $D_{f^*}(y' \| y) \leq \frac{1}{2}\|y-y'\|_{*}^2$ for all $y, y'$.
 \end{lemma}
@@
 \begin{equation}
-\sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}).
+\sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}).
 \end{equation}
@@
 \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) 
-&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) \\
-&\quad + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}). \nonumber
+&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}).
 \end{align}
@@
 \begin{proof}
-If we invoke Lemma~\ref{lem:bregman} with $f=R$ and $z_t=-u_t$, then we get 
+If we invoke Lemma~\ref{lem:bregman} with $f=R$ and $z_t=-u_t$, then for all $\theta$ we get 
 the inequality
@@
 \frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta_t) 
-&\leq \frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta^*) \\
+&\leq \frac{1}{A_T} \sum_{t=1}^T -\alpha_t L(u_t, \theta) \\
 &\quad - \frac{1}{A_T} \sum_{t=1}^T A_tD_R(\theta_t \| \theta_{t+1}). \nonumber
@@
 \begin{align}
-\frac{1}{A_T }\sum_{t=1}^T \alpha_t L(u_t, \theta^*) 
+\frac{1}{A_T }\sum_{t=1}^T \alpha_t L(u_t, \theta) 
 &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
@@
 
-% Convergence rate theorem
-\begin{theorem}
-\label{thm:method-2}
-Suppose that $R$ is strongly convex with respect to a norm $\|\cdot\|$ 
-and let $r = \sup_{u} \|u\|_{*}$. Then 
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
-\end{theorem}
-\begin{corollary}
-\label{cor:method-2}
-Under the hypotheses of Theorem \ref{thm:method-2}, for $\alpha_t = 1$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2(\log(T) + 1)}{T} \]
-and for $\alpha_t = t$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T} \]
-\end{corollary}
-\begin{proof}[Proofs]
+\begin{proof}[Proofs of Theorem~\ref{thm:method-2} and Corollary~\ref{cor:method-2}]
 The proofs are identical to those for \primal.

### 2013-12-02 nips/paper.tex
small fixes to nips paper

@@
 is to take $\hat{\theta}_T$ at the end of the algorithm instead of $\hat{u}_T$. 
-(This can be seeing by defining the loss function $L_2(\theta, u) \eqdef -L(u, \theta)$, 
+(This can be seeing by defining the loss function $L'(\theta, u) \eqdef -L(u, \theta)$, 
 which inverts the role of $u$ and $\theta$ but yields the same updates as above.)
@@
 \paragraph{Acknowledgements.}
-Thanks to Simon Lacoste-Julien, Martin Jaggi, Percy Liang, Arun Chaganty, and Sida Wang 
+Thanks to Cameron Freer, Simon Lacoste-Julien, Martin Jaggi, Percy Liang, Arun Chaganty, and Sida Wang 
 for helpful discussions. Thanks also to the anonymous reviewer for suggesting improvements 
-to the paper. This work was supported by the Hertz Foundation and by the NDSEG graduate 
+to the paper.
+JR was supported by the Hertz Foundation.
+JHH was supported by the U.S.
+Government under FA9550-11-C-0028 and awarded by the Department of Defense, Air Force Office of Scientific Research, National Defense Science and Engineering Graduate (NDSEG) Fellowship, 32 CFR 168a.
 fellowship.

### 2013-12-02 nips/paper.tex
1/T -> O(1/T)

@@
 Convergence now hinges upon strong convexity of $R$ rather than 
-$h$, and we obtain the same $1/T$ convergence 
+$h$, and we obtain the same $O(1/T)$ convergence 
 rate (see Corollary~\ref{cor:method-2}):
@@
 optimal amongst algorithms with sparse solutions. This 
-suggests that the $1/T$ convergence rate obtained in this paper cannot be 
+suggests that the $O(1/T)$ convergence rate obtained in this paper cannot be 
 improved without further assumptions. For further discussion of the properties 

### 2013-12-04 nips/paper.tex
adding slides

@@
 ``proximal gradient surrogates'' that they suggest using, their Frank-Wolfe algorithm 
-does indeed deal with lower bounds). Another difference is that their analysis focuses 
+does implicitly use lower bounds). Another difference is that their analysis focuses 
 on the Lipschitz properties of the objective function, whereas we use the more general 
@@
 to the paper.
-JR was supported by the Hertz Foundation.
+JS was supported by the Hertz Foundation.
 JHH was supported by the U.S.
 Government under FA9550-11-C-0028 and awarded by the Department of Defense, Air Force Office of Scientific Research, National Defense Science and Engineering Graduate (NDSEG) Fellowship, 32 CFR 168a.
-fellowship.
 

### 2013-12-04 nips-poster/nips-poster.tex
added logos and resized poster
first pass at structuring the NIPS poster

@@
 
-\usepackage[orientation=landscape,size=custom,width=129,height=110,scale=1.4]{beamerposter}
+%\usepackage[orientation=landscape,size=a0,scale=1.4]{beamerposter}
+\usepackage[orientation=landscape,size=custom,width=121.92,height=91.44,scale=1.4]{beamerposter} % 48in x 36in
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% BEGIN INTRODUCTION BLOCK %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{block}{\large Introduction}
-Recent work has shown many connections between conditional gradient 
-and other first-order optimization methods, such as herding \cite{Bach:2012a} 
-and subgradient descent \cite{Bach:2012b}. By considering a type of 
-\emph{proximal conditional method}, which we call boosted mirror 
-descent (\bmd), we are able to unify all of these algorithms into 
-a single framework, which can be interpreted as taking successive 
-arg-mins of a sequence of surrogate functions. Using a standard 
-online learning analysis based on Bregman divergences, we are 
-able to demonstrate an $O(1/T)$ convergence rate for all algorithms 
-in this class. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% BEGIN MOTIVATION BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{block}{\large Motivation}
+JS
 \end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% END MOTIVATION BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% END INTRODUCTION BLOCK %%%
+%%% BEGIN TOO GREEDY BLOCK %%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{block}{\large Being (Too) Greedy}
+JHH
+\end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% END TOO GREEDY BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{block}{\large The Boosted Mirror Descent Algorithm}
+\begin{block}{\large Being Greedy Enough: Boosted Mirror Descent}
+JHH
 
-More concretely, suppose that we are given a function 
-$L : U \times \Theta \to \bR$ defined by
-\[ L(u, \theta) = h(u) + \ip{u}{\theta} - R(\theta) \]
-and wish to find the \emph{saddle point} 
-\[ L_* \eqdef \min_{u} \max_{\theta} L(u, \theta). \]
-We should think of $u$ as the primal variables and $\theta$ 
-as the dual variables; we will assume throughout that 
-$h$ and $R$ are both convex. We will also abuse notation 
-and define $L(u) \eqdef \max_{\theta} L(u, \theta)$; we 
-can equivalently write $L(u)$ as 
-\[ L(u) = h(u) + R^*(u), \]
-where $R^*$ is the Fenchel conjugate of $R$. Note that 
-$L(u)$ is a convex function. Moreover, since 
-$R \leftrightarrow R^*$ is a one-to-one mapping, for 
-\emph{any} convex function $L$ and \emph{any} decomposition 
-of $L$ into convex functions $h$ and $R^*$, we get a 
-corresponding two-argument function $L(u, \theta)$.
-
-\begin{columns}[t]
-%%% PRIMAL ALGORITHM SUB-BLOCK %%%
-\begin{column}{0.45\linewidth}
-\begin{block}{Primal Algorithm}
-Given the function $L(u, \theta)$, we define the following 
-optimization procedure, which will generate a sequence of 
-points $(u_1,\theta_1), (u_2, \theta_2), \ldots$ converging 
-to a saddle point of $L$. First, take a sequence of weights
-$\alpha_1, \alpha_2, \ldots$, and for notational convenience 
-define
-\(
-\hat{u}_t = \frac{\sum_{s=1}^t \alpha_su_s}{\sum_{s=1}^t \alpha_s} 
-\quad \text{and} \quad 
-\hat{\theta}_t = \frac{\sum_{s=1}^t \alpha_s\theta_s}{\sum_{s=1}^t \alpha_s}.
-\)
-Then 
-the {\em primal boosted mirror descent} (\primal) algorithm is:
-\begin{enumerate}
-\item $u_1 \in \arg\min_u h(u)$
-\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, u_t \rangle - R(\theta) = \partial R^{*}(u_{t})$
-\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \hat{\theta}_t, u \rangle = \partial h^{*}(-\hat\theta_{t})$
-\end{enumerate}
-As long as $h$ is strongly convex, for the proper choice of $\alpha_{t}$ we 
-obtain the bound (see Corollary~\ref{cor:method-1}):
-\begin{equation}
-\sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq L_{*} + O(1/T).
-\end{equation}
-
-As an example, suppose that we are given a $\gamma$-strongly convex function 
-$f$: that is, $f(x) = \frac{\gamma}{2}\|x\|_2^2 + f_0(x)$, where $f_0$ is convex. 
-Then we let $h(x) = \frac{\gamma}{2}\|x\|_2^2$, $R^*(x) = f_0(x)$, and obtain 
-the updates:
-\begin{enumerate}
-\item $u_1 = 0$
-\item $\theta_t = \partial f_0(u_t)$
-\item $u_{t+1} = -\frac{1}{\gamma}\hat{\theta}_t = -\frac{\sum_{s=1}^t \alpha_s\partial f_0(u_s)}{\gamma \sum_{s=1}^t \alpha_s}$
-\end{enumerate}
-We therefore obtain a variant on subgradient descent where $u_{t+1}$ 
-is a weighted average of the first $t$ subgradients (times a step size 
-$1/\gamma$). Note that these are the subgradients of $f_0$, which are related 
-to the subgradients of $f$ by $\partial f_0(x) = \partial f(x) - \gamma x$.
-\end{block}
-\end{column}
-
-%%% DUAL ALGORITHM SUB-BLOCK %%%
-\begin{column}{0.45\linewidth}
-\begin{block}{Dual Algorithm}
-We can also consider the dual form of our mirror descent algorithm 
-({\em dual boosted mirror descent}, or \dual):
-\begin{enumerate}
-\item $\theta_1 \in \arg\min_{\theta} R(\theta)$a
-\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle = \partial h^{*}(-\theta_{t})$
-\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \hat{u}_t \rangle - R(\theta) = \partial R^{*}(\hat u_{t})$
-\end{enumerate}
-Convergence now hinges upon strong convexity of $R$ rather than 
-$h$, and we obtain the same $1/T$ convergence 
-rate (see Corollary~\ref{cor:method-2}):
-\[ \sup_{\theta \in \Theta} L(\hat{u}_T, \theta) \leq L_{*} + O(1/T). \]
-An important special case is $h(u) = \left\{ \begin{array}{ccl} 0 & : & u \in K \\ \infty & : & u \not\in K \end{array} \right.$, where $K$ is some 
-compact set. Also let $R^* = f$, where $f$ is an arbitrary strongly convex 
-function. Then we are minimizing $f$ over the compact set $K$, and we obtain 
-the updates from conditional gradient:
-\begin{enumerate}
-\item $\theta_1 = \partial f(0)$
-\item $u_t \in \arg\min_{u \in K} \langle \theta_t, u \rangle$
-\item $\theta_{t+1} = \partial f(\hat{u}_t)$
-\end{enumerate}
-Our notation is slightly different from previous presentations in 
-that we use linear weights ($\alpha_t$) instead of geometric weights 
-(often denoted $\rho_t$, as in \cite{Bach:2012b}). However, under the 
-mapping $\alpha_t = \rho_t/\prod_{s=1}^t (1-\rho_s)$, we obtain an 
-algorithm equivalent to the usual formulation.
-\end{block}
-
-\end{column}
-\end{columns}
+This block would contain the statement of the algorithms and the geometric intuition 
 \end{block}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% BEGIN APPLICATIONS BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{block}{\large Special Cases and Applications}
+JS?
+\begin{columns}[t]
+\begin{column}{0.45\linewidth}
+
+\begin{block}{Subgradient descent}
+
+\end{block}
+
+\begin{block}{Conditional gradient descent}
+\end{block}
+
+\begin{block}{Thresholded gradient}
+\end{block}
+
+\end{column}
+\begin{column}{0.45\linewidth}
+\begin{block}{AHK low-rank SDP solver}
+\end{block}
+
+\begin{block}{$q$-herding}
+\end{block}
+
+\end{column}
+\end{columns}
+\end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% END APPLICATIONS BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{columns}[t]
-%%% PRIMAL CONVERGENCE SUB-BLOCK %%%
 \begin{column}{0.45\linewidth}
-\begin{block}{Primal Algorithm}
 \begin{theorem}
-\label{thm:method-1}
-Suppose that $h$ is strongly convex with respect to a norm $\|\cdot\|$ 
-and let $r = \sup_{\theta} \|\theta\|_{*}$. Then 
+Consider the primal (respectively the dual) algorithm. Suppose that $h$ ($R$) is strongly convex with respect to a norm $\|\cdot\|$ 
+and let $r = \sup_{\theta} \|\theta\|_{*}$ ($r = \sup_{u} \|u\|_{*}$). Then
 \[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
 \end{theorem}
+\end{column}
+\begin{column}{0.45\linewidth}
 \begin{corollary} 
-\label{cor:method-1}
-Under the hypotheses of Theorem \ref{thm:method-1}, for $\alpha_{t} = 1$ we have
+Under the hypotheses of the Theorem, for $\alpha_{t} = 1$ we have
 \[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2 (\log (T) + 1)}{T}. \]
@@
 \end{corollary}
-\end{block}
-\end{column}
-
-%%% DUAL CONVERGENCE SUB-BLOCK %%%
-\begin{column}{0.45\linewidth}
-\begin{block}{Dual Convergence}
-\begin{theorem}
-\label{thm:method-2}
-Suppose that $R$ is strongly convex with respect to a norm $\|\cdot\|$ 
-and let $r = \sup_{u} \|u\|_{*}$. Then 
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
-\end{theorem}
-\begin{corollary}
-\label{cor:method-2}
-Under the hypotheses of Theorem \ref{thm:method-2}, for $\alpha_t = 1$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2(\log(T) + 1)}{T} \]
-and for $\alpha_t = t$ we have
-\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T} \]
-\end{corollary}
-\end{block}
 \end{column}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% BEGIN AHK APPLICATION BLOCK %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{block}{\large Application: Solving SDPs}
-
-\end{block}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% END AHK APPLICATION BLOCK %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% BEGIN Q-HERDING APPLICATION BLOCK %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{block}{\large Application: $q$-herding}
-
-\end{block}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% END Q-HERDING APPLICATION BLOCK %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%% BEGIN AHK APPLICATION BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{block}{\large Application: Solving SDPs}
+%
+%
+%\end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%% END AHK APPLICATION BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%% BEGIN Q-HERDING APPLICATION BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{block}{\large Application: $q$-herding}
+%
+%
+%\end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%% END Q-HERDING APPLICATION BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{block}{\large Acknowledgements}
-We thank Cameron Freer for helpful conversations that led to this work.
+\begin{block}{\small Acknowledgements}
+{\footnotesize Thanks to Cameron Freer, Simon Lacoste-Julien, Martin Jaggi, Percy Liang, Arun Chaganty, and Sida Wang 
+for helpful discussions. Thanks also to the anonymous reviewer for suggesting improvements 
+to the paper.
+JS was supported by the Hertz Foundation.
 JHH was supported by the U.S.
-Government under FA9550-11-C-0028 and awarded by the Department of Defense, Air Force Office of Scientific Research, National Defense Science and Engineering Graduate (NDSEG) Fellowship, 32 CFR 168a.
+Government under FA9550-11-C-0028 and awarded by the Department of Defense, Air Force Office of Scientific Research, National Defense Science and Engineering Graduate (NDSEG) Fellowship, 32 CFR 168a.}
 \end{block}

### 2013-12-04 nips-poster/nips-poster.tex
poster

@@
 \begin{block}{\large Motivation}
-JS
+Suppose we want to solve the following \emph{saddle point} problem:
+\[ \min_u \max_{\theta} L(u, \theta), \]
+where $L(u,\theta) = h(u) + \langle u, \theta \rangle - R(\theta)$. 
+We assume that $h$ and $R$ are both convex and that $\langle \cdot, \cdot \rangle$ 
+is an arbitrary bilinear map.
+
+\textbf{Intuition:} $u$ is the ``primal'', $\theta$ is the ``dual''.
+
+Tie-in with optimization: can recast as minimizing
+\[ L(u) \eqdef \max_{\theta} L(u, \theta) = h(u) + R^*(u). \]
+
+Here $R^*(u) \eqdef \max_{\theta} u^T\theta - R(\theta)$ is the 
+\emph{Fenchel conjugate} of $\theta$.
+\end{block}
+
+\begin{block}{\large Fenchel conjugate properties}
+Fenchel conjugates are a key component of \emph{Legendre-Fenchel duality} 
+and also show up in mirror descent. It turns out that they will also be 
+important for generalizing the Frank-Wolfe algorithm.
+
+\textbf{Property 1:} If $R$ is convex then $R^{**} = R$.
+
+\textbf{Property 2:} $\partial R^*(u) = \argmin_{\theta} u^T\theta - R(\theta)$.
 \end{block}
@@
 \begin{block}{\large Special Cases and Applications}
-JS?
+Many particular algorithms can be cast into our framework.
 \begin{columns}[t]
@@
 
-\begin{block}{Subgradient descent}
-
+\begin{block}{Conditional gradient (Frank-Wolfe)}
+Let $h \equiv 0$. Then we have the updates
+\begin{align*}
+u_t &= \argmin_u u^T\theta_t \\
+ &= \argmin_u u^T\partial R^*(\hat{u}_{t-1}),
+\end{align*}
+which is the Frank-Wolfe update for $R^*$.
 \end{block}
 
-\begin{block}{Conditional gradient descent}
+\begin{block}{Thresholded Frank-Wolfe}
+Suppose we want to minimize $\|u\|_1 + R^*(u)$. Then 
+we can consider the following 
+\emph{thresholded Frank-Wolfe algorithm}:
+\begin{align*}
+u_t &= \argmin_u \|u\|_1 + u^T\partial R^*(\hat{u}_{t-1}),
+\end{align*}
+which converges at the same rate as conditional gradient 
+but also imposes sparsity.
 \end{block}
 
-\begin{block}{Thresholded gradient}
+\begin{block}{(Modified) subgradient}
+Let $h(u) = \frac{\gamma}{2} \|u\|_2^2$. Then the primal version of the 
+algorithm gives the updates:
+\begin{align*}
+u_{t+1} &= \argmin_{\theta} \frac{\gamma}{2}\|u\|_2^2 + u^T\hat{\theta}_t \\
+ &= -\frac{1}{\gamma}\hat{\theta}_t \\
+ &= -\frac{1}{\gamma} \frac{\sum_{s=1}^t \alpha_s \partial R^*(u_s)}{\sum_{s=1}^t \alpha_s}.
+\end{align*}
 \end{block}
@@
 \begin{block}{AHK low-rank SDP solver}
+Suppose we want to solve the SDP
+\begin{align*}
+\text{maximize }   & \Tr(A^TX) \\
+\text{subject to } & X \succeq 0 \\
+                  & X_{ii} \leq 1 \, \forall i
+\end{align*}
+Set \small{$L(X,y) = \Tr(A^TX) + \sum_{i=1}^n \left[y_i(X_{ii}-1) - \eta y_i\log y_i\right]$}.
+Then updating $X$ is an eigenvalue problem and each update is rank $1$. This 
+yields a variant of the Arora-Hazan-Kale fast SDP solver.
 \end{block}
@@
 \begin{block}{$q$-herding}
+The \emph{herding} algorithm finds a distribution $\mu$ over 
+$\sX$ such that $\bE_{x \sim \mu}[\phi(x)] \approx \bar{\phi}$ 
+for a given collection of moments $\phi$. It is equivalent to 
+the dual version of our algorithm for
+\[ L(\mu,\theta) = \bE_{x \sim \mu}\left[\theta^T[\phi(x)-\bar{\phi}]\right] - \frac{1}{2}\|\theta\|_2^2, \]
+which hinges upon $\|\bar{\phi}\|_2$ existing. In infinite 
+dimensions this may not be the case and we can instead use the 
+\emph{$q$-herding} algorithm based on the updates
+\[ L(\mu,\theta) = \bE_{x \sim \mu}\left[\theta^T[\phi(x)-\bar{\phi}]\right] - \frac{1}{q}\|\theta\|_q^q, \]
+which only requires that $\|\bar{\phi}\|_p$ exists, where $\frac{1}{p} + \frac{1}{q} = 1$.
 \end{block}

### 2013-12-04 nips-poster/nips-poster.tex
add sections I was responsible for; minor edits elsewhere

@@
 \begin{columns}
-\begin{column}{0.48\linewidth}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{column}{0.46\linewidth}
 
@@
 is an arbitrary bilinear map.
-
+\newline\newline
 \textbf{Intuition:} $u$ is the ``primal'', $\theta$ is the ``dual''.
-
-Tie-in with optimization: can recast as minimizing
+\newline\newline
+\textbf{Tie-in with optimization:} can recast as minimizing
 \[ L(u) \eqdef \max_{\theta} L(u, \theta) = h(u) + R^*(u). \]
@@
 important for generalizing the Frank-Wolfe algorithm.
-
+\newline\newline
 \textbf{Property 1:} If $R$ is convex then $R^{**} = R$.
-
+\qquad\qquad
 \textbf{Property 2:} $\partial R^*(u) = \argmin_{\theta} u^T\theta - R(\theta)$.
@@
 \begin{block}{\large Being (Too) Greedy}
-JHH
+\begin{columns}[t]
+\begin{column}{.48\linewidth}
+We might try the following ``iterative best response'' updates:
+\begin{align*}
+u_t\,\,\,\,\,\, &= \argmin_{u} L(u,\theta_t) \\
+\theta_{t+1}   &= \argmax_{\theta} L(u_t, \theta)
+\end{align*}
+\end{column}
+\begin{column}{.48\linewidth}
+{\bf Problem:} let $u,\theta \in \bR$, $h(u) = \frac{1}{2}u^2$, $R(\theta) = \frac{1}{2}\theta^2$. 
+Then:
+\begin{align*}
+u_t\,\,\,\,\,\, &= \argmin_u \left[\frac{1}{2}u^2 + u\theta\right]             = -\theta_t \\
+\theta_{t+1}    &= \argmax_{\theta} \left[u\theta - \frac{1}{2}\theta^2\right] = u_t
+\end{align*}
+\begin{center}
+\textcolor{red}{OSCILLATION}
+\end{center}
+\end{column}
+\end{columns}
 \end{block}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{block}{\large Being Greedy Enough: Boosted Mirror Descent}
-JHH
-
-This block would contain the statement of the algorithms and the geometric intuition 
+\begin{columns}[t]
+\begin{column}{.43\linewidth}
+\begin{block}{\large Being (Less) Greedy: \\ Boosted Mirror Descent}
+Replace $u_t$ with a weighted average
+\begin{align*}
+\hat{u}_t &\eqdef \frac{1}{A_{t}} \sum_{s=1}^t \alpha_{s} u_s & A_{t} = \sum_{s=1}^{t} \alpha_{s}
+\end{align*}
+New updates (``dual'' algorithm):
+\begin{align*}
+u_t\,\,\,\,\,\, &= \argmin_{u} L(u,\theta_t)  = && \partial h^{*}(-\theta_{t}) \\
+\theta_{t+1}    &= \argmax_{\theta} L(\hat{u}_t, \theta)  = &&\partial R^{*}(\hat u_{t})
+\end{align*}
+Alternatively, replace $\theta_{t}$ with a weighted average (``primal'' algorithm):
+\begin{align*}
+u_t\,\,\,\,\,\, &= \argmin_{u} L(u,\hat\theta_t) = && \partial h^{*}(-\hat\theta_{t}) \\
+\theta_{t+1}    &= \argmax_{\theta} L(u_t, \theta) = &&\partial R^{*}(u_{t})
+\end{align*}
 \end{block}
+\end{column}
+\begin{column}{.53\linewidth}
+\begin{block}{\large A Geometric Picture (Primal)}
+\begin{center}
+\includegraphics[width=10in]{figures/primal-demo.pdf}
+\end{center}
+For the primal algorithm,
+\begin{align*}
+u_t\,\,\,\,\,\, &= \partial h^{*}(-\hat\theta_{t}) \text{ is a point where $-\hat\theta_{t} \in \partial h$} \\ 
+\theta_{t+1}    &= \partial R^{*}(u_{t}) \text{ is the gradient of $R^{*}$ at $u_{t}$}
+\end{align*}
+So $u_{t}$ is chosen such that the gradient contributed by $h$ at $u_{t}$ cancels out the gradient contributed by $R^{*}$. 
+\end{block}
+\end{column}
+\end{columns}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{column}{0.48\linewidth}
+\begin{column}{0.50\linewidth}
 
@@
 \begin{columns}[t]
-\begin{column}{0.45\linewidth}
+\begin{column}{0.01\linewidth}\end{column}
+\begin{column}{0.47\linewidth}
 
@@
 \end{column}
-\begin{column}{0.45\linewidth}
+\begin{column}{0.47\linewidth}
 \begin{block}{AHK low-rank SDP solver}
@@
 \begin{columns}[t]
-\begin{column}{0.45\linewidth}
+\begin{column}{0.01\linewidth}\end{column}
+\begin{column}{0.47\linewidth}
 \begin{theorem}
@@
 \end{column}
-\begin{column}{0.45\linewidth}
+\begin{column}{0.47\linewidth}
 \begin{corollary} 
@@
 \begin{block}{\small Acknowledgements}
-{\footnotesize Thanks to Cameron Freer, Simon Lacoste-Julien, Martin Jaggi, Percy Liang, Arun Chaganty, and Sida Wang 
-for helpful discussions. Thanks also to the anonymous reviewer for suggesting improvements 
-to the paper.
+{\tiny Thanks to Cameron Freer, Simon Lacoste-Julien, Martin Jaggi, Percy Liang, Arun Chaganty, and Sida Wang 
+for helpful discussions.
 JS was supported by the Hertz Foundation.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% BEGIN BIBLIOGRAPHY BLOCK %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{block}{\small References}
-\begin{footnotesize}
-\bibliographystyle{plainnat}
-\bibliography{herding}
-\end{footnotesize}
-\end{block}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% END BIBLIOGRAPHY BLOCK %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%% BEGIN BIBLIOGRAPHY BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{block}{\small References}
+%\begin{footnotesize}
+%\bibliographystyle{plainnat}
+%\bibliography{herding}
+%\end{footnotesize}
+%\end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%% END BIBLIOGRAPHY BLOCK %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2013-12-25 lp/l1l1-regret.tex
added l1l1

\documentclass[11pt]{article}
\usepackage{import}
\input{../../latex-utils/defs.tex}
\usepackage{fullpage}

\title{$l^1$-$l^1$ regret bound}
\begin{document}
\maketitle

Suppose that we want to minimize the regret 
\[ \sum_{t=1}^T x_t^Tz_t - \inf_{x \in \sC} \sum_{t=1}^T x^Tz_t. \]
\begin{proposition}
If $\|x\|_1 \leq 1$ for all $x \in \sC$ and 
$\|z_t\|_1 \leq 1$ for all $t$, then there is 
an algorithm whose regret 
after $T$ rounds is $\sqrt{\frac{2T}{n}}$.
\end{proposition}
\begin{proof}
Consider a matrix $Q$ such that $|Q_{ij}| \leq 1/\sqrt{n}$ 
for all $i,j$, and such that $Q^TQ = I$; one such matrix would 
be any Hadamard matrix, which exists whenever $n$ is a power of 
$2$. Then take the regularizer $h(x) = \frac{\eta}{2}x^TQx$, 
and apply the mirror descent algorithm. Note that 
$|x^TQX| \leq 1/\sqrt{n}$ whenever $\|x\|_1 \leq 1$, by H\"{o}lder's 
inequality. Then the OMD regret is:
\begin{align*}
\sup_{x \in \sC} \frac{\eta}{2}x^TQx - \inf_{x \in \sC} \frac{\eta}{2}x^TQx + \sum_{t=1}^T \frac{1}{2\eta} z_t^TQ^Tz_t &\leq \frac{\eta}{2\sqrt{n}} + \frac{\eta}{2\sqrt{n}} + \frac{T}{2\eta \sqrt{n}}.
\end{align*}
If we set $\eta$ to $\sqrt{T/2}$, then the above quantity is equal to 
$\sqrt{\frac{2T}{n}}$, which proves the result.
\end{proof}

\section{Extension}
Suppose there was a matrix $A$ such that:
\begin{align*}
\left|\frac{1}{2}x^TAx\right| &\leq n^{\frac{1}{2}-\frac{1}{p}} \|x\|_p^2. \\
\left|\frac{1}{2}y^TA^{-1}y\right| &\leq n^{\frac{1}{2}-\frac{1}{q}} \|x\|_q^2.
\end{align*}
(we conjecture that such a matrix exists for $p, q \leq 2$). Then we could 
obtain the following interesting extension of the above:
\begin{proposition}[Conjecture]
Suppose that $p, q \leq 2$. If $\|x\|_p \leq 1$ for all $x \in \sC$ 
and $\|z_t\|_q \leq 1$ for all $t$, then we can obtain regret 
$\sqrt{Tn^{1-1/p-1/q}}$.
\end{proposition}
\begin{proof}
Use the $A$ matrix that is conjectured to exist, and apply the same argument as above.
\end{proof}
\end{document}

### 2014-01-01 lp/framework.tex
added online learning framework

\documentclass[11pt]{article}
\usepackage{import}
\input{../icml/latex-defs.tex}
\usepackage{fullpage}

\title{Online Learning Framework}
\author{Jacob Steinhardt}
\begin{document}
\maketitle
We are interested in the following learning problem: our experts 
consist of vectors $w \in \sC_P$, where $\sC_P$ is some convex 
set; similarly, our adversary can play any move $z_t$ with 
$z_t \in \sC_D$. A typical example is $\sC_P$ is the $n$-dimensional 
simplex and $\sC_D$ is the unit $l^{\infty}$ ball. Our end goal is 
to obtain a regret bound of the following type:
\[ \sum_{t=1}^T w_t^Tz_t \leq \frac{\rho}{\epsilon} + \epsilon \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t. \]
If we let $\sigma = \sup_{w \in \sC_P, z \in \sC_D} |w|^T|z|$, then the regret is bounded by
$\frac{\rho}{\epsilon} + \epsilon \sigma T$, which is equal to $2\sqrt{\rho \sigma T}$ for 
$\epsilon = \sqrt{\frac{\rho}{\sigma T}}$. This is the type of regret bound that is typical 
in most online learning settings. However, our bound can lead to stronger results in some 
situations.

We propose the following general style of algorithm for obtaining such 
bounds. The key innovation is an auxiliary variable $x_t$ that lets us 
relate $w_t^Tz_t$ and $w^Tz_t$. The updates are as follows:
\begin{align*}
x_{t+1} &= f(x_t, \eta z_t) \\
w_{t+1} &= \argmax_{w \in \sC_P} w^Tx_{t+1} - \psi(w) \\
\Phi_{t+1} &= w_{t+1}^Tx_{t+1} - \psi(w_{t+1})  \\ &= \psi^*(x_{t+1}).
\end{align*}

To see why these updates are a good idea, we will make a ``physicist's approximation'' 
and suppose that $f(x_t, \eta z_t) \approx x_t - \eta z_t$. Then we would have
\[ x_{T+1} \approx x_1 - \eta \sum_{t=1}^T z_t \] and thus, by definition, 
\[ \Phi_{T+1} \gtrsim w^Tx_1 - \psi(w) - \eta \sum_{t=1}^T w^Tz_t. \]
On the other hand, 
\begin{align*}
\Phi_{t+1} &\approx \Phi_t + w_t^T(x_{t+1} - x_t) \\
           &= \Phi_t - \eta w_t^T z_t,
\end{align*}
so we have 
\[ \Phi_{T+1} \approx \Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t. \]
Combining this with the preceding bound on $\Phi_{T+1}$ yields
$\Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t \gtrsim w^Tx_1 - \psi(w) - \eta \sum_{t=1}^T w^Tz_t$, 
or
\[ \sum_{t=1}^T w_t^Tz_t \lesssim \frac{\Phi_1 + \psi(w) - w^Tx_1}{\eta} + \sum_{t=1}^T w^Tz_t, \]
which gives us an even better regret bound than we had sought (the $\sum_{t=1}^T |w|^T|z_t|$ term 
will show up when we track the errors made in each of the approximations).

\paragraph{Formal statement.} To turn these approximations into a rigorous bound, we will 
need $f$ and $\psi$ to satisfy the following properties:
\begin{itemize}
\item Convexity of $\psi^* \circ f$: $\psi^*(f(x,\eta z)) \leq \psi^*(x) - \eta w_x^Tz$, where $w_x = \partial \psi^*(x)$.
\item Smoothness of $f$: $w^T(x-\eta z - f(x,\eta z)) \leq \eta^2 \alpha |w|^T|z|$ for all $w$.
\end{itemize}
(For some intuition, any $f$ satisfying the two properties above 
must satisfy $f(x,0) = x$ and $f(x,\eta z) = x-\eta z + O(\eta^2)$.)
\begin{theorem}
\label{thm:main}
Suppose that $\psi^*$ and $f$ satisfy the two properties above. Then, 
for $\eta = \frac{\epsilon}{\alpha}$, we have the bound
\[ \sum_{t=1}^T w_t^Tz_t \leq \alpha \frac{\Phi_1 + \psi(w) - w^Tx_1}{\epsilon} + \epsilon \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t. \]
\end{theorem}
\begin{proof}
By the first condition, we have
\begin{align*}
\Phi^{t+1} &= \psi^*(x_{t+1}) \\
 &= \psi^*(f(x_t, \eta z_t)) \\
 &\leq \psi^*(x_t) - \eta \left(\frac{\partial \psi^*}{\partial x_t}\right)^T z_t \\
 &= \Phi_t - \eta w_t^T z_t,
\end{align*}
which by induction implies that
\[ \Phi_{T+1} \leq \Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t. \]
By the second condition, we have
\begin{align*}
w^Tx_{t+1} &= w^Tf(x, \eta z) \\
 &\geq w^Tx_t - \eta w^Tz_t - \eta^2 \alpha |w|^T|z_t|,
\end{align*}
which by induction implies that
\[ w^Tx_{T+1} \geq w^Tx_1 - \eta^2\alpha \sum_{t=1}^T |w|^T|z_t| - \eta \sum_{t=1}^T w^Tz_t. \]
Combining this with the first inequality yields
\begin{align*}
\Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t &\geq \Phi_{T+1} \\
 &\geq w^Tx_{T+1} - \psi(w) \\
 &\geq w^Tx_1 - \psi(w) - \eta^2\alpha \sum_{t=1}^T |w|^T|z_t| - \eta \sum_{t=1}^T w^Tz_t.
\end{align*}
Re-arranging yields
\[ \sum_{t=1}^T w_t^Tz_t \leq \frac{\Phi_1 + \psi(w) - w^Tx_1}{\eta} + \eta\alpha \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t, \]
which gives the desired result.
\end{proof}
\section{Examples}
\paragraph{Multiplicative updates.} If we let $\sC_P$ be the simplex and $\sC_D$ be the 
$l^{\infty}$ ball of radius $r$, then a standard algorithm for minimizing regret is the 
\emph{multiplicative weights update method}. We recover this by letting $f(x,z) = x + \log(1-z)$ 
and $\psi(w) = \sum_{i=1}^n w_i\log w_i$. Then $\psi^*(x) = \log\left(\sum_{i=1}^n \exp(x_i)\right)$, 
and $w_{t+1,i} = \exp(x_{t+1,i})/\sum_{j=1}^n \exp(x_{t+1,j})$.
\begin{align*}
\psi^*(f(x,\eta z)) &= \log\left(\sum_{i=1}^n \exp(x_i + \log(1-\eta z_i))\right) \\
 &= \log\left(\sum_{i=1}^n \exp(x_i)(1-\eta z_i)\right) \\
 &= \log\left(\sum_{i=1}^n \exp(x_i) - \eta \sum_{i=1}^n \exp(x_i) z_i\right) \\
 &\leq \log\left(\sum_{i=1}^n \exp(x_i)\right) - \eta \frac{\sum_{i=1}^n \exp(x_i) z_i}{\sum_{i=1}^n \exp(x_i)} \\
 &= \psi^*(x) - \eta w^Tz,
\end{align*}
so condition 1 is satisfied.

Also, using the fact that $\log(1-\eta z_i) \geq -\eta z_i - \eta^2 z_i^2$, we have
\begin{align*}
w^T(x-\eta z - f(x, \eta z)) &= w^T(x-\eta z - (x + \log(1-\eta z))) \\
 &\leq \eta^2 \sum_{i=1}^n w_iz_i^2 \\
 &\leq \eta^2 r \sum_{i=1}^n |w_i||z_i|,
\end{align*}
where in the last step we used the fact that $|z_i| \leq r$. These updates 
thus fulfill the conditions of Theorem~\ref{thm:main} for $\alpha$ equal to 
$r$.

\end{document}

### 2014-01-05 lp/framework.tex
writeup with lp, sum of pointed sets

@@
 We are interested in the following learning problem: our experts 
-consist of vectors $w \in \sC_P$, where $\sC_P$ is some convex 
+consist of vectors $w \in \sH$, where $\sH$ is some convex 
 set; similarly, our adversary can play any move $z_t$ with 
-$z_t \in \sC_D$. A typical example is $\sC_P$ is the $n$-dimensional 
-simplex and $\sC_D$ is the unit $l^{\infty}$ ball. Our end goal is 
+$z_t \in \sA$. A typical example is $\sH$ is the $n$-dimensional 
+simplex and $\sA$ is the unit $l^{\infty}$ ball. Our end goal is 
 to obtain a regret bound of the following type:
 \[ \sum_{t=1}^T w_t^Tz_t \leq \frac{\rho}{\epsilon} + \epsilon \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t. \]
-If we let $\sigma = \sup_{w \in \sC_P, z \in \sC_D} |w|^T|z|$, then the regret is bounded by
+If we let $\sigma = \sup_{w \in \sH, z \in \sA} |w|^T|z|$, then the regret is bounded by
 $\frac{\rho}{\epsilon} + \epsilon \sigma T$, which is equal to $2\sqrt{\rho \sigma T}$ for 
@@
 in most online learning settings. However, our bound can lead to stronger results in some 
-situations.
+situations, for instance if $\sA$ is very asymmetrical about the origin, in which case 
+$|w|^T|z_t|$ will tend to be close to $w^Tz_t$ and we get bound
+\[ \sum_{t=1}^T w_t^Tz_t \leq \frac{\rho}{\epsilon} + (1+\epsilon)\sum_{t=1}^T w^Tz_t, \]
+which is potentially much better, for instance if there is an expert that achieves 
+very small loss.
 
@@
 x_{t+1} &= f(x_t, \eta z_t) \\
-w_{t+1} &= \argmax_{w \in \sC_P} w^Tx_{t+1} - \psi(w) \\
+w_{t+1} &= \argmax_{w \in \sH} w^Tx_{t+1} - \psi(w) \\
 \Phi_{t+1} &= w_{t+1}^Tx_{t+1} - \psi(w_{t+1})  \\ &= \psi^*(x_{t+1}).
@@
 \begin{itemize}
-\item Convexity of $\psi^* \circ f$: $\psi^*(f(x,\eta z)) \leq \psi^*(x) - \eta w_x^Tz$, where $w_x = \partial \psi^*(x)$.
+\item Concavity of $\psi^* \circ f$: $\psi^*(f(x,\eta z)) \leq \psi^*(x) - \eta w_x^Tz$, where $w_x = \partial \psi^*(x)$.
 \item Smoothness of $f$: $w^T(x-\eta z - f(x,\eta z)) \leq \eta^2 \alpha |w|^T|z|$ for all $w$.
@@
 \end{proof}
+
+\paragraph{Relationship to mirror descent.} Mirror descent can be cast in the same framework as above, 
+and corresponds to the exact additive update $x_{t+1} = x_t - \eta z_t$. Thus while our algorithm 
+makes sure that $\psi^*$ decreases at a rate of $w_t^Tz_t$ and then bounds the error between 
+$x_{t+1}$ and $x_t - \eta z_t$, in contrast mirror descent sets the error between $x_{t+1}$ and 
+$x_t - \eta z_t$ to zero, and then relies on the smoothness of $\psi^*$ to bound the error between 
+$\psi^*(x-\eta z)$ and $\psi^*(x) - \eta w_x^Tz$.
 \section{Examples}
-\paragraph{Multiplicative updates.} If we let $\sC_P$ be the simplex and $\sC_D$ be the 
+\paragraph{Multiplicative updates.} If we let $\sH$ be the simplex and $\sA$ be the 
 $l^{\infty}$ ball of radius $r$, then a standard algorithm for minimizing regret is the 
@@
 
+\paragraph{Sum of pointed convex sets.} Suppose that our hypothesis class 
+$\sH \subseteq \bR^n$ can be decomposed as $\sH = \sH_+ + \sH_-$, where $\sH_+$ consists 
+only of non-negative vectors and $\sH_-$ consists only of non-positive vectors. 
+We can translate this into a problem n $\bR^{2n}$, by letting $\sH' \eqdef \sH_+ \times (-\sH_-)$ 
+and $\sA' \eqdef \{(z, -z) \mid z \in \sA\}$. Then $(w')^T(z') = w^Tz$. Now 
+let $(w_+, w_-)$ be the decomposition of $w'$ into its two components. We will 
+define
+\[ \psi(w') = \left\{ \begin{array}{ccl} \sum_{i=1}^n \log((w_+)_i + (w_-)_i) & : & w' \in \sH \\ \infty & : & \mathrm{else}. \end{array} \right. \]
+Now let 
+\[ R \eqdef \sup\limits_{w_+ \in \sH_+, w_- \in \sH_-, z \in \sA}  |w_+^Tz| + |w_-^Tz|, \]
+and consider the update $x_{t+1} = x_t - \eta z'_t - 2\eta^2 R |z'_t|$. Then we have
+\begin{align*}
+\psi^*(x_{t+1}) &= \psi^*(x_t) - \eta (w'_t)^T(z'_t) - 2\eta^2R (w'_t)^T|z'_t| + D_{\psi^*}(x_{t+1} \| x_t) \\
+ &= \psi^*(x_t) - \eta (w'_t)^Tz_t - 2\eta^2R (w'_t)^T|z'_t| + \frac{1}{2}((w'_t)^T|x_{t+1}-x_t|)^2 \\
+ &= \psi^*(x_t) - \eta (w'_t)^Tz_t - 2\eta^2R (w'_t)^T|z'_t| + \frac{1}{2}(\eta (w'_t)^T|z'_t| + 2\eta^2R (w'_t)^T|z'_t|)^2 \\
+ &= \psi^*(x_t) - \eta (w'_t)^Tz_t
+\end{align*}
+as long as $2\eta R \leq 1$, or $\eta \leq \frac{1}{2R}$. Thus the first condition is 
+satisfied. Also, by construction we have 
+\[ w^T(x_t - \eta z_t - x_{t+1}) = 2\eta^2 R w^T|z'_t|, \]
+so that the second condition is satisfied for $\alpha = 2R$. Combining these, and 
+setting $x_1 = 0$, we obtain the bound
+\[ \sum_{t=1}^T w_t^Tz_t \leq \frac{2R[\psi(w) - \inf_{\tilde{w}} \psi(\tilde{w})]}{\epsilon} + \epsilon \sum_{t=1}^T (w_+ + w_-)^T|z_t| + \sum_{t=1}^T w^Tz_t. \]
+This is a quite general result, but suffers from the fact that $\psi(w)$ can grow very large. 
+Finding ways to deal with this would thus be very interesting.
+
+\paragraph{Inflated $l^p$ norm.} By abuse of notation, let $\alpha$ denote the vector 
+whose coordinates are all equal to $\alpha$. Let $\sH = B_{\alpha}^p(\rho)$, where 
+\[ B_{\alpha}^p(\rho) \eqdef \{w \mid w \geq \alpha, \|w-\alpha\|_p \leq \rho \}. \]
+Note that this is the positive $l^p$ ball of radius $\rho$, translated by $\alpha$. 
+(We will show below that we can embed the unconstrained $l^p$ ball into $B_{\alpha}^p(\rho)$ 
+at the cost of increasing the dimension by a factor of 2, implying that 
+$B_{\alpha}^p(\rho)$ is a rich hypothesis class.)
+
+We will let $\psi(w) = \left\{ \begin{array}{ccl} \frac{1}{2(p-1)}\|w-\alpha\|_p^2 & : & w \geq \alpha \\ \infty & : & w \not\geq \alpha. \end{array} \right.$. 
+Then 
+\[ \psi^*(x) = \alpha \sum_{i=1}^n x_i + \frac{1}{2(q-1)}\|x_+\|_q^2, \]
+where $x_+$ is the positive part of $x$ and $\frac{1}{p} + \frac{1}{q} = 1$.
+
+Now suppose we use the update $x_{t+1,i} = x_{t,i} - \eta z_{t,i} - 2\eta^2 \frac{\|z\|_q^2}{\|z\|_1} |z_{t,i}| / \alpha$. 
+Note that the Bregman divergence $D_{\psi^*}(x_{t+1} \| x_t)$ is bounded above by the Bregman divergence 
+of $\frac{1}{2(q-1)}\|x\|_q^2$ (since it can be obtained from $\psi^*$ by adding a convex function), 
+which is in turn bounded above by $\frac{1}{2}\|x_{t+1}-x_t\|_q^2$. Using these facts, together with the fact that 
+$w \geq \alpha$, we have
+\begin{align*}
+\psi^*(x_{t+1}) &\leq \psi^*(x_t) - \eta w_t^Tz_t - \frac{2\eta^2\|z\|_q^2}{\alpha \|z\|_1} w_t^T|z_t| + D_{\psi^*}(x_{t+1} \| x_t) \\
+ &\leq \psi^*(x_t) - \eta w_t^Tz_t - \frac{2\eta^2\|z\|_q^2}{\alpha \|z\|_1} w_t^T|z_t| + \frac{1}{2}\left\|\eta z_t + \frac{2\eta^2\|z\|_q^2}{\alpha \|z\|_1} |z_t|\right\|_q^2 \\
+ &\leq \psi^*(x_t) - \eta w_t^Tz_t - 2\eta^2 \|z\|_q^2 + \frac{1}{2} \left(\eta \|z_t\|_q + \frac{2\eta^2\|z_t\|_q^3}{\alpha \|z_t\|_1}\right)^2 \\
+ &\leq \psi^*(x_t) - \eta w_t^Tz_t - 2\eta^2 \|z\|_q^2 + \frac{1}{2} \left(\eta \|z_t\|_q + 2\eta^2 \|z_t\|_{q}^2 / \alpha\right)^2 \\
+ &\leq \psi^*(x_t) - \eta w_t^Tz_t
+\end{align*}
+as long as $2\eta \|z_t\|_q/\alpha \leq 1$, which holds as long as $\eta \leq \frac{\alpha}{2r_q}$. In addition, we have that
+\begin{align*}
+w^T(x_{t} - \eta z_t - x_{t+1}) &= -\frac{2\eta^2\|z_t\|_q^2}{\alpha\|z_t\|_1}w^T|z_t| \\
+ &\leq \eta^2 \frac{2\|z_t\|_q}{\alpha} |w|^T|z_t| \\
+ &\leq \eta^2 \frac{2r_q}{\alpha} |w|^T|z_t|.
+\end{align*}
+So, both conditions of Theorem~\ref{thm:main} are satisfied. 
+If we set $x_1 = 0$, then $\psi^*(x_1) = 0$ as well.
+Therefore, for $\eta = \frac{\epsilon \alpha}{2r_q}$, we obtain a regret bound of 
+\[ \frac{2r_q\psi(w)}{\epsilon \alpha} + \epsilon \sum_{t=1}^T |w|^T|z_t|. \]
+Note that the regret gets smaller as $\alpha$ grows larger. This might seem too good 
+to be true, until we notice the following: if $\alpha$ is very large, then all 
+vectors with a small value of $\|w-\alpha\|_p$ will point in approximately the same 
+direction, and thus get similar (multiplicative) values for the regret.
+\paragraph{Embedding the $l^p$ ball into $B_{\alpha}^p$.} We now show how to turn problems 
+involving the $l^p$ ball into problems involving $B_{\alpha}^p$. Given a vector 
+$w \in \bR^n$, we can create a vector $w' \in \bR^{2n}$ via the embedding 
+$w \mapsto \left[ \begin{array}{c} \alpha + \max(w, 0) \\ \alpha - \min(w,0) \end{array} \right]$. 
+This embedding satisfies the property that $\|w\|_p = \|w'-\alpha\|_p$. Also, $w^Tz = w'^T\left[ \begin{array}{c} z \\ -z \end{array} \right]$. 
+If we plug into the regret bound above, the only difference is that the value of $|w|^T|z_t|$ increases, yielding 
+a regret of
+\[ \frac{r_q \|w\|_p^2}{\epsilon \alpha (p-1)} + \epsilon \sum_{t=1}^T \left(|w|^T|z_t| + 2\alpha \|z_t\|_1\right). \]
+Unfortunately, this final bound is quite bad, because the $\|z_t\|_1$ term grows with $T$ and depends on the 
+$l^1$ norm, which is the biggest of all the $l^p$ norms.
 \end{document}

### 2014-01-16 lp/adaptive-regularization.tex
added first half of adaptive regularization writeup

\documentclass[11pt]{article}
\usepackage{import}
\input{../icml/latex-defs.tex}
\DeclareMathOperator{\Regret}{Regret}
\usepackage{fullpage}

\title{Exponentiated Gradient with Adaptive Regularization}
\author{Jacob Steinhardt}
\begin{document}
\maketitle
\section{Introduction}
Standard online learning algorithms obtain regret bounds that depend on the 
norm of the gradients $z_t$; for instance, exponentiated gradient has a regret bound 
of the form
\[ \Regret \leq \frac{\log(n)}{\eta} + \eta \sum_{t=1}^T \|z_t\|_{\infty}^2. \]
By using an analysis based on local norms, we can improve this to a bound
that also depends on the learner $w_t$:
\[ \Regret \leq \frac{\log(n)}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_{t,i} z_{t,i}^2. \]
In this paper, we present a variant on the exponentiated gradient that instead 
obtains a bound depending on the competitor $w$:
\[ \Regret(w) \leq \frac{\log(n)}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_{i}z_{t,i}^2. \]
This bound yields improvements when a component $i$ of the gradient is 
very noisy but has mean (across time) equal to zero. In this case 
$w_{t,i}z_{t,i}^2$ may be large but we can set $w_i$ to zero and thus $w_iz_{t,i}^2$ 
will be zero.

Our algorithm has close connections to the multiplicative weights update method, 
which is a variant of exponentiated gradient that is sometimes used in the theory 
community to achieve better asymptotic loss bounds. We present a slight variant 
of this algorithm together with a novel analysis based on 
\emph{adaptively regularized mirror descent}, which clarifies the connection between 
exponentiated gradient and multiplicative updates and also explains why 
multiplicative updates sometimes yield better performance 
(because of the adaptive nature of the regularizer, versus a fixed regularizer for exponentiated gradient).

\section{Summary of Algorithms}
In this paper we compare three similar algorithms: exponentiated gradient, 
multiplicative updated, and adaptive exponentiated gradient. They are presented 
in Table~\ref{fig:alg1} in the case that all weights are constrained to the 
simplex. We later generalize to the unconstrained case but present the constrained 
versions here in order to contrast the three algorithms.

\begin{table}
\begin{tabular}{|c|c|c|} \hline
        Algorithm & Update & Regret \\ \hline
        Exponentiated gradient & $w_{t+1,i} \propto w_{t,i}\exp(-\eta z_{t,i})$ & $\frac{\log(n)}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_{t,i}z_{t,i}^2$ \\ \hline
        Multiplicative updates & $w_{t+1,i} \propto w_{t,i}(1-\eta z_{t,i})$ & $\frac{\log(n)}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_{i}z_{t,i}^2$ \\ \hline
        Adaptive exponentiated gradient & $w_{t+1,i} \propto w_{t,i}\exp(-\eta z_{t,i} - \eta^2 z_{t,i}^2)$ & $\frac{\log(n)}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_iz_{t,i}^2$ \\ \hline
\end{tabular}
\caption{Comparison of exponentiated gradient, multiplicative updates, and adaptive exponentiated gradient algorithms on the simplex.}
\label{fig:alg1}
\end{table}

Each algorithm works by maintaining a weight vector $w_t \in \Delta_n$ and 
performing updates of the form $w_{t+1, i} \propto w_{t,i} \cdot f(z_{t,i})$, 
for some function $f$. The three cases are:
\begin{enumerate}
        \item \textbf{Exponentiated gradient:} $w_{t+1, i} \propto w_{t,i} \exp(-\eta z_{t,i})$ ($f(z) = \exp(-\eta z)$)
        \item \textbf{Multiplicative updates:} $w_{t+1, i} \propto w_{t,i} (1-\eta z_{t,i})$ ($f(z) = 1-\eta z$)
        \item \textbf{Adaptive exponentiated gradient:} $w_{t+1, i} \propto w_{t,i} \exp(-\eta z_{t,i}-\eta^2 z_{t,i}^2)$ ($f(z) = \exp(-\eta z - \eta^2 z^2)$)
\end{enumerate}
The first of these achieves regret that depends on the learner ($w_{t,i}z_{t,i}^2$) while the last two achieve regret that depends on the expert ($w_iz_{t,i}^2$).
Note that adaptive exponentiated gradient will tend to downweight any experts with high variance.

\section{Simple Example}
Suppose that there are $N$ experts, such that one has loss $-\epsilon$ always and the others 
each have loss alternating between $1$ and $-1$. 
Figure~\ref{fig:simple} shows the performance of all three 
algorithms in this setting, for $\epsilon = 0.1$, $T = 1000$, and $\eta = \sqrt{\frac{\log(N)}{T}}$. Note that the 
total regret of exponentiated gradient is $58$ in contrast to a regret of $39$ for adaptive 
exponentiated gradient. This is because the adaptive algorithm downweights all the high-variance 
experts due to the $\eta^2z_{t,i}^2$ term in the update.

We can ask what happens in the opposite case: when the optimal expert is high-variance and all 
the rest are low variance (for instance, the first expert has loss alternating bewteen $1-\epsilon$ 
and $-1-\epsilon$ while the rest have loss $0$ always). In this case being adaptive hurts us because 
we downweight the high-variance strategy even though it is optimal. Our regret is $88$ in contrast 
to the regret of $58$ of exponentiated gradient. In general, we should expect to do well when the optimal 
expert has low variance and poorly when the optimal expert has high variance. This is reflected 
in the regret bounds given in Table~\ref{fig:alg1}.

\begin{figure}
        \begin{center}
                \includegraphics[width=0.4\textwidth]{simple.pdf} \includegraphics[width=0.4\textwidth]{tricky.pdf}
        \end{center}
        \caption{Regret as a function of time for exponentiated gradient (blue), multiplicative updates (green), and adaptive exponentiated gradient (red). Left: 
        the suboptimal strategies have high variance and the optimal strategy has low variance. Right: the suboptimal strategies have low variance and the 
        optimal strategy has high variance.}
        \label{fig:simple}
\end{figure}

\section{General Algorithms and Theoretical Analysis}
We now present the unconstrained versions of our three algorithms, together 
with an analysis based on adaptive regularizers. In each case, the weight 
vector $w$ is necessarily non-negative. We can deal with this constraint by 
maintaining two weight vectors $w_+$ and $w_-$ and predicting with $w_+-w_-$ 
(this is a standard idea, going back at least to the original exponentiated 
gradient paper by Kivinen and Warmuth). This is equivalent to replacing the 
gradient $z$ with the expanded vector $\left[ \begin{array}{cc} z \\ -z \end{array} \right]$. 
For simplicity of exposition, we will simply assume that $w$ is constrained 
to be non-negative in our algorithms and analysis.

Each algorithm can be interpreted as (adaptive) mirror descent with an appropriate 
regularizer. The regularizers are as follows:
\begin{enumerate}
        \item \textbf{Exponentiated gradient:} $\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1]$.
        \item \textbf{Multiplicative updates:} $\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1-\sum_{t=1}^T (\eta z_{t,i} + \log(1-\eta z_{t,i}))]$.
        \item \textbf{Adaptive exponentiated gradient:} $\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1+\sum_{t=1}^T \eta^2 z_{t,i}^2 w_i]$.
\end{enumerate}
The parameter $\lambda$ is a hyperparameter controlling the size of $w$. Typically it should be on the order of $\frac{1}{n}$ 
(as we will see more explicitly when we prove a formal regret bound). Note that in the limit of small $\eta$, 
$-\eta z_{t,i} - \log(1-\eta z_{t,i}) \approx \frac{1}{2}\eta z_{t,i}^2$, so that 
the latter two regularizers are in fact very similar. The mirror descent updates corresponding 
to each of the above regularizers are:
\begin{enumerate}
        \item \textbf{Exponentiated gradient:} $w_{t,i} = \lambda \cdot \exp\left(-\eta \sum_{t=1}^T z_{t,i}\right)$
        \item \textbf{Multiplicative updates:} $w_{t,i} = \lambda \cdot \prod_{t=1}^T (1- \eta z_{t,i})$
        \item \textbf{Adaptive exponentiated gradient:} $w_{t,i} = \lambda \cdot \exp\left(- \sum_{t=1}^T \eta z_{t,i} + \eta^2 z_{t,i}^2\right)$
\end{enumerate}
Finally, we base our regret bounds on the following standard bound on the regret of mirror descent:
\begin{equation}
        \label{eqn:bound}
        \Regret(w) \leq \psi_T(w) + \sum_{t=1}^T D_{\psi_t^*}(\theta_{t+1} \| \theta_t).
\end{equation}
Our idea is to pick $\psi_{t+1}-\psi_t$ to cancel out the Bregman divergence term. The intuition is 
that we are absorbing all of our loss into the regularizer, therefore giving us a bound that depends 
on $w$ rather than $w_t$. This strategy is non-standard, although has been used before, for instance in 
various papers by Francesco Orabona. Note that in all three cases, $\psi_t(w)$ differs from $\psi_0(w)$ 
only in a linear term, and hence $D_{\psi_t^*} \equiv D_{\psi_0^*}$, and in particular 
$D_{\psi_t^*}(\theta_{t+1} \| \theta_t) \leq \sum_{i=1}^n w_{t,i}z_{t,i}^2$ by analysis via a local norm.

%%% We are interested in the following learning problem: our experts 
%%% consist of vectors $w \in \sH$, where $\sH$ is some convex 
%%% set; similarly, our adversary can play any move $z_t$ with 
%%% $z_t \in \sA$. A typical example is $\sH$ is the $n$-dimensional 
%%% simplex and $\sA$ is the unit $l^{\infty}$ ball. Our end goal is 
%%% to obtain a regret bound of the following type:
%%% \[ \sum_{t=1}^T w_t^Tz_t \leq \frac{\lambda}{\epsilon} + \epsilon \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t. \]
%%% If we let $\sigma = \sup_{w \in \sH, z \in \sA} |w|^T|z|$, then the regret is bounded by
%%% $\frac{\lambda}{\epsilon} + \epsilon \sigma T$, which is equal to $2\sqrt{\lambda \sigma T}$ for 
%%% $\epsilon = \sqrt{\frac{\lambda}{\sigma T}}$. This is the type of regret bound that is typical 
%%% in most online learning settings. However, our bound can lead to stronger results in some 
%%% situations, for instance if $\sA$ is very asymmetrical about the origin, in which case 
%%% $|w|^T|z_t|$ will tend to be close to $w^Tz_t$ and we get bound
%%% \[ \sum_{t=1}^T w_t^Tz_t \leq \frac{\lambda}{\epsilon} + (1+\epsilon)\sum_{t=1}^T w^Tz_t, \]
%%% which is potentially much better, for instance if there is an expert that achieves 
%%% very small loss.
%%% 
%%% We propose the following general style of algorithm for obtaining such 
%%% bounds. The key innovation is an auxiliary variable $x_t$ that lets us 
%%% relate $w_t^Tz_t$ and $w^Tz_t$. The updates are as follows:
%%% \begin{align*}
%%% x_{t+1} &= f(x_t, \eta z_t) \\
%%% w_{t+1} &= \argmax_{w \in \sH} w^Tx_{t+1} - \psi(w) \\
%%% \Phi_{t+1} &= w_{t+1}^Tx_{t+1} - \psi(w_{t+1})  \\ &= \psi^*(x_{t+1}).
%%% \end{align*}
%%% 
%%% To see why these updates are a good idea, we will make a ``physicist's approximation'' 
%%% and suppose that $f(x_t, \eta z_t) \approx x_t - \eta z_t$. Then we would have
%%% \[ x_{T+1} \approx x_1 - \eta \sum_{t=1}^T z_t \] and thus, by definition, 
%%% \[ \Phi_{T+1} \gtrsim w^Tx_1 - \psi(w) - \eta \sum_{t=1}^T w^Tz_t. \]
%%% On the other hand, 
%%% \begin{align*}
%%% \Phi_{t+1} &\approx \Phi_t + w_t^T(x_{t+1} - x_t) \\
%%%            &= \Phi_t - \eta w_t^T z_t,
%%% \end{align*}
%%% so we have 
%%% \[ \Phi_{T+1} \approx \Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t. \]
%%% Combining this with the preceding bound on $\Phi_{T+1}$ yields
%%% $\Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t \gtrsim w^Tx_1 - \psi(w) - \eta \sum_{t=1}^T w^Tz_t$, 
%%% or
%%% \[ \sum_{t=1}^T w_t^Tz_t \lesssim \frac{\Phi_1 + \psi(w) - w^Tx_1}{\eta} + \sum_{t=1}^T w^Tz_t, \]
%%% which gives us an even better regret bound than we had sought (the $\sum_{t=1}^T |w|^T|z_t|$ term 
%%% will show up when we track the errors made in each of the approximations).
%%% 
%%% \paragraph{Formal statement.} To turn these approximations into a rigorous bound, we will 
%%% need $f$ and $\psi$ to satisfy the following properties:
%%% \begin{itemize}
%%% \item Concavity of $\psi^* \circ f$: $\psi^*(f(x,\eta z)) \leq \psi^*(x) - \eta w_x^Tz$, where $w_x = \partial \psi^*(x)$.
%%% \item Smoothness of $f$: $w^T(x-\eta z - f(x,\eta z)) \leq \eta^2 \alpha |w|^T|z|$ for all $w$.
%%% \end{itemize}
%%% (For some intuition, any $f$ satisfying the two properties above 
%%% must satisfy $f(x,0) = x$ and $f(x,\eta z) = x-\eta z + O(\eta^2)$.)
%%% \begin{theorem}
%%% \label{thm:main}
%%% Suppose that $\psi^*$ and $f$ satisfy the two properties above. Then, 
%%% for $\eta = \frac{\epsilon}{\alpha}$, we have the bound
%%% \[ \sum_{t=1}^T w_t^Tz_t \leq \alpha \frac{\Phi_1 + \psi(w) - w^Tx_1}{\epsilon} + \epsilon \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t. \]
%%% \end{theorem}
%%% \begin{proof}
%%% By the first condition, we have
%%% \begin{align*}
%%% \Phi^{t+1} &= \psi^*(x_{t+1}) \\
%%%  &= \psi^*(f(x_t, \eta z_t)) \\
%%%  &\leq \psi^*(x_t) - \eta \left(\frac{\partial \psi^*}{\partial x_t}\right)^T z_t \\
%%%  &= \Phi_t - \eta w_t^T z_t,
%%% \end{align*}
%%% which by induction implies that
%%% \[ \Phi_{T+1} \leq \Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t. \]
%%% By the second condition, we have
%%% \begin{align*}
%%% w^Tx_{t+1} &= w^Tf(x, \eta z) \\
%%%  &\geq w^Tx_t - \eta w^Tz_t - \eta^2 \alpha |w|^T|z_t|,
%%% \end{align*}
%%% which by induction implies that
%%% \[ w^Tx_{T+1} \geq w^Tx_1 - \eta^2\alpha \sum_{t=1}^T |w|^T|z_t| - \eta \sum_{t=1}^T w^Tz_t. \]
%%% Combining this with the first inequality yields
%%% \begin{align*}
%%% \Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t &\geq \Phi_{T+1} \\
%%%  &\geq w^Tx_{T+1} - \psi(w) \\
%%%  &\geq w^Tx_1 - \psi(w) - \eta^2\alpha \sum_{t=1}^T |w|^T|z_t| - \eta \sum_{t=1}^T w^Tz_t.
%%% \end{align*}
%%% Re-arranging yields
%%% \[ \sum_{t=1}^T w_t^Tz_t \leq \frac{\Phi_1 + \psi(w) - w^Tx_1}{\eta} + \eta\alpha \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t, \]
%%% which gives the desired result.
%%% \end{proof}
%%% 
%%% \paragraph{Relationship to mirror descent.} Mirror descent can be cast in the same framework as above, 
%%% and corresponds to the exact additive update $x_{t+1} = x_t - \eta z_t$. Thus while our algorithm 
%%% makes sure that $\psi^*$ decreases at a rate of $w_t^Tz_t$ and then bounds the error between 
%%% $x_{t+1}$ and $x_t - \eta z_t$, in contrast mirror descent sets the error between $x_{t+1}$ and 
%%% $x_t - \eta z_t$ to zero, and then relies on the smoothness of $\psi^*$ to bound the error between 
%%% $\psi^*(x-\eta z)$ and $\psi^*(x) - \eta w_x^Tz$.
%%% \section{Examples}
%%% \paragraph{Multiplicative updates.} If we let $\sH$ be the simplex and $\sA$ be the 
%%% $l^{\infty}$ ball of radius $r$, then a standard algorithm for minimizing regret is the 
%%% \emph{multiplicative weights update method}. We recover this by letting $f(x,z) = x + \log(1-z)$ 
%%% and $\psi(w) = \sum_{i=1}^n w_i\log w_i$. Then $\psi^*(x) = \log\left(\sum_{i=1}^n \exp(x_i)\right)$, 
%%% and $w_{t+1,i} = \exp(x_{t+1,i})/\sum_{j=1}^n \exp(x_{t+1,j})$.
%%% \begin{align*}
%%% \psi^*(f(x,\eta z)) &= \log\left(\sum_{i=1}^n \exp(x_i + \log(1-\eta z_i))\right) \\
%%%  &= \log\left(\sum_{i=1}^n \exp(x_i)(1-\eta z_i)\right) \\
%%%  &= \log\left(\sum_{i=1}^n \exp(x_i) - \eta \sum_{i=1}^n \exp(x_i) z_i\right) \\
%%%  &\leq \log\left(\sum_{i=1}^n \exp(x_i)\right) - \eta \frac{\sum_{i=1}^n \exp(x_i) z_i}{\sum_{i=1}^n \exp(x_i)} \\
%%%  &= \psi^*(x) - \eta w^Tz,
%%% \end{align*}
%%% so condition 1 is satisfied.
%%% 
%%% Also, using the fact that $\log(1-\eta z_i) \geq -\eta z_i - \eta^2 z_i^2$, we have
%%% \begin{align*}
%%% w^T(x-\eta z - f(x, \eta z)) &= w^T(x-\eta z - (x + \log(1-\eta z))) \\
%%%  &\leq \eta^2 \sum_{i=1}^n w_iz_i^2 \\
%%%  &\leq \eta^2 r \sum_{i=1}^n |w_i||z_i|,
%%% \end{align*}
%%% where in the last step we used the fact that $|z_i| \leq r$. These updates 
%%% thus fulfill the conditions of Theorem~\ref{thm:main} for $\alpha$ equal to 
%%% $r$.
%%% 
%%% \paragraph{Sum of pointed convex sets.} Suppose that our hypothesis class 
%%% $\sH \subseteq \bR^n$ can be decomposed as $\sH = \sH_+ + \sH_-$, where $\sH_+$ consists 
%%% only of non-negative vectors and $\sH_-$ consists only of non-positive vectors. 
%%% We can translate this into a problem n $\bR^{2n}$, by letting $\sH' \eqdef \sH_+ \times (-\sH_-)$ 
%%% and $\sA' \eqdef \{(z, -z) \mid z \in \sA\}$. Then $(w')^T(z') = w^Tz$. Now 
%%% let $(w_+, w_-)$ be the decomposition of $w'$ into its two components. We will 
%%% define
%%% \[ \psi(w') = \left\{ \begin{array}{ccl} \sum_{i=1}^n \log((w_+)_i + (w_-)_i) & : & w' \in \sH \\ \infty & : & \mathrm{else}. \end{array} \right. \]
%%% Now let 
%%% \[ R \eqdef \sup\limits_{w_+ \in \sH_+, w_- \in \sH_-, z \in \sA}  |w_+^Tz| + |w_-^Tz|, \]
%%% and consider the update $x_{t+1} = x_t - \eta z'_t - 2\eta^2 R |z'_t|$. Then we have
%%% \begin{align*}
%%% \psi^*(x_{t+1}) &= \psi^*(x_t) - \eta (w'_t)^T(z'_t) - 2\eta^2R (w'_t)^T|z'_t| + D_{\psi^*}(x_{t+1} \| x_t) \\
%%%  &= \psi^*(x_t) - \eta (w'_t)^Tz_t - 2\eta^2R (w'_t)^T|z'_t| + \frac{1}{2}((w'_t)^T|x_{t+1}-x_t|)^2 \\
%%%  &= \psi^*(x_t) - \eta (w'_t)^Tz_t - 2\eta^2R (w'_t)^T|z'_t| + \frac{1}{2}(\eta (w'_t)^T|z'_t| + 2\eta^2R (w'_t)^T|z'_t|)^2 \\
%%%  &= \psi^*(x_t) - \eta (w'_t)^Tz_t
%%% \end{align*}
%%% as long as $2\eta R \leq 1$, or $\eta \leq \frac{1}{2R}$. Thus the first condition is 
%%% satisfied. Also, by construction we have 
%%% \[ w^T(x_t - \eta z_t - x_{t+1}) = 2\eta^2 R w^T|z'_t|, \]
%%% so that the second condition is satisfied for $\alpha = 2R$. Combining these, and 
%%% setting $x_1 = 0$, we obtain the bound
%%% \[ \sum_{t=1}^T w_t^Tz_t \leq \frac{2R[\psi(w) - \inf_{\tilde{w}} \psi(\tilde{w})]}{\epsilon} + \epsilon \sum_{t=1}^T (w_+ + w_-)^T|z_t| + \sum_{t=1}^T w^Tz_t. \]
%%% This is a quite general result, but suffers from the fact that $\psi(w)$ can grow very large. 
%%% Finding ways to deal with this would thus be very interesting.
%%% 
%%% \paragraph{Inflated $l^p$ norm.} By abuse of notation, let $\alpha$ denote the vector 
%%% whose coordinates are all equal to $\alpha$. Let $\sH = B_{\alpha}^p(\lambda)$, where 
%%% \[ B_{\alpha}^p(\lambda) \eqdef \{w \mid w \geq \alpha, \|w-\alpha\|_p \leq \lambda \}. \]
%%% Note that this is the positive $l^p$ ball of radius $\lambda$, translated by $\alpha$. 
%%% (We will show below that we can embed the unconstrained $l^p$ ball into $B_{\alpha}^p(\lambda)$ 
%%% at the cost of increasing the dimension by a factor of 2, implying that 
%%% $B_{\alpha}^p(\lambda)$ is a rich hypothesis class.) 
%%% 
%%% We will let $\psi(w) = \left\{ \begin{array}{ccl} \frac{1}{2(p-1)}\|w-\alpha\|_p^2 & : & w \geq \alpha \\ \infty & : & w \not\geq \alpha. \end{array} \right.$. 
%%% Then 
%%% \[ \psi^*(x) = \alpha \sum_{i=1}^n x_i + \frac{1}{2(q-1)}\|x_+\|_q^2, \]
%%% where $x_+$ is the positive part of $x$ and $\frac{1}{p} + \frac{1}{q} = 1$.
%%% 
%%% Now suppose we use the update $x_{t+1,i} = x_{t,i} - \eta z_{t,i} - 2\eta^2 \frac{\|z\|_q^2}{\|z\|_1} |z_{t,i}| / \alpha$. 
%%% Note that the Bregman divergence $D_{\psi^*}(x_{t+1} \| x_t)$ is bounded above by the Bregman divergence 
%%% of $\frac{1}{2(q-1)}\|x\|_q^2$ (since it can be obtained from $\psi^*$ by adding a convex function), 
%%% which is in turn bounded above by $\frac{1}{2}\|x_{t+1}-x_t\|_q^2$. Using these facts, together with the fact that 
%%% $w \geq \alpha$, we have
%%% \begin{align*}
%%% \psi^*(x_{t+1}) &\leq \psi^*(x_t) - \eta w_t^Tz_t - \frac{2\eta^2\|z\|_q^2}{\alpha \|z\|_1} w_t^T|z_t| + D_{\psi^*}(x_{t+1} \| x_t) \\
%%%  &\leq \psi^*(x_t) - \eta w_t^Tz_t - \frac{2\eta^2\|z\|_q^2}{\alpha \|z\|_1} w_t^T|z_t| + \frac{1}{2}\left\|\eta z_t + \frac{2\eta^2\|z\|_q^2}{\alpha \|z\|_1} |z_t|\right\|_q^2 \\
%%%  &\leq \psi^*(x_t) - \eta w_t^Tz_t - 2\eta^2 \|z\|_q^2 + \frac{1}{2} \left(\eta \|z_t\|_q + \frac{2\eta^2\|z_t\|_q^3}{\alpha \|z_t\|_1}\right)^2 \\
%%%  &\leq \psi^*(x_t) - \eta w_t^Tz_t - 2\eta^2 \|z\|_q^2 + \frac{1}{2} \left(\eta \|z_t\|_q + 2\eta^2 \|z_t\|_{q}^2 / \alpha\right)^2 \\
%%%  &\leq \psi^*(x_t) - \eta w_t^Tz_t
%%% \end{align*}
%%% as long as $2\eta \|z_t\|_q/\alpha \leq 1$, which holds as long as $\eta \leq \frac{\alpha}{2r_q}$. In addition, we have that
%%% \begin{align*}
%%% w^T(x_{t} - \eta z_t - x_{t+1}) &= -\frac{2\eta^2\|z_t\|_q^2}{\alpha\|z_t\|_1}w^T|z_t| \\
%%%  &\leq \eta^2 \frac{2\|z_t\|_q}{\alpha} |w|^T|z_t| \\
%%%  &\leq \eta^2 \frac{2r_q}{\alpha} |w|^T|z_t|.
%%% \end{align*}
%%% So, both conditions of Theorem~\ref{thm:main} are satisfied. 
%%% If we set $x_1 = 0$, then $\psi^*(x_1) = 0$ as well.
%%% Therefore, for $\eta = \frac{\epsilon \alpha}{2r_q}$, we obtain a regret bound of 
%%% \[ \frac{2r_q\psi(w)}{\epsilon \alpha} + \epsilon \sum_{t=1}^T |w|^T|z_t|. \]
%%% Note that the regret gets smaller as $\alpha$ grows larger. This might seem too good 
%%% to be true, until we notice the following: if $\alpha$ is very large, then all 
%%% vectors with a small value of $\|w-\alpha\|_p$ will point in approximately the same 
%%% direction, and thus get similar (multiplicative) values for the regret.
%%% \paragraph{Embedding the $l^p$ ball into $B_{\alpha}^p$.} We now show how to turn problems 
%%% involving the $l^p$ ball into problems involving $B_{\alpha}^p$. Given a vector 
%%% $w \in \bR^n$, we can create a vector $w' \in \bR^{2n}$ via the embedding 
%%% $w \mapsto \left[ \begin{array}{c} \alpha + \max(w, 0) \\ \alpha - \min(w,0) \end{array} \right]$. 
%%% This embedding satisfies the property that $\|w\|_p = \|w'-\alpha\|_p$. Also, $w^Tz = w'^T\left[ \begin{array}{c} z \\ -z \end{array} \right]$. 
%%% If we plug into the regret bound above, the only difference is that the value of $|w|^T|z_t|$ increases, yielding 
%%% a regret of
%%% \[ \frac{r_q \|w\|_p^2}{\epsilon \alpha (p-1)} + \epsilon \sum_{t=1}^T \left(|w|^T|z_t| + 2\alpha \|z_t\|_1\right). \]
%%% Unfortunately, this final bound is quite bad, because the $\|z_t\|_1$ term grows with $T$ and depends on the 
%%% $l^1$ norm, which is the biggest of all the $l^p$ norms.
\end{document}

### 2014-01-17 lp/adaptive-regularization.tex
basic draft, but need much better notation

@@
         \item \textbf{Exponentiated gradient:} $\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1]$.
-        \item \textbf{Multiplicative updates:} $\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1-\sum_{t=1}^T (\eta z_{t,i} + \log(1-\eta z_{t,i}))]$.
-        \item \textbf{Adaptive exponentiated gradient:} $\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1+\sum_{t=1}^T \eta^2 z_{t,i}^2 w_i]$.
+        \item \textbf{Multiplicative updates:} $\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1-\sum_{s=1}^t (\eta z_{s,i} + \log(1-\eta z_{s,i}))]$.
+        \item \textbf{Adaptive exponentiated gradient:} $\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1+\sum_{s=1}^t \eta^2 z_{s,i}^2 w_i]$.
 \end{enumerate}
@@
 \end{enumerate}
-Finally, we base our regret bounds on the following standard bound on the regret of mirror descent:
+Finally, we base our regret bounds on an adaptation of the mirror descent analysis. In the analysis of 
+mirror descent based on Fenchel conjugacy, we have the following inequality for any regularizer $\psi$; 
+letting $\theta_t \eqdef -\sum_{s=1}^T z_s$ and $\psi_t$ denote the regularizer at time $t$:
 \begin{equation}
-        \label{eqn:bound}
-        \Regret(w) \leq \psi_T(w) + \sum_{t=1}^T D_{\psi_t^*}(\theta_{t+1} \| \theta_t).
+        \label{eqn:bound1}
+        -\psi_T(w) - \eta \sum_{t=1}^T w^T z_t \leq \psi_0^*(0) + \sum_{t=1}^T \psi_t^*(\theta_t) - \psi_{t-1}^*(\theta_{t-1}).
 \end{equation}
-Our idea is to pick $\psi_{t+1}-\psi_t$ to cancel out the Bregman divergence term. The intuition is 
-that we are absorbing all of our loss into the regularizer, therefore giving us a bound that depends 
-on $w$ rather than $w_t$. This strategy is non-standard, although has been used before, for instance in 
-various papers by Francesco Orabona. Note that in all three cases, $\psi_t(w)$ differs from $\psi_0(w)$ 
-only in a linear term, and hence $D_{\psi_t^*} \equiv D_{\psi_0^*}$, and in particular 
-$D_{\psi_t^*}(\theta_{t+1} \| \theta_t) \leq \sum_{i=1}^n w_{t,i}z_{t,i}^2$ by analysis via a local norm.
+Usually the regularizer is fixed, in which case $\psi^*(\theta_t) - \psi^*(\theta_{t-1}) = -\eta w_t^Tz_t + D_{\psi^*}(\theta_{t+1} \| \theta_t)$. 
+Note that $w_t^Tz_t$ is the loss suffered in round $t$; the Bregman divergence term $D_{\psi^*}$ measures the increase in regret for that round. 
+In particular, re-arranging gives
+\begin{equation}
+        \label{eqn:bound2}
+        \Regret(w) = \sum_{t=1}^T w_t^Tz_t - \sum_{t=1}^T w^Tz_t \leq \frac{1}{\eta}\left[ \psi^*(0) + \psi(w) + \sum_{t=1}^T D_{\psi^*}(\theta_{t+1} \| \theta_t)\right].
+\end{equation}
+One can then use an analysis based on strong convexity of $\psi$ to bound $\psi^*$; for instance, in the case of exponentiated gradient, 
+we have the bound $D_{\psi^*}(\theta_{t+1} \| \theta_t) \leq \eta^2 \sum_{i=1}^n w_{t,i}z_{t,i}^2$, which leads to our claimed regret bound 
+for exponentiated gradient.
+
+In contrast, our analysis based on adaptive regularizers will choose $\psi_t^*$ such that $\psi_t^*(\theta_t) - \psi_{t-1}^*(\theta_{t-1}) \leq -\eta w_t^Tz_t$. 
+Assuming we succeed in picking such a $\psi_t^*$, the $\sum_{t=1}^T \psi_t^*(\theta_t) - \psi_{t-1}^*(\theta_{t-1})$ term can be replaced with 
+$-\eta \sum_{t=1}^T w_t^Tz_t$. We can then re-arrange to obtain the regret bound
+\begin{equation}
+        \label{eqn:bound3}
+        \Regret(w) = \sum_{t=1}^T w_t^Tz_t - \sum_{t=1}^T w^Tz_t \leq \frac{\psi_0^*(0) + \psi_T(w)}{\eta}.
+\end{equation}
+This is quite interesting as we have pushed the entire regret bound into an increase in the size 
+of the regularizer.
+
+We will now verify the property $\psi_t^*(\theta_t) - \psi_{t-1}^*(\theta_{t-1}) \leq -\eta w_t^Tz_t$ 
+for both multiplicative updates and adaptive EG. It will be useful to know the following fact about 
+$\psi_t^*$:
+\begin{lemma}
+        \label{lem:fenchel}
+        If $\psi(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1+s_i]$, 
+        then $\psi^*(\theta) = \lambda \sum_{i=1}^n \exp(\theta_i - s_i)$.
+\end{lemma}
+Now, we can easily verify the property for multiplicative updates:
+\begin{align}
+        \psi_t^*(\theta_t) &= \psi_t^*(\theta_{t-1}- \eta z_t) \\
+                           &= \lambda \sum_{i=1}^n \exp\left(\theta_{t-1,i} - \eta z_{t,i} + \sum_{s=1}^t \eta z_{s,i} + \log(1 - \eta z_{s,i})\right) \\
+                           &= \lambda \sum_{i=1}^n \exp\left(\theta_{t-1,i} + \log(1- \eta z_{t,i}) + \sum_{s=1}^{t-1} \eta z_{s,i} + \log(1 - \eta z_{s,i})\right) \\
+                           &= \lambda \sum_{i=1}^n \exp\left(\theta_{t-1,i} + \sum_{s=1}^{t-1} \eta z_{s,i} + \log(1 - \eta z_{s,i})\right)(1 - \eta z_{t,i}) \\
+                           &= \psi_{t-1}^*(\theta_{t-1}) - \eta \lambda \sum_{i=1}^n \exp\left(\theta_{t-1,i} + \sum_{s=1}^{t-1} \eta z_{s,i} + \log(1- \eta z_{s,i})\right) z_{t,i} \\
+                           &= \psi_{t-1}^*(\theta_{t-1}) - \eta w_t^Tz_t,
+\end{align}
+hence $\psi_t^*(\theta_t) - \psi_{t-1}^*(\theta_{t-1}) = -\eta w_t^Tz_t$, i.e. the property holds with equality. We \
+thus have the regret bound
+\begin{align}
+        \Regret(w) &\leq \frac{\psi_0*(0) + \psi_T(w)}{\eta} \\
+                   &= \frac{\lambda n + \sum_{i=1}^n w_i[\log(w_i/\lambda)-1-\sum_{t=1}^T \eta z_{s,i} + \log(1 - \eta z_{t,i})]}{\eta} \\
+                   &\leq \frac{\lambda n + \sum_{i=1}^n w_i[\log(w_i/\lambda)-1] + \eta^2 \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2}{\eta} \\
+                   &= \frac{\lambda n + \sum_{i=1}^n w_i[\log(w_i/\lambda)-1]}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_i z_{t,i}^2.
+\end{align}
+Setting $\lambda$ to $1/n$ yields a bound of
+\[ \Regret(w) \leq \frac{1+(\log(n)-1)\|w\|_1 + \sum_{i=1}^n w_i\log(w_i)}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_i z_{t,i}^2. \]
+The proof for adaptive exponentiated gradient is similar.
+\section{Experiments}
+TODO.
+
+% the following standard bound on the regret of mirror descent:
+% \begin{equation}
+%         \label{eqn:bound}
+%         \Regret(w) \leq \psi_T(w) + \sum_{t=1}^T D_{\psi_t^*}(\theta_{t+1} \| \theta_t).
+% \end{equation}
+% Our idea is to pick $\psi_{t+1}-\psi_t$ to cancel out the Bregman divergence term. The intuition is 
+% that we are absorbing all of our loss into the regularizer, therefore giving us a bound that depends 
+% on $w$ rather than $w_t$. This strategy is non-standard, although has been used before, for instance in 
+% various papers by Francesco Orabona. Note that in all three cases, $\psi_t(w)$ differs from $\psi_0(w)$ 
+% only in a linear term, and hence $D_{\psi_t^*} \equiv D_{\psi_0^*}$, and in particular 
+% $D_{\psi_t^*}(\theta_{t+1} \| \theta_t) \leq \sum_{i=1}^n w_{t,i}z_{t,i}^2$ by analysis via a local norm.
 

### 2014-01-17 lp/adaptive-regularization.tex
better notation

--- 2014-01-16 lp/adaptive-regularization.tex
@@
 \section{General Algorithms and Theoretical Analysis}
+
+\begin{table}
+\begin{tabular}{|c|c|c|} \hline
+        Algorithm & Regret \\ \hline
+        Exponentiated gradient & $\frac{\lambda n + \sum_{i=1}^n w_i[\log(w_i/\lambda)-1]}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_{t,i}z_{t,i}^2$ \\ \hline
+        Multiplicative updates & $\frac{\lambda n + \sum_{i=1}^n w_i[\log(w_i/\lambda)-1]}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_{i}z_{t,i}^2$ \\ \hline
+        Adaptive exponentiated gradient & $\frac{\lambda n + \sum_{i=1}^n w_i[\log(w_i/\lambda)-1]}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_iz_{t,i}^2$ \\ \hline
+\end{tabular}
+\caption{Comparison of exponentiated gradient, multiplicative updates, and adaptive exponentiated gradient algorithms in the unconstrained case.}
+\label{fig:unconstrained}
+\end{table}
+
 We now present the unconstrained versions of our three algorithms, together 
-with an analysis based on adaptive regularizers. In each case, the weight 
+with an analysis based on adaptive regularizers (the bounds can be found summarized 
+in Table~\ref{fig:unconstrained}). In each case, the weight 
 vector $w$ is necessarily non-negative. We can deal with this constraint by 
@@
 Each algorithm can be interpreted as (adaptive) mirror descent with an appropriate 
-regularizer. The regularizers are as follows:
+regularizer. Each regularizer will (potentially) vary with the round $t$, and will be a 
+member of the family $\psi_u(w)$, which for $u \in \bR^n$ is defined as
+\[ \psi_u(w) \eqdef \sum_{i=1}^n w_i[\log(w_i/\lambda)-1] + u^{\top}w. \]
+The parameter $\lambda$ is a hyperparameter controlling the size of $w$. Typically it should be on the order of $\frac{1}{n}$ 
+(as we will see more explicitly when we prove a formal regret bound). 
+Before continuing, it will be useful to point out a few properties of $\psi_u$, which we 
+state in Lemma~\ref{lem:properties} and prove in the Appendix.
+\begin{lemma}
+\label{lem:properties}
+For $\psi_u(w) \eqdef \sum_{i=1}^n w_i[\log(w_i/\lambda)-1] + u^{\top}w$, we have:
+\begin{align}
+\psi_u^*(\theta) &= \lambda \sum_{i=1}^n \exp(\theta_i - u_i). \\
+\frac{\partial \psi_i^*}{\partial \theta_i} &= \lambda \exp(\theta_i - u_i). 
+\end{align}
+\end{lemma}
+Now, as noted above, each of the three algorithms can be expressed in terms of a 
+time-varying regularizer $\psi_t(w) \equiv \psi_{u_t}(w)$. The particular choices of $u_t$ are:
 \begin{enumerate}
-        \item \textbf{Exponentiated gradient:} $\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1]$.
-        \item \textbf{Multiplicative updates:} $\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1-\sum_{t=1}^T (\eta z_{t,i} + \log(1-\eta z_{t,i}))]$.
-        \item \textbf{Adaptive exponentiated gradient:} $\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1+\sum_{t=1}^T \eta^2 z_{t,i}^2 w_i]$.
+        \item \textbf{Exponentiated gradient:} $u_{t,i} = 0$ (non-varying regularizer) %$\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1]$.
+        \item \textbf{Multiplicative updates:} $u_{t,i} = -\sum_{s=1}^t \eta z_{s,i} + \log(1-\eta z_{s,i})$ %$\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1-\sum_{s=1}^t (\eta z_{s,i} + \log(1-\eta z_{s,i}))]$.
+        \item \textbf{Adaptive exponentiated gradient:} $u_{t,i} = \eta \sum_{s=1}^t z_{s,i}^2$ %$\psi_t(w) = \sum_{i=1}^n w_i[\log(w_i/\lambda)-1+\sum_{s=1}^t \eta^2 z_{s,i}^2 w_i]$.
 \end{enumerate}
-The parameter $\lambda$ is a hyperparameter controlling the size of $w$. Typically it should be on the order of $\frac{1}{n}$ 
-(as we will see more explicitly when we prove a formal regret bound). Note that in the limit of small $\eta$, 
+Note that in the limit of small $\eta$, 
 $-\eta z_{t,i} - \log(1-\eta z_{t,i}) \approx \frac{1}{2}\eta z_{t,i}^2$, so that 
@@
 \end{enumerate}
-Finally, we base our regret bounds on the following standard bound on the regret of mirror descent:
+Finally, we base our regret bounds on an adaptation of the mirror descent analysis. In the analysis of 
+mirror descent based on Fenchel conjugacy, we have the following inequality for any regularizer $\psi$; 
+letting $\theta_t \eqdef -\sum_{s=1}^T z_s$ and $\psi_t$ denote the regularizer at time $t$:
 \begin{equation}
-        \label{eqn:bound}
-        \Regret(w) \leq \psi_T(w) + \sum_{t=1}^T D_{\psi_t^*}(\theta_{t+1} \| \theta_t).
+        \label{eqn:bound1}
+        -\psi_T(w) - \eta \sum_{t=1}^T w^{\top} z_t \leq \psi_0^*(0) + \sum_{t=1}^T \psi_t^*(\theta_t) - \psi_{t-1}^*(\theta_{t-1}).
 \end{equation}
-Our idea is to pick $\psi_{t+1}-\psi_t$ to cancel out the Bregman divergence term. The intuition is 
-that we are absorbing all of our loss into the regularizer, therefore giving us a bound that depends 
-on $w$ rather than $w_t$. This strategy is non-standard, although has been used before, for instance in 
-various papers by Francesco Orabona. Note that in all three cases, $\psi_t(w)$ differs from $\psi_0(w)$ 
-only in a linear term, and hence $D_{\psi_t^*} \equiv D_{\psi_0^*}$, and in particular 
-$D_{\psi_t^*}(\theta_{t+1} \| \theta_t) \leq \sum_{i=1}^n w_{t,i}z_{t,i}^2$ by analysis via a local norm.
+Usually the regularizer is fixed, in which case $\psi^*(\theta_t) - \psi^*(\theta_{t-1}) = -\eta w_t^{\top}z_t + D_{\psi^*}(\theta_{t+1} \| \theta_t)$. 
+Note that $w_t^{\top}z_t$ is the loss suffered in round $t$; the Bregman divergence term $D_{\psi^*}$ measures the increase in regret for that round. 
+In particular, re-arranging gives
+\begin{equation}
+        \label{eqn:bound2}
+        \Regret(w) = \sum_{t=1}^T w_t^{\top}z_t - \sum_{t=1}^T w^{\top}z_t \leq \frac{1}{\eta}\left[ \psi^*(0) + \psi(w) + \sum_{t=1}^T D_{\psi^*}(\theta_{t+1} \| \theta_t)\right].
+\end{equation}
+One can then use an analysis based on strong convexity of $\psi$ to bound $\psi^*$; for instance, in the case of exponentiated gradient, 
+we have the bound $D_{\psi^*}(\theta_{t+1} \| \theta_t) \leq \eta^2 \sum_{i=1}^n w_{t,i}z_{t,i}^2$, which leads to our claimed regret bound 
+for exponentiated gradient.
 
-%%% We are interested in the following learning problem: our experts 
-%%% consist of vectors $w \in \sH$, where $\sH$ is some convex 
-%%% set; similarly, our adversary can play any move $z_t$ with 
-%%% $z_t \in \sA$. A typical example is $\sH$ is the $n$-dimensional 
-%%% simplex and $\sA$ is the unit $l^{\infty}$ ball. Our end goal is 
-%%% to obtain a regret bound of the following type:
-%%% \[ \sum_{t=1}^T w_t^Tz_t \leq \frac{\lambda}{\epsilon} + \epsilon \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t. \]
-%%% If we let $\sigma = \sup_{w \in \sH, z \in \sA} |w|^T|z|$, then the regret is bounded by
-%%% $\frac{\lambda}{\epsilon} + \epsilon \sigma T$, which is equal to $2\sqrt{\lambda \sigma T}$ for 
-%%% $\epsilon = \sqrt{\frac{\lambda}{\sigma T}}$. This is the type of regret bound that is typical 
-%%% in most online learning settings. However, our bound can lead to stronger results in some 
-%%% situations, for instance if $\sA$ is very asymmetrical about the origin, in which case 
-%%% $|w|^T|z_t|$ will tend to be close to $w^Tz_t$ and we get bound
-%%% \[ \sum_{t=1}^T w_t^Tz_t \leq \frac{\lambda}{\epsilon} + (1+\epsilon)\sum_{t=1}^T w^Tz_t, \]
-%%% which is potentially much better, for instance if there is an expert that achieves 
-%%% very small loss.
-%%% 
-%%% We propose the following general style of algorithm for obtaining such 
-%%% bounds. The key innovation is an auxiliary variable $x_t$ that lets us 
-%%% relate $w_t^Tz_t$ and $w^Tz_t$. The updates are as follows:
-%%% \begin{align*}
-%%% x_{t+1} &= f(x_t, \eta z_t) \\
-%%% w_{t+1} &= \argmax_{w \in \sH} w^Tx_{t+1} - \psi(w) \\
-%%% \Phi_{t+1} &= w_{t+1}^Tx_{t+1} - \psi(w_{t+1})  \\ &= \psi^*(x_{t+1}).
-%%% \end{align*}
-%%% 
-%%% To see why these updates are a good idea, we will make a ``physicist's approximation'' 
-%%% and suppose that $f(x_t, \eta z_t) \approx x_t - \eta z_t$. Then we would have
-%%% \[ x_{T+1} \approx x_1 - \eta \sum_{t=1}^T z_t \] and thus, by definition, 
-%%% \[ \Phi_{T+1} \gtrsim w^Tx_1 - \psi(w) - \eta \sum_{t=1}^T w^Tz_t. \]
-%%% On the other hand, 
-%%% \begin{align*}
-%%% \Phi_{t+1} &\approx \Phi_t + w_t^T(x_{t+1} - x_t) \\
-%%%            &= \Phi_t - \eta w_t^T z_t,
-%%% \end{align*}
-%%% so we have 
-%%% \[ \Phi_{T+1} \approx \Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t. \]
-%%% Combining this with the preceding bound on $\Phi_{T+1}$ yields
-%%% $\Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t \gtrsim w^Tx_1 - \psi(w) - \eta \sum_{t=1}^T w^Tz_t$, 
-%%% or
-%%% \[ \sum_{t=1}^T w_t^Tz_t \lesssim \frac{\Phi_1 + \psi(w) - w^Tx_1}{\eta} + \sum_{t=1}^T w^Tz_t, \]
-%%% which gives us an even better regret bound than we had sought (the $\sum_{t=1}^T |w|^T|z_t|$ term 
-%%% will show up when we track the errors made in each of the approximations).
-%%% 
-%%% \paragraph{Formal statement.} To turn these approximations into a rigorous bound, we will 
-%%% need $f$ and $\psi$ to satisfy the following properties:
-%%% \begin{itemize}
-%%% \item Concavity of $\psi^* \circ f$: $\psi^*(f(x,\eta z)) \leq \psi^*(x) - \eta w_x^Tz$, where $w_x = \partial \psi^*(x)$.
-%%% \item Smoothness of $f$: $w^T(x-\eta z - f(x,\eta z)) \leq \eta^2 \alpha |w|^T|z|$ for all $w$.
-%%% \end{itemize}
-%%% (For some intuition, any $f$ satisfying the two properties above 
-%%% must satisfy $f(x,0) = x$ and $f(x,\eta z) = x-\eta z + O(\eta^2)$.)
-%%% \begin{theorem}
-%%% \label{thm:main}
-%%% Suppose that $\psi^*$ and $f$ satisfy the two properties above. Then, 
-%%% for $\eta = \frac{\epsilon}{\alpha}$, we have the bound
-%%% \[ \sum_{t=1}^T w_t^Tz_t \leq \alpha \frac{\Phi_1 + \psi(w) - w^Tx_1}{\epsilon} + \epsilon \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t. \]
-%%% \end{theorem}
-%%% \begin{proof}
-%%% By the first condition, we have
-%%% \begin{align*}
-%%% \Phi^{t+1} &= \psi^*(x_{t+1}) \\
-%%%  &= \psi^*(f(x_t, \eta z_t)) \\
-%%%  &\leq \psi^*(x_t) - \eta \left(\frac{\partial \psi^*}{\partial x_t}\right)^T z_t \\
-%%%  &= \Phi_t - \eta w_t^T z_t,
-%%% \end{align*}
-%%% which by induction implies that
-%%% \[ \Phi_{T+1} \leq \Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t. \]
-%%% By the second condition, we have
-%%% \begin{align*}
-%%% w^Tx_{t+1} &= w^Tf(x, \eta z) \\
-%%%  &\geq w^Tx_t - \eta w^Tz_t - \eta^2 \alpha |w|^T|z_t|,
-%%% \end{align*}
-%%% which by induction implies that
-%%% \[ w^Tx_{T+1} \geq w^Tx_1 - \eta^2\alpha \sum_{t=1}^T |w|^T|z_t| - \eta \sum_{t=1}^T w^Tz_t. \]
-%%% Combining this with the first inequality yields
-%%% \begin{align*}
-%%% \Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t &\geq \Phi_{T+1} \\
-%%%  &\geq w^Tx_{T+1} - \psi(w) \\
-%%%  &\geq w^Tx_1 - \psi(w) - \eta^2\alpha \sum_{t=1}^T |w|^T|z_t| - \eta \sum_{t=1}^T w^Tz_t.
-%%% \end{align*}
-%%% Re-arranging yields
-%%% \[ \sum_{t=1}^T w_t^Tz_t \leq \frac{\Phi_1 + \psi(w) - w^Tx_1}{\eta} + \eta\alpha \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t, \]
-%%% which gives the desired result.
-%%% \end{proof}
-%%% 
-%%% \paragraph{Relationship to mirror descent.} Mirror descent can be cast in the same framework as above, 
-%%% and corresponds to the exact additive update $x_{t+1} = x_t - \eta z_t$. Thus while our algorithm 
-%%% makes sure that $\psi^*$ decreases at a rate of $w_t^Tz_t$ and then bounds the error between 
-%%% $x_{t+1}$ and $x_t - \eta z_t$, in contrast mirror descent sets the error between $x_{t+1}$ and 
-%%% $x_t - \eta z_t$ to zero, and then relies on the smoothness of $\psi^*$ to bound the error between 
-%%% $\psi^*(x-\eta z)$ and $\psi^*(x) - \eta w_x^Tz$.
-%%% \section{Examples}
-%%% \paragraph{Multiplicative updates.} If we let $\sH$ be the simplex and $\sA$ be the 
-%%% $l^{\infty}$ ball of radius $r$, then a standard algorithm for minimizing regret is the 
-%%% \emph{multiplicative weights update method}. We recover this by letting $f(x,z) = x + \log(1-z)$ 
-%%% and $\psi(w) = \sum_{i=1}^n w_i\log w_i$. Then $\psi^*(x) = \log\left(\sum_{i=1}^n \exp(x_i)\right)$, 
-%%% and $w_{t+1,i} = \exp(x_{t+1,i})/\sum_{j=1}^n \exp(x_{t+1,j})$.
-%%% \begin{align*}
-%%% \psi^*(f(x,\eta z)) &= \log\left(\sum_{i=1}^n \exp(x_i + \log(1-\eta z_i))\right) \\
-%%%  &= \log\left(\sum_{i=1}^n \exp(x_i)(1-\eta z_i)\right) \\
-%%%  &= \log\left(\sum_{i=1}^n \exp(x_i) - \eta \sum_{i=1}^n \exp(x_i) z_i\right) \\
-%%%  &\leq \log\left(\sum_{i=1}^n \exp(x_i)\right) - \eta \frac{\sum_{i=1}^n \exp(x_i) z_i}{\sum_{i=1}^n \exp(x_i)} \\
-%%%  &= \psi^*(x) - \eta w^Tz,
-%%% \end{align*}
-%%% so condition 1 is satisfied.
-%%% 
-%%% Also, using the fact that $\log(1-\eta z_i) \geq -\eta z_i - \eta^2 z_i^2$, we have
-%%% \begin{align*}
-%%% w^T(x-\eta z - f(x, \eta z)) &= w^T(x-\eta z - (x + \log(1-\eta z))) \\
-%%%  &\leq \eta^2 \sum_{i=1}^n w_iz_i^2 \\
-%%%  &\leq \eta^2 r \sum_{i=1}^n |w_i||z_i|,
-%%% \end{align*}
-%%% where in the last step we used the fact that $|z_i| \leq r$. These updates 
-%%% thus fulfill the conditions of Theorem~\ref{thm:main} for $\alpha$ equal to 
-%%% $r$.
-%%% 
-%%% \paragraph{Sum of pointed convex sets.} Suppose that our hypothesis class 
-%%% $\sH \subseteq \bR^n$ can be decomposed as $\sH = \sH_+ + \sH_-$, where $\sH_+$ consists 
-%%% only of non-negative vectors and $\sH_-$ consists only of non-positive vectors. 
-%%% We can translate this into a problem n $\bR^{2n}$, by letting $\sH' \eqdef \sH_+ \times (-\sH_-)$ 
-%%% and $\sA' \eqdef \{(z, -z) \mid z \in \sA\}$. Then $(w')^T(z') = w^Tz$. Now 
-%%% let $(w_+, w_-)$ be the decomposition of $w'$ into its two components. We will 
-%%% define
-%%% \[ \psi(w') = \left\{ \begin{array}{ccl} \sum_{i=1}^n \log((w_+)_i + (w_-)_i) & : & w' \in \sH \\ \infty & : & \mathrm{else}. \end{array} \right. \]
-%%% Now let 
-%%% \[ R \eqdef \sup\limits_{w_+ \in \sH_+, w_- \in \sH_-, z \in \sA}  |w_+^Tz| + |w_-^Tz|, \]
-%%% and consider the update $x_{t+1} = x_t - \eta z'_t - 2\eta^2 R |z'_t|$. Then we have
-%%% \begin{align*}
-%%% \psi^*(x_{t+1}) &= \psi^*(x_t) - \eta (w'_t)^T(z'_t) - 2\eta^2R (w'_t)^T|z'_t| + D_{\psi^*}(x_{t+1} \| x_t) \\
-%%%  &= \psi^*(x_t) - \eta (w'_t)^Tz_t - 2\eta^2R (w'_t)^T|z'_t| + \frac{1}{2}((w'_t)^T|x_{t+1}-x_t|)^2 \\
-%%%  &= \psi^*(x_t) - \eta (w'_t)^Tz_t - 2\eta^2R (w'_t)^T|z'_t| + \frac{1}{2}(\eta (w'_t)^T|z'_t| + 2\eta^2R (w'_t)^T|z'_t|)^2 \\
-%%%  &= \psi^*(x_t) - \eta (w'_t)^Tz_t
-%%% \end{align*}
-%%% as long as $2\eta R \leq 1$, or $\eta \leq \frac{1}{2R}$. Thus the first condition is 
-%%% satisfied. Also, by construction we have 
-%%% \[ w^T(x_t - \eta z_t - x_{t+1}) = 2\eta^2 R w^T|z'_t|, \]
-%%% so that the second condition is satisfied for $\alpha = 2R$. Combining these, and 
-%%% setting $x_1 = 0$, we obtain the bound
-%%% \[ \sum_{t=1}^T w_t^Tz_t \leq \frac{2R[\psi(w) - \inf_{\tilde{w}} \psi(\tilde{w})]}{\epsilon} + \epsilon \sum_{t=1}^T (w_+ + w_-)^T|z_t| + \sum_{t=1}^T w^Tz_t. \]
-%%% This is a quite general result, but suffers from the fact that $\psi(w)$ can grow very large. 
-%%% Finding ways to deal with this would thus be very interesting.
-%%% 
-%%% \paragraph{Inflated $l^p$ norm.} By abuse of notation, let $\alpha$ denote the vector 
-%%% whose coordinates are all equal to $\alpha$. Let $\sH = B_{\alpha}^p(\lambda)$, where 
-%%% \[ B_{\alpha}^p(\lambda) \eqdef \{w \mid w \geq \alpha, \|w-\alpha\|_p \leq \lambda \}. \]
-%%% Note that this is the positive $l^p$ ball of radius $\lambda$, translated by $\alpha$. 
-%%% (We will show below that we can embed the unconstrained $l^p$ ball into $B_{\alpha}^p(\lambda)$ 
-%%% at the cost of increasing the dimension by a factor of 2, implying that 
-%%% $B_{\alpha}^p(\lambda)$ is a rich hypothesis class.) 
-%%% 
-%%% We will let $\psi(w) = \left\{ \begin{array}{ccl} \frac{1}{2(p-1)}\|w-\alpha\|_p^2 & : & w \geq \alpha \\ \infty & : & w \not\geq \alpha. \end{array} \right.$. 
-%%% Then 
-%%% \[ \psi^*(x) = \alpha \sum_{i=1}^n x_i + \frac{1}{2(q-1)}\|x_+\|_q^2, \]
-%%% where $x_+$ is the positive part of $x$ and $\frac{1}{p} + \frac{1}{q} = 1$.
-%%% 
-%%% Now suppose we use the update $x_{t+1,i} = x_{t,i} - \eta z_{t,i} - 2\eta^2 \frac{\|z\|_q^2}{\|z\|_1} |z_{t,i}| / \alpha$. 
-%%% Note that the Bregman divergence $D_{\psi^*}(x_{t+1} \| x_t)$ is bounded above by the Bregman divergence 
-%%% of $\frac{1}{2(q-1)}\|x\|_q^2$ (since it can be obtained from $\psi^*$ by adding a convex function), 
-%%% which is in turn bounded above by $\frac{1}{2}\|x_{t+1}-x_t\|_q^2$. Using these facts, together with the fact that 
-%%% $w \geq \alpha$, we have
-%%% \begin{align*}
-%%% \psi^*(x_{t+1}) &\leq \psi^*(x_t) - \eta w_t^Tz_t - \frac{2\eta^2\|z\|_q^2}{\alpha \|z\|_1} w_t^T|z_t| + D_{\psi^*}(x_{t+1} \| x_t) \\
-%%%  &\leq \psi^*(x_t) - \eta w_t^Tz_t - \frac{2\eta^2\|z\|_q^2}{\alpha \|z\|_1} w_t^T|z_t| + \frac{1}{2}\left\|\eta z_t + \frac{2\eta^2\|z\|_q^2}{\alpha \|z\|_1} |z_t|\right\|_q^2 \\
-%%%  &\leq \psi^*(x_t) - \eta w_t^Tz_t - 2\eta^2 \|z\|_q^2 + \frac{1}{2} \left(\eta \|z_t\|_q + \frac{2\eta^2\|z_t\|_q^3}{\alpha \|z_t\|_1}\right)^2 \\
-%%%  &\leq \psi^*(x_t) - \eta w_t^Tz_t - 2\eta^2 \|z\|_q^2 + \frac{1}{2} \left(\eta \|z_t\|_q + 2\eta^2 \|z_t\|_{q}^2 / \alpha\right)^2 \\
-%%%  &\leq \psi^*(x_t) - \eta w_t^Tz_t
-%%% \end{align*}
-%%% as long as $2\eta \|z_t\|_q/\alpha \leq 1$, which holds as long as $\eta \leq \frac{\alpha}{2r_q}$. In addition, we have that
-%%% \begin{align*}
-%%% w^T(x_{t} - \eta z_t - x_{t+1}) &= -\frac{2\eta^2\|z_t\|_q^2}{\alpha\|z_t\|_1}w^T|z_t| \\
-%%%  &\leq \eta^2 \frac{2\|z_t\|_q}{\alpha} |w|^T|z_t| \\
-%%%  &\leq \eta^2 \frac{2r_q}{\alpha} |w|^T|z_t|.
-%%% \end{align*}
-%%% So, both conditions of Theorem~\ref{thm:main} are satisfied. 
-%%% If we set $x_1 = 0$, then $\psi^*(x_1) = 0$ as well.
-%%% Therefore, for $\eta = \frac{\epsilon \alpha}{2r_q}$, we obtain a regret bound of 
-%%% \[ \frac{2r_q\psi(w)}{\epsilon \alpha} + \epsilon \sum_{t=1}^T |w|^T|z_t|. \]
-%%% Note that the regret gets smaller as $\alpha$ grows larger. This might seem too good 
-%%% to be true, until we notice the following: if $\alpha$ is very large, then all 
-%%% vectors with a small value of $\|w-\alpha\|_p$ will point in approximately the same 
-%%% direction, and thus get similar (multiplicative) values for the regret.
-%%% \paragraph{Embedding the $l^p$ ball into $B_{\alpha}^p$.} We now show how to turn problems 
-%%% involving the $l^p$ ball into problems involving $B_{\alpha}^p$. Given a vector 
-%%% $w \in \bR^n$, we can create a vector $w' \in \bR^{2n}$ via the embedding 
-%%% $w \mapsto \left[ \begin{array}{c} \alpha + \max(w, 0) \\ \alpha - \min(w,0) \end{array} \right]$. 
-%%% This embedding satisfies the property that $\|w\|_p = \|w'-\alpha\|_p$. Also, $w^Tz = w'^T\left[ \begin{array}{c} z \\ -z \end{array} \right]$. 
-%%% If we plug into the regret bound above, the only difference is that the value of $|w|^T|z_t|$ increases, yielding 
-%%% a regret of
-%%% \[ \frac{r_q \|w\|_p^2}{\epsilon \alpha (p-1)} + \epsilon \sum_{t=1}^T \left(|w|^T|z_t| + 2\alpha \|z_t\|_1\right). \]
-%%% Unfortunately, this final bound is quite bad, because the $\|z_t\|_1$ term grows with $T$ and depends on the 
-%%% $l^1$ norm, which is the biggest of all the $l^p$ norms.
+In contrast, our analysis based on adaptive regularizers will choose $\psi_t^*$ such that $\psi_t^*(\theta_t) - \psi_{t-1}^*(\theta_{t-1}) \leq -\eta w_t^{\top}z_t$. 
+Assuming we succeed in picking such a $\psi_t^*$, the $\sum_{t=1}^T \psi_t^*(\theta_t) - \psi_{t-1}^*(\theta_{t-1})$ term can be replaced with 
+$-\eta \sum_{t=1}^T w_t^{\top}z_t$. We can then re-arrange to obtain the regret bound
+\begin{equation}
+        \label{eqn:bound3}
+        \Regret(w) = \sum_{t=1}^T w_t^{\top}z_t - \sum_{t=1}^T w^{\top}z_t \leq \frac{\psi_0^*(0) + \psi_T(w)}{\eta}.
+\end{equation}
+This is quite interesting as we have pushed the entire regret bound into an increase in the size 
+of the regularizer.
+
+We will now verify the property $\psi_t^*(\theta_t) - \psi_{t-1}^*(\theta_{t-1}) \leq -\eta w_t^{\top}z_t$ 
+for both multiplicative updates and adaptive EG. Our analysis relies on Lemma~\ref{lem:properties}.
+For multiplicative updates:
+\begin{align}
+        \psi_t^*(\theta_t) &= \psi_t^*(\theta_{t-1}- \eta z_t) \\
+                           &= \lambda \sum_{i=1}^n \exp\left(\theta_{t-1,i} - \eta z_{t,i} - u_{t,i}\right) \\
+                           &= \lambda \sum_{i=1}^n \exp\left(\theta_{t-1,i} + \log(1- \eta z_{t,i}) - u_{t-1,i}\right) \\
+                           &= \lambda \sum_{i=1}^n \exp\left(\theta_{t-1,i} - u_{t-1,i}\right)(1 - \eta z_{t,i}) \\
+                           &= \psi_{t-1}^*(\theta_{t-1}) - \eta \lambda \sum_{i=1}^n \exp\left(\theta_{t-1,i} - u_{t-1,i}\right) z_{t,i} \\
+                           &= \psi_{t-1}^*(\theta_{t-1}) - \eta w_t^{\top}z_t,
+\end{align}
+hence $\psi_t^*(\theta_t) - \psi_{t-1}^*(\theta_{t-1}) = -\eta w_t^{\top}z_t$, i.e. the property holds with equality. Using the 
+fact that $\eta z + \log(1-\eta z) \leq \eta^2 z^2$ for $\eta \leq \frac{1}{2z}$, we have the regret bound
+\begin{align}
+        \Regret(w) &\leq \frac{\psi_0*(0) + \psi_T(w)}{\eta} \\
+                   &= \frac{\lambda n + \sum_{i=1}^n w_i[\log(w_i/\lambda)-1] + u_T^{\top}w}{\eta} \\
+                   &\leq \frac{\lambda n + \sum_{i=1}^n w_i[\log(w_i/\lambda)-1] + \eta^2 \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2}{\eta} \\
+                   &= \frac{\lambda n + \sum_{i=1}^n w_i[\log(w_i/\lambda)-1]}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_i z_{t,i}^2,
+\end{align}
+which holds whenever $\eta \leq \frac{1}{2 \max_{t=1}^T \|z_t\|_{\infty}}$.
+
+Setting $\lambda$ to $1/n$ yields a bound of
+\[ \Regret(w) \leq \frac{1+(\log(n)-1)\|w\|_1 + \sum_{i=1}^n w_i\log(w_i)}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_i z_{t,i}^2. \]
+The proof for adaptive exponentiated gradient is similar; we have:
+\begin{align}
+\psi_t^*(\theta_t) &= \psi_t^*(\theta_{t-1} - \eta z_t) \\
+ &= \lambda \sum_{i=1}^n \exp\left(\theta_{t-1,i} - \eta z_{t,i} - u_{t,i}\right) \\
+ &= \lambda \sum_{i=1}^n \exp\left(\theta_{t-1,i} - \eta z_{t,i} - \eta^2 z_{t,i}^2 - u_{t-1,i}\right) \\
+ &= \lambda \sum_{i=1}^n \exp\left(\theta_{t-1,i} - u_{t-1,i}\right)\exp(-\eta z_{t,i} - \eta^2 z_{t,i}^2) \\
+ &\leq \lambda \sum_{i=1}^n \exp\left(\theta_{t-1,i} - u_{t-1,i}\right)(1-\eta z_{t,i}) \\
+ &= \psi_{t-1}^*(\theta_{t-1}) - \eta w_t^{\top}z_t,
+\end{align}
+where the last step re-uses the analysis for multiplicative updates. We thus have
+\begin{align}
+\Regret(w) &\leq \frac{\psi_0^*(0) + \psi_T(w)}{\eta} \\
+ &= \frac{\lambda n + \sum_{i=1}^n w_i[\log(w_i/\lambda)-1]}{\eta} + \eta \sum_{t=1}^T \sum_{i=1}^n w_i z_{t,i}^2,
+\end{align}
+which is the same regret bound as before.
+\section{Experiments}
+TODO.
+
 \end{document}

### 2014-01-17 lp/code/EGLearner.java
initial java code (not yet compiled)

@@ -1,0 +1,14 @@
+public class EGLearner extends Learner {
+  public EGLearner(Info info){
+    super(info);
+  }
+
+  public void updateTheta(Example example){
+    for(Integer index : example.features.keySet()){
+      double value = example.get(index);
+      theta[index][0] -= eta * value;
+      theta[index][1] += eta * value;
+    }
+  }
+}
+

### 2014-01-17 lp/code/Example.java

@@ -1,0 +1,13 @@
+public class Example {
+  int label, mult;
+  Map<Integer, Double> features;
+  public Example(int label){
+    features = new HashMap<Integer, Double>();
+    this.label = label;
+    this.mult = label == 2 ? 1 : -1;
+  }
+  void add(int index, double value){
+    features.put(index, value * mult);
+  }
+
+}

### 2014-01-17 lp/code/Learner.java

@@ -1,0 +1,40 @@
+public abstract class Learner {
+  double[] theta;
+  double lambda, eta;
+  public Learner(Info info){
+    if(info.numLabels() != 2){
+      throw new RuntimeException("expected binary classification problem");
+    } else {
+      lambda = 1.0/info.numFeatures();
+      eta = Math.sqrt(-Math.log(lambda)/info.numExamples());
+      theta = new double[info.numFeatures()][2]; // two sets of weights (positive and negative)
+      for(int i = 0; i < info.numFeatures(); i++){
+        theta[i][0] = theta[i][1] = Math.log(lambda);
+      }
+    }
+  }
+
+  double weights(int index){
+    return Math.exp(theta[index][0]) - Math.exp(theta[index][1]);
+  }
+
+  // hinge loss
+  double loss(double y){
+    return Math.min(1+y, 0);
+  }
+
+  abstract void updateTheta(Example example);
+
+  public double predict(Example example){
+    double y = 0.0;
+    for(Integer index : example.features.keySet()){
+      double value = example.get(index);
+      y += weights(index) * value;
+    }
+    double ret = loss(y);
+    if(y > -1){
+      updateTheta(example);
+    }
+  }
+
+}

### 2014-01-17 lp/code/Main.java

@@ -1,0 +1,42 @@
+import java.util.*;
+import java.io.*;
+import fig.basic.LogInfo;
+import fig.basic.Option;
+import fig.basic.StatFig;
+import fig.exec.Execution;
+
+public class Main implements Runnable {
+    @Option(gloss="experiment name (for easier tracking)")
+		public static String experimentName = "UNKNOWN";
+    @Option(required=true, gloss="dataset")
+		public static String dataset;
+
+    public void run() {
+        try {
+            runWithException();
+        } catch(Exception e) {
+						e.printStackTrace();
+            throw new RuntimeException();
+        }
+    }
+
+    void runWithException() throws Exception {
+        LogInfo.logs("EXPERIMENT NAME: %s", experimentName);
+        Info info = new Info("../"+dataset+"/info");
+        IncrementalReader reader = new IncrementalReader("../"+dataset+"/shuffled");
+        Learner learner = new EGLearner(info);
+        int T = info.numExamples();
+        double[] losses = new double[T];
+        LogInfo.begin_track("Predicting");
+        for(int t = 0; t < T; t++){
+          loss += learner.predict(reader.nextExample());
+          LogInfo.logs("loss[%d] = %f", t, loss);
+          losses[t] = loss;
+        }
+        LogInfo.end_Track();
+    }
+
+    public static void main(String[] args){
+        Execution.run(args, new Main());
+    }
+}

### 2014-01-18 lp/code/EGLearner.java
compiles, but crappy results

@@ -5,7 +5,7 @@ public class EGLearner extends Learner {
 
   public void updateTheta(Example example){
     for(Integer index : example.features.keySet()){
-      double value = example.get(index);
+      double value = example.features.get(index);
       theta[index][0] -= eta * value;
       theta[index][1] += eta * value;
     }

### 2014-01-18 lp/code/Example.java

@@ -1,3 +1,4 @@
+import java.util.*;
 public class Example {
   int label, mult;
   Map<Integer, Double> features;

### 2014-01-18 lp/code/IncrementalReader.java

@@ -1,6 +1,8 @@
+import java.io.*;
+import java.util.*;
 public class IncrementalReader {
   Scanner scanner;
-  public IncrementalReader(String filename){
+  public IncrementalReader(String filename) throws Exception {
     scanner = new Scanner(new File(filename));
   }
 
@@ -10,7 +12,7 @@ public class IncrementalReader {
     Example ret = new Example(label);
     for(int i = 1; i < tokens.length; i++){
       String[] tokens2 = tokens[i].split(":");
-      ret.add(Integer.parseInt(tokens2[0]), Double.parseDouble(tokens2[1]));
+      ret.add(Integer.parseInt(tokens2[0])-1, Double.parseDouble(tokens2[1]));
     }
     return ret;
   }

### 2014-01-18 lp/code/Info.java

@@ -1,10 +1,11 @@
 import java.io.*;
+import java.util.*;
 public class Info {
   Map<String, String> map;
-  public Info(String filename){
+  public Info(String filename) throws Exception {
     map = new HashMap<String, String>();
     Scanner scanner = new Scanner(new File(filename));
-    while(scanner.hasNextLine(){
+    while(scanner.hasNextLine()){
       String[] line = scanner.nextLine().split("\\s");
       map.put(line[0], line[1]);
     }

### 2014-01-18 lp/code/Learner.java

@@ -1,5 +1,5 @@
 public abstract class Learner {
-  double[] theta;
+  double[][] theta;
   double lambda, eta;
   public Learner(Info info){
     if(info.numLabels() != 2){
@@ -20,7 +20,7 @@ public abstract class Learner {
 
   // hinge loss
   double loss(double y){
-    return Math.min(1+y, 0);
+    return Math.max(1+y, 0);
   }
 
   abstract void updateTheta(Example example);
@@ -28,12 +28,17 @@ public abstract class Learner {
   public double predict(Example example){
     double y = 0.0;
     for(Integer index : example.features.keySet()){
-      double value = example.get(index);
+      double value = example.features.get(index);
       y += weights(index) * value;
     }
     double ret = loss(y);
     if(y > -1){
       updateTheta(example);
+    }
+    if(Main.zeroOne){
+      return ret > .999 ? 1 : 0;
+    } else {
+      return ret;
     }
   }
 

### 2014-01-18 lp/code/Main.java

@@ -10,6 +10,10 @@ public class Main implements Runnable {
 		public static String experimentName = "UNKNOWN";
     @Option(required=true, gloss="dataset")
 		public static String dataset;
+    @Option(required=true, gloss="learner type")
+    public static int learnerType; // 0 = EG, 1 = MW, 2 = AEG
+    @Option(gloss="zero-one vs hinge")
+    public static boolean zeroOne = false;
 
     public void run() {
         try {
@@ -22,18 +26,25 @@ public class Main implements Runnable {
 
     void runWithException() throws Exception {
         LogInfo.logs("EXPERIMENT NAME: %s", experimentName);
-        Info info = new Info("../"+dataset+"/info");
-        IncrementalReader reader = new IncrementalReader("../"+dataset+"/shuffled");
-        Learner learner = new EGLearner(info);
+        Info info = new Info("../data/"+dataset+"/info");
+        IncrementalReader reader = new IncrementalReader("../data/"+dataset+"/shuffled");
+        Learner learner;
+        switch(learnerType){
+          case 0 : learner = new EGLearner(info); break;
+          case 1 : learner = new MWLearner(info); break;
+          case 2 : learner = new AEGLearner(info); break;
+          default : throw new RuntimeException("invalid learnerType");
+        }
         int T = info.numExamples();
         double[] losses = new double[T];
+        double loss = 0.0;
         LogInfo.begin_track("Predicting");
         for(int t = 0; t < T; t++){
           loss += learner.predict(reader.nextExample());
-          LogInfo.logs("loss[%d] = %f", t, loss);
+          LogInfo.logs("loss[%d] = %f / %f", t, loss/(t+1), 0.01*(loss-losses[Math.max(0, t-100)]));
           losses[t] = loss;
         }
-        LogInfo.end_Track();
+        LogInfo.end_track();
     }
 
     public static void main(String[] args){

### 2014-01-18 lp/code/AEGLearner.java
adding AEG/MW learners

@@ -1,0 +1,14 @@
+public class AEGLearner extends Learner {
+  public AEGLearner(Info info){
+    super(info);
+  }
+
+  public void updateTheta(Example example){
+    for(Integer index : example.features.keySet()){
+      double value = example.features.get(index);
+      theta[index][0] -= eta * value - eta*eta * value*value;
+      theta[index][1] += eta * value - eta*eta * value*value;
+    }
+  }
+}
+

### 2014-01-18 lp/code/MWLearner.java

@@ -1,0 +1,14 @@
+public class MWLearner extends Learner {
+  public MWLearner(Info info){
+    super(info);
+  }
+
+  public void updateTheta(Example example){
+    for(Integer index : example.features.keySet()){
+      double value = example.features.get(index);
+      theta[index][0] += Math.log(1.0 - eta * value);
+      theta[index][1] -= Math.log(1.0 - eta * value);
+    }
+  }
+}
+

### 2014-01-24 lp/summary-of-results.tex
added summary of results

\documentclass[11pt]{article}
\usepackage{import}
\input{../icml/latex-defs.tex}
\DeclareMathOperator{\Regret}{Regret}
\usepackage{fullpage}

\begin{document}
\section{Summary of Results}
This is a summary of the main results for the upcoming ICML paper. I'm writing them here because it 
will hopefully help us organize the paper and we can also brainstorm the best way to present the results.

The first result is the adaptive mirror descent result from before, with the ability to add in 
``hints'' $s_t$ that guess what the next value of $z_t$ will be (and decrease the loss if the guess is 
accurate). This is important because it will later let us achieve bounds based on the variance of the $z_t$ 
rather than the second moment.
\begin{theorem}
For any sequence $s_1, s_2, \ldots, s_T$, consider the following prediction algorithm (we assume 
that $\|z_t\|_{\infty} \leq 1$ and $\eta \leq \frac{1}{2}$):
\begin{enumerate}
\item Predict with $w_{t,i} \propto \exp(x_{t,i} - \eta s_{t,i})$.
\item Receive loss $w_t^{\top}z_t$.
\item Update $x_{t+1,i} = x_{t,i} - \eta z_{t,i} - \eta^2 (z_{t,i} - s_{t,i})^2$.
\end{enumerate}
Then the regret relative to a fixed weight vector $w$ is bounded by
\[ \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i}-s_{t,i})^2. \]
\end{theorem}

Our major corollary uses a particular choice of $s_t$ to get bounds in terms of the variance. Interestingly, 
the proof technique involves analyzing the $s_t$ as the moves played by a FTRL-style algorithm, so we can 
view this algorithm as ``doubly-adaptive''.

Let $\Var_i \eqdef \sum_{t=1}^T (z_{t,i}-\mu_i)^2$, with $\mu \eqdef \frac{1}{T} \sum_{t=1}^T z_t$.
\begin{corollary}
In the above setting, for the choice $s_t = \frac{1}{t} \sum_{s=1}^{t-1} z_s$, the 
regret relative to $w$ is bounded by
\[ \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i\Var_i + O(\log(T)). \]
\end{corollary}

These same results hold in the matrix case. Now we assume that $-I \preceq Z_t \preceq I$ 
and continue to assume that $\eta \leq \frac{1}{2}$:
\begin{theorem}
For a sequence $S_1, \ldots, S_T$ of symmetric matrices, consider the following 
prediction algorithm:
\begin{enumerate}
\item Predict $W_t \propto \exp(X_t - \eta S_t)$.
\item Receive loss $\Tr(W_t^{\top}Z_t)$.
\item Update $X_{t+1} = X_t - \eta Z_t - \eta^2 (Z_t-S_t)^2$.
\end{enumerate}
Then the regret relative to a fixed matrix $W$ is bounded by
\[ \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \Tr((Z_t-S_t)^{\top}W(Z_t-S_t)). \]
\end{theorem}
We now define $\Var$ to be the matrix $\sum_{t=1}^T (Z_t-M)^2$, with $M \eqdef \frac{1}{T} \sum_{t=1}^T Z_t$.
\begin{corollary}
In the matrix setting, for the cohice $S_t = \frac{1}{t} \sum_{s=1}^{t-1} Z_s$, the 
regret relative to $w$ is bounded by
\[ \frac{\log(n)}{\eta} + \eta \Tr(W^{\top}\Var) + O(\log(T)). \]
\end{corollary}
It is important to note that in all of these bounds, the $O(\log(T))$ term does not 
contain any hidden factors of $n$ or $\eta$. In fact we could take $O(\log(T)) = 8\log(T) + 2$.

Finally, we show an interesting separation between our adaptive algorithm and vanilla mirror 
descent even in the case where $s_t = 0$ for all $t$. For simplicity, we restrict our attention to the vector 
setting (the result extends trivially to the matrix setting). We consider the ``quasi-realizable'' setting, 
where one of the experts has identically zero loss (i.e. $z_{t,i} = 0$ for all $t$ for some $i$), and all 
other experts have non-negative cumulative loss. This is different from the realizable setting because 
losses can be both positive and negative.

\begin{proposition}
In the quasi-realizable setting, for any fixed step size $\eta$ there 
exists a sequence of losses such that mirror descent suffers regret $\Omega(\sqrt{T})$. 
In contrast, for any quasi-realizable sequence of losses and step size of $\eta = \frac{1}{2}$, 
adaptive mirror descent has regret $O(1)$ (where the $O(1)$ suppresses a $\log(n)$ dependence 
on dimension). These same results also hold in the matrix setting.
\end{proposition}

\end{document}

### 2014-01-25 lp/algorithms.tex
initial paper

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Algorithms} 
\label{sec:algorithms}
We now state our two main algorithms and their associated 
regret bounds.

\end{document}

### 2014-01-25 lp/paper_icml.tex
initial paper

\documentclass{article}
\usepackage{import, subfiles}
\usepackage{hyperref, icml2014, times}
\usepackage{tabularx}
\input{latex-defs.tex}

%\allowdisplaybreaks

\icmltitlerunning{Adaptive Exponentiated Gradient}

\begin{document}

\twocolumn[
\icmltitle{Adaptive Exponentiated Gradient}

\icmlauthor{Jacob Steinhardt}{jsteinhardt@cs.stanford.edu}
\icmladdress{Stanford University,
             353 Serra Street, Stanford, CA 94305 USA}
\icmlauthor{Jonathan Huggins}{jhuggins@mit.edu}
\icmladdress{MIT CSAIL,
             77 Massachusetts Avenue, Cambridge, MA 02139 USA}
\icmlauthor{Percy Liang}{pliang@cs.stanford.edu}
\icmladdress{Stanford University,
             353 Serra Street, Stanford, CA 94305 USA}

% You may provide any keywords that you 
% find helpful for describing your paper; these are used to populate 
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{online learning, optimistic, adaptive, matrix multiplicative weights}

\vskip 0.3in
]

\begin{abstract} 
We present an adaptive variant of the exponentiated gradient algorithm, 
achieving regret that depends only on the variance of the best retrospective 
expert. This generalizes and improves a result by 
\cite{hazan2010variation}. Our analysis is conceptually clean and brings together 
two recent developments in online learning, as well as a novel construction 
using an auxiliary instance of follow the regularized leader to tune a regret 
bound. Using the machinery developed, we are also able to extend our results 
to the matrix case, presenting an \emph{adaptive matrix exponentiated gradient} 
algorithm. Our proof in this case involves a novel analysis tool generalizing 
follow the regularized leader to vector-valued payoffs, which may be of 
independent interest.
\end{abstract}

\subfile{introduction.tex}
\subfile{algorithms.tex}

\bibliography{adaptive}
\bibliographystyle{icml2014}

\end{document}

### 2014-01-25 lp/summary-of-results.tex
initial paper

@@
 \end{corollary}
+This result is a slight strengthening of one of the main results in the Hazan-Kale paper I showed you today. 
+The places where it is stronger are that the constant is much better (additive $\log(T)$ rather than 
+multiplicative factor of $16$) and also the regret is just in terms of the variance of the best expert, 
+whereas before it was a more complicated metric in terms of largest cumulative variance of best expert 
+at any point in time.
 
@@
 contain any hidden factors of $n$ or $\eta$. In fact we could take $O(\log(T)) = 8\log(T) + 2$.
+As far as I'm aware, both the theorem and the corollary in the matrix case are new. But 
+a lot of the work done on this is in the theory community and I haven't reviewed 
+references on matrix multiplicative weights nearly as thoroughly as for online learning.
 

### 2014-01-25 lp/mw12.tex
added in part of mw section

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Two Types of Updates} 
\label{sec:mw12}
Our point of departure is the two different types of multiplicative updates 
mentioned in the introduction. For simplicity we will consider the setting of 
learning from experts, although our later theorems will hold in greater 
generality. In this setting there are $n$ experts, and the learner maintains a 
probability distribution $w_t \in \Delta_n$ over the experts, which gets 
updated in each of $T$ rounds. After selecting the distribution $w_t$, 
the loss vector $w_t \in [-1,1]^n$ of each of the experts is revealed, and the 
learner incurs loss $w_t^{\top}z_t$. The learner's goal is to minimize 
the regret $\sup_{w \in \Delta_n} \Regret(w)$, where
\[ \Regret(w) \eqdef \sum_{t=1}^T w_t^{\top}z_t - \sum_{t=1}^T w^{\top}z_t. \]
We consider two types of updates for the weight vector $w_t$. In both cases 
$w_1$ is initialized as $w_{1,i} = \frac{1}{n}$ for $1 \leq i \leq n$, but 
the updates differ, as in (\ref{eqn:mw1}) and (\ref{eqn:mw2}) below:
\begin{align}
\label{eqn:mw1}
w_{t+1,i} &\propto w_{t,i}\exp(-\eta z_{t,i})\\
\label{eqn:mw2}
w_{t+1,i} &\propto w_{t,i}(1-\eta z_{t,i})
\end{align}
The regret bounds for each of (\ref{eqn:mw1}) and (\ref{eqn:mw2}) are 
well-known \cite{people}, but we include them below for completeness.
\begin{theorem}
\label{thm:mw12}
For any $0 < \eta \leq \frac{1}{2}$, the updates (\ref{eqn:mw1}) and 
(\ref{eqn:mw2}) enjoy regret, respectively,
\begin{align}
\label{eqn:mw1-regret}
\Regret(w) &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \sum_{t=1}^T w_{t,i}z_{t,i}^2 \\
\label{eqn:mw2-regret}
\Regret(w) &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2
\end{align}
\end{theorem}
To understand why (\ref{eqn:mw2-regret}) may be a better bound than 
(\ref{eqn:mw1-regret}), suppose that the best expert has a loss identically 
equal to zero.\footnote{Of course, if we knew that this was the case ahead of 
time, there would be far better algorithms; we use this scenario purely for 
illustrative purposes.} Then the optimal $w$ places all mass on this expert, 
and (\ref{eqn:mw2-regret}) reduces to $\frac{\log(n)}{\eta} = 2\log(n)$ for 
$\eta = \frac{1}{2}$.

More formally, define a sequence of losses $z_t$ to be \emph{quasi-realizable} 
if one of the experts has identically zero loss and all other experts 
have non-negative cumulative loss, i.e. $\sum_{t=1}^T z_{t,i} \geq 0$. It is 
apparent by the argument above that (\ref{eqn:mw2}) achieves asymptotically 
constant (as a function of $T$) regret for any quasi-realizable sequence. In 
contrast, for \emph{any} step size $\eta$, there exists a quasi-realizable 
sequence such that (\ref{eqn:mw1}) suffers regret $\Omega(\sqrt{T})$:
\begin{proposition}
\label{prop:separation}
If the updates (\ref{eqn:mw1}) are used with a step size of $\eta$, then for 
any $T$ there exists a quasi-realizable sequence of losses and a vector $w$ 
such that $\Regret(w) = \Omega(\sqrt{T})$.
\end{proposition}
This establishes that the apparent separation between (\ref{eqn:mw1}) and 
(\ref{eqn:mw2}) is real and not an artifact of the analysis. We remark that 
this separation does \emph{not} exist in the realizable case (when all losses 
are non-negative). In this case both (\ref{eqn:mw1}) and (\ref{eqn:mw2}) enjoy 
$O(1)$ regret as a function of $T$.

Finally, note that (\ref{eqn:mw2}) cannot be realized as mirror descent for any 
fixed regularizer. This is because, for any mirror descent algorithm, the 
prediction at time $t+1$ must be a function of $\sum_{t=1}^T z_t$, which is 
not the case for (\ref{eqn:mw2}). Instead, we obtain (\ref{eqn:mw2}) in terms of 
an \emph{adaptive regularizer} $\psi_t(w)$. The mirror descent updates for 
such a regularizer are given by 
\[ w_{t} = \nabla \psi_t^*\left(-\eta \sum_{s=1}^{t-1} z_s\right). \]
See \cite{orabona2013general} for a more complete exposition. For convenience, 
define $\theta_t \eqdef -\eta \sum_{s=1}^{t-1} \theta_s$. The important 
observation in our context is that, by definition of $\psi_{T+1}^*$, 
\begin{align}
- \eta w^{\top} \sum_{t=1}^T z_t &\leq \psi_{T+1}(w) + \psi_{T+1}^*(\theta_{T+1}) \\
 &= \psi_{T+1}(w) + \psi_1^*(\theta_1) + \sum_{t=1}^T \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t).
\end{align}
Motivated by this, we note that if we could choose $\psi_{t+1}$ so that 
$\psi_{t+1}^*(\theta_{t+1}) \leq \psi_{t}^*(\theta_t) - \eta w_t^{\top}z_t$, 
then re-arranging would yield a regret bound of $\frac{\psi_1^*(\theta_1) + \psi_{T+1}(w)}{\eta}$.
\begin{proposition}
Define $\lambda_{t,i} \eqdef \sum_{s=1}^{t-1} \log(1-\eta z_{s,i})$ and let
\[ \psi_t(w) \eqdef \sum_{i=1}^n w_i\log(w_i) + (\theta_t - \lambda_t)^{\top}w. \]
Then adaptive mirror descent with regularizer $\psi_t$ corresponds exactly to 
the updates (\ref{eqn:mw2}), and 
$\psi_{t+1}^*(\theta_{t+1}) \leq \psi_t^*(\theta_t) - \eta w_t^{\top}z_t$. 
The corresponding regret bound is
\begin{align}
\Regret(w) &\leq \frac{\psi_1^*(\theta_1) + \psi_{T+1}(w)}{\eta} \\
% &= \frac{\log(n)+\sum_{i=1}^n w_i\log(w_i)}{\eta} + \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i} + \frac{\log(1-\eta z_{t,i})}{\eta} \\
 &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2.
\end{align}
\end{proposition}
\begin{proof}
By standard properties of Fenchel conjugates, we have
\begin{align}
\nabla \psi_t^*(\theta_t) &= \argmin_w \psi_t(w) - \theta_t^{\top}w \\
 &= \argmin_w \sum_{i=1}^n w_i\log(w_i) - \lambda_t^{\top}w.
\end{align}
From here we see that 
$w_{t,i} \propto \exp(\lambda_{t,i}) = \prod_{s=1}^{t-1} (1-\eta z_{s,i})$, 
so that $w_{t,i}$ does indeed correspond to the update in (\ref{eqn:mw2}). 
\end{proof}
\end{document}

### 2014-01-25 lp/algorithms.tex

@@
 
+\begin{theorem}
+Suppose that $\|z_t\|_{\infty} \leq 1$ and $\eta \leq \frac{1}{2}$. 
+Consider the updates and predictions given by:
+\begin{itemize}
+\item $x_{1,i} = -\log(n)$
+\item $m_{t} = \frac{1}{t} \sum_{s=1}^{t-1} z_s$
+\item $w_{t,i} = \exp(x_{t,i} - \eta m_{t,i})$
+\item $x_{t+1,i} = x_{t,i} - \eta z_{t,i} - \eta^2 (z_{t,i} - m_{t,i})^2$
+\end{itemize}
+Then 
+\begin{align}
+\nonumber
+\lefteqn{\sum_{t=1}^T w_t^Tz_t \leq} \\
+ & \sum_{t=1}^T w^Tz_t + \frac{\log(n)}{\eta} + \eta \left[ 4\log(T+1) + \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_i)^2 \right].
+\end{align}
+\end{theorem}
+Note that this algorithm constraints $w$ to be non-negative. If this is 
+undesirable, we can overcome this restriction by writing $w = w_{+} - w_{-}$ 
+and making separate updates to $w_{+}$ and $w_{-}$. (The $w_{-}$ updates behave 
+as if they had observed a gradient of $-z_t$ rather than $z_t$. This is explained 
+in more detail in \cite{kivinen1997}.)
+
+Our second algorithm pertains to the matrix case:
+\begin{theorem}
+Suppose that $Z_t$ is symmetric and has eigenvalues bounded between $-1$ and $1$. 
+Let $\eta \leq \frac{1}{2}$. Consider the updates and predictions given by:
+\begin{itemize}
+\item $X_1 = -\log(n)I$
+\item $M_t = \frac{1}{t} \sum_{s=1}^{t-1} Z_s$
+\item $W_t = \exp(X_t - \eta M_t)$
+\item $X_{t+1} = X_t - \eta Z_t - \eta^2 (Z_t - M_t)$
+\end{itemize}
+Then
+\begin{align}
+\sum_{t=1}^T \Tr(W_t^{\top}Z_t) \leq & \sum_{t=1}^T \Tr(W^{\top}Z_t) + \frac{\log(n)}{\eta} \\
+ & + \eta \left[ 4\log(T+1) + \sum_{t=1}^T \Tr(W(Z_t-M)^2)\right].
+\end{align}
+\end{theorem}
+Moreover, both of these bounds continue to hold for the normalized variants of the 
+algorithms, e.g. if we replace $w_t$ with $w_t/\|w_t\|_1$ or $W_t$ with 
+$W_t/\Tr(W_t)$.
+
 \end{document}

### 2014-01-25 lp/introduction.tex

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Introduction} 
\label{sec:intro}
The exponentiated gradient algorithm is a powerful tool for performing online 
learning in the presence of many irrelevant features 
\cite{kivinen1997, littlestone1988}. One common application of exponentiated 
gradient is to the learning from experts setting, in which it is also known 
as the weighted majority algorithm \cite{wm} and entertains regret bounds of the 
form 
\[ \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \sum_{t=1}^T w_{t,i}z_{t,i}^2, \]
where $w_t$ is the weight vector of the learner at time $t$ and $z_t$ is the 
vector of losses. Such bounds can be obtained under the mirror descent 
framework using a local norms analysis \cite{shalev2011}.

In contrast, \cite{cesa2007} present a variant 
of this algorithm based on a multiplicative update of $(1 - \eta z_{t,i})$ rather 
than the usual update of $\exp(-\eta z_{t,i})$. This algorithm cannot be 
straightforwardly cast in the mirror descent framework, yet achieves an 
improved  regret bound of
\begin{equation}
\label{eqn:mw-regret}
\frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2.
\end{equation}
The difference is that the regret is in terms of the competing expert $w$ instead 
of the learner's strategy $w_t$. This latter bound can sometimes be much 
stronger; we show in Proposition~\ref{prop:separation} that there is a 
$\Theta(\sqrt{T})$ separation in worst-case regret in what we refer to as the 
``quasi-realizable'' setting. The difference between these two types of updates 
is also discussed in \cite{mw-survey}. % TODO: should we actually cite this?

The fact that an algorithm achieving a better regret bound cannot be cast in the 
mirror descent framework is a bit unsettling. Does this mean that we should 
abandon mirror descent as the gold standard for online learning, despite 
theorems asserting its optimality \cite{srebro2011}? We answer this question 
in the negative: the $1-\eta z_{t,i}$ can be understood as an \emph{adaptive 
mirror descent algorithm}: one where the regularizer changes in each round 
in response to previously observed vectors $z_t$. We obtain a 
particularly natural interpretation of the updates as performing a form of 
second-order correction on the gradient step.

By exploiting this connection, we show that the perhaps more natural update 
of $\exp(-\eta z_{t,i} - \eta^2 z_{t,i}^2)$ achieves the same regret bound 
as in (\ref{eqn:mw-regret}), and that both updates continue to enjoy this 
regret even when generalized from the simplex to an arbitrary convex set 
(including all of $\bR^n$, which corresponds to unconstrained exponentiated 
gradient).

Examining (\ref{eqn:mw-regret}) more closely, we see that we should expect 
to perform well when the best expert $i$ has losses consistently close to zero: 
then we set $w_j$ to $\delta_{ij}$ and the second term in (\ref{eqn:mw-regret}) 
becomes $\sum_{t=1}^T z_{t,i}^2 \approx 0$. However, this assumption may be 
unrealistic and a more realistic assumption is that the best expert's losses 
have low variance: $\sum_{t=1}^T (z_{t,i} - m)^2$ is small for some $m$. More 
generally, we would like to obtain a regret bound of the form
\begin{equation}
\label{eqn:variance-regret}
\frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i}-m_i)^2
\end{equation}
with $m$ chosen to be the optimal value of 
$m^* \eqdef \frac{1}{T} \sum_{t=1}^T z_t$.

We achieve such a regret bound by combining two ideas: the first is 
Rakhlin and Sridharan's framework of \emph{optimistic learning} 
\cite{rakhlin2012}, which in our setting yields bounds of the form 
\begin{equation}
\label{eqn:optimistic-regret}
\frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_{t,i})^2.
\end{equation}
If we could set $m_t$ to be $m^*$, then we would achieve the bound in 
(\ref{eqn:variance-regret}). Unfortunately, $m^*$ is only known after observing 
the full sequence of gradients. To overcome this, we learn $m^*$ itself in an 
online fashion, choosing the $m_t$ according to a Follow the Regularized Leader 
(FTRL) algorithm \cite{ftrl}. The regret bound thus obtained is given in 
Theorem~\ref{thm:variance-vector}, which improves a result by 
\cite{hazan2010variation}.

Finally, we extend all these results to the matrix setting, where $W_t$ is now 
a positive semidefinite matrix, possibly constrained to have trace $1$ 
(in analogy with the simplex). This setting has already been extensively studied 
\cite{lotsofpeople} and is important in obtaining online bounds for various 
combinatorial optimization problems \cite{morepeople}. Using the machinery so 
far developed, all of our results extend straightforwardly to the matrix setting, 
with the exception of the auxiliary FTRL algorithm for estimating $M^*$. In this 
case we need a new algorithm and analysis tool: a variant of FTRL for 
\emph{vector-valued} losses ordered relative to some cone $\sK$. This generalized 
FTRL lemma is given in Lemma~\ref{lem:ftrl-k} and may be of independent interest.

For convenience, we summarize our contributions below:
\begin{itemize}
\item An interpretation of the multiplicative weights update as 
      exponentiated gradient with an adaptive regularizer.
\item An improved exponentiated gradient algorithm based on adaptive 
      updates to an entropic regularizer.
\item Further improvements to exponentiated gradient based on optimistic 
      predictions and online learning of the mean gradient.
\item An adaptive matrix exponentiated gradient, with similar improvements 
      based on learning the mean gradient.
\item A generalization of Follow the Regularized Leader to vector-valued 
      loss functions.
\end{itemize}

\paragraph{Related work.} Our work contributes to an already rich literature 
on using adaptive regularizers to obtain 
better regret bounds for online learning algorithms. A common setting is 
adaptive learning of a quadratic regularizer, as in the AROW 
\cite{crammer2009arow}, AdaGrad \cite{duchi2011adagrad}, and online 
preconditioning \cite{streeter2010} algorithms. 
Other work includes dimension-free 
exponentiated gradient \cite{orabona2013dimension} (which, however, 
still performs essentially additive updates) and whitened perceptron 
\cite{cesa2005perceptron}. The non-stationary setting has been explored 
in \cite{vaits2013}, and \cite{mcmahan2010adaptive} obtain regret bounds 
relative to an entire family of regularizers. More recently, many of these 
algorithms have been unified into a single framework by \cite{orabona2013general}.
To our knowledge, adaptively regularized exponentiated gradient has not 
been explored, though \cite{hazan2010variation} provide an ad hoc algorithm 
that is similar to one of ours.

Other notions of ``adaptive'' weights have also been explored; though unrelated 
to the current work, we briefly mention them for completeness. The first notion 
notion of adaptivity pertains to time-varying experts: here the desire is to 
obtain regret relative to the best \emph{schedule} of experts, where the number 
of switches in the schedule is bounded. This is explored in \cite{switching}. 
The other slightly more related notion is an adaptive step size, where the step 
size $\eta$ changes as a function of time (often at a fixed schedule of 
$1/\sqrt{T}$, but sometimes in a more sophisticated fashion). This idea is used 
extensively throughout the literature but a few particularly relevant papers are 
\cite{adaptive-step}.

Finally, we note that the exponentiated gradient itself has been analyzed 
and applied in various settings. One common setting, using a hinge-like loss, 
is the Winnow Algorithm \cite{winnow}, which has been carefully analyzed in 
\cite{winnow-careful}.

\end{document}

### 2014-01-25 lp/matrix-theorem.tex

\documentclass[11pt]{article}
\usepackage{import}
\input{latex-defs.tex}
\DeclareMathOperator{\Regret}{Regret}
\usepackage{fullpage}

\begin{document}
\section{Proof for Matrix Setting}
We assume that $-I \preceq Z_t \preceq I$  and that $\eta \leq \frac{1}{2}$. We 
consider $W$ bounded so that $W \succeq 0$, $\Tr(W) \leq 1$. Thus when we write 
$W \propto A$ we mean that $W = \frac{A}{\Tr{A}}$.
\begin{theorem}
\label{thm:matrix}
For a sequence $S_1, \ldots, S_T$ of symmetric matrices, consider the following 
prediction algorithm:
\begin{enumerate}
\item Predict $W_t \propto \exp(X_t - \eta S_t)$.
\item Receive loss $\Tr(W_t^{\top}Z_t)$.
\item Update $X_{t+1} = X_t - \eta Z_t - \eta^2 (Z_t-S_t)^2$.
\end{enumerate}
Then the regret relative to a fixed matrix $W$ is bounded by
\[ \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \Tr((Z_t-S_t)^{\top}W(Z_t-S_t)). \]
\end{theorem}
% To prove the theorem, we will make use of the following lemma:
% \begin{lemma}
% \label{lem:main}
% Let $\psi(W)$ be a convex function that is bounded below, and 
% let $\psi^*(X) \eqdef \sup_{W} \Tr(W^TX) - \psi(W)$. Let $S$ be any convex set. 
% 
% Suppose that 
% we generate a sequence $X_t$ via $X_0 = 0$ and 
% $X_{t+1} = X_t - \eta Z_t - \eta^2 U_t$, which satisfies the 
% property 
% \[ \psi^*(X_t - \eta Z_t - \eta^2 U_t) \leq \psi^*(X_t) - \eta \partial \psi^*(X_t)^TZ_t. \]
% Then, defining $W_t \eqdef \arg\max_{\tilde{W} \in S} \Tr(\tilde{W}^TX) - \psi(\tilde{W})$, for any $W \in S$ we have 
% the bound
% \[ \sum_{t=1}^T \Tr(W_t^TX_t) - \sum_{t=1}^T \Tr(W^TX_t) \leq \frac{\psi(W) + \psi^*(0)}{\eta} + \eta \sum_{t=1}^T \Tr(W^TU_t). \]
% \end{lemma}

\begin{proof}[Proof of Theorem~\ref{thm:matrix}]
Define $\psi(W) = \Tr(W[\log(W)-I])$ and $\psi^*(X) \eqdef \sup_{W \succeq 0} \Tr(W^{\top}X) - \psi(W) = \Tr(\exp(X))$. 
Note that $\partial \psi^*(X)^{\top}Z = \Tr(\exp(X)Z)$.

We will prove a pair of bounds that will then allow us to obtain a regret bound by telescoping $\psi^*(X_t)$. 
Our first bound is:
\begin{equation}
\label{eqn:bound-1}
\Tr(\exp(X_t - \eta Z_t - \eta^2 (Z_t - S_t)^2 )) \leq \Tr(\exp(X_t - \eta S_t)) - \eta \Tr(\exp(X_t - \eta S_t)(Z_t - S_t)).
\end{equation}
This bound is the crux of our argument (the rest of the manipulations are standard and don't make use of 
any specific properties of our setting other than convexity of $\psi$). We can verify the bound 
using the Golden-Thompson inequality $\Tr(\exp(A+B)) \leq \Tr(\exp(A)\exp(B))$ together 
with the fact that $-X-X^2 \preceq \log(I-X)$ for $-\frac{1}{2}I \preceq X \preceq \frac{1}{2}I$:
\begin{align}
\Tr(\exp(X_t - \eta Z_t - \eta^2 (Z_t - S_t)^2)) &= \Tr(\exp((X_t - \eta S_t) - \eta (Z_t - S_t) - \eta^2 (Z_t - S_t)^2)) \\
 &\leq \Tr(\exp(X_t - \eta S_t)\exp(-\eta (Z_t - S_t) - \eta^2 (Z_t - S_t)^2)) \\
 &\leq \Tr(\exp(X_t - \eta S_t)\exp(\log(I - \eta (Z_t - S_t)))) \\
 &= \Tr(\exp(X_t - \eta S_t)(I - \eta (Z_t - S_t)) \\
 &= \Tr(\exp(X_t - \eta S_t)) - \eta \Tr(\exp(X_t - \eta S_t)(Z_t - S_t)).
\end{align}
Our second bound is
\begin{equation}
\label{eqn:bound-2}
\Tr(\exp(X_t - \eta S_t)) \leq \Tr(\exp(X_t)) - \eta \Tr(\exp(X_t - \eta S_t)S_t).
\end{equation}
This is just saying that $\psi^*(X_t - \eta S_t) \leq \psi^*(X_t) - \eta \langle \partial \psi^*(X_t - \eta S_t), S_t \rangle$, 
which follows directly from convexity of $\psi^*$.

Combining (\ref{eqn:bound-1}) and (\ref{eqn:bound-2}) and defining $\tilde{W}_t \eqdef \exp(X_t - \eta S_t)$ yields
\begin{equation}
\label{eqn:bound-3}
\Tr(\exp(X_{t+1})) \leq \Tr(\exp(X_t)) - \eta \Tr(\tilde{W}_t^{\top}Z_t).
\end{equation}
We thus have
\begin{align}
\sum_{t=1}^T \Tr(\tilde{W}_t^{\top}Z_t) & \leq \frac{\psi^*(0) - \psi^*(X_{T+1})}{\eta} \\
 &\leq \frac{\psi^*(0) + \psi(W) + \sum_{t=1}^T \Tr(W^{\top}(\eta Z_t + \eta^2 (Z_t - S_t)^2))}{\eta} \\
 &\leq \frac{\psi^*(0) + \psi(W)}{\eta} + \sum_{t=1}^T \Tr(W^{\top}Z) + \eta \sum_{t=1}^T \Tr(W^{\top}(Z_t - S_t)^2) \\
 &= \frac{\psi^*(0) + \psi(W)}{\eta} + \sum_{t=1}^T \Tr(W^{\top}Z) + \eta \sum_{t=1}^T \Tr((Z_t-S_t)^{\top}W(Z_t - S_t)).
\end{align}
We thus see that, had we predicted with $\tilde{W}_t$, we would have essentially the desired regret bound. 
However, we instead predicted with $W_t = \frac{\tilde{W}_t}{\Tr(\tilde{W}_t)}$, and so we require one additional step.

This additional step is captured in the following lemma:
%% \begin{lemma}
%% \label{lem:project}
%% Define $\psi_{\sH}^*(X) \eqdef \sup_{W \in \sH} \Tr(W^{\top}X) - \psi(W)$. Then
%% \[ \psi_{\sH}^*(X
(TODO: add in additional step, may not be easily stated as a separate lemma unfortunately)
\end{proof}

\section{FTRL-$\sK$}

In order to fully exploit Theorem~\ref{thm:matrix}, we need a 
generalization of the Follow the Regularized Leader lemma, where 
losses are vector-valued and are ordered by some cone, rather than 
via the usual total ordering on $\bR$. To deal with the cone structure, 
we define the notion of a \emph{global minimizer}:
\begin{definition}
Let $\sK$ be a cone in $\bR^d$ and let $f : \bR^n \to \bR^d$ be a 
function. We call $X$ a \emph{global minimizer} of $f$ if $f(X) \leq_{\sK} f(X')$ 
for all $X'$.
\end{definition}

We prove the following:
\begin{lemma}
\label{lem:ftrl-k}
Suppose that for all $1 \leq t \leq T+1$, there 
exists a global minimizer $M_t$ of 
$\psi(X) + \sum_{s=1}^{t-1} f_s(X)$. Then for all $M$,
\[ \sum_{t=1}^T f_t(M_t) \leq_{\sK} \psi(M) - \psi(M_1) + \sum_{t=1}^T f_t(M) + \sum_{t=1}^T f_t(M_{t}) - f_t(M_{t+1}). \]
\end{lemma}
\begin{proof}
We will prove the lemma by induction on $T$. Note that the lemma is equivalent to showing that 
\[ \psi(M_1) + \sum_{t=1}^T f_t(M_{t+1}) \leq_{\sK} \psi(M) + \sum_{t=1}^T f_t(M) \]
for all $M$.

In the base case $T = 0$, we have
\[ \psi(M_1) \leq_{\sK} \psi(M), \]
which follows from the fact that $M_1$ is a global minimizer of $\psi$ and 
hence $\psi(M_1) \leq_{\sK} \psi(M)$ for all $M$. For the inductive step, suppose that
\[ \sum_{t=1}^{T-1} f_t(M_{t+1}) \leq_{\sK} \psi(M) \sum_{t=1}^{T-1} f_t(M) \]
for all $M$, and invoke this for the particular choice $M = M_{T+1}$. Then we have
\begin{align}
\psi(M_1) + \sum_{t=1}^T f_t(M_{t+1}) &= \psi(M_1) + \sum_{t=1}^{T-1} f_t(M_{t+1}) + f_T(M_{T+1}) \\
 &\leq_{\sK} \psi(M_{T+1}) + \sum_{t=1}^{T-1} f_t(M_{T+1}) + f_T(M_{T+1}) \\
 &= \psi(M_{T+1}) + \sum_{t=1}^T f_t(M_{T+1}) \\
 &\leq_{\sK} \psi(M) + \sum_{t=1}^T f_t(M)
\end{align}
for all $M$, where we use the fact that $M_{T+1}$ is a global minimizer of $\psi(M) + \sum_{t=1}^T f_t(M)$ for 
the last inequality. This completes the inductive hypothesis and hence the proof.
\end{proof}

The condition that a global minimizer always exists may seem somewhat strong, so we now invoke it for a concrete case.
\begin{corollary}
\label{cor:ftrl-matrix}
Let $\sK$ be the semidefinite cone, $\psi(M) = M^{\top}M$, and $f_t(M) = (M - Z_t)^{\top}(M - Z_t)$. 
Then the unique global minimizer of $\psi(M) + \sum_{s=1}^{t-1} f_t(M)$ is $M_t \eqdef \frac{1}{t} \sum_{s=1}^{t-1} Z_s$. 
In particular, 
\[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq \sum_{t=1}^T (Z_t - M_{T+1})^2 + 4\log(T+1)I. \]
\end{corollary}

\begin{proof}
To show that $M_t$ is the global minimizer, consider some other value $\tilde{M}_t = M_t + D_t$. Then we have 
\begin{align}
\lefteqn{\tilde{M}_t^{\top}\tilde{M}_t + \sum_{s=1}^{t-1} (Z_s - \tilde{M}_t)^{\top}(Z_s - \tilde{M}_t)} \\
 &= (M_t + D_t)^{\top}(M_t + D_t) + \sum_{s=1}^{t-1} (Z_s - M_t - D_t)^{\top}(Z_s - M_t - D_t) \\
 &= M_t^{\top}M_t + \sum_{s=1}^{t-1} (Z_s - M_t)^{\top}(Z_s - M_t) + tD_t^{\top}D_t + D_t^{\top}\left[ M_t + \sum_{s=1}^{t-1} (Z_s - M_t) \right] + \left[ M_t + \sum_{s=1}^{t-1} (Z_s - M_t)\right]^{\top} D_t \\
 &= M_t^{\top}M_t + \sum_{s=1}^{t-1} (Z_s - M_t)^{\top}(Z_s - M_t) + tD_t^{\top}D_t \\
 &\succeq M_t^{\top}M_t + \sum_{s=1}^{t-1} (Z_s - M_t)^{\top}(Z_s - M_t).
\end{align}
Since this holds for all $D_t$, $M_t$ is indeed a global minimizer. Thus, by Lemma~\ref{lem:ftrl-k}, we have
\[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq \sum_{t=1}^T (Z_t - M_{T+1})^2 + \sum_{t=1}^T (Z_t-M_t)^2 - (Z_t-M_{t+1})^2. \]
To complete the corollary, we need only show that $(Z_t-M_t)^2 - (Z_t-M_{t+1})^2 \preceq \frac{4}{t+1}I + M_t^2 - M_{t+1}^2$ 
(since the $M_t^2 - M_{t+1}^2$ terms will telescope to $M_1^2 - M_{T+1}^2 = -M_{T+1}^2 \preceq 0$).
To show this, note that $M_{t+1} - M_t = \frac{1}{t+1}(Z_t - M_t)$ and hence
\begin{align}
(Z_t - M_t)^2 - (Z_t - M_{t+1})^2 &= (M_{t+1}-M_t)Z_t + Z_t(M_{t+1}-M_t) + M_t^2 - M_{t+1}^2 \\
  &= \frac{1}{t+1}\left[(Z_t - M_t)Z_t + Z_t(Z_t-M_t)\right] + M_t^2 - M_{t+1}^2 \\
  &\preceq \frac{4}{t+1}I + M_t^2 - M_{t+1}^2.
\end{align}
This completes the corollary.
\end{proof}

% \begin{corollary}
% If we set $S_t = \frac{1}{t} \sum_{s=1}^{t-1} Z_s$ (so $S_1 = 0$, $S_2 = \frac{1}{2}Z_1$, $S_3 = \frac{1}{3}(Z_1+Z_2)$, etc.), 
% then we obtain the bound
% \[ \sum_{t=1}^T \Tr(W_t^TZ_t) - \sum_{t=1}^T \Tr(W^TZ_t) \leq \frac{\log(n)+\Tr(W\log(W))}{\eta} + \eta \sum_{t=1}^T \Tr((Z_t-M)^TW(Z_t-M)) + O(1), \]
% where $M \eqdef \frac{1}{T} \sum_{t=1}^T Z_t$.
% \end{corollary}
% \begin{proof}
% We will show that $\sum_{t=1}^T \Tr((Z_t-S_t)^2) \leq \sum_{t=1}^T \Tr((Z_t-M)^2) + O(1)$. (This follows from the fact that 
% $S_{t} = \argmin_S \Tr(S^2) + \sum_{s=1}^{t-1} \Tr((Z_s-S)^2)$, and hence can be interpreted as Follow the Regularized Leader.)
% 
% Next, we use the fact that, 
% \end{proof}

\end{document}

### 2014-01-25 lp/machinery.tex
done with mw12 draft

--- 2014-01-25 lp/introduction.tex
@@
 
-\section{Introduction} 
-\label{sec:intro}
-The exponentiated gradient algorithm is a powerful tool for performing online 
-learning in the presence of many irrelevant features 
-\cite{kivinen1997, littlestone1988}. One common application of exponentiated 
-gradient is to the learning from experts setting, in which it is also known 
-as the weighted majority algorithm \cite{wm} and entertains regret bounds of the 
-form 
-\[ \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \sum_{t=1}^T w_{t,i}z_{t,i}^2, \]
-where $w_t$ is the weight vector of the learner at time $t$ and $z_t$ is the 
-vector of losses. Such bounds can be obtained under the mirror descent 
-framework using a local norms analysis \cite{shalev2011}.
-
-In contrast, \cite{cesa2007} present a variant 
-of this algorithm based on a multiplicative update of $(1 - \eta z_{t,i})$ rather 
-than the usual update of $\exp(-\eta z_{t,i})$. This algorithm cannot be 
-straightforwardly cast in the mirror descent framework, yet achieves an 
-improved  regret bound of
-\begin{equation}
-\label{eqn:mw-regret}
-\frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2.
-\end{equation}
-The difference is that the regret is in terms of the competing expert $w$ instead 
-of the learner's strategy $w_t$. This latter bound can sometimes be much 
-stronger; we show in Proposition~\ref{prop:separation} that there is a 
-$\Theta(\sqrt{T})$ separation in worst-case regret in what we refer to as the 
-``quasi-realizable'' setting. The difference between these two types of updates 
-is also discussed in \cite{mw-survey}. % TODO: should we actually cite this?
-
-The fact that an algorithm achieving a better regret bound cannot be cast in the 
-mirror descent framework is a bit unsettling. Does this mean that we should 
-abandon mirror descent as the gold standard for online learning, despite 
-theorems asserting its optimality \cite{srebro2011}? We answer this question 
-in the negative: the $1-\eta z_{t,i}$ can be understood as an \emph{adaptive 
-mirror descent algorithm}: one where the regularizer changes in each round 
-in response to previously observed vectors $z_t$. We obtain a 
-particularly natural interpretation of the updates as performing a form of 
-second-order correction on the gradient step.
-
-By exploiting this connection, we show that the perhaps more natural update 
-of $\exp(-\eta z_{t,i} - \eta^2 z_{t,i}^2)$ achieves the same regret bound 
-as in (\ref{eqn:mw-regret}), and that both updates continue to enjoy this 
-regret even when generalized from the simplex to an arbitrary convex set 
-(including all of $\bR^n$, which corresponds to unconstrained exponentiated 
-gradient).
-
-Examining (\ref{eqn:mw-regret}) more closely, we see that we should expect 
-to perform well when the best expert $i$ has losses consistently close to zero: 
-then we set $w_j$ to $\delta_{ij}$ and the second term in (\ref{eqn:mw-regret}) 
-becomes $\sum_{t=1}^T z_{t,i}^2 \approx 0$. However, this assumption may be 
-unrealistic and a more realistic assumption is that the best expert's losses 
-have low variance: $\sum_{t=1}^T (z_{t,i} - m)^2$ is small for some $m$. More 
-generally, we would like to obtain a regret bound of the form
-\begin{equation}
-\label{eqn:variance-regret}
-\frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i}-m_i)^2
-\end{equation}
-with $m$ chosen to be the optimal value of 
-$m^* \eqdef \frac{1}{T} \sum_{t=1}^T z_t$.
-
-We achieve such a regret bound by combining two ideas: the first is 
-Rakhlin and Sridharan's framework of \emph{optimistic learning} 
-\cite{rakhlin2012}, which in our setting yields bounds of the form 
-\begin{equation}
-\label{eqn:optimistic-regret}
-\frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_{t,i})^2.
-\end{equation}
-If we could set $m_t$ to be $m^*$, then we would achieve the bound in 
-(\ref{eqn:variance-regret}). Unfortunately, $m^*$ is only known after observing 
-the full sequence of gradients. To overcome this, we learn $m^*$ itself in an 
-online fashion, choosing the $m_t$ according to a Follow the Regularized Leader 
-(FTRL) algorithm \cite{ftrl}. The regret bound thus obtained is given in 
-Theorem~\ref{thm:variance-vector}, which improves a result by 
-\cite{hazan2010variation}.
-
-Finally, we extend all these results to the matrix setting, where $W_t$ is now 
-a positive semidefinite matrix, possibly constrained to have trace $1$ 
-(in analogy with the simplex). This setting has already been extensively studied 
-\cite{lotsofpeople} and is important in obtaining online bounds for various 
-combinatorial optimization problems \cite{morepeople}. Using the machinery so 
-far developed, all of our results extend straightforwardly to the matrix setting, 
-with the exception of the auxiliary FTRL algorithm for estimating $M^*$. In this 
-case we need a new algorithm and analysis tool: a variant of FTRL for 
-\emph{vector-valued} losses ordered relative to some cone $\sK$. This generalized 
-FTRL lemma is given in Lemma~\ref{lem:ftrl-k} and may be of independent interest.
-
-For convenience, we summarize our contributions below:
-\begin{itemize}
-\item An interpretation of the multiplicative weights update as 
-      exponentiated gradient with an adaptive regularizer.
-\item An improved exponentiated gradient algorithm based on adaptive 
-      updates to an entropic regularizer.
-\item Further improvements to exponentiated gradient based on optimistic 
-      predictions and online learning of the mean gradient.
-\item An adaptive matrix exponentiated gradient, with similar improvements 
-      based on learning the mean gradient.
-\item A generalization of Follow the Regularized Leader to vector-valued 
-      loss functions.
-\end{itemize}
-
-\paragraph{Related work.} Our work contributes to an already rich literature 
-on using adaptive regularizers to obtain 
-better regret bounds for online learning algorithms. A common setting is 
-adaptive learning of a quadratic regularizer, as in the AROW 
-\cite{crammer2009arow}, AdaGrad \cite{duchi2011adagrad}, and online 
-preconditioning \cite{streeter2010} algorithms. 
-Other work includes dimension-free 
-exponentiated gradient \cite{orabona2013dimension} (which, however, 
-still performs essentially additive updates) and whitened perceptron 
-\cite{cesa2005perceptron}. The non-stationary setting has been explored 
-in \cite{vaits2013}, and \cite{mcmahan2010adaptive} obtain regret bounds 
-relative to an entire family of regularizers. More recently, many of these 
-algorithms have been unified into a single framework by \cite{orabona2013general}.
-To our knowledge, adaptively regularized exponentiated gradient has not 
-been explored, though \cite{hazan2010variation} provide an ad hoc algorithm 
-that is similar to one of ours.
-
-Other notions of ``adaptive'' weights have also been explored; though unrelated 
-to the current work, we briefly mention them for completeness. The first notion 
-notion of adaptivity pertains to time-varying experts: here the desire is to 
-obtain regret relative to the best \emph{schedule} of experts, where the number 
-of switches in the schedule is bounded. This is explored in \cite{switching}. 
-The other slightly more related notion is an adaptive step size, where the step 
-size $\eta$ changes as a function of time (often at a fixed schedule of 
-$1/\sqrt{T}$, but sometimes in a more sophisticated fashion). This idea is used 
-extensively throughout the literature but a few particularly relevant papers are 
-\cite{adaptive-step}.
-
-Finally, we note that the exponentiated gradient itself has been analyzed 
-and applied in various settings. One common setting, using a hinge-like loss, 
-is the Winnow Algorithm \cite{winnow}, which has been carefully analyzed in 
-\cite{winnow-careful}.
+\section{Adaptive Optimistic Learning} 
+\label{sec:machinery}
 

### 2014-01-25 lp/mw12.tex
done with mw12 draft

@@
 - \eta w^{\top} \sum_{t=1}^T z_t &\leq \psi_{T+1}(w) + \psi_{T+1}^*(\theta_{T+1}) \\
+\label{eqn:telescope}
  &= \psi_{T+1}(w) + \psi_1^*(\theta_1) + \sum_{t=1}^T \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t).
@@
 \begin{proposition}
+\label{prop:correspondence}
 Define $\lambda_{t,i} \eqdef \sum_{s=1}^{t-1} \log(1-\eta z_{s,i})$ and let
@@
 \end{proposition}
+We will be able to prove this more easily with the machinery developed in 
+Section~\ref{sec:machinery}, but we also prove it here directly.
 \begin{proof}
@@
 so that $w_{t,i}$ does indeed correspond to the update in (\ref{eqn:mw2}).
+
+Next, we exploit the property that the Fenchel conjugate of $\psi(w) + c^{\top}w$ 
+is $\psi^*(\theta - c)$, together with the fact that the Fenchel conjugate of 
+negentropy is logsumexp, to obtain
+\begin{align}
+\psi_{t+1}^*(\theta_{t+1}) &= \log\left(\sum_{i=1}^n \exp(\theta_{t+1,i} - (\theta_{t+1,i}-\lambda_{t+1,i}))\right) \\
+ &= \log\left(\sum_{i=1}^n \exp(\lambda_{t+1,i})\right) \\
+ &= \log\left(\sum_{i=1}^n \exp(\lambda_{t,i} + \log(1 - \eta z_{t,i}))\right) \\
+ &= \log\left(\sum_{i=1}^n \exp(\lambda_{t,i})(1 - \eta z_{t,i})\right) \\
+ &= \log\left(\sum_{i=1}^n \exp(\lambda_{t,i}) - \eta \sum_{i=1}^n \exp(\lambda_{t,i})z_{t,i}\right) \\
+ &\leq \log\left(\sum_{i=1}^n \exp(\lambda_{t,i})\right) - \eta \frac{\sum_{i=1}^n \exp(\lambda_{t,i})z_{t,i}}{\sum_{i=1}^n \exp(\lambda_{t,i})} \\
+ &= \psi_t^*(\theta_t) - \eta w_t^{\top}z_t.
+\end{align}
+Thus by the observations following (\ref{eqn:telescope}), $\Regret(w)$ is bounded by 
+$\frac{\psi_1^*(\theta_1) + \psi_{T+1}(w)}{\eta}$. Now using the fact that 
+$-\eta z - \log(1-\eta z) \leq \eta^2 z^2$ when $|\eta z| \leq \frac{1}{2}$, and also the 
+fact that $\sum_i w_i\log(w_i) \leq 0$, we have
+\begin{align}
+\Regret(w) &\leq \frac{\psi_1^*(\theta_1) + \psi_{T+1}(w)}{\eta} \\
+ &= \frac{\log(n) + \sum_{i=1}^n w_i\log(w_i) + (\theta_{T+1}-\lambda_{T+1})^{\top}w}{\eta} \\
+ &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2,
+\end{align}
+which completes the proposition.
 \end{proof}
+In the next section, we will present a more general framework for \emph{adaptive updates}, 
+also incorporating \emph{optimistic prediction} \cite{rakhlin2012}, which will both simplify 
+subsequent proofs and also allow us to obtain better bounds.
 \end{document}

### 2014-01-25 lp/machinery.tex
done with machinery

@@
 \label{sec:machinery}
+In this section we present a framework for \emph{adaptive optimistic learning}. 
+This combines the framework of adaptive learning presented in \cite{orabona2013general} with 
+the framework of optimistic learning in \cite{rakhlin2012}. While the combination is 
+straightforward, we will show that in our case the combination of the two yields meaningfully 
+better results than using either one in isolation. The adaptive optimistic mirror descent 
+algorithm thus obtained is given in Algorithm~\ref{alg:adaptive-optimistic}.
+
+\begin{algorithm}
+\caption{Adaptive Optimistic Mirror Descent}
+\label{alg:adaptive-optimistic}
+\begin{algorithmic}
+\STATE $\theta_1 = 0$
+\FOR{$t=1$ {\bfseries to} $T$}
+  \STATE Choose $w_t = \nabla \psi_t^*(\theta_t - \eta m_t)$
+  \STATE Observe $z_t$ and suffer loss $w_t^{\top}z_t$
+  \STATE Update $\theta_{t+1} = \theta_t - \eta z_t$
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
+
+The key regret bound for Algorithm~\ref{alg:adaptive-optimistic} is 
+given in Theorem~\ref{thm:adaptive-optimistic}:
+\begin{theorem}
+Algorithm~\ref{alg:adaptive-optimistic} achieves regret bounded by
+\begin{equation}
+\frac{1}{\eta}\left(\psi_1^*(\theta_1) + \psi_{T+1}(w) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t - \eta m_t) + \eta w_t^{\top}m_t\right).
+\end{equation}
+\end{theorem}
+\begin{proof}
+TODO
+\end{proof}
+As an immediate consequence, we have
+\begin{corollary}
+\label{cor:a-o1}
+Suppose that $\psi_{t+1}$ is chosen so that $\psi_{t+1}^*(\theta_t - \eta z_t) \leq \psi_t^*(\theta_t - \eta m_t) - \eta w_t^{\top}(z_t - m_t)$. Then
+\begin{equation}
+\Regret(w) \leq \frac{\psi_1^*(\theta_1) + \psi_{T+1}(w)}{\eta}.
+\end{equation}
+\end{corollary}
+\begin{proof}
+TODO
+\end{proof}
+Finally, we consider a simplified version of Algorithm~\ref{alg:adaptive-optimistic}, 
+which corresponds to choosing $\psi_t(w) = \psi(w) + \eta^2 w^{\top} \sum_{s=1}^{t-1} u_s$ 
+for some fixed function $\psi$.
+
+\begin{algorithm}
+\caption{Adaptive optimistic mirror distance (additive version)}
+\label{alg:a-o2}
+
+\begin{algorithmic}
+\STATE Initialize $\theta_1$ arbitrarily
+\FOR{$t=1$ {\bfseries to} $T$}
+  \STATE Choose $w_t = \nabla \psi^*(\theta_t - \eta m_t)$
+  \STATE Observe $z_t$ and suffer loss $w_t^{\top}z_t$
+  \STATE Update $\theta_{t+1} = \theta_t - \eta z_t - \eta^2 u_t$
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
+
+\begin{corollary}
+\label{cor:a-o2}
+Suppose that $u_t$ is chosen so that $\psi^*(\theta_t - \eta z_t - \eta^2 u_t) \leq \psi^*(\theta_t - \eta m_t) - \eta w_t^{\top}(z_t - m_t)$. 
+Then 
+\begin{equation}
+\Regret(w) \leq \frac{\psi^*(\theta_1)}{\eta} + \eta w^{\top} \sum_{t=1}^T u_t.
+\end{equation}
+\end{corollary}
+\begin{proof}
+TODO
+\end{proof}
+To give some intuition for the condition in Corollary~\ref{cor:a-o2}, note that 
+$w_t = \nabla \psi^*(z_t - \eta m_t)$, and so, since $\psi^*$ is convex, we have
+$\psi^*(\theta_t - \eta z_t) \geq \psi^*(\theta_t - \eta m_t) - \eta w_t^{\top}(z_t - m_t)$. 
+The term $\eta^2 u_t$ is thus a second-order correction to flip the sign of the inequality 
+(by taking a Taylor expansion, one can show that $\eta^2$ is the right order of magnitude for 
+the correction term).
+
+\paragraph{Application to exponentiated gradient.} We now show how to apply 
+Corollary~\ref{cor:a-o2} to obtain an adaptive exponentiated gradient algorithm.
+\begin{proposition}[Adaptive Exponentiated Gradient]
+\label{prop:aeg}
+Consider the updates given by $\theta_{1,i} = -\log(n)$ and 
+$\theta_{t+1,i} = \theta_{t,i} - \eta z_{t,i} - \eta^2 (z_{t,i} - m_{t,i})^2$, 
+with prediction $w_{t,i} = \exp(\theta_{t,i} - \eta m_{t,i})$. For this 
+algorithm,
+\begin{equation}
+\Regret(w) \leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_{t,i})^2.
+\end{equation}
+\end{proposition}
+\begin{proof}
+TODO
+\end{proof}
+Note that in the case that $m_{t,i} = 0$, the update is 
+$\theta_{t+1,i} = \theta_{t,i} - \eta z_{t,i} - \eta^2 z_{t,i}^2$, whereas multiplicative 
+updates by $1 - \eta z_{t,i}$ would correspond to 
+$\theta_{t+1,i} = \theta_{t,i} + \log(1 - \eta z_{t,i})$. Since 
+$-\eta z_{t,i} - \eta^2 z_{t,i}^2 \leq \log(1 - \eta z_{t,i})$ when 
+$|\eta z_{t,i}| \leq \frac{1}{2}$, we can think of these updates as a second-order 
+under-approximation to (\ref{eqn:mw2}), which however can more naturally incorporate 
+the optimistic updates based on $m_t$.
+
+In the next section, we will show how to adaptively choose $m_t$ to optimize the regret bound 
+in Proposition~\ref{prop:aeg}.
 

### 2014-01-26 lp/ftrl-aux.tex
done with machinery
done with ftrl-aux

--- 2014-01-25 lp/machinery.tex
@@
 
-\section{Adaptive Optimistic Learning} 
-\label{sec:machinery}
+\section{Choosing $m_t$ via Follow the Regularized Leader} 
+\label{sec:ftrl-aux}
+In the previous section we introduced an adaptive exponentiated 
+gradient algorithm that achieved a regret bound of 
+\[ \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_{t,i})^2 \]
+for any sequence $m_t$. We now turn our attention to the task of 
+optimizing $m_t$. Clearly, the optimal value of $m_t$ is $m_t = z_t$. However, 
+this would require perfect knowledge of the future. Suppose we are less ambitious 
+and only want to be competitive with the best fixed value of $m_t$. This optimal value 
+is $m^* \eqdef \frac{1}{T} \sum_{t=1}^T z_t$, but we still have the issue that 
+we do not $m^*$ ahead of time. However, we have now cast the problem in the familiar 
+regret minimzation framework, and can use the machinery of follow the regularized 
+leader:
+\begin{proposition}
+\label{prop:ftrl-aux}
+Suppose that we choose $m_{t,i} = \frac{1}{t} \sum_{s=1}^{t-1} z_{s,i}$. 
+Then, for all $i$, 
+\[ \sum_{t`=1}^T (z_{t,i} - m_{t,i})^2 \leq \sum_{t=1}^T (z_{t,i} - m^*_i)^2 + 4\log(T+1). \]
+\end{proposition}
+\begin{proof}
+TODO
+\end{proof}
+Combining Proposition~\ref{prop:ftrl-aux} with Proposition~\ref{prop:aeg} yields 
+the algorithm titled EG-Variance in Table~\ref{tab:algorithms}. This algorithm has the 
+same updates as an algorithm given by \cite{hazan2010variation}, but uses the 
+optimistic prediction $\exp(\theta_t - \eta m_t)$ rather than $\exp(\theta_t)$. 
+As a straightforward consequence of Propositions~\ref{prop:aeg} and~\ref{prop:ftrl-aux}, we 
+obtain the following regret bound, which improves the bound presented in 
+\cite{hazan2010variation}:
+\begin{corollary}
+The EG-Variance algorithm achieves a regret bound of
+%\[ \Regret(w) \leq \frac{1 + (\log(n)-1)\|w\|_1 + \sum_{i=1}^n w_i\log(w_i)}{\eta} + \eta \left[ \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_i^*)^2 + 4\|w\|_1\log(T) \right]. \]
+\[ \Regret(w) \leq \frac{\log(n)}{\eta} + \eta \left[ \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_i^*)^2 + 4\log(T) \right]. \]
+\end{corollary}
 

### 2014-01-26 lp/ftrl-aux.tex
done with matrix section
fixed missing references

@@
 regret minimzation framework, and can use the machinery of follow the regularized 
-leader:
+leader \cite{hazan2011ftrl}:
 \begin{proposition}
@@
 Then, for all $i$, 
-\[ \sum_{t`=1}^T (z_{t,i} - m_{t,i})^2 \leq \sum_{t=1}^T (z_{t,i} - m^*_i)^2 + 4\log(T+1). \]
+\[ \sum_{t=1}^T (z_{t,i} - m_{t,i})^2 \leq \sum_{t=1}^T (z_{t,i} - m^*_i)^2 + 4\log(T+1). \]
 \end{proposition}
@@
 \begin{corollary}
+\label{cor:EG-variance}
 The EG-Variance algorithm achieves a regret bound of
-%\[ \Regret(w) \leq \frac{1 + (\log(n)-1)\|w\|_1 + \sum_{i=1}^n w_i\log(w_i)}{\eta} + \eta \left[ \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_i^*)^2 + 4\|w\|_1\log(T) \right]. \]
-\[ \Regret(w) \leq \frac{\log(n)}{\eta} + \eta \left[ \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_i^*)^2 + 4\log(T) \right]. \]
+%\[ \Regret(w) \leq \frac{1 + (\log(n)-1)\|w\|_1 + \sum_{i=1}^n w_i\log(w_i)}{\eta} + \eta \left[ \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_i^*)^2 + 4\|w\|_1\log(T+1) \right]. \]
+\[ \Regret(w) \leq \frac{\log(n)}{\eta} + \eta \left[ \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_i^*)^2 + 4\log(T+1) \right]. \]
 \end{corollary}

### 2014-01-26 lp/introduction.tex
fixed missing references

@@
 gradient is to the learning from experts setting, in which it is also known 
-as the weighted majority algorithm \cite{wm} and entertains regret bounds of the 
+as the weighted majority algorithm \cite{littlestone1989wm} and entertains regret bounds of the 
 form 
@@
 online fashion, choosing the $m_t$ according to a Follow the Regularized Leader 
-(FTRL) algorithm \cite{ftrl}. The regret bound thus obtained is given in 
-Theorem~\ref{thm:variance-vector}, which improves a result by 
+(FTRL) algorithm \cite{hazan2011ftrl}. The regret bound thus obtained is given in 
+Corollary~\ref{cor:EG-variance}, which improves a result by 
 \cite{hazan2010variation}.
@@
 (in analogy with the simplex). This setting has already been extensively studied 
-\cite{lotsofpeople} and is important in obtaining online bounds for various 
-combinatorial optimization problems \cite{morepeople}. Using the machinery so 
+\cite{tsuda2005matrix, arora2007combinatorial} and is important in 
+obtaining online and approximation bounds for various combinatorial optimization 
+problems \cite{arora2007combinatorial, hazan2012near}. Using the machinery so 
 far developed, all of our results extend straightforwardly to the matrix setting, 
@@
 Other notions of ``adaptive'' weights have also been explored; though unrelated 
-to the current work, we briefly mention them for completeness. The first notion 
-notion of adaptivity pertains to time-varying experts: here the desire is to 
-obtain regret relative to the best \emph{schedule} of experts, where the number 
-of switches in the schedule is bounded. This is explored in \cite{switching}. 
+to the current work, we briefly mention them for completeness. The first 
+notion, which has occasionally been referred to as adaptivity, pertains to time-varying experts: here the desire is to obtain regret relative to the best \emph{schedule} of experts, 
+where the schedule is in some way restricted to drift slowly over time 
+\cite{auer1998tracking, kleinberg2010regret, chiang2012online}. 
 The other slightly more related notion is an adaptive step size, where the step 
@@
 extensively throughout the literature but a few particularly relevant papers are 
-\cite{adaptive-step}.
+\cite{hazan2007adaptive, orabona2013dimension}.
 
-Finally, we note that the exponentiated gradient itself has been analyzed 
+Finally, we note that the exponentiated gradient algorithm itself has been analyzed 
 and applied in various settings. One common setting, using a hinge-like loss, 
-is the Winnow Algorithm \cite{winnow}, which has been carefully analyzed in 
-\cite{winnow-careful}.
+is the Winnow Algorithm \cite{littlestone1988}, which has been carefully analyzed in 
+\cite{sabato2012learning}.
 

### 2014-01-26 lp/mw12.tex
fixed missing references

@@
 The regret bounds for each of (\ref{eqn:mw1}) and (\ref{eqn:mw2}) are 
-well-known \cite{people}, but we include them below for completeness.
+well-known \cite{shalev2011, cesa2007}, but we include them below for completeness.
 \begin{theorem}

### 2014-01-26 lp/matrix.tex
matrix.tex

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Extension to Matrices} 
\label{sec:matrix}
We now extend our results to the matrix setting. Here $W$ is a positive semidefinite 
matrix. The analogue to the simplex is constrainint $\Tr(W)$ to be $1$. Instead of a vector 
of losses $z_t$, we have a matrix of losses $Z_t$, which is assumed to be symmetric and to 
have all eigenvalues between $-1$ and $1$. The loss in round $t$ is $\Tr(W_t^{\top}Z_t)$. 
Since all matrices involves are symmetric, we will typically suppress the transpose and 
just write $\Tr(W_tZ_t)$. Note that we can embed the vector setting in the matrix setting 
by restricting to diagonal matrices.

We start with an extension of Proposition~\ref{prop:aeg} to the matrix setting:
\begin{proposition}
\label{prop:maeg}
Consider the updates given by $\Theta_1 = -\log(n)I$ and 
$\Theta_{t+1} = \Theta_t - \eta Z_t - \eta^2 (Z_t - M_t)^2$, 
with prediction $W_t = \exp(\Theta_t - \eta M_t)$. For this algorithm,
\begin{equation}
\Regret(w) \leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \Tr(W(Z_t-M_t)^2).
\end{equation}
\end{proposition}
\begin{proof}
TODO
\end{proof}
We would now, as in Section~\ref{sec:ftrl-aux}, like to learn $M_t$ adaptively 
to be competitive with the best fixed value $M^*$. However, it isn't even clear 
that such a best fixed value exists (since the optimal value may depend on $W$). 
Lemma~\ref{lem:global-opt} addresses this issue (proved in slightly greater 
generality for later convenience):
\begin{lemma}
\label{lem:global-opt}
For any $\delta \geq 0$, define $M^* \eqdef \frac{1}{T+\delta} \sum_{t=1}^T Z_t$. Then, for 
any $M' \neq M^*$, we have
\[ \delta (M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2 \preceq \delta (M')^2 + \sum_{t=1}^T (Z_t-M')^2. \]
\end{lemma}
\begin{proof}
Write $M' = M^* + D$. Then we have
\begin{align}
\lefteqn{\delta(M')^2 + \sum_{t=1}^T (Z_t - M')^2} \\
 &= \delta(M^*+D)^2 + \sum_{t=1}^T (Z_t - M^* - D)^2 \\
 &= \delta(M^*)^2 + \sum_{t=1}^T (Z_t - M^*)^2 \\
\nonumber
 &\phantom{==} + \left[\delta M^* + \sum_{t=1}^T M^*-Z_t\right]D \\
\nonumber
 &\phantom{==} + D\left[\delta M^* + \sum_{t=1}^T M^*-Z_t\right] + TD^2 \\
 &= \delta(M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2 + TD^2 \\
 &\succeq \delta(M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2,
\end{align}
which completes the lemma.
\end{proof}
Setting $\delta$ to $0$, we now see that $\frac{1}{T} \sum_{t=1}^T Z_t$ is the 
optimal value of $M^*$ for any $W \succeq 0$, in analogy with the vector case. We 
remark that Lemma~\ref{lem:global-opt} is almost purely algebraic, and only relies 
on the property that $D^2 \succeq 0$ for any symmetric matrix $D$. We could obtain 
a similar lemma for any partially ordered algebra equipped with an adjoint operator 
$*$ such that $a^{*}a \succeq 0$ for all elements $a$ of the algebra.

Lemma~\ref{lem:global-opt} gives us a target value for the $M_t$, but we cannot 
simply apply the standard Follow the Regularized Leader lemma. This is because 
we need a result of the form
\[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq \sum_{t=1}^T (Z_t - M^*) + \epsilon I, \]
which cannot be straightforwardly expressed as a regret bound. We deal with this 
by deriving a generalization of the FTRL algorithm, which we call FTRL-$\sK$. 
This is an algorithm in which losses are vector-valued and we obtain regret relative 
to a partial ordering defined by a cone $\sK$.

An important notion is that of a \emph{global optimum}. For a function $f : X \to \bR^n$ 
and a cone $\sK \subset \bR^n$, we say that $f$ has a global optimum relative to 
$\sK$ if there exists an $x$ such that $f(x) \leq_{\sK} f(y)$ for all $y \in X$.

\begin{lemma}
\label{lem:ftrl-k}
Suppose that for all $1 \leq t \leq T+1$, there 
exists a global minimizer $M_t$ of 
$\psi(X) + \sum_{s=1}^{t-1} f_s(X)$. Then for all $M$,
\[ \sum_{t=1}^T f_t(M_t)-f_t(M) \leq_{\sK} \psi(M) - \psi(M_1) + \sum_{t=1}^T f_t(M_{t}) - f_t(M_{t+1}). \]
\end{lemma}
\begin{proof}
We will prove the lemma by induction on $T$. Note that the lemma is equivalent to showing that 
\[ \psi(M_1) + \sum_{t=1}^T f_t(M_{t+1}) \leq_{\sK} \psi(M) + \sum_{t=1}^T f_t(M) \]
for all $M$.

In the base case $T = 0$, we have
\[ \psi(M_1) \leq_{\sK} \psi(M), \]
which follows from the fact that $M_1$ is a global minimizer of $\psi$ and 
hence $\psi(M_1) \leq_{\sK} \psi(M)$ for all $M$. For the inductive step, suppose that
\[ \sum_{t=1}^{T-1} f_t(M_{t+1}) \leq_{\sK} \psi(M) \sum_{t=1}^{T-1} f_t(M) \]
for all $M$, and invoke this for the particular choice $M = M_{T+1}$. Then we have
\begin{align}
\lefteqn{\psi(M_1) + \sum_{t=1}^T f_t(M_{t+1})} \\
 &= \psi(M_1) + \sum_{t=1}^{T-1} f_t(M_{t+1}) + f_T(M_{T+1}) \\
 &\leq_{\sK} \psi(M_{T+1}) + \sum_{t=1}^{T-1} f_t(M_{T+1}) + f_T(M_{T+1}) \\
 &= \psi(M_{T+1}) + \sum_{t=1}^T f_t(M_{T+1}) \\
 &\leq_{\sK} \psi(M) + \sum_{t=1}^T f_t(M)
\end{align}
for all $M$, where we use the fact that $M_{T+1}$ is a global minimizer of $\psi(M) + \sum_{t=1}^T f_t(M)$ for 
the last inequality. This completes the inductive hypothesis and hence the proof.
\end{proof}

As a corollary, we have (compare to Proposition~\ref{prop:ftrl-aux}):
\begin{corollary}
\label{cor:ftrl-k-aux}
Suppose that we choose $M_t = \frac{1}{t} \sum_{s=1}^{t-1} Z_s$. Then
\[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq \sum_{t=1}^T (Z_t - M^*)^2 + 4\log(T+1)I, \]
for $M^* \eqdef \frac{1}{T} \sum_{t=1}^T Z_t$.
\end{corollary}
Combining Proposition~\ref{prop:maeg} with Corollary~\ref{cor:ftrl-k-aux} gives a 
regret bound for the MEG-Variance algorithm, whose updates are given in 
Table~\ref{tab:algorithms}.
\begin{corollary}
\label{cor:maeg-variance}
The MEG-Variance algorithm achieves a regret bound of
\[ \Regret(w) \leq \frac{\log(n)}{\eta} + \eta \left[ \sum_{t=1}^T \Tr(W(Z_t-M^*)^2) + 4\log(T+1) \right]. \]
\end{corollary}

\end{document}

### 2014-01-26 lp/conclusion.tex
conc

--- 2013-10-04 icml/conclusion.tex
@@
-\documentclass[paper.tex]{subfiles}
+\documentclass[paper_icml.tex]{subfiles}
 \begin{document}
@@
 
-In this paper we have introduced the primal and dual boosted mirror descent algorithms in order to investigate the properties of herding and a number of novel classes of generalized herding algorithms.
-$\alpha_{t}$-weighted herding is very similar to standard herding, but the pseudosamples are not required to have equal weights.
-We also defined $q$-herding.
-The $q$-herding algorithm minimizes the distance between the sample estimator and the target moments as measured by the $\ell^{p}$ norm (with $p$ the conjugate of $q$, $p \ge 2$).
-For $p = 2$, we get herding.
-But in infinite dimensional space, $q$-herding with $p > 2$ may some times be applied when standard herding is inapplicable.
-
-Using \bmd we also place the convergence results of \citet{Chen:2010} and \citet{Bach:2010} into a single framework.
-Furthermore, we were able to extend those convergence results to certain classes of herding-like algorithms including $\alpha_{t}$-weighted herding, $q$-herding, and \NA{chen} herding.
-
-Our results thus extend the applicability and flexibility of deterministic algorithms for generating pseudosamples.
-While not universally applicable, these deterministic algorithms offer an increasingly attractive alternative to random sampling for certain types of problems and model classes because the can offer faster convergence rates, convergence guarantees without ``with high probability'' conditions, and simple implementations.
-
 \end{document}

### 2014-01-27 lp/introduction.tex
mostly fixed equation formatting

@@
 
-By exploiting this connection, we show that the perhaps more natural update 
-of $\exp(-\eta z_{t,i} - \eta^2 z_{t,i}^2)$ achieves the same regret bound 
-as in (\ref{eqn:mw-regret}), and that both updates continue to enjoy this 
-regret even when generalized from the simplex to an arbitrary convex set 
-(including all of $\bR^n$, which corresponds to unconstrained exponentiated 
-gradient).
+% By exploiting this connection, we show that the perhaps more natural update 
+% of $\exp(-\eta z_{t,i} - \eta^2 z_{t,i}^2)$ achieves the same regret bound 
+% as in (\ref{eqn:mw-regret}), and that both updates continue to enjoy this 
+% regret even when generalized from the simplex to an arbitrary convex set 
+% (including all of $\bR^n$, which corresponds to unconstrained exponentiated 
+% gradient).
 
@@
 \item An interpretation of the multiplicative weights update as 
-      exponentiated gradient with an adaptive regularizer.
-\item An improved exponentiated gradient algorithm based on adaptive 
-      updates to an entropic regularizer.
-\item Further improvements to exponentiated gradient based on optimistic 
-      predictions and online learning of the mean gradient.
-\item An adaptive matrix exponentiated gradient, with similar improvements 
-      based on learning the mean gradient.
+      exponentiated gradient with an adaptive regularizer (Section~\ref{sec:mw12}).
+%\item An improved exponentiated gradient algorithm based on adaptive 
+%      updates to an entropic regularizer.
+\item An improved exponentiated gradient algorithm based on an adaptive 
+      regularizer, optimistic predictions, and online learning of the 
+      mean gradient (Sections~\ref{sec:machinery} and~\ref{sec:ftrl-aux}).
+\item An adaptive matrix exponentiated gradient algorithm, 
+      with similar improvements based on learning the mean gradient 
+      (Section~\ref{sec:matrix}).
 \item A generalization of Follow the Regularized Leader to vector-valued 
-      loss functions.
+      loss functions (Lemma~\ref{lem:ftrl-k}).
 \end{itemize}

### 2014-01-27 lp/machinery.tex
fixed missing references
mostly fixed equation formatting

@@
 \end{algorithm}
+Note that Algorithm~\ref{alg:adaptive-optimistic} does not contain a 
+projection step back onto the feasible domain $S$. If $\psi_t$ is defined to be 
+infinite outside of $S$, then $w_t$ will always lie in $S$ and hence the 
+projection step is unnecessary. (However, modifying $\psi_t$ in this way may 
+complicate the analysis, i.e. by giving $\psi_t^*$ a more complicated form.)
 
-The key regret bound for Algorithm~\ref{alg:adaptive-optimistic} is 
-given in Theorem~\ref{thm:adaptive-optimistic}:
+The regret bound for Algorithm~\ref{alg:adaptive-optimistic} is 
+given in Theorem~\ref{thm:adaptive-optimistic}. A key observation is that 
+the regret is in terms of $z_t - m_t$ rather than $z_t$ (so the closer 
+$m_t$ is to $z_t$, the smaller the regret; this will be more apparent 
+in various corollaries).
 \begin{theorem}
+\label{thm:adaptive-optimistic}
 Algorithm~\ref{alg:adaptive-optimistic} achieves regret bounded by
-\begin{equation}
-\frac{1}{\eta}\left(\psi_1^*(\theta_1) + \psi_{T+1}(w) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t - \eta m_t) + \eta w_t^{\top}m_t\right).
-\end{equation}
+\begin{dmath}
+\label{eqn:regret-1}
+\frac{1}{\eta}\left(\psi_1^*(\theta_1) + \psi_{T+1}(w) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t - \eta m_t) + \eta w_t^{\top}(z_t-m_t)\right).
+\end{dmath}
 \end{theorem}
 \begin{proof}
-TODO
+First note that $\psi_t^*$ is convex and that 
+$w_t = \nabla \psi_t^*(\theta_t - \eta m_t)$. It thus follows that 
+$\psi_t^*(\theta_t) \geq \psi_t^*(\theta_t - \eta m_t) + \eta m_t^{\top}z_t$. 
+Then, by definition of the Fenchel conjugate together with telescoping sums, we 
+have, for any $w$, 
+\begin{align*}
+\lefteqn{\hskip -0.1in w^{\top}\theta_{T+1} - \psi_{T+1}(w)} \\
+\hskip 0.1in  &\leq \psi_{T+1}^*(\theta_{T+1}) \\
+  &= \psi_1^*(\theta_1) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t) \\
+  &\leq \psi_1^*(\theta_1) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t-\eta m_t) - \eta w_t^{\top}m_t.
+\end{align*}
+Re-arranging and expanding $\theta_{T+1}$, we have
+\begin{dmath}
+-\eta w^{\top} \sum_{t=1}^T z_t \leq \psi_{T+1}(w) + \psi_1^*(\theta_1) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t-\eta m_t) - \eta w_t^{\top}m_t.
+\end{dmath}
+Adding $\eta \sum_{t=1}^T w_t^{\top}z_t$ to both sides, we obtain
+\begin{dmath}
+\eta \Regret(w) \leq \psi_{T+1}(w) + \psi_1^*(\theta_1) + \sum_{t=1}^T \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t) + \eta w_t^{\top}(z_t-m_t),
+\end{dmath}
+which yields the desired result.
 \end{proof}
@@
 \begin{proof}
-TODO
+Since $\theta_{t+1} = \theta_t - \eta z_t$, if the condition in the corollary 
+holds then the sum in the regret bound (\ref{eqn:regret-1}) is bounded above 
+term-wise by zero. 
+Hence we obtain $\Regret(w) \leq \frac{1}{\eta}\left(\psi_1^*(\theta_1) + \psi_{T+1}(w)\right)$, as was to be shown.
 \end{proof}
@@
 which corresponds to choosing $\psi_t(w) = \psi(w) + \eta^2 w^{\top} \sum_{s=1}^{t-1} u_s$ 
-for some fixed function $\psi$.
+for some fixed function $\psi$. In this case, we define 
+$\psi_S^*(w) \eqdef \sup_{w \in S} w^{\top}x - \psi(w)$, which 
+will allow us to more easily handle restrictions on the domain $S$.
 
@@
 \FOR{$t=1$ {\bfseries to} $T$}
-  \STATE Choose $w_t = \nabla \psi^*(\theta_t - \eta m_t)$
+  \STATE Choose $w_t = \nabla \psi_S^*(\theta_t - \eta m_t)$
   \STATE Observe $z_t$ and suffer loss $w_t^{\top}z_t$
@@
 \begin{equation}
-\Regret(w) \leq \frac{\psi^*(\theta_1)}{\eta} + \eta w^{\top} \sum_{t=1}^T u_t.
+\Regret(w) \leq \frac{\psi_S^*(\theta_1)}{\eta} + \eta w^{\top} \sum_{t=1}^T u_t.
 \end{equation}
 \end{corollary}
+Note that, importantly, the conditions of Corollary~\ref{cor:a-o2} are in terms 
+of $\psi^*$, while the regret bound and the choice of $w_t$ are in terms of 
+$\psi_S^*$. This means that, once we construct an algorithm for the 
+unconstrained case, we then achieve a regret bound under arbitrary convex 
+constraints.
 \begin{proof}
@@
 Corollary~\ref{cor:a-o2} to obtain an adaptive exponentiated gradient algorithm. 
+We will state it here in the case of the simplex, since the bounds in this 
+case are particularly easy to interpret. Note, however, that our analysis 
+goes via Corollary~\ref{cor:a-o2}, and hence applies to arbitrary convex 
+constraints.
 \begin{proposition}[Adaptive Exponentiated Gradient]
 \label{prop:aeg}
-Consider the updates given by $\theta_{1,i} = -\log(n)$ and 
+Consider the updates given by $\theta_{1,i} = 0$ and 
 $\theta_{t+1,i} = \theta_{t,i} - \eta z_{t,i} - \eta^2 (z_{t,i} - m_{t,i})^2$, 
-with prediction $w_{t,i} = \exp(\theta_{t,i} - \eta m_{t,i})$. For this 
-algorithm,
+with prediction $w_{t,i} \propto \exp(\theta_{t,i} - \eta m_{t,i})$. Then, 
+assuming $\|z_t\|_{\infty} \leq 1$, $\|m_t\|_{\infty} \leq 1$ and 
+$0 < \eta \leq \frac{1}{4}$, we have
 \begin{equation}
@@
 \begin{proof}
-TODO
+We note that, for $S$ equal to the simplex $\Delta_n$ and 
+$\psi(w) = \sum_{i=1}^n w_i\log(w_i)$, the $w_t$ is equal to 
+$\nabla \psi_S^*(\theta_t - \eta m_t)$. It therefore suffices 
+to check that the condition of Corollary~\ref{cor:a-o2} is satisfied with 
+$u_{t,i} = (z_{t,i}-m_{t,i})^2$. 
+Note that $\psi^*(\theta) = \sum_{i=1}^n \exp(\theta_i)$. We then have
+\begin{dgroup*}
+\begin{dmath*} \lefteqn{\psi^*(\theta_{t} - \eta z_t - \eta^2 u_t)} 
+\end{dmath*}\begin{dmath*}  = \sum_{i=1}^n \exp(\theta_{t,i} - \eta z_{t,i} - \eta^2 (z_{t,i} - m_{t,i})^2) 
+\end{dmath*}\begin{dmath*}  = \sum_{i=1}^n \exp(\theta_{t,i} - \eta m_{t,i})\exp(-\eta (z_{t,i} - m_{t,i}) - \eta^2 (z_{t,i} - m_{t,i})^2) 
+\end{dmath*}\begin{dmath*}  \leq \sum_{i=1}^n \exp(\theta_{t,i} - \eta m_{t,i})(1 - \eta (z_{t,i} - m_{t,i})) 
+\end{dmath*}\begin{dmath*}  = \sum_{i=1}^n \exp(\theta_{t,i} - \eta m_{t,i}) - \eta \sum_{i=1}^n \exp(\theta_{t,i} - \eta m_{t,i}) (z_{t,i} - m_{t,i}) 
+\end{dmath*}\begin{dmath*}  = \psi^*(\theta_t - \eta m_t) - \eta \nabla \psi^*(\theta_t - \eta m_t)^{\top} (z_t - m_t).
+\end{dmath*}
+\end{dgroup*}
+The one inequality we made use of was $\exp(-x-x^2) \leq 1-x$ for $|x| < \frac{1}{2}$.
+This verifies the condition of Corollary~\ref{cor:a-o2}, yielding a regret bound 
+of $\frac{\psi_S^*(0) + \psi(w)}{\eta} + \eta \sum_{i=1}^n w^{\top} u_t$. 
+Finally, we note that $\psi_S^*(0) = \log(n)$, 
+$\psi(w) = \sum_{i=1}^n w_i\log(w_i) \leq 0$, and 
+$u_{t,i} = (z_{t,i} - m_{t,i})^2$, which completes the proof.
 \end{proof}

### 2014-01-27 lp/matrix.tex
mostly fixed equation formatting

@@
 $\Theta_{t+1} = \Theta_t - \eta Z_t - \eta^2 (Z_t - M_t)^2$, 
-with prediction $W_t = \exp(\Theta_t - \eta M_t)$. For this algorithm,
+with prediction $W_t = \frac{\exp(\Theta_t - \eta M_t)}{\Tr(\exp(\Theta_t - \eta M_t))}$. For this algorithm,
 \begin{equation}
@@
 \begin{proof}
-TODO
+When $\psi(W) = \Tr(W\log(W))$ and $S$ consists of positive semidefinite 
+matrices with trace $1$, the prediction $W_t$ corresponds to 
+$\nabla \psi_S^*(\Theta_t - \eta M_t)$. So, again, it suffices to check 
+that the condition of Corolarry~\ref{cor:a-o2} is satisfies for 
+$U_t = (Z_t - M_t)^2$. To do so, we need to make use of the 
+Golden-Thompson inequality $\Tr(\exp(A+B)) \preceq \Tr(\exp(A)\exp(B))$, 
+together with the fact that $-X-X^2 \preceq \log(I-X)$ for 
+$-\frac{1}{2}I \preceq X \preceq \frac{1}{2}I$. We then have
+\begin{align*}
+\lefteqn{\psi^*(\Theta_t - \eta Z_t - \eta^2 U_t)} \\
+ &= \Tr(\exp(\Theta_t - \eta Z_t - \eta^2 (Z_t - M_t)^2)) \\
+ &\leq \Tr(\exp(\Theta_t - \eta M_t)\exp(-\eta (Z_t - M_t) - \eta^2 (Z_t - M_t)^2)) \\
+ &\leq \Tr(\exp(\Theta_t - \eta M_t)(I - \eta (Z_t - M_t))) \\
+ &= \Tr(\exp(\Theta_t - \eta M_t)) - \eta \Tr(\exp(\Theta_t - \eta M_t)(Z_t - M_t)) \\
+ &= \psi^*(\Theta_t - \eta M_t) - \eta \langle \nabla \psi^*(\Theta_t - \eta M_t), Z_t - M_t \rangle.
+\end{align*}
+This verifies the condition of Corolarry~\ref{cor:a-o2}, so that we have 
+a regret bound of $\frac{\psi_S^*(0) + \psi(W)}{\eta} + \eta \sum_{t=1}^T \Tr(WU_t)$. 
+Finally, noting that $\psi_S^*(0) = \log(n)$, $\psi(W) = \Tr(W\log(W)) \leq 0$, and 
+$U_t = (Z_t - M_t)^2$ completes the proof.
 \end{proof}
@@
 any $M' \neq M^*$, we have
-\[ \delta (M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2 \preceq \delta (M')^2 + \sum_{t=1}^T (Z_t-M')^2. \]
+\begin{equation*} \delta (M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2 \preceq \delta (M')^2 + \sum_{t=1}^T (Z_t-M')^2. \end{equation*}
 \end{lemma}
@@
 Write $M' = M^* + D$. Then we have
-\begin{align}
-\lefteqn{\delta(M')^2 + \sum_{t=1}^T (Z_t - M')^2} \\
- &= \delta(M^*+D)^2 + \sum_{t=1}^T (Z_t - M^* - D)^2 \\
- &= \delta(M^*)^2 + \sum_{t=1}^T (Z_t - M^*)^2 \\
-\nonumber
- &\phantom{==} + \left[\delta M^* + \sum_{t=1}^T M^*-Z_t\right]D \\
-\nonumber
- &\phantom{==} + D\left[\delta M^* + \sum_{t=1}^T M^*-Z_t\right] + TD^2 \\
- &= \delta(M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2 + TD^2 \\
- &\succeq \delta(M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2,
-\end{align}
+\begin{dgroup*}
+\begin{dmath*} \delta(M')^2 + \sum_{t=1}^T (Z_t - M')^2
+\end{dmath*}\begin{dmath*}  = \delta(M^*+D)^2 + \sum_{t=1}^T (Z_t - M^* - D)^2 
+\end{dmath*}\begin{dmath*}  = \delta(M^*)^2 + \sum_{t=1}^T (Z_t - M^*)^2 + \left[\delta M^* + \sum_{t=1}^T M^*-Z_t\right]D + D\left[\delta M^* + \sum_{t=1}^T M^*-Z_t\right] + TD^2 
+\end{dmath*}\begin{dmath*}  = \delta(M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2 + TD^2 
+\end{dmath*}\begin{dmath*}  \succeq \delta(M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2,
+\end{dmath*}
+\end{dgroup*}
 which completes the lemma.
@@
 exists a global minimizer $M_t$ of 
-$\psi(X) + \sum_{s=1}^{t-1} f_s(X)$. Then for all $M$,
-\[ \sum_{t=1}^T f_t(M_t)-f_t(M) \leq_{\sK} \psi(M) - \psi(M_1) + \sum_{t=1}^T f_t(M_{t}) - f_t(M_{t+1}). \]
+$\psi(M) + \sum_{s=1}^{t-1} f_s(M)$. Then for all $M$,
+\begin{dmath} \sum_{t=1}^T f_t(M_t)-f_t(M) \leq_{\sK} \psi(M) - \psi(M_1) + \sum_{t=1}^T f_t(M_{t}) - f_t(M_{t+1}). \end{dmath}
 \end{lemma}
@@
 for all $M$, and invoke this for the particular choice $M = M_{T+1}$. Then we have
-\begin{align}
+\begin{align*}
 \lefteqn{\psi(M_1) + \sum_{t=1}^T f_t(M_{t+1})} \\
@@
  &\leq_{\sK} \psi(M) + \sum_{t=1}^T f_t(M)
-\end{align}
+\end{align*}
 for all $M$, where we use the fact that $M_{T+1}$ is a global minimizer of $\psi(M) + \sum_{t=1}^T f_t(M)$ for 
@@
 Suppose that we choose $M_t = \frac{1}{t} \sum_{s=1}^{t-1} Z_s$. Then
-\[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq \sum_{t=1}^T (Z_t - M^*)^2 + 4\log(T+1)I, \]
+\[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq 2\sum_{t=1}^T (Z_t - M^*)^2 + 6I, \]
 for $M^* \eqdef \frac{1}{T} \sum_{t=1}^T Z_t$.
 \end{corollary}
+\begin{proof}
+The key observation is the \emph{matrix Young's inequality}: 
+$AB + BA \preceq \frac{1}{\gamma}A^2 + \gamma B^2$ for all symmetric $A$, $B$ and 
+all $\gamma > 0$. (This follows immediately upon expanding 
+$(A/\sqrt{\gamma} - \sqrt{\gamma} B)^2 \succeq 0$.) We then note that 
+$M_t$ obeys Lemma~\ref{lem:ftrl-k} with $\psi(M) = M^2$ and $f_t(M) = (M-Z_t)^2$. 
+Hence we have
+\begin{dgroup*}
+\begin{dmath*} \sum_{t=1}^T (Z_t - M_t)^2 - (Z_t - M^*)^2
+\end{dmath*}\begin{dmath*} = (M^*)^2 + \sum_{t=1}^T (Z_t - M_t)^2 - (Z_t - M_{t+1})^2 
+\end{dmath*}\begin{dmath*} = (M^*)^2 + \sum_{t=1}^T Z_t(M_{t+1} - M_t) + (M_{t+1}-M_t)Z_t + M_t^2 - M_{t+1}^2 
+\end{dmath*}\begin{dmath*} = (M^*)^2 + M_1^2 - M_{T+1}^2 + \sum_{t=1}^T \frac{1}{t+1}\left[ Z_t(Z_t-M_t) + (Z_t-M_t)Z_t\right] 
+\end{dmath*}\begin{dmath*} \leq I + \sum_{t=1}^T \frac{Z_t^2}{\gamma (t+1)^2} + \gamma (Z_t - M_t)^2 
+\end{dmath*}\begin{dmath*} \leq I + \frac{I}{\gamma} + \gamma \sum_{t=1}^T (Z_t - M_t)^2.
+\end{dmath*}
+\end{dgroup*}
+Re-arranging yields
+\begin{equation*}
+\sum_{t=1}^T (Z_t - M_t)^2 \leq \frac{1}{1-\gamma} \left( \frac{1+\gamma}{\gamma}I + \sum_{t=1}^T (Z_t - M^*)^2\right),
+\end{equation*}
+which for $\gamma = \frac{1}{2}$ again yields the desired result.
+\end{proof}
 Combining Proposition~\ref{prop:maeg} with Corollary~\ref{cor:ftrl-k-aux} gives a 
@@
 The MEG-Variance algorithm achieves a regret bound of
-\[ \Regret(w) \leq \frac{\log(n)}{\eta} + \eta \left[ \sum_{t=1}^T \Tr(W(Z_t-M^*)^2) + 4\log(T+1) \right]. \]
+\begin{dmath}\Regret(w) \leq \frac{\log(n)}{\eta} + \eta \left[ \sum_{t=1}^T \Tr(W(Z_t-M^*)^2) + 4\log(T+1) \right].\end{dmath} 
 \end{corollary}

### 2014-01-27 lp/mw12.tex
mostly fixed equation formatting

--- 2014-01-25 lp/mw12.tex
@@
 \label{eqn:mw1}
-w_{t+1,i} &\propto w_{t,i}\exp(-\eta z_{t,i})\\
+w_{t+1,i} &\propto w_{t,i}\exp(-\eta z_{t,i}) \tag{MW1}\\
 \label{eqn:mw2}
-w_{t+1,i} &\propto w_{t,i}(1-\eta z_{t,i})
+w_{t+1,i} &\propto w_{t,i}(1-\eta z_{t,i}) \tag{MW2}
 \end{align}
 The regret bounds for each of (\ref{eqn:mw1}) and (\ref{eqn:mw2}) are 
-well-known \cite{people}, but we include them below for completeness.
+well-known \cite{shalev2011, cesa2007}, but we include them below for completeness.
 \begin{theorem}
@@
 \label{eqn:mw1-regret}
-\Regret(w) &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \sum_{t=1}^T w_{t,i}z_{t,i}^2 \\
+\Regret(w) &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \sum_{t=1}^T w_{t,i}z_{t,i}^2 %\tag{MW1-regret}
+ \\
 \label{eqn:mw2-regret}
-\Regret(w) &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2
+\Regret(w) &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2 %\tag{MW2-regret}
 \end{align}
@@
 have non-negative cumulative loss, i.e. $\sum_{t=1}^T z_{t,i} \geq 0$. It is 
-apparent by the argument above that (\ref{eqn:mw2}) achieves asymptotically 
+apparent by the preceding paragraph that (\ref{eqn:mw2}) achieves asymptotically 
 constant (as a function of $T$) regret for any quasi-realizable sequence. In 
@@
 \end{proposition}
+We will be able to prove this more easily with the machinery developed in 
+Section~\ref{sec:machinery}, but we prove it here directly.
 \begin{proof}
@@
 \begin{align}
-\nabla \psi_t^*(\theta_t) &= \argmin_w \psi_t(w) - \theta_t^{\top}w \\
- &= \argmin_w \sum_{i=1}^n w_i\log(w_i) - \lambda_t^{\top}w.
+\nabla \psi_t^*(\theta_t) &= \argmin_{w \in \Delta_n} \psi_t(w) - \theta_t^{\top}w \\
+ &= \argmin_{w \in \Delta_n} \sum_{i=1}^n w_i\log(w_i) - \lambda_t^{\top}w.
 \end{align}
@@
 so that $w_{t,i}$ does indeed correspond to the update in (\ref{eqn:mw2}).
+
+Next, we exploit the property that the Fenchel conjugate of $\psi(w) + c^{\top}w$ 
+is $\psi^*(\theta - c)$, together with the fact that the Fenchel conjugate of 
+$\psi_1(w) = \sum_{i=1}^n w_i\log(w_i)$ is $\psi_1^*(\theta) = \log\left(\sum_{i=1}^n \exp(\theta_i)\right)$, 
+to obtain
+%\begin{small}
+\begin{align*}
+\psi_{t+1}^*(\theta_{t+1}) &= \psi_1^*(\theta_{t+1}-(\theta_{t+1}-\lambda_{t+1})) \\
+ &= \psi_1^*(\lambda_{t+1}) \\
+ &= \log(\sum_{i=1}^n \exp(\lambda_{t+1,i})) \\
+ &= \log(\sum_{i=1}^n \exp(\lambda_{t,i} + \log(1 - \eta z_{t,i}))) \\
+ &= \log(\sum_{i=1}^n \exp(\lambda_{t,i})(1 - \eta z_{t,i})) \\
+ &= \log(\sum_{i=1}^n \exp(\lambda_{t,i}) - \eta \sum_{i=1}^n \exp(\lambda_{t,i})z_{t,i}) \\
+ &\leq \log(\sum_{i=1}^n \exp(\lambda_{t,i})) - \eta \frac{\sum_{i=1}^n \exp(\lambda_{t,i})z_{t,i}}{\sum_{i=1}^n \exp(\lambda_{t,i})} \\
+ &= \psi_t^*(\theta_t) - \eta w_t^{\top}z_t.
+\end{align*}
+%\end{small}
+Thus by the observations following (\ref{eqn:telescope}), $\Regret(w)$ is bounded by 
+$\frac{\psi_1^*(\theta_1) + \psi_{T+1}(w)}{\eta}$. Now using the fact that 
+$-\eta z - \log(1-\eta z) \leq \eta^2 z^2$ when $|\eta z| \leq \frac{1}{2}$, and also the 
+fact that $\sum_i w_i\log(w_i) \leq 0$, we have
+\begin{align*}
+\lefteqn{\hskip -0.1in \eta\Regret(w)} \\
+\hskip 0.1in &\leq \psi_1^*(\theta_1) + \psi_{T+1}(w) \\
+ &= \log(n) + \sum_{i=1}^n w_i\log(w_i) %\nonumber \\ &\phantom{==}
+    + (\theta_{T+1}-\lambda_{T+1})^{\top}w \\
+ &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2,
+\end{align*}
+which completes the proposition.
 \end{proof}
+In the next section, we will present a more general framework for \emph{adaptive updates}, 
+also incorporating \emph{optimistic prediction} \cite{rakhlin2012}, which will both simplify 
+subsequent proofs and also allow us to obtain better bounds.
 \end{document}

### 2014-01-27 lp/ftrl-aux.tex
mostly fixed equation formatting
proofread a first draft, now just need to add a conclusion

@@
 is $m^* \eqdef \frac{1}{T} \sum_{t=1}^T z_t$, but we still have the issue that 
-we do not $m^*$ ahead of time. However, we have now cast the problem in the familiar 
+we do not know $m^*$ ahead of time. However, we have now cast the problem in the familiar 
 regret minimzation framework, and can use the machinery of follow the regularized 
-leader \cite{hazan2011ftrl}:
+leader \cite{hazan2011ftrl} to prove:
 \begin{proposition}
 \label{prop:ftrl-aux}
-Suppose that we choose $m_{t,i} = \frac{1}{t} \sum_{s=1}^{t-1} z_{s,i}$. 
-Then, for all $i$, 
-\[ \sum_{t=1}^T (z_{t,i} - m_{t,i})^2 \leq \sum_{t=1}^T (z_{t,i} - m^*_i)^2 + 4\log(T+1). \]
+Suppose that we choose $m_{t,i} = \frac{1}{t} \sum_{s=1}^{t-1} z_{s,i}$ and that 
+$\|z_s\|_{\infty} \leq 1$. Then for all $i$ we have
+\[ \sum_{t=1}^T (z_{t,i} - m_{t,i})^2 \leq 2\sum_{t=1}^T (z_{t,i} - m^*_i)^2 + 6. \]
 \end{proposition}
 \begin{proof}
-TODO
+Note that $m_{t,i}$ is the minimizer of $m_i^2 + \sum_{s=1}^{t-1} (z_{s,i} - m_i)^2$. 
+Therefore, by the FTRL lemma, we have
+\begin{align*}
+\lefteqn{\sum_{t=1}^T (z_{t,i}-m_{t,i})^2 - (z_{t,i} - m^*_i)^2} \\
+ &\leq (m^*_i)^2 + \sum_{t=1}^T (z_{t,i}-m_{t,i})^2 - (z_{t,i} - m_{t+1,i})^2 \\
+ &= (m^*_i)^2 + \sum_{t=1}^T 2z_{t,i}(m_{t+1,i}-m_{t,i}) + m_{t,i}^2 - m_{t+1,i}^2 \\
+ &= (m^*_i)^2 + m_{1,i}^2 - m_{T+1,i}^2 + \sum_{t=1}^T \frac{2}{t+1}z_{t,i}(z_{t,i} - m_{t,i}) \\
+ &\leq 1 + \sum_{t=1}^T \frac{z_{t,i}^2}{\gamma (t+1)^2} + \gamma(z_{t,i} - m_{t,i})^2 \\
+ &\leq 1 + \frac{1}{\gamma} + \gamma \sum_{t=1}^T (z_{t,i} - m_{t,i})^2.
+\end{align*}
+Re-arranging yields
+\begin{equation*}
+\sum_{t=1}^T (z_{t,i} - m_{t,i})^2 \leq \frac{1}{1-\gamma}\left(\frac{1+\gamma}{\gamma} + \sum_{t=1}^T (z_{t,i} - m_i^*)^2\right),
+\end{equation*}
+which for $\gamma = \frac{1}{2}$ yields the desired result.
 \end{proof}
@@
 the algorithm titled EG-Variance in Table~\ref{tab:algorithms}. This algorithm has the 
-same updates as an algorithm given by \cite{hazan2010variation}, but uses the 
+same updates as the algorithm given by \cite{hazan2010variation}, but uses the 
 optimistic prediction $\exp(\theta_t - \eta m_t)$ rather than $\exp(\theta_t)$. 
@@
 \label{cor:EG-variance}
-The EG-Variance algorithm achieves a regret bound of
+For $0 < \eta \leq \frac{1}{4}$ and $\|z_t\|_{\infty} \leq 1$, the EG-Variance 
+algorithm achieves a regret bound of
 %\[ \Regret(w) \leq \frac{1 + (\log(n)-1)\|w\|_1 + \sum_{i=1}^n w_i\log(w_i)}{\eta} + \eta \left[ \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_i^*)^2 + 4\|w\|_1\log(T+1) \right]. \]
-\[ \Regret(w) \leq \frac{\log(n)}{\eta} + \eta \left[ \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_i^*)^2 + 4\log(T+1) \right]. \]
+\begin{equation*}\Regret(w) \leq \frac{\log(n)}{\eta} + \eta \left[ 2\sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_i^*)^2 + 6 \right].\end{equation*} 
 \end{corollary}

### 2014-01-27 lp/machinery.tex
proofread a first draft, now just need to add a conclusion

@@
 \label{eqn:regret-1}
-\frac{1}{\eta}\left(\psi_1^*(\theta_1) + \psi_{T+1}(w) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t - \eta m_t) + \eta w_t^{\top}(z_t-m_t)\right).
+\frac{1}{\eta}\left(\psi_1^*(\theta_1) + \psi_{T+1}(w) + \sum_{t=1}^{T} \left[\psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t - \eta m_t) + \eta w_t^{\top}(z_t-m_t)\right]\right).
 \end{dmath}
@@
 \begin{dmath}
--\eta w^{\top} \sum_{t=1}^T z_t \leq \psi_{T+1}(w) + \psi_1^*(\theta_1) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t-\eta m_t) - \eta w_t^{\top}m_t.
+-\eta w^{\top} \sum_{t=1}^T z_t \leq \psi_{T+1}(w) + \psi_1^*(\theta_1) + \sum_{t=1}^{T} [\psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t-\eta m_t) - \eta w_t^{\top}m_t].
 \end{dmath}
@@
 \begin{dmath}
-\eta \Regret(w) \leq \psi_{T+1}(w) + \psi_1^*(\theta_1) + \sum_{t=1}^T \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t) + \eta w_t^{\top}(z_t-m_t),
+\eta \Regret(w) \leq \psi_{T+1}(w) + \psi_1^*(\theta_1) + \sum_{t=1}^T [\psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t) + \eta w_t^{\top}(z_t-m_t)],
 \end{dmath}
@@
 We note that, for $S$ equal to the simplex $\Delta_n$ and 
-$\psi(w) = \sum_{i=1}^n w_i\log(w_i)$, the $w_t$ is equal to 
+$\psi(w) = \sum_{i=1}^n w_i\log(w_i)$, $w_t$ is equal to 
 $\nabla \psi_S^*(\theta_t - \eta m_t)$. It therefore suffices 

### 2014-01-27 lp/matrix.tex
proofread a first draft, now just need to add a conclusion

@@
 We now extend our results to the matrix setting. Here $W$ is a positive semidefinite 
-matrix. The analogue to the simplex is constrainint $\Tr(W)$ to be $1$. Instead of a vector 
+matrix. The analogue to the simplex is constraining $\Tr(W)$ to be $1$. Instead of a vector 
 of losses $z_t$, we have a matrix of losses $Z_t$, which is assumed to be symmetric and to 
 have all eigenvalues between $-1$ and $1$. The loss in round $t$ is $\Tr(W_t^{\top}Z_t)$. 
-Since all matrices involves are symmetric, we will typically suppress the transpose and 
+Since all matrices involved are symmetric, we will typically suppress the transpose and 
 just write $\Tr(W_tZ_t)$. Note that we can embed the vector setting in the matrix setting 
@@
 \label{prop:maeg}
-Consider the updates given by $\Theta_1 = -\log(n)I$ and 
+Consider the updates given by $\Theta_1 = 0$ and 
 $\Theta_{t+1} = \Theta_t - \eta Z_t - \eta^2 (Z_t - M_t)^2$, 
-with prediction $W_t = \frac{\exp(\Theta_t - \eta M_t)}{\Tr(\exp(\Theta_t - \eta M_t))}$. For this algorithm,
+with prediction $W_t = \frac{\exp(\Theta_t - \eta M_t)}{\Tr(\exp(\Theta_t - \eta M_t))}$. For $0 < \eta \leq \frac{1}{4}$, $\|Z_t\|_{op} \leq 1$, and $\|M_t\|_{op} \leq 1$, we have
 \begin{equation}
-\Regret(w) \leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \Tr(W(Z_t-M_t)^2).
+\Regret(w) \leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \Tr(W(Z_t-M_t)^2)
 \end{equation}
+for all $W \succeq 0$ with $\Tr(W) = 1$.
 \end{proposition}
@@
 $\nabla \psi_S^*(\Theta_t - \eta M_t)$. So, again, it suffices to check 
-that the condition of Corolarry~\ref{cor:a-o2} is satisfies for 
+that the condition of Corollary~\ref{cor:a-o2} is satisfies for 
 $U_t = (Z_t - M_t)^2$. To do so, we need to make use of the 
-Golden-Thompson inequality $\Tr(\exp(A+B)) \preceq \Tr(\exp(A)\exp(B))$, 
+Golden-Thompson inequality $\Tr(\exp(A+B)) \preceq \Tr(\exp(A)\exp(B))$ \cite{golden, thompson}, 
 together with the fact that $-X-X^2 \preceq \log(I-X)$ for 
@@
 We would now, as in Section~\ref{sec:ftrl-aux}, like to learn $M_t$ adaptively 
-to be competitive with the best fixed value $M^*$. However, it isn't even clear 
+to be competitive with the best fixed value $M^*$. However, it is no longer clear 
 that such a best fixed value exists (since the optimal value may depend on $W$). 
@@
 \label{lem:global-opt}
-For any $\delta \geq 0$, define $M^* \eqdef \frac{1}{T+\delta} \sum_{t=1}^T Z_t$. Then, for 
-any $M' \neq M^*$, we have
+For any $\delta \geq 0$, define $M^* \eqdef \frac{1}{T+\delta} \sum_{t=1}^T Z_t$. Then, for any symmetric matrix $M'$, we have
 \begin{equation*} \delta (M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2 \preceq \delta (M')^2 + \sum_{t=1}^T (Z_t-M')^2. \end{equation*}
@@
 we need a result of the form
-\[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq \sum_{t=1}^T (Z_t - M^*) + \epsilon I, \]
+\[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq \sum_{t=1}^T (Z_t - M^*)^2 + \epsilon I, \]
 which cannot be straightforwardly expressed as a regret bound. We deal with this 
@@
 
-An important notion is that of a \emph{global optimum}. For a function $f : X \to \bR^n$ 
-and a cone $\sK \subset \bR^n$, we say that $f$ has a global optimum relative to 
-$\sK$ if there exists an $x$ such that $f(x) \leq_{\sK} f(y)$ for all $y \in X$.
+An important notion is that of a \emph{global minimizer}. For a function $f : X \to \bR^n$ 
+and a cone $\sK \subset \bR^n$, we say that $x$ is a global minimizer of $f$ 
+relative to $\sK$ if $f(x) \leq_{\sK} f(y)$ for all $y \in X$.
 
-\begin{lemma}
+\begin{lemma}[FTRL-$\sK$]
 \label{lem:ftrl-k}
@@
 \[ \psi(M_1) + \sum_{t=1}^T f_t(M_{t+1}) \leq_{\sK} \psi(M) + \sum_{t=1}^T f_t(M) \]
-for all $M$.
-
-In the base case $T = 0$, we have
+for all $M$. In the base case $T = 0$, we have
 \[ \psi(M_1) \leq_{\sK} \psi(M), \]
@@
 \lefteqn{\psi(M_1) + \sum_{t=1}^T f_t(M_{t+1})} \\
- &= \psi(M_1) + \sum_{t=1}^{T-1} f_t(M_{t+1}) + f_T(M_{T+1}) \\
- &\leq_{\sK} \psi(M_{T+1}) + \sum_{t=1}^{T-1} f_t(M_{T+1}) + f_T(M_{T+1}) \\
+ &= \psi(M_1) + \left[\sum_{t=1}^{T-1} f_t(M_{t+1})\right] + f_T(M_{T+1}) \\
+ &\leq_{\sK} \psi(M_{T+1}) + \left[\sum_{t=1}^{T-1} f_t(M_{T+1})\right] + f_T(M_{T+1}) \\
  &= \psi(M_{T+1}) + \sum_{t=1}^T f_t(M_{T+1}) \\
@@
 for all $M$, where we use the fact that $M_{T+1}$ is a global minimizer of $\psi(M) + \sum_{t=1}^T f_t(M)$ for 
-the last inequality. This completes the inductive hypothesis and hence the proof.
+the last inequality. This completes the induction and hence the proof.
 \end{proof}
@@
 \label{cor:ftrl-k-aux}
-Suppose that we choose $M_t = \frac{1}{t} \sum_{s=1}^{t-1} Z_s$. Then
+Suppose that we choose $M_t = \frac{1}{t} \sum_{s=1}^{t-1} Z_s$. Then, assuming 
+$\|Z_t\|_{op} \leq 1$ for all $t$, we have
 \[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq 2\sum_{t=1}^T (Z_t - M^*)^2 + 6I, \]
@@
 \begin{dmath*} \sum_{t=1}^T (Z_t - M_t)^2 - (Z_t - M^*)^2
-\end{dmath*}\begin{dmath*} = (M^*)^2 + \sum_{t=1}^T (Z_t - M_t)^2 - (Z_t - M_{t+1})^2 
-\end{dmath*}\begin{dmath*} = (M^*)^2 + \sum_{t=1}^T Z_t(M_{t+1} - M_t) + (M_{t+1}-M_t)Z_t + M_t^2 - M_{t+1}^2 
+\end{dmath*}\begin{dmath*} \preceq (M^*)^2 + \sum_{t=1}^T (Z_t - M_t)^2 - (Z_t - M_{t+1})^2 
+\end{dmath*}\begin{dmath*} = (M^*)^2 + \sum_{t=1}^T \left[Z_t(M_{t+1} - M_t) + (M_{t+1}-M_t)Z_t + M_t^2 - M_{t+1}^2 \right]
 \end{dmath*}\begin{dmath*} = (M^*)^2 + M_1^2 - M_{T+1}^2 + \sum_{t=1}^T \frac{1}{t+1}\left[ Z_t(Z_t-M_t) + (Z_t-M_t)Z_t\right] 
-\end{dmath*}\begin{dmath*} \leq I + \sum_{t=1}^T \frac{Z_t^2}{\gamma (t+1)^2} + \gamma (Z_t - M_t)^2 
-\end{dmath*}\begin{dmath*} \leq I + \frac{I}{\gamma} + \gamma \sum_{t=1}^T (Z_t - M_t)^2.
+\end{dmath*}\begin{dmath*} \preceq I + \sum_{t=1}^T \frac{Z_t^2}{\gamma (t+1)^2} + \gamma (Z_t - M_t)^2 
+\end{dmath*}\begin{dmath*} \preceq I + \frac{I}{\gamma} + \gamma \sum_{t=1}^T (Z_t - M_t)^2.
 \end{dmath*}
@@
 \label{cor:maeg-variance}
-The MEG-Variance algorithm achieves a regret bound of
+For $0 < \eta \leq \frac{1}{4}$ and $\|Z_t\|_{op} \leq 1$, the MEG-Variance algorithm achieves a regret bound of
 \begin{dmath}\Regret(w) \leq \frac{\log(n)}{\eta} + \eta \left[ \sum_{t=1}^T \Tr(W(Z_t-M^*)^2) + 4\log(T+1) \right].\end{dmath} 

### 2014-01-27 lp/paper_icml.tex
done with mw12 draft
done with machinery
done with ftrl-aux
done with matrix section
added algorithms-table
mostly fixed equation formatting
proofread a first draft, now just need to add a conclusion

@@
 \usepackage{tabularx}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 \input{latex-defs.tex}
+\usepackage{breqn}
 
@@
              353 Serra Street, Stanford, CA 94305 USA}
-\icmlauthor{Jonathan Huggins}{jhuggins@mit.edu}
-\icmladdress{MIT CSAIL,
-             77 Massachusetts Avenue, Cambridge, MA 02139 USA}
+%\icmlauthor{Jonathan Huggins}{jhuggins@mit.edu}
+%\icmladdress{MIT CSAIL,
+%             77 Massachusetts Avenue, Cambridge, MA 02139 USA}
 \icmlauthor{Percy Liang}{pliang@cs.stanford.edu}
@@
 to the matrix case, presenting an \emph{adaptive matrix exponentiated gradient} 
-algorithm. Our proof in this case involves a novel analysis tool generalizing 
+algorithm. Our proof in this case involves an analysis tool generalizing 
 follow the regularized leader to vector-valued payoffs, which may be of 
@@
 \subfile{introduction.tex}
-\subfile{algorithms.tex}
+\subfile{algorithms-table.tex}
+\subfile{mw12.tex}
+\subfile{machinery.tex}
+\subfile{ftrl-aux.tex}
+\subfile{matrix.tex}
+\subfile{conclusion.tex}
 

### 2014-01-27 lp/conclusion.tex
done with complete draft

@@
 
-\section{Conclusion} 
+\section{Discussion} 
 \label{sec:conclusion}
+We have presented an adaptive exponentiated gradient algorithm, 
+which attains regret bounded by the variance of the best expert in retrospect. 
+Recently \cite{duchi2011adagrad} have proposed an adaptive subgradient algorithm, 
+and it is interesting to compare to their result. They define the quantity 
+$\gamma_{T,i} \eqdef \sqrt{\sum_{t=1}^T z_{t,i}^2}$, and obtain regret in terms 
+of $\|w\|_{\infty} \sum_{i=1}^n \gamma_{T,i}$. For an appropriately chosen 
+$\eta$ our algorithm achieves a similar regret bound on the simplex of
+\[ \sqrt{\log(n) \sum_{i=1}^n w_i \gamma_{T,i}^2} \leq \sqrt{\vphantom{\sum_{i=1}^n}\log(n)\|w\|_{\infty}}\sqrt{\sum_{i=1}^n \gamma_{T,i}^2}, \]
+which is worse due to the $\log(n)$ factor but better because we pay in terms of 
+$\|\gamma_T\|_2$ rather than $\|\gamma_T\|_1$. Our method may therefore perform 
+better when the cumulative updates are relatively dense and homogeneous.
+
+When we move beyond the simplex, we obtain a similar regret bound but the 
+$\log(n)$ term is replaced by (roughly) a $\log(n) + \|w\|_1\log(\|w\|_1)$ term. 
+%The overall regret is then $\sqrt{(\log(n) + \|w\|_1\log(\|w\|_1))\|w\|_{\infty}}\|\gamma_T\|_2$, 
+%in comparison to the adagrad regret of $\|w\|_{\infty}\|\gamma_T\|_1$. 
+In addition to the $\log(n)$ term that was there before, we now pay for (ignoring 
+the $\log(\|w\|_1)$ term) $\sqrt{\|w\|_1\|w\|_{\infty}}$ as opposed to 
+$\|w\|_{\infty}$, but are again compensated by measuring $\gamma_T$ in the 
+$2$-norm rather than the $1$-norm. In analogy with the preceding paragraph, 
+we may expect to perform well when 
+the optimal predictor $w$ is sparse but the cumulative squared gradient 
+$\gamma_T$ is dense. It would be interesting to perform an empirical comparison 
+of the two algorithms to understand how they perform in various settings.
+
+Finally, we think that the idea of treating the regret bound as a quantity to 
+itself be optimized in an online fashion is quite interesting. While the bounds 
+thus obtained are particularly nice for adaptive exponentiated gradient 
+(because we obtain guarantees in terms of the best fixed expert), there are no 
+obstacles to applying the same techniques to other settings as well, such 
+as trying to minimize $\sum_{t=1}^T \|z_t - m_t\|_*^2$ in the standard mirror 
+descent bound. We believe that such ``auxiliary online optimization'' techniques, 
+when applied under the optimistic learning framework, have the potential to greatly improve the performance of online learning algorithms.
 

### 2014-02-03 trevisan-long.tex
added trevisan stuff

\documentclass[11pt]{article}
\usepackage{import}
\input{../latex-utils/defs.tex}
\usepackage{fullpage}

\title{Computational Indistinguishability via Boosting}
\author{Jacob Steinhardt\footnote{Based on Trevisan et al, 2010.}}
\begin{document}
\maketitle

\section{Setup}
Suppose we have a family $\sF$ of functions $f : X \to [0,1]$. We 
say that $h$ has complexity $C$ relative to $f$ if it can be obtained 
from elements of $f$ via at most $C$ of the following operations:
\begin{enumerate}
\item addition: $f,g \mapsto f+g$
\item multiplication: $f,g \mapsto f\cdot g$
\item multiplication by scalars: $c,f \mapsto c \cdot f$
\item thresholding: $f \mapsto \bI[f(x) \leq t]$
\end{enumerate}
Note that two-sided thresholding such as $\bI[s \leq f(x) \leq t]$ can be 
expressed as $\bI[f(x) \leq t] \cdot \bI[-f(x) \leq -s]$ and so has complexity 
$4$ relative to $\sF$. We will denote by $\sF[C]$ the space of $h : X \to [0,1]$ 
with complexity at most $C$ relative to $\sF$.

Given a distribution $\mu$ on $X$, we say that $g$ and $h$ are 
\emph{$\epsilon$-indistinguishable} relative to $\sF$ if 
$|\bE_{\mu}[f(x)(g(x)-h(x))] \leq \epsilon$ for all $f \in \sF$.

\section{Main Result}
The main result is the following:

\begin{theorem}
\label{thm:luca}
Let $\sF$ be any family of functions $f : X \to [0,1]$, let $\mu$ be a 
distribution on $X$, and let $g$ be any function $g : X \to [0,1]$. Then 
there exists an $h$ with complexity $O(1/\epsilon^2)$ relative to $\sF$ that is 
$\epsilon$-indistinguishable from $g$ relative to $\sF$. Furthermore, $\bE_{\mu}[g(x)] = \bE_{\mu}[h(x)]$.
\end{theorem}

\begin{proof}
Assume WLOG that $\sF = -\sF$, so that we need only prove indistinguishability 
in one direction. Also, assume that $1 \in \sF$, so that $\bE_{\mu}[g(x)] \approx \bE_{\mu}[h(x)]$ (we can 
make them exactly equal by slightly adjusting $h$ post-hoc).

We use the following stochastic gradient descent algorithm:
\begin{itemize}
\item $h_0 = 0$
\item $f_i \in \argmax_{f \in \sF} \bE_{\mu}[f(x)(h_i(x)-g(x))]$
\item $h_{i+1} = \min(\eta \sum_{j \leq i} f_j, 1)$
\end{itemize}
This is online gradient descent with regularizer $\frac{1}{2\eta}\bE_{\mu}[h(x)^2]$, so we get the regret bound:
\begin{align*}
\sum_{i=0}^{k-1} \bE_{\mu}[f_i(x)(h_i(x)-g(x))] &\leq \sum_{i=0}^k \bE_{\mu}[f_i(x)(g(x)-g(x))] + \frac{1}{2\eta}\bE_{\mu}[g(x)^2] + \frac{\eta}{2}\sum_{i=0}^{k-1} \bE_{\mu}[f_i(x)^2] \\
 &\leq \frac{1}{2\eta} + \frac{k\eta}{2}.
\end{align*}
Setting $\eta = \frac{1}{\sqrt{k}}$ yields 
\[ \sum_{i=0}^k \bE_{\mu}[(h_i(x)-g(x))f_i(x)] \leq \sqrt{k}, \]
meaning that $\bE_{\mu}[(h_i(x)-g(x))f_i(x)] \leq \frac{1}{\sqrt{k}}$ for some $i < k$. 
Since $f_i$ is by construction the maximizer of $\bE[(h_i(x)-g(x))f(x)]$, we have that 
$\bE[(h_i(x)-g(x))f(x)] \leq \frac{1}{\sqrt{k}}$ for all $f \in \sF$ for some $i < k$. Letting 
$k = \frac{1}{\epsilon^2}$ then gives us $\epsilon$-distinguishability relative to $\sF$.

Since $\min(\cdot, 1)$ can be implemented in $O(1)$ steps and $k = \frac{1}{\epsilon^2}$, we 
then have that $h_i$ has complexity $O\left(\frac{1}{\epsilon^2}\right)$ relative to $\sF$, which 
proves the theorem.
\end{proof}

\begin{remark}
Note that the theorem still holds (possibly with different constants) if we replace $[0,1]$ by any 
bounded set; in particular, with $[-1,1]$ (this fact is used in Trevisan et al. although here we will 
only use $[0,1]$).
\end{remark}

%We can also prove a slightly strengthened version of the theorem:
%\begin{theorem}
%\label{thm:luca-strong}
%Under the same conditions as Theorem~\ref{thm:luca}, we can obtain a 
%function $h : X \to \mathbb{R}_{\geq 0}$ that is a non-negative \emph{linear} 
%combination of $O(1/\epsilon^2)$ members of $\sF$, and such that $g$, $h$, 
%and $\min(h,1)$ are all mutually $\epsilon$-indistinguishable. Furthermore, 
%$\bE_{\mu}[g(x)] = \bE_{\mu}[\min(h(x),1)]$.
%\end{theorem}
%\begin{remark}
%In this version of the theorem, the proof will no longer hold if we replace $[0,1]$ by $[-1,1]$: 
%the non-negativity of the co-domain is important.
%\end{remark}
%\begin{proof}
%Perform stochastic gradient descent as before, but now with the loss function 
%$l_f(h) = \bE_{\mu}\left[f(x) \cdot \frac{h(x) + \min(h(x),1)}{2}\right]$. Also, for this 
%version, do not project $h$ but simply perform the gradient update $h_{i+1} = h_i + \eta \partial_h l_f(h)$. 
%If we actually compute this update, we get
%\[ h_{i+1}(x) = f(x) \cdot \frac{1 + \bI[h_i(x) < 1]}{2}. \]
%Note that $h$ and $f$ are both non-negative, so we can conclude the following:
%\begin{itemize}
%\item If $l_f(h) < \frac{\epsilon}{2}$, then $\bE_{\mu}[f(x)h(x)] < \epsilon$ and 
%      $\bE_{\mu}[f(x)\min(h(x),1)] < \epsilon$. Therefore, 
%\end{itemize}
%\end{proof}

\section{Corollaries}
\subsection{Dense Model Theorem}
Theorem~\ref{thm:luca} has many interesting corollaries. The first is the dense model theorem. Call 
a distribution $p$ \emph{$\delta$-dense} relative to $q$ if $p(x) \leq \frac{q(x)}{\delta}$: that is, 
$p$ is $\delta$-dense if it is never too much more likely than $q$. This implies in particular that 
there is a rejection sampler for $p$ with proposal distribution $q$ and acceptance rate at least $\delta$.

\begin{theorem}[Dense Model Theorem]
Suppose that $X$ is a finite set, let $p_U$ be the uniform distribution 
on $X$, and let $\sF$ be a family of functions $f : X \to [0,1]$. 
Let $p_D$ and $p_R$ be distributions such that $p_R$ and $p_U$ have moments in $\sF[C]$ differing 
by at most $\epsilon$: $|\bE_{p_R}[f(x)]-\bE_{p_U}[f(x)]| \leq \epsilon$ for all $f \in \sF[C]$.
Suppose also that $p_D$ is $\delta$-dense in $p_R$. Then, if $C \geq O(1/\epsilon^2)$, 
there is a distribution $p_M$ that is $(\delta-\epsilon)$-dense in $p_U$, and whose moments in $\sF$ 
differ from $p_D$ by at most $\frac{3\epsilon}{\delta}$.
\end{theorem}

\begin{proof}
Let $g(x) = 1-\delta\frac{p_D(x)}{p_R(x)}$. Let $h(x)$ be an $\epsilon$-indistinguishable function 
with respect to the measure $p_R$ and the family $\sF$, and define $p_M(x) \propto 1-h(x)$.

First, let us compute the normalization constant of $p_M$. We have:
\begin{align*}
\sum_{x} p_M(x) &= |X|\bE_{p_U}[(1-h(x))] \\
 &= |X|(\bE_{p_R}[(1-h(x))] \pm \epsilon) \\
 &= |X|(\bE_{p_R}[(1-g(x))] \pm \epsilon) \\
 &= |X|(\delta \pm \epsilon).
\end{align*}

It follows that $p_M(x) \leq \frac{1-h(x)}{|X|(\delta-\epsilon)} \leq \frac{1}{|X|(\delta-\epsilon)}$ and 
hence $p_M$ is $(\delta-\epsilon)$-dense in $p_U$.

To get indistinguishability, we need to show that $|\bE_{p_D}[f(x)]-\bE_{p_M}[f(x)]|$ is always small. 
We do this by showing that $p_D(x)$ and $p_M(x)$ are both close to $\frac{1-h(x)}{\delta}$. Indeed, we have
\begin{align*}
\left|\bE_{p_D}[f(x)]-\bE_{p_U}\left[\frac{1-h(x)}{\delta}f(x)\right]\right| &\leq \left|\bE_{p_D}[f(x)]-\bE_{p_R}\left[\frac{1-h(x)}{\delta}f(x)\right]\right| + \frac{\epsilon}{\delta} \\
 &\leq \left|\bE_{p_D}[f(x)]-\bE_{p_R}\left[\frac{1-g(x)}{\delta}f(x)\right]\right| + \frac{2\epsilon}{\delta} \\
 &= \left|\bE_{p_D}[f(x)]-\bE_{p_R}\left[\frac{p_D(x)}{p_R(x)}f(x)\right]\right| + \frac{2\epsilon}{\delta} \\
 &= \frac{2\epsilon}{\delta},
\end{align*}
as well as
\begin{align*}
\left|\bE_{p_M}[f(x)] - \bE_{p_U}\left[\frac{1-h(x)}{\delta}f(x)\right]\right| &= \left|\bE_{p_M}[f(x)] - \frac{\sum_{x} 1-h(x)}{\delta|X|} \bE_{p_M}[f(x)]\right| \\
 &= \left|\bE_{p_M}[f(x)] - \frac{(\delta \pm \epsilon)|X|}{\delta |X|} \bE_{p_M}[f(x)] \right| \\
 &\leq \frac{\epsilon}{\delta} |\bE_{p_M}[f(x)]| \\
 &\leq \frac{\epsilon}{\delta}.
\end{align*}
These two inequalities complete the proof.
\end{proof}

\begin{remark}
Note that the construction also shows that $p_M \in \sF[C]$. Since $p_M$ has high density relative to $p_U$, 
we can build a rejection sampler for $p_M$ using the uniform distribution as a proposal, and hence sample $p_M$ 
efficiently. If we set $R = U$, then this implies in particular that every $\delta$-dense distribution has 
an indistinguishable distribution that can be efficiently sampled.
\end{remark}

%\begin{remark}
%Note that we only need to multiply functions in $\sF$ once, when we consider $\frac{1-h(x)}{\delta} f(x)$. 
%So suppose that $\sF$ is defined by some collection $\phi : X \to \{0,1\}^d$ of binary features. Also 
%imagine that $R$ and $U$ have identical \emph{second} moments with respect to $\phi$: 
%$\bE_{p_U}[\phi(x)\phi(x)^T] = \bE_{p_R}[\phi(x)\phi(x)^T]$. (We'll also assume that one of the features 
%is identically equal to $1$, so that this also implies that $R$ and $U$ have identical first moments.)
%
%By the construction in Theorem~\ref{thm:luca}, $h(x) = \min(\eta \sum_{i=1}^k \phi_i(x), 1)$. 
%For the result of the dense model theorem to hold, we need $\bE_{p_U}[h(x)\phi(x)] = \bE_{p_R}[h(x)\phi(x)]$, 
%or equivalently
%\[ \bE_{p_U}[\min(\eta \sum_{i=1}^k \phi_i(x), 1)\phi(x)] = \bE_{p_R}[\min(\eta \sum_{i=1}^k \phi_i(x), 1)\phi(x)]. \]
%Without the $\min$, we would be in good shape. It turns out we can modify our construction so that the $\min$ affects 
%things by at most $\epsilon$, so that these expectations are indeed within $\epsilon$ of each other.
%\end{remark}
\subsection{Szemeredi Regularity Lemma}
\begin{theorem}[Szemeredi]
Let $G$ be a graph with edge set $E$. Then there exists a partition of $V$ into $M$ components $V_1,\ldots,V_M$ and a 
function $w : \{1,\ldots,M\} \times \{1,\ldots,M\} \to [0,1]$ such that if $A \subset V_i$ and $B \subset V_j$ then 
the number of edges $e(A,B)$ between $A$ and $B$ satisfies $|A||B|w(i,j)-\epsilon|V|^2 \leq e(A,B) \leq |A||B|w(i,j) + \epsilon|V|^2$. Furthermore, 
we can take $M = 2^{O(1/\epsilon^2)}$.
\end{theorem}

\begin{proof}
Let $\sF$ consist of all functions from $V \times V$ to $\{0,1\}$ of the form $f_{A,B}(u,v) = \bI[u \in A] \bI[v \in B]$. Let $g(u,v) = \bI[(u,v) \in E]$. 
Then there exists some $h = \sum_{i=1}^k c_i f_{A_i,B_i}$ with $k = O(1/\epsilon^2)$ that is $\epsilon$-indistinguishable from $g$ with respect to $\sF$. 
We can then partition $V$ into $2^{2k}$ sets $V_1,\ldots,V_{2^{2k}}$ such that $h$ is piecewise constant on each $V_i \times V_j$ (just let the $V_j$ consist of all 
intersections of the $A_i$ and the $B_i$ or their complements). We can then define $w(i,j)$ to be the constant value that $h$ takes on $V_i \times V_j$.
In particular, indistinguishability gives us $|\sum_{u \in A, v \in B} h(u,v) - g(u,v)| \leq \epsilon |V|^2$, 
or equivalently $|w(i,j)|A||B|-e(A,B)| \leq \epsilon |V|^2$, which proves the theorem.
\end{proof}

\subsection{Impaggliazo Hard-Core Set Lemma}
\begin{theorem}
Let $X$ be a set and $p_U$ the uniform distribution over that 
set. Let $\sF$ be a family of boolean functions and let 
$g$ be a boolean function such that $p_U[f(x)=g(x)] \leq 1-\delta$ 
for all $f \in \sF[C]$, with $C \geq O(1/(\delta\epsilon)^2)$. Then 
there exists a distribution $p_M \in \sF[C]$ such that $p_M(x) \leq \frac{1}{\delta |X|}$ 
for all $x \in X$, and 
$p_M[f(x)=g(x)] \leq \frac{1}{2}+\epsilon$ for all $f \in \sF$.
\end{theorem}
Intuitively, this means that if $g$ is somewhat hard to match by any function 
with complexity $C$ relative to $\sF$, then there is some ``core'' of $X$ that 
is hard for any element of $\sF$ to match at all.
\begin{proof}
Let $h \in \sF[C]$ be $(\delta\epsilon)$-indistinguishable from $g$ 
relative to $\sF$. Let $p_M(x) \propto |g(x)-h(x)|$. First, we note that 
$\bE_{p_U}[|g(x)-h(x)|] \geq \delta$, or else $\bI[h(x) \leq t]$ would match 
$g(x)$ on more than $1-\delta$ of the inputs for some $t$. Therefore, 
$\sum_{x \in X} |g(x)-h(x)| \geq \delta |X|$ and hence $p_M(x) \leq \frac{|g(x)-h(x)|}{\delta |X|} \leq \frac{1}{\delta |X|}$, 
which establishes the first condition.

To finish, note the identity 
\begin{align*}
|g(x)-h(x)| \bI[f(x)=g(x)] &= \frac{1}{2}|g(x)-h(x)| + \frac{1}{2}|g(x)-h(x)|(-1)^{f(x)-g(x)} \\
 &= \frac{1}{2}|g(x)-h(x)| - \frac{1}{2}(g(x)-h(x))(-1)^{f(x)-g(x)+g(x)} \\ 
 &= \frac{1}{2}|g(x)-h(x)| - \frac{1}{2}(g(x)-h(x))(-1)^{f(x)} \\
 &= \frac{1}{2}|g(x)-h(x)| + \left(f(x)-\frac{1}{2}\right) (g(x)-h(x)).
\end{align*}
It then follows that 
\begin{align*}
p_M[\bI[f(x)=g(x)]] &= \frac{\sum_{x} \frac{1}{2}|g(x)-h(x)| + \sum_{x} \left(f(x)-\frac{1}{2}\right) (g(x)-h(x))}{\sum_{x} |g(x)-h(x)|} \\
 &= \frac{1}{2} + \frac{\sum_{x} f(x)(g(x)-h(x))}{\sum_{x} |g(x)-h(x)|} \\
 &\leq \frac{1}{2} \frac{\epsilon\delta}{\delta} \\
 &= \frac{1}{2} + \epsilon,
\end{align*}
which yields the desired result.
\end{proof}
\end{document}

### 2014-02-03 trevisan.tex
added trevisan stuff

\documentclass[11pt]{article}
\usepackage{import}
\input{../latex-utils/defs.tex}
\usepackage{fullpage}

\title{Shorter Proof of The Indistinguishability Theorem}
\begin{document}
\maketitle

\section{Conditional Gradient}
Suppose we have a bilinear form $\langle \cdot, \cdot \rangle$ on 
$\sX \times \sY$ together with convex functions $\phi(x)$ and $\psi(y)$. 
We can consider the two-argument loss function
\[ L(x,y) = \phi(x) + \langle x, y \rangle - \psi(y). \]
We know by the minimax theorem that $L$ has a saddle point $(x^*,y^*)$; 
conditional gradient obtains for us an approximate saddle point with 
desired properties. Given a sequence of weights $\alpha_t$, 
let $\hat{x}_t$ be the weighted average of $x_{1:t}$ with respect 
to the weights $\alpha_{1:t}$:
\[ \hat{x}_t \eqdef \frac{\sum_{s=1}^t \alpha_sx_s}{\sum_{s=1}^t \alpha_s}. \]
Then the conditional gradient algorithm is:
\begin{align*}
    x_t &\in \argmin_x \phi(x) + \langle x, y_t \rangle \\
y_{t+1} &\in \argmax_y \langle \hat{x}_t, y \rangle - \psi(y).
\end{align*}
Assuming that $\phi$ is strongly convex with respect to a norm $\|\cdot\|$, 
and letting $\|\cdot\|_*$ denote the dual norm and $r \eqdef \sup_{y} \|y\|_*$, 
we have the bound (for $\alpha_t = t$):
\[ \sup_{y} L(\hat{x}_t, y) \leq \sup_{y} L(x^*, y) + \frac{8r^2}{T}. \]

\section{Indistinguishability}
To apply conditional gradient to indistinguishability, let $\sX$ be the space 
of functions $f : S \to [0,1]$ and let $\sY = \operatorname{Conv}(\sF)$. Then 
define 
\[ L(h,f) = \frac{\eta}{2} \bE_{\mu}[h(s)^2] + \bE_{\mu}[h(s)f(s)] - \bE_{\mu}[g(s)f(s)]. \]
This corresponds to setting $\phi(h) = \frac{\eta}{2} \bE_{\mu}[h(s)^2]$, $\langle h, f \rangle = \bE_{\mu}[h(s)f(s)]$, 
and $\psi(f) = \bE_{\mu}[g(s)f(s)]$. If we let $D(g,h)$ denote 
the discrepancy between $g$ and $h$, then note that
\[ D(g,h) \leq \sup_{f} L(h,f) \leq D(g,h) + \frac{\eta}{2}. \]
Also note that $f_t \in \sF$ for all $t$ and that $h_t = \max(\min(\frac{1}{\eta}f_t,1),0)$. Finally, 
$\phi$ is strongly convex with respect to the norm $\|h\| = \sqrt{\eta \bE_{\mu}[h(s)^2]}$, which 
has the dual norm $\|f\|_* = \sqrt{\eta^{-1} \bE_{\mu}[f(s)^2]}$. Since $0 \leq f(s) \leq 1$, we then have 
$r^2 \leq \frac{1}{\eta}$ and so our bound yields
\begin{align*}
D(g,\hat{h}_T) &\leq \sup_f L(\hat{h}_t, f) \\
 &\leq \sup_f L(h^*, f) + \frac{8}{\eta T} \\
 &\leq D(g, h^*) + \eta + \frac{8}{\eta T} \\
 &\leq \eta + \frac{8}{\eta T}.
\end{align*}
Letting $\eta = \sqrt{8/T}$ and $T = 32/\epsilon^2$, 
we re-obtain the indistinguishability theorem.

\end{document}

### 2014-02-05 arXiv/boosted-mirror-descent.tex
created new arXiv version and updated notation

\documentclass[paper.tex]{subfiles}
\begin{document}

\section{Boosted Mirror Descent}
\label{sec:algorithm}

Our results primarily rely on a pair of novel convex optimization algorithms we call primal and dual boosted mirror descent. They are inspired by the decomposition used in \cite{Bach:2012b}.

\subsection{Convex Optimization}

Before introducing the boosted mirror descent (\bmd) algorithms, we establish notation and recall some basic definitions related to convex optimization.

Let $f$ be a real function whose domain is a normed inner product space. The {\em convex conjugate} $f^{*}$ of $f$ is defined as
\[
f^{*}(x') \eqdef \max_{x} \ip{x}{x'} - f(x).
\]
If $f$ is proper, lower semi-continuous, and convex, then by the 
Fenchel-Moreau theorem $f = f^{**}$, i.e.
\[
f(x) = \max_{x'} \ip{x}{x'} - f^{*}(x').
\]
Throughout the paper, we will assume without further notice that functions are proper and lower semi-continuous when required. The (sub)gradients of $f^{*}$ and $f$ can be defined as any argument that maximizes the previous two equations:
\[
\partial f^{*}(x') &\in \argmax_{x} \ip{x}{x'} - f(x) \\
\partial f(x) &\in \argmax_{x'} \ip{x}{x'} - f^{*}(x').
\]

In order to prove fast convergence, we will require additional assumptions beyond convexity. 
Define the \emph{Bregman divergence} of a convex function to be
\[
D_f(x \| x') = f(x)-\ip{\partial f(x')}{x'-x} - f(x'), 
\]
which is a measure of ``how convex'' $f$ is between $x$ and $x'$. 
We say that a function $f$ is \emph{strongly convex with respect 
to a norm $\|\cdot\|$} if
\[
D_f(x \| x') \geq \frac{1}{2}\|x-x'\|^2
\]
A related notion is \emph{strong smoothness}: a convex function $f$ 
is \emph{strongly smooth with respect to $\|\cdot\|$} if
\[ D_f(x \| x') \leq \frac{1}{2}\|x-x'\|^2 \]
An important result is that $f$ is strongly convex with respect to $\|\cdot\|$ 
if and only if its Fenchel conjugate $f^*$ is strongly smooth with respect 
to the dual norm $\|\cdot\|_*$.

\subsection{The Boosted Mirror Descent Algorithms}

We can now define the \bmd  algorithms. Let $X$ and $Y$ be convex subsets 
of the Banach spaces $\mcX$ and $\mcY$, respectively, and let 
$\blf{\cdot}{\cdot}$ be a bilinear map on $\mcX \times \mcY$; typically 
$\blf{\cdot}{\cdot}$ will correspond to an inner product.
% but we will use the bracket notation for convenience even when this is not the case.
Consider the two-argument loss function $L : X \times Y \to \reals$ 
defined as
\[
L(x,y) = \psi(x) + \blf{x}{y} - R(y)
\]
for which we wish to find arguments that obtain $L_{*} = \min_{x}\max_{y} L(x, y)$. Taking only the maximum, we have the one-argument loss function
\[
L(x)
&= \psi(x) + \max_{y \in \Theta} \left\{\blf{x}{y} - R(y) \right\} \\
&= \psi(x) + R^{*}(x).
\]
We can think of $\psi$ as a \emph{primal regularizer} and $R$ as a \emph{dual regularizer}.  The pair of regularization functions $\psi$ and $R$ define a flexible class of functions to optimize, as seen from the following examples:
\begin{enumerate}
\item Let $X$ be the probability simplex over a space $\sS$, let $\phi : \sS \to \reals^d$ be a collection 
      of statistics, and let $Y$ be the unit $L^1$ ball in $\reals^d$. Let $\psi(x) = 0$, 
      $\blf{x}{y} = y^T\bE_{v \sim x}[\phi(v)]$, and 
      $R(y) = \blf{x_0}{y}$. Then $L(x)$ is the 
      \emph{maximum mean discrepancy} between $x$ and $x_0$ relative to $\phi$.
\item \NA{JHH: I don't remember what this example what supposed to be}
	  Let $\psi(x) = 0$ and $R(y) = S^*(i^{-1}(y))$ where $S^*$ is the 
      convex conjugate of any strongly convex function $S : U \to \bR$. Then $L(x) = S(x)$ 
      as long as $i^{-1}(\Theta) = \sB_{U}$.
\item Let $X = Y = \reals^d$. Let $\blf{x}{y} = \ip{x}{y}$, $\psi(x) = \|x\|_1$, $R(x) = \blf{x_0}{y} + \frac{1}{2} \|y\|_2^2$. 
      Then $L(x) = \|x\|_1 + \frac{1}{2} \|x-x_0\|_2^2$. 
\end{enumerate}

We will assume 
throughout that $\psi$ and $R$ are both convex functions, and furthermore that 
$\arg\max_{x} \psi(x) + \blf{x}{y}$ can be efficiently 
computed for all values of $y$.

%In this paper we will consider a family of methods for minimizing $L(x)$, 
%which are based on Bach's generalization of the Frank-Wolfe algorithm and can 
%be interpreted as boosted mirror descent.

Mirror descent, combined with boosting, yields an algorithm for finding 
``saddle points'' of $L(x, y)$. 
For notational convenience, for a sequence of weights $\alpha_1,\alpha_2,\ldots$ 
let 
\(
\bar{x}_t = \frac{\sum_{s=1}^t \alpha_{s}x_{s}}{A_{t}} 
\quad \text{and} \quad 
\bar{y}_t = \frac{\sum_{s=1}^t \alpha_{s}y_{s}}{A_{t}},
\quad \text{where} \quad 
A_{t} \defined \sum_{s=1}^t \alpha_s.
\)

Then 
the {\em primal boosted mirror descent} (\primal) algorithm is:
\begin{enumerate}
\item $x_1 \in \arg\min_x \psi(x)$
\item $y_{t} \in \arg\max_{y \in \Theta} \blf{y}{x_t} - R(y) = \partial R^{*}(x_{t})$
\item $x_{t+1} \in \arg\min_{x} \psi(x) + \blf{x}{\bar{y}_t} = \partial \psi^{*}(-\bar y_{t})$
\end{enumerate}
As long as $\psi$ is strongly convex, for the proper choice of $\alpha_{t}$ we obtain the 
bound (see Corollary~\ref{cor:method-1}):
\begin{equation}
\sup_{y \in \Theta} L(\bar{x}_T, y) \leq L_{*} + O(1/T).
\end{equation}
In other words, $\bar{x}_T$ is close to being a global minimum of $L(x)$.
However, it is often the case that $\psi$ is not strongly convex but $R$ is strongly convex. As we shall see, this will be the case for our application to herding.
In this case we may wish to use the following slightly different algorithm, which we call {\em dual boosted mirror descent} (\dual):
\begin{enumerate}
\item $y_1 \in \arg\max_{y} R(y)$
\item $x_t \in \arg\min_{x} \psi(x) + \blf{x}{y_t} = \partial \psi^{*}(-y_{t})$
\item $y_{t+1} \in \arg\max_{y \in \Theta} \langle y, \bar{x}_t \rangle - R(y) = \partial R^{*}(\bar x_{t})$
\end{enumerate}
We then obtain the following bound 
when $R$ is strongly convex (see Corollary~\ref{cor:method-2}):
\[ \sup_{y \in \Theta} L(\bar{x}_T, y) \leq L_{*} + O(1/T). \]

\subsection{Connections to Other Gradient Descent Algorithms}

As suggested by their names, the \bmd algorithms are closely related: they are dual to each other in the sense that 
performing \primal on $L(x,y)$ is the same as performing \dual on $L'(y, x) \eqdef -L(x,y)$. This relationship is obvious by inspection of the update equations:
\(
\text{Primal} && \text{Dual} \\
y_{t} &= \partial R^{*}(x_{t}) & y_{t+1} &= \partial R^{*}(\bar x_{t}) \\
x_{t+1} &= \partial \psi^{*}(-\bar y_{t}) & x_t &= \partial \psi^{*}(-y_{t})
\)
This duality is analogous to that between mirror descent and conjugate gradient descent \citep{Bach:2012b}. \md minimizes $g(x) + f(x)$ and requires that $g$ be $\alpha$-strongly convex while \cgd minimizes $f^{*}(x) + g^{*}(x)$ and requires that $g$ be $\alpha$-strongly convex. The update equations for \md are
\(
y_{t} &= \partial f(x_{t})  \\
\bar y_{t} &=  (1 - \rho_{t+1})\partial g(x_{t}) - \rho_{t+1}y_{t}  \\
x_{t+1} &=  \partial g^{*}(\bar y_{t}) 
\)
and for \cgd are
\(
 y_{t+1} &= \partial f(x_{t}) \\
\bar y_{t+1} &= (1 - \rho_{t})\bar y_{t} + \rho_{t} y_{t+1}  \\
x_{t} &= \partial g^{*}(-\bar y_{t}).
\)
In fact, the \bmd algorithms are very closely related to \cgd and \md: \cgd can be obtained from \dual and \md obtained from \primal. The relationships are given in Table~\ref{tab:connections}. The table also shows how $\alpha_{t}$-weighted herding and $q$-herding --- which are discussed in the next section --- are obtained from \bmd.

\renewcommand{\arraystretch}{1.5}
\begin{center}
\begin{table}[t]
\begin{adjustwidth}{-.5in}{-.5in}
\caption{Formulation of herding and convex optimization algorithms in terms of \bmd. Full details relating \md and \cgd to \bmd are given in the Supplementary Material. See also \citet{Bach:2012b}.}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}
\begin{tabularx}{1.15\textwidth}{L|lllcL}
Algorithm 			& $\psi(x)$ 	& $R(y)$ & $\alpha_{t}$ & \bmd Algorithm & Notes 
\\ \hline
herding 			& 0			& $\blf{\bar x}{y} + \frac{1}{2} \| y \|_{\sH}^{2}$ 	& 1				 	& \dual & 
\\
$\alpha_{t}$-weighted herding 	& 0	& $\blf{\bar x}{y} + \frac{1}{2} \| y \|_{\sH}^{2}$ 	& unconstrained 	& \dual & $\sum_{s =1}^{t} \alpha_{t} > 0~\forall t$
\\ 
$q$-herding			& 0			& $\blf{\bar\phi}{y} + \frac{1}{q} \| y \|_{q}^{q}$ 	& unconstrained 	& \dual & $1 \le q \le 2$
\\ 
\md					& $g(x)$	& $f^{*}(y)$				& $\rho_{t}/\prod_{s=1}^{t} (1 - \rho_{s})$				& \primal & 
\\ 
\cgd				& $f^{*}(y)$& $g(x)$					& $\rho_{t}/\prod_{s=1}^{t} (1 - \rho_{s})$				& \dual &
\\
\end{tabularx}
\label{tab:connections}
\end{adjustwidth}
\end{table}
\end{center}
\end{document}

### 2014-02-05 arXiv/chen-proofs.tex
created new arXiv version and updated notation

\documentclass[paper.tex]{subfiles}

\begin{document}
\section{An Upper Bound in Infinite Dimensions}
\label{sec:chen-proofs}

An intriguing aspect of classical herding is that it is known to have a convergence rate of 
$O(1/T^2)$ under suitable conditions, as was shown in \citet{Chen:2010a}. In this section, 
we will need to assume that $\psi(x)$ is the indicator function for $X$ (i.e. $0$ on $X$ and 
$\infty$ outside of it) and that $R(y) = \ip{x_0}{y} + R_0(y)$, where 
$R_0$ exhibits \emph{quadratic scaling}: $R_0(sy) = s^2R_0(y)$. Note that 
this implies that $R_0^*(x-x_0) = R^*(x)$.
For simplicity 
we will focus on the case $\alpha_t = 1$. Then we have 
the following bound:
\begin{theorem}
\label{thm:chen}
Suppose that $R$ is strongly convex with respect to the norm $\|\cdot\|$ and 
let $r_* \eqdef \sup_{x \in U} \|x\|_*$. Also suppose that $\psi \equiv 0$ and 
that $R_0$ exhibits quadratic scaling as above. Define 
\[ \gamma^* \eqdef \inf_{y \in y} \sup_{x \in U} \frac{\ip{-y}{x-x_0}}{\sqrt{R(y)}}. \]
Then $R(y_{t+1}) \leq \frac{2r_*^2+(2r_*^2/\gamma_*)^2}{t^2}$ as long 
as $\gamma^* > 0$.
\end{theorem}
This corresponds to Proposition 1 of \citet{Chen:2010b} for $R_{0}(y) = \frac{1}{2}\|y\|_2^2$. Note that if there is an $\ell^2$ ball of radius $\epsilon$ 
around $x_0$ that is contained in $u$ then $\gamma^* \geq \frac{\epsilon}{r_*}$ 
in this case, which is how Chen's result follows from ours.
\begin{proof}[Proof of Theorem~\ref{thm:chen}]
The key quantity we will analyze is $(t-1)^2R(y_t)$. Note that we have
\begin{align*}
\lefteqn{(t-1)^2R(y_t) - t^2R(y_{t+1})} \\
 &= (t-1)^2R^*(\bar{x}_{t-1})-t^2R^*(\bar{x}_t) \\
 &= (t-1)^2R^*(\bar{x}_{t-1})-t^2R_0^*(\bar{x}_t-x_{0}) \\
 &= (t-1)^2\left[R^*(\bar{x}_{t-1})-R_0^*((t/(t-1))(\bar{x}_t-x_{0}))\right] \\
 &= (t-1)^2\left[R^*(\bar{x}_{t-1})-R^*\left(\frac{t\bar{x}_t-x_{0}}{t-1}\right)\right] \\
 &= (t-1)^2\left[\ip{\partial R^*(\bar{x}_{t-1})}{\bar{x}_{t-1}-\frac{t\bar{x}_t-x_{0}}{t-1}}-D_{R^*}\left(\frac{t\bar{x}_t-x_{0}}{t-1} \middle\| \bar{x}_{t-1}\right)\right] \\
 &\geq (t-1)^2\left[\ip{y_t}{-\frac{1}{t-1}(x_t-x_{0})}-\frac{1}{2}\left\|\frac{x_t-x_{0}}{t-1}\right\|_*^2\right] \\
 &= -(t-1)\ip{y_t}{x_t-x_{0}}-\frac{1}{2}\|x_t-x_{0}\|_*^2 \\
 &= -(t-1)\left[\inf_{x \in U} \ip{y_t}{x} - \ip{y_t}{x_{0}}\right]-\frac{1}{2}\|x_t-x_{0}\|_*^2 \\
 &\geq (t-1)\sqrt{R(y_t)}\left[\inf_{y} \sup_{x \in U} \frac{\ip{-y}{x-x_0}}{\sqrt{R(y_t)}}\right]-2r_*^2 \\
 &= \sqrt{(t-1)^2R(y_t)}\gamma^*-2r_*^2.
\end{align*}
There are now two possibilities for each $t$:
\begin{enumerate}
\item $t^2R(y_{t+1}) < (t-1)^2R(y_t)$, or
\item $(t-1)^2R(y_t) \leq \frac{4r_*^4}{(\gamma^*)^2}$.
\end{enumerate}
In particular, consider the value of $t$ for which 
$t^2R(y_{t+1})$ is maximized. Then by assumption 
case $1$ cannot hold and so $(t-1)^2R(y_t) \leq \frac{4r_*^4}{(\gamma^*)^2}$.
But then by applying the above sequence of inequalities 
we also have 
\begin{align*}
\lefteqn{t^2R(y_{t+1}) - (t-1)^2R(y_t)} \\
 &\leq 2r_*^2 - \sqrt{(t-1)^2R(y_t)}\gamma^* \\
 &\leq 2r_*^2,
\end{align*}
hence $t^2R(y_{t+1}) \leq 2r_*^2 + \frac{4r_*^4}{(\gamma_*)^2}$. 
Since $t^2R(y_{t+1})$ was maximal by assumption, we have
$R(y_{t+1}) \leq \frac{2r_*^2+(2r_*/\gamma_*)^2}{t^2}$, which 
completes the proof.
\end{proof}

\end{document}

### 2014-02-05 arXiv/herding-connection.tex
created new arXiv version and updated notation

\documentclass[paper.tex]{subfiles}
\begin{document}
\section{Application to Herding}
\label{sec:herding}

We now derive herding as a special case of \dual. The standard setup for herding is as follows: we are given a 
space $\sS$ and a feature mapping $\phi : \sS \to \sH$ for some 
reproducing kernel Hilbert space $\sH$ \cite{Hofmann:2008}. 
We define $\Delta_{\sS}$ to be the simplex 
of probability distributions over $\sS$ and define the marginal polytope 
$\sM \defined \{\bE_{v \sim x}[\phi(v)] \mid x \in \Delta_{\sS}\}$.

The goal of herding is to construct samples $v_1,\ldots,v_T \in \sS$ 
such that $\|\frac{1}{T}\sum_{t=1}^T \phi(v_t)-\bar{\phi}\|_{\sH}^2$ 
is minimized for some given $\bar{\phi} \in \sM$. One can typically
think of $\bar \phi$ as being the expected value of $\phi$ under some 
distribution $x_{0} \in \Delta_{\sS}$.

The herding algorithm uses the following updates\footnote{The original 
algorithm includes an additional $\bar{\phi}$ term in the third bullet 
point and indexes $x_t$ as $x_{t+1}$.} \citep[][equations (1) and (2)]{Chen:2010a}:
\begin{enumerate}
\item $w_1 = \bar{\phi}$
\item $v_t \in \arg\max_{v \in \sS} \ip{w_t}{\phi(v)}$
\item $w_{t+1} = \sum_{s=1}^t [\bar{\phi}-\phi(v_s)]$
\end{enumerate}
If we define $y_t \defined -w_t/t$, then these updates are equivalent to
\begin{enumerate}
\item $y_1 = -\bar{\phi}$
\item $x_t \in \arg\min_{x \in \Delta_{\sS}} \ip{\bE_{x}\phi}{y_t}$
\item $y_{t+1} = \frac{1}{t} \sum_{s=1}^t \bE_{x}\phi - \bar{\phi}$
\end{enumerate}
Note that we have replaced the point $v_t$ with a distribution $x_t$. 
This does not change the algorithm as $\ip{\bE_{x}\phi}{y_t}$ will 
always be minimized at an extreme point of $\Delta_{\sS}$, which corresponds 
to a point mass.

The above updates correspond to \dual for $X = \Delta_{\sS}$, $Y = \sH$, 
$\blf{x}{y} = \ip{\bE_{x}\phi}{y}_{\sH}$, $\alpha_t = 1$, $\psi \equiv 0$, and
$R(y) = \ip{\bar{\phi}}{y}_{\sH} + \frac{1}{2}\|y\|_{\sH}^2$.
To see this, note that 
\begin{align}
\arg\min_{y \in \sH} \blf{x}{y} - R(y)
 &= \arg\min_{y \in \sH} \ip{\bE_{x}\phi-\bar{\phi}}{y}_{\sH} - \frac{1}{2}\|y\|_{\sH}^2 \\
 &= \bE_u[\phi(x)] - \bar{\phi}.
\end{align}
For $x = 0$, this gives us the first bullet point. For $x = \bar{x}_t$, 
this gives us the third bullet point. Finally, the second bullet point 
is already in the form given in \dual.

Note that $R$ is strongly convex with respect to $\|\cdot\|_{\sH}$ and that 
$R^*(u) = \frac{1}{2}\|\bar{\phi}-\bE_{x}\phi\|_{\sH}^2$. We can 
therefore apply Corollary~\ref{thm:method-2} to obtain the bound
\begin{align*}
\lefteqn{\frac{1}{2}\|\bar{\phi}-\frac{1}{T}\sum_{t=1}^T \phi(x_t)\|_2^2} \\
 &\leq \inf_{u^*} \frac{1}{2}\|\bar{\phi}-\bE_{u^*}[\phi(x)]\|_2^2 + \frac{2r^2\log(T+1)}{T} \\
 &= \frac{2r^2\log(T+1)}{T},
\end{align*}
where $r \defined \sup_{m \in \sM} \|m\|_2$. This is a slightly weaker 
version of the typical herding bound on MMD.

If we do not restrict $\alpha_{t}$ to be 1, we obtain a class of algorithms
we call {\em $\alpha_{t}$-weighted herding}. For the $\alpha_t = t$ we obtain 
a stronger bound of $\frac{8r^2}{T}$.

\subsection{$q$-herding}
\label{sec:infinite-case}

The convergence bound for herding implicitly assumes that 
$\sM$ is bounded in norm. Assuming that $\sS$ is compact 
and that $\sH$ is finite-dimensional, this will always be true. However, in 
the infinite-dimensional case this may no longer be true. 
For instance, consider $\sS = \{1,\ldots,N\}$ and let 
$\phi : \sS \to \ell^3(\bZ_{\geq 0})$ be
defined by $\phi(n)_i = \frac{1}{\sqrt{n+i}}$. Then 
$\|\phi(n)\|_2 = \infty$ for each $n \in \sS$, and so herding 
yields no convergence guarantees. Indeed, the algorithm 
does not even necessarily make sense: since $y$ is 
no longer guaranteed to have finite $\ell^2$ norm, the 
inner product $\ip{\phi(v)}{y}$ may not be defined.

Note, however, that in this case
$\|\phi(n)\|_3 = \sqrt[3]{\sum_{i=0}^{\infty} (n+i)^{-3/2}} < \infty$.
(In fact, the maximum is equal to
$\|\phi(1)\|_3 = \sqrt[3]{\zeta(3/2)} < 1.378$.) This motivates 
a generalization of herding in which we minimize 
$\|\bE_{x}[\phi]-\bar{\phi}\|_3^3$ instead of 
$\|\bE_{x}[\phi]-\bar{\phi}\|_2^2$.

More generally, for any 
$p \geq 2$ and $q$ such that $\frac{1}{p} + \frac{1}{q} = 1$, we can take 
$h \equiv 0$ and $R(y) = \ip{\bar{\phi}}{y} + \frac{1}{q}\|y\|_q^q$. 
Then $R^*(u) = \frac{1}{p}\|\bE_{x}\phi-\bar{\phi}\|_p^p$. And 
$\left(\frac{\partial R^*}{\partial v}\right)_i = \sign(v_i) \cdot |v_i|^{p-1}$. 
Therefore, define the ``correction function'' $C_p$ by
\begin{equation}
C_p(u)_i \defined \sign(v_i) \cdot |v_i|^{p-1}.
\end{equation}
We obtain the following \emph{$q$-herding} updates:
\begin{itemize}
\item $y_{1} = C_p(-\bar{\phi})$
\item $v_{t} \in \arg\min_{v \in \sS} \ip{\phi(v)}{y_t}$
\item $y_{t+1} = C_p\left(\frac{1}{t} \sum_{s=1}^t \phi(v_s)-\bar{\phi}\right)$
\end{itemize}
Note that $\|C_p(v)\|_q^q = \sum_{i} |v_i|^{q(p-1)} = \sum_{i} |v_i|^p = \|u\|_p^p$. 
Therefore, $C_p(v) \in \ell^q(\bZ_{\geq 0})$ whenever $v \in \ell^p(\bZ_{\geq 0})$. By 
H\"{o}lder's inequality, it follows that $\ip{u}{y_t}$ exists for all 
$v \in \ell^p(\bZ_{\geq 0})$ and hence the $q$-herding updates are well-defined.

We now show that $q$-herding converges at a rate of $O(T^{-1/p})$. By Proposition~\ref{prop:method-2}, we know that
\begin{align}
\frac{1}{p}\|\bar{\phi}-\frac{1}{T}\sum_{t=1}^T \phi(v_t)\|_p^p \leq \frac{1}{A_T} \sum_{t=1}^T A_tD_{R^*}(\bar{v}_{t+1}\|\bar{v}_t).
\end{align}
We can no longer use Corollary~\ref{cor:method-2} since $\|\cdot\|_p^p$ is not strongly convex. 
Instead, we use the following two inequalities (proofs are given in the Supplementary Material):
\begin{lemma}
Let $x, y \in \bR_{\geq 0}$. Then:
\begin{equation*}
x^p-py^{p-1}x+(p-1)y^p \leq \binom{p}{2}(x-y)^2\max(x,y)^{p-2}
\end{equation*}
and
\begin{equation*}
(y-z)^2\max(x,y)^{p-2} \leq 4\max(x,y,z)^p
\end{equation*}
\end{lemma}
Using these, we obtain the bound
\(
\lefteqn{D_{R^{*}}(\bar{v}_{t+1} \| \bar{v}_t)} \\
 &= \frac{1}{p} \sum_{i} \bar{v}_{t+1,i}^p - p\bar{v}_{t+1,i}\bar{v}_{t,i}^{p-1} + (p-1)\bar{v}_{t,i}^{p} \\
 &\leq \frac{1}{p}\binom{p}{2} \sum_i (\bar{v}_{t+1,i}-\bar{v}_{t,i})^2\max(\bar{v}_{t+1,i},\bar{v}_{t,i})^{p-2} \\
 &= \frac{1}{p}\binom{p}{2}\frac{\alpha_{t+1}^2}{A_{t+1}^2} \sum_i (u_{t+1,i}-\bar{v}_{t})^2\max(\bar{v}_{t+1,i},\bar{v}_{t,i})^{p-2} \\
 &\leq 2(p-1)\frac{\alpha_{t+1}^2}{A_{t+1}^2} \sum_i \max(\bar{v}_{t,i},\bar{v}_{t+1,i},u_{t+1,i})^p \\
 &\leq 2(p-1)\frac{\alpha_{t+1}^2}{A_{t+1}^2} \sum_i \bar{v}_{t,i}^p+\bar{v}_{t+1,i}^p+u_{t+1,i}^p \\
 &\leq 6(p-1)\frac{\alpha_{t+1}^2}{A_{t+1}^2} r_p^p
\)
Plugging back in yields a convergence rate of $\frac{6(p-1)r_p^p(\log(T)+1)}{T}$ for $\alpha_t = 1$ 
and $\frac{24(p-1)r_p^p}{T}$ for $\alpha_t = t$. We therefore obtain a convergence rate of $O(T^{-1/p})$ 
in the $\ell^p$ norm, even if $\sM$ is not bounded in the $\ell^2$ norm.

\subsection{Herding and Maximum Entropy}
\label{sec:max-ent}

\citet{Welling:2009a} and \citet{Bach:2012b} have observed that herding appears to be approximately building 
samples from the maximum entropy distribution for a given collection of moment 
constraints. In this section we present a herding-style algorithm that more 
explicitly yields a maximum-entropy distribution. Our hope is that comparing 
the two algorithms may help to make the connection between herding and maximum 
entropy more explicit.

The key idea is to use the primal regularizer $\psi(x) = -\eta H(x)$, where 
$H(x)$ is the entropy of $x$ and $\eta \in \bR_{>0}$. We will still use 
$R(y) = \ip{\bar{\phi}}{y} + \frac{1}{2} \|y\|_{\sH}^2$.
In this case, the \dual updates are
\begin{itemize}
\item $y_1 = -\bar{\phi}$
\item $x_t(v) \propto \exp\left(-\frac{1}{\eta}\ip{\phi(v)}{y_t}\right)$
\item $y_{t+1} = \bE_{\bar{x}_t}\phi - \bar\phi$.
\end{itemize}
Corollary~\ref{cor:method-2} implies that, for 
$\bar{x}_T = \frac{1}{T} \sum_{t=1}^T x_t$, we have
\begin{align*}
\lefteqn{-\eta H(\bar{x}_T) + \|\bar{\phi}-\bE_{\bar{x}_T}[\phi(x)]\|_{\sH}^2} \\
 &\leq \inf_{x} -\eta H(x) + \|\bar{\phi}-\bE_{x}\phi\|_{\sH}^2 + \frac{2r^2\log(T+1)}{T} \\
 &\leq -\eta H_{\max}(\bar{\phi}) + \frac{2r^2\log(T+1)}{T},
\end{align*}
where $H_{\max}(\bar{\phi})$ is the maximum entropy of any distribution whose 
moments are $\bar{\phi}$.

Note that as $\eta \to 0$, $u_{t}$ concentrates around its mode, converging to a delta function at the herding sample location. It is thus possible that, under the proper regularity conditions, it can be shown using entropically regularized \bmd that herding converges to a distribution that is close to the maximum entropy distribution. However, we leave this line of investigation for future work.

\end{document}

### 2014-02-05 arXiv/introduction.tex
created new arXiv version and updated notation

--- 2013-10-04 icml/introduction.tex
@@
 
-In order to define herding, let $\sX$ be an observation space, $\phi : \sX \to \sB_{\Theta}$ a 
-feature map to an inner product space, and $\bar\phi = \bE_{x \sim p}[\phi(x)]$ the empirical 
+In order to define herding, let $\sX$ be an observation space, $\phi : \sX \to \sF$ a 
+feature map to an inner product feature space, and $\bar\phi = \bE_{x \sim p}[\phi(x)]$ the empirical 
 moments of some distribution. Then the herding algorithm is the following simple procedure to 
@@
 \[
-x_{t} &\in \argmin_{x \in \sX} \ip{\theta_{t}}{\phi(x)} \\
-\theta_{t} &= \theta_{t-1}  + \bar \phi - \phi(x_{t}).
+x_{t} &\in \argmin_{x \in \sX} \ip{\phi(x)}{y_{t-1}} \\
+y_{t} &= y_{t-1}  + \bar \phi - \phi(x_{t}).
 \]
-In this paper we generalize the herding algorithm to minimize $h(h) + R^{*}(u)$ via the following set of updates, which 
-require only the convexity of $h$ and $R$:
+In this paper we generalize the herding algorithm to minimize $\psi(x) + R^{*}(x)$ via the following set of updates, which 
+require only the convexity of $\psi$ and $R$:
 \[
-u_t &\in \argmin_{u} h(u) + \ip{\theta_t}{u} \\
-\theta_{t+1} &\in \argmax_{\theta} \ip{\theta}{\hat{u}_t} - R(\theta),
+x_t &\in \argmin_{x} \psi(x) + \ip{x}{y_{t-1}} \\
+y_{t} &\in \argmax_{y} \ip{\bar{x}_t}{y} - R(y),
 \]
-where $\hat{u}_t$ is a certain weighted average of $u_1,\ldots,u_t$. Our 
+where $\bar{x}_t$ is a certain weighted average of $x_1,\ldots,x_t$. Our 
 algorithm is described in more detail in Section~\ref{sec:algorithm}. This 
@@
 as special cases. A full table outlining the relationships for various choices of 
-$h$ and $R$ is given in Table~\ref{tab:connections}.
+$\psi$ and $R$ is given in Table~\ref{tab:connections}.
 

### 2014-02-05 arXiv/lower-bounds.tex
created new arXiv version and updated notation

--- 2013-10-04 icml/lower-bounds.tex
@@
 \[
-\theta_{t} = \sum_{s =1}^{t}w_{t} \phi(x_{t}),
+y_{t} = \sum_{s =1}^{t}w_{t} \phi(x_{t}),
 \]
@@
 Without loss of generality we can assume that $\alpha$ is monotonically increasing.
-We will consider the standard $\ell^{2}$ herding loss function $L(\theta) = \|\theta - \bar \phi\|_2^2$.
+We will consider the standard $\ell^{2}$ herding loss function $L(y) = \|y - \bar \phi\|_2^2$.
 
@@
 \[
-\|\theta_{t}- \bar \phi\|_2^2 \ge c^{2}\sum_{n > t} 1/\alpha(n)^{2} = \Theta(1/\beta(t)) \label{eq:convergence-lb}
+\|y_{t}- \bar \phi\|_2^2 \ge c^{2}\sum_{n > t} 1/\alpha(n)^{2} = y(1/\beta(t)) \label{eq:convergence-lb}
 \]
@@
 Then we can choose $\alpha = \sqrt \gamma \in \omega(t)$ to satisfy \eqref{eq:convergence-lb}.
-Therefore, any sparse update method must have convergence rate no faster than $\Theta(1/\beta(t))$ for all superlinear $\beta$, and hence no faster than $\Theta(1/t)$.
+Therefore, any sparse update method must have convergence rate no faster than $y(1/\beta(t))$ for all superlinear $\beta$, and hence no faster than $y(1/t)$.
 

### 2014-02-05 arXiv/paper.tex
created new arXiv version and updated notation

--- 2013-10-04 icml/paper.tex
@@
-\documentclass{article}
+\documentclass[reqno,oneside,a4paper]{amsart}
 \usepackage{import, subfiles}
-\usepackage{hyperref, icml2014}
+\usepackage{natbib}
+\usepackage{hyperref}
 \usepackage{tabularx}
+\usepackage{chngpage}
 \input{latex-defs.tex}
 
-\icmltitlerunning{Herding and Boosted Mirror Descent}
+\allowdisplaybreaks
 
@@
 
-\twocolumn[
-\icmltitle{Herding and its Generalizations via Boosted Mirror Descent}
+\title{Conjugate Gradient, Boosting, and Herding: \\ 
+	   Connections and Application}
 
-\icmlauthor{Jonathan Huggins \NA{add footnote on authorship}}{jhuggins@mit.edu}
-\icmladdress{MIT CSAIL,
-             77 Massachusetts Avenue, Cambridge, MA 02139 USA}
-\icmlauthor{Jacob Steinhardt}{jsteinhardt@cs.stanford.edu}
-\icmladdress{Stanford University,
-             353 Serra Street, Stanford, CA 94305 USA}
+\thanks{The authors contributed equally to this work.}
 
-% You may provide any keywords that you 
-% find helpful for describing your paper; these are used to populate 
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{boosting, optimization, mirror descent, frank-wolfe, conjugate gradient descent, herding}
+\author[J.~H.~Huggins]{Jonathan H.~Huggins}
+\address{Massachusetts Institute of Technology}
+\urladdr{http://jhuggins.org/}
+\email{jhuggins@mit.edu}
 
-\vskip 0.3in
-]
+\author[J.~Steinhardt]{Jacob Steinhardt}
+\address{Stanford University}
+\urladdr{http://cs.stanford.edu/~jsteinhardt/}
+\email{jsteinhardt@cs.stanford.edu}
+
+\keywords{boosting, optimization, mirror descent, Frank-Wolfe, conjugate gradient descent, herding, graph sparsification}
 
@@
 We introduce a unifying framework for understanding herding and 
-related algorithms. The framework is based on pair of optimization 
+related algorithms. The framework is based on a pair of optimization 
 algorithms that we call primal and dual boosted mirror descent.
@@
 
+\maketitle
+
 \subfile{introduction.tex}
@@
 \subfile{lower-bounds.tex}
-%\subfile{chen-proofs.tex}
+\subfile{chen-proofs.tex}
 \subfile{conclusion.tex}
@@
 
-\end{document}
+\end{document} \maketitle

### 2014-02-05 arXiv/boosted-mirror-descent.tex
notation fixes

@@
 \[
-f^{*}(x') \eqdef \max_{x} \ip{x}{x'} - f(x).
+f^{*}(x') \eqdef \sup_{x} \ip{x}{x'} - f(x).
 \]
@@
 \[
-f(x) = \max_{x'} \ip{x}{x'} - f^{*}(x').
+f(x) = \sup_{x'} \ip{x}{x'} - f^{*}(x').
 \]
@@
 \[
-\partial f^{*}(x') &\in \argmax_{x} \ip{x}{x'} - f(x) \\
-\partial f(x) &\in \argmax_{x'} \ip{x}{x'} - f^{*}(x').
+\partial f^{*}(x') &\in \arg\sup_{x} \ip{x}{x'} - f(x) \\
+\partial f(x) &\in \arg\sup_{x'} \ip{x}{x'} - f^{*}(x').
 \]
@@
 \]
-for which we wish to find arguments that obtain $L_{*} = \min_{x}\max_{y} L(x, y)$. Taking only the maximum, we have the one-argument loss function
+for which we wish to find arguments that obtain $L_{*} = \inf_{x}\sup_{y} L(x, y)$. Taking only the maximum, we have the one-argument loss function
 \[
 L(x)
-&= \psi(x) + \max_{y \in \Theta} \left\{\blf{x}{y} - R(y) \right\} \\
+&= \psi(x) + \sup_{y \in Y} \left\{\blf{x}{y} - R(y) \right\} \\
 &= \psi(x) + R^{*}(x).
@@
       convex conjugate of any strongly convex function $S : U \to \bR$. Then $L(x) = S(x)$ 
-      as long as $i^{-1}(\Theta) = \sB_{U}$.
+      as long as $i^{-1}(Y) = \sB_{U}$.
 \item Let $X = Y = \reals^d$. Let $\blf{x}{y} = \ip{x}{y}$, $\psi(x) = \|x\|_1$, $R(x) = \blf{x_0}{y} + \frac{1}{2} \|y\|_2^2$. 
@@
 throughout that $\psi$ and $R$ are both convex functions, and furthermore that 
-$\arg\max_{x} \psi(x) + \blf{x}{y}$ can be efficiently 
+$\arg\sup_{x} \psi(x) + \blf{x}{y}$ can be efficiently 
 computed for all values of $y$.
@@
 \begin{enumerate}
-\item $x_1 \in \arg\min_x \psi(x)$
-\item $y_{t} \in \arg\max_{y \in \Theta} \blf{y}{x_t} - R(y) = \partial R^{*}(x_{t})$
-\item $x_{t+1} \in \arg\min_{x} \psi(x) + \blf{x}{\bar{y}_t} = \partial \psi^{*}(-\bar y_{t})$
+\item $x_1 \in \arg\inf_x \psi(x)$
+\item $y_{t} \in \arg\sup_{y \in Y} \blf{y}{x_t} - R(y) = \partial R^{*}(x_{t})$
+\item $x_{t+1} \in \arg\inf_{x} \psi(x) + \blf{x}{\bar{y}_t} = \partial \psi^{*}(-\bar y_{t})$
 \end{enumerate}
@@
 \begin{equation}
-\sup_{y \in \Theta} L(\bar{x}_T, y) \leq L_{*} + O(1/T).
+\sup_{y \in Y} L(\bar{x}_T, y) \leq L_{*} + O(1/T).
 \end{equation}
@@
 \begin{enumerate}
-\item $y_1 \in \arg\max_{y} R(y)$
-\item $x_t \in \arg\min_{x} \psi(x) + \blf{x}{y_t} = \partial \psi^{*}(-y_{t})$
-\item $y_{t+1} \in \arg\max_{y \in \Theta} \langle y, \bar{x}_t \rangle - R(y) = \partial R^{*}(\bar x_{t})$
+\item $y_1 \in \arg\sup_{y} R(y)$
+\item $x_t \in \arg\inf_{x} \psi(x) + \blf{x}{y_t} = \partial \psi^{*}(-y_{t})$
+\item $y_{t+1} \in \arg\sup_{y \in Y} \langle y, \bar{x}_t \rangle - R(y) = \partial R^{*}(\bar x_{t})$
 \end{enumerate}
@@
 when $R$ is strongly convex (see Corollary~\ref{cor:method-2}):
-\[ \sup_{y \in \Theta} L(\bar{x}_T, y) \leq L_{*} + O(1/T). \]
+\[ \sup_{y \in Y} L(\bar{x}_T, y) \leq L_{*} + O(1/T). \]
 

### 2014-02-18 arXiv/paper.tex
updated title, abstract

@@
 \usepackage{chngpage}
+\usepackage{fullpage}
 \input{latex-defs.tex}
@@
 
-\title{Conjugate Gradient, Boosting, and Herding: \\ 
-	   Connections and Application}
+\title{Boosting, Sparsity, and Conditional Gradient: A Tutorial}
+%Conjugate Gradient, Boosting, and Herding: \\ 
+%	   Connections and Application}
 
@@
 \begin{abstract} 
-Much has been learned about herding since it was first proposed.
-We introduce a unifying framework for understanding herding and 
-related algorithms. The framework is based on a pair of optimization 
-algorithms that we call primal and dual boosted mirror descent.
-We define two classes of generalized herding algorithms, 
-including one for infinite dimensional spaces
-that may apply even when herding does not. A number 
-of convergence bounds that apply to herding and generalized
-herding are obtained. Along the way we provide
-a more cohesive picture of the connections between herding, 
-conditional gradient descent, and maximum entropy. 
+Many recent papers in machine learning and theoretical computer science have 
+focused on the idea of sparse or low-rank ``approximations'' to some object: 
+for instance, graph sparsification \cite{?}, low-rank SDP approximation \cite{?}, 
+and pseudosampling \cite{?}. In this paper we show that, whenever presented with 
+a strongly convex optimization problem whose dual problem is linear, we can obtain 
+sparse approximate solutions, and use this to re-derive many of the known 
+sparsification results.
+% Much has been learned about herding since it was first proposed.
+% We introduce a unifying framework for understanding herding and 
+% related algorithms. The framework is based on a pair of optimization 
+% algorithms that we call primal and dual boosted mirror descent.
+% We define two classes of generalized herding algorithms, 
+% including one for infinite dimensional spaces
+% that may apply even when herding does not. A number 
+% of convergence bounds that apply to herding and generalized
+% herding are obtained. Along the way we provide
+% a more cohesive picture of the connections between herding, 
+% conditional gradient descent, and maximum entropy. 
 \end{abstract}
@@
 \bibliography{herding}
-\bibliographystyle{icml2014}
+\bibliographystyle{plainnat} %icml2014}
 

### 2014-02-18 arXiv/introduction.tex
added a bit of the intro

--- 2013-10-04 icml/introduction.tex
@@
 
-Many algorithms, such as SMC \citep{smc} and MCMC \citep{mcmc}, approximate a 
-distribution by samples. Most such algorithms require access to a full posterior and 
-may be prone to becoming stuck in local modes. In contrast, the herding algorithm introduced in 
-\citet{Welling:2009a} generates samples whose moments are guaranteed to converge 
-to any given collection of feasible moments. Moreover, \emph{only} the moments 
-are required; a full posterior is not needed.
+The purpose of this paper is to serve as a tutorial on sparsification. While 
+most of the results presented are known, they are often distributed across the 
+theory, machine learning, and optimization literatures, and we hope to bring 
+them all into one place, as well as to provide a unifying perspective. To start, 
+we define a generalized notion of sparsity that encompasses both sparsity and 
+low-rank conditions as special cases. Throughout, we will let $S$ denote a 
+compact subset of $\bR^n$ and let $\Conv(S)$ denote the convex hull of $S$ 
+(which will always be a closed set due to compactness of $S$). Two particularly 
+important examples are:
+\begin{example}
+\label{ex:simplex}
+Let $S$ consist of the $n$ standard unit vectors $e_1, \ldots, e_n$. Then 
+$\Conv(S)$ is the $n$-dimensional simplex $\Delta_n$.
+\end{example}
 
-Since its introduction, herding has been shown to have many desirable properties.
-For instance, \citet{Chen:2010a} showed that the moments converge at a rate of $O(1/T^{2})$ for 
-$T$ samples, which is even faster than the $O(1/T)$ rate that would be obtained 
-by sampling independently from the exact posterior. Furthermore, \citet{Bach:2012a} show that herding is 
-equivalent to conditional gradient, and \citet{Huszar:2012} show that 
-herding is equivalent to Bayesian quadrature. Herding was originally conceived of as a method 
-for inference in Markov random fields, and this line of work has been further explored in 
-\citet{Welling:2009a}, \citet{Gelfand:2010}, and \citet{Bornn:2013}.
+\begin{example}
+\label{ex:sdp}
+Let $S = \{vv^{\top} \mid \|v\|_2 = 1\}$. An alternative characterization 
+of $S$ is all rank-$1$, positive semidefinite matrices with trace $1$. The 
+convex hull of $S$ is $\Conv(S) = \{V \mid V \succeq 0, \Tr(V) = 1\}$.
+\end{example}
 
-In order to define herding, let $\sX$ be an observation space, $\phi : \sX \to \sB_{\Theta}$ a 
-feature map to an inner product space, and $\bar\phi = \bE_{x \sim p}[\phi(x)]$ the empirical 
-moments of some distribution. Then the herding algorithm is the following simple procedure to 
-generate pseudosamples from $\sX$:
-\[
-x_{t} &\in \argmin_{x \in \sX} \ip{\theta_{t}}{\phi(x)} \\
-\theta_{t} &= \theta_{t-1}  + \bar \phi - \phi(x_{t}).
-\]
-In this paper we generalize the herding algorithm to the following set of updates, which 
-require only the convexity of $h$ and $R$:
-\[
-u_t &\in \argmin_{u} h(u) + \ip{\theta_t}{u} \\
-\theta_{t+1} &\in \argmax_{\theta} \ip{\theta}{\hat{u}_t} - R(\theta),
-\]
-where $\hat{u}_t$ is a certain weighted average of $u_1,\ldots,u_t$. Our 
-algorithm is described in more detail in Section~\ref{sec:algorithm}. This 
-generalization has several advantages:
-\begin{itemize}
-\item It allows us to cast herding in the framework of mirror descent \cite{Beck:2003}, 
-      which lets us better understand its convergence properties.
-\item It allows us to obtain generalizations of herding for infinite 
-      dimensional spaces that apply in cases where standard herding does not.
-\item It makes the known connections between herding, subgradient, and 
-      conditional gradient methods more clear. 
-\end{itemize}
-In particular, while it was known that herding was a form of conditional 
-gradient descent \cite{Bach:2012a} and hence (via the relationship in \citet{Bach:2012b}) also 
-a form of subgradient descent, our framework exposes all three algorithms 
-as special cases. A full table outlining the relationships for various choices of 
-$h$ and $R$ is given in Table~\ref{tab:connections}.
+\begin{definition}
+Given a set $S \subseteq \bR^n$, a \emph{$(S, k)$-sparse} point is a point 
+$x \in \Conv(S)$ that is a convex combination of at most $k$ extreme 
+points of $\Conv(S)$.
+\end{definition}
 
-In summary, we introduce a novel gradient descent algorithm to investigate herding and its generalizations, and make the following contributions:
-\begin{itemize}
-\item In Section \ref{sec:algorithm} we introduce a pair of convex optimization algorithms, which we call primal and dual boosted mirror descent (\primal and \dual).
+Recall that an extreme point is a point that cannot be written as a convex 
+combination of other points in $\Conv(S)$. Importantly, every extreme point 
+of $\Conv(S)$ lies in $S$. So, in Example~\ref{ex:simplex}, every 
+$(S,k)$-sparse point is also $k$-sparse in the usual sense. In 
+Example~\ref{ex:sdp}, every $(S,k)$-sparse point has rank at most $k$, since 
+it is in the convex hull of $k$ rank-$1$ matrices.
+
+\begin{example}
+\label{ex:marginal-polytope}
+Given a space $\sX$ and a function $\phi : \sX \to \bR^d$, let 
+$S = \{\phi(x) \mid x \in \sX\}$. Then $\Conv(S)$ is called the 
+\emph{marginal polytope} of $\sX$ with respect to $\phi$. The 
+marginal polytope consists of all possible expected values of 
+$\phi(x)$ with respect to an arbitrary probability distribution 
+on $\sX$. A $(S,k)$-sparse point in $\Conv(S)$ is a point that 
+can be written as a weighted average $\sum_{i=1}^k w_i \phi(x_i)$. 
+In this case, we call the $x_i$ \emph{pseudosamples}.
+\end{example}
+
+\begin{example}
+\label{ex:graph}
+TODO: graph sparsification example
+\end{example}
+
+%% Many algorithms, such as SMC \citep{smc} and MCMC \citep{mcmc}, approximate a 
+%% distribution with samples. Most such algorithms require access to a full posterior and 
+%% may be prone to becoming stuck in local modes. In contrast, the herding algorithm introduced in 
+%% \citet{Welling:2009a} generates samples whose moments are guaranteed to converge 
+%% to any given collection of feasible moments. Moreover, \emph{only} the moments 
+%% are required; a full posterior is not needed.
+%% 
+%% Since its introduction, herding has been shown to have many desirable properties.
+%% For instance, \citet{Chen:2010a} showed that the moments converge at a rate of $O(1/T^{2})$ for 
+%% $T$ samples, which is even faster than the $O(1/T)$ rate that would be obtained 
+%% by sampling independently from the exact posterior. Furthermore, \citet{Bach:2012a} show that herding is 
+%% equivalent to conditional gradient, and \citet{Huszar:2012} show that 
+%% herding is equivalent to Bayesian quadrature. Herding was originally conceived of as a method 
+%% for inference in Markov random fields, and this line of work has been further explored in 
+%% \citet{Welling:2009a}, \citet{Gelfand:2010}, and \citet{Bornn:2013}.
+%% 
+%% In order to define herding, let $\sX$ be an observation space, $\phi : \sX \to \sF$ a 
+%% feature map to an inner product feature space, and $\bar\phi = \bE_{x \sim p}[\phi(x)]$ the empirical 
+%% moments of some distribution. Then the herding algorithm is the following simple procedure to 
+%% generate pseudosamples from $\sX$:
+%% \[
+%% x_{t} &\in \argmin_{x \in \sX} \ip{\phi(x)}{y_{t-1}} \\
+%% y_{t} &= y_{t-1}  + \bar \phi - \phi(x_{t}).
+%% \]
+%% In this paper we generalize the herding algorithm to minimize $\psi(x) + R^{*}(x)$ via the following set of updates, which 
+%% require only the convexity of $\psi$ and $R$:
+%% \[
+%% x_t &\in \argmin_{x} \psi(x) + \ip{x}{y_{t-1}} \\
+%% y_{t} &\in \argmax_{y} \ip{\bar{x}_t}{y} - R(y),
+%% \]
+%% where $\bar{x}_t$ is a certain weighted average of $x_1,\ldots,x_t$. Our 
+%% algorithm is described in more detail in Section~\ref{sec:algorithm}. This 
+%% generalization has several advantages:
+%% \begin{itemize}
+%% \item It allows us to cast herding in the framework of mirror descent \cite{Beck:2003}, 
+%%       which lets us better understand its convergence properties.
+%% \item It allows us to obtain generalizations of herding for infinite 
+%%       dimensional spaces that apply in cases where standard herding does not.
+%% \item It makes the known connections between herding, subgradient, and 
+%%       conditional gradient methods more clear. 
+%% \end{itemize}
+%% In particular, while it was known that herding was a form of conditional 
+%% gradient descent \cite{Bach:2012a} and hence (via the relationship in \citet{Bach:2012b}) also 
+%% a form of subgradient descent, our framework exposes all three algorithms 
+%% as special cases. A full table outlining the relationships for various choices of 
+%% $\psi$ and $R$ is given in Table~\ref{tab:connections}.
+%% 
+%% In summary, we introduce a novel gradient descent algorithm to investigate herding and its generalizations, and make the following contributions:
+%% \begin{itemize}
+%% \item In Section \ref{sec:algorithm} we introduce a pair of convex optimization algorithms, which we call primal and dual boosted mirror descent (\primal and \dual).
 We show how \primal and \dual relate to conditional gradient (\cgd) and mirror descent (\md).
-\item We then use \dual in Section \ref{sec:herding} to define two classes of generalized herding algorithms.
+%% \item We then use \dual in Section \ref{sec:herding} to define two classes of generalized herding algorithms.
 Connections with the ``learning'' interpretation of herding are made by showing how to obtain herding as the limit of an entropically normalized case of \dual.
-\item Section \ref{sec:proofs} presents convergence rates for \primal and \dual.
+%% \item Section \ref{sec:proofs} presents convergence rates for \primal and \dual.
 As special cases, we recover a $O(\log T/ T)$ convergence rate for classical herding and a $O(1/T)$ rate for weighted herding.
 Both of these results apply to the infinite dimensional case.
-\item In Section \ref{sec:lower-bounds} we prove a matching lower bound to the upper bound given in Section \ref{sec:proofs}, showing that any ``sparse'' herding-like method cannot converge faster than $O(1/T)$ in the the infinite dimensional case.% and that herding converges at the slower $O(\log T / T)$ rate.
-\item Finally, in section \ref{sec:chen-proofs} we generalize the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \bmd. 
-\end{itemize}
+%% \item In Section \ref{sec:lower-bounds} we prove a matching lower bound to the upper bound given in Section \ref{sec:proofs}, showing that any ``sparse'' herding-like method cannot converge faster than $O(1/T)$ in the the infinite dimensional case.% and that herding converges at the slower $O(\log T / T)$ rate.
+%% %\item Finally, in section \ref{sec:chen-proofs} we generalize the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \bmd. 
+%% \end{itemize}
 

### 2014-02-19 arXiv/introduction.tex
fleshed out intro a bit more

@@
 theory, machine learning, and optimization literatures, and we hope to bring 
-them all into one place, as well as to provide a unifying perspective. To start, 
-we define a generalized notion of sparsity that encompasses both sparsity and 
-low-rank conditions as special cases. Throughout, we will let $S$ denote a 
+them all into one place, as well as to provide a unifying perspective.
+
+By a ``sparsification result'', we mean, loosely, a method for approximating an 
+object (such as a matrix or a probability distribution) by a sparse linear 
+combination of a pre-defined set of atoms. For instance, \emph{graph 
+sparsification} consists of approximating the adjacency matrix of a graph by a 
+weighted matrix that contains only a small number of edges \cite{?}. 
+\emph{Low-rank matrix approximation} consists of approximating a matrix by a 
+linear combination of a small number of rank-$1$ matrices \cite{?}. 
+\emph{Pseudo-sampling} (also known as \emph{herding} \cite{?}) consists of 
+approximating the moments of a probability distribution using a weighted 
+combination of a small number of ``pseudo-samples'', e.g. 
+$\bE_{x \sim \pi}[\phi(x)] \approx \sum_{i=1}^k w_i \phi(x_i)$.
+
+Other sparsification results include:  approximating the 
+Nash equilibrium of a game by a weighted mixture of a small number of pure 
+strategies \cite{?}; finding low-complexity circuits that are indistinguishable 
+from a high-complexity circuit relative to a given oracle \cite{?}; and 
+approximating the maximum flow by a small number of resistive flows \cite{?}.
+
+In this paper, we will define a generalized notion of sparsity that captures 
+all of the above examples, and provide general sparsification algorithms that 
+can then be applied to each of these settings as special cases. However, in some 
+cases these general results are not enough to yield the best known bounds; we will 
+state when this is the case and cite the best bounds of which we are currently 
+aware.
+
+In writing this tutorial we are indebted to several authors who have already 
+published unifying results between different notions of sparsity.
+
+\section{A Generalized Sparsity Result}
+Throughout, we will let $S$ denote a 
 compact subset of $\bR^n$ and let $\Conv(S)$ denote the convex hull of $S$ 

### 2014-02-19 arXiv/introduction.tex
minor tweaks to abstract and intro

--- 2014-02-18 arXiv/introduction.tex
@@
 
-The purpose of this paper is to serve as a tutorial on sparsification. While 
-most of the results presented are known, they are often distributed across the 
-theory, machine learning, and optimization literatures, and we hope to bring 
-them all into one place, as well as to provide a unifying perspective. To start, 
+The purpose of this paper is to serve as a tutorial on sparsification as it arises 
+in machine learning, computer science theory, and optimization. While most of the 
+results presented are known, they are often distributed across the these 
+literatures. Our objective is to bring them all into one place, 
+as well as to provide a unifying perspective. To start, 
 we define a generalized notion of sparsity that encompasses both sparsity and 

### 2014-02-19 arXiv/paper.tex
minor tweaks to abstract and intro

@@
 
-\keywords{boosting, optimization, mirror descent, Frank-Wolfe, conjugate gradient descent, herding, graph sparsification}
+\keywords{boosting, optimization, mirror descent, Frank-Wolfe, conditional gradient descent, herding, graph sparsification}
 
@@
 a strongly convex optimization problem whose dual problem is linear, we can obtain 
-sparse approximate solutions, and use this to re-derive many of the known 
-sparsification results.
+sparse approximate solutions, and use this to re-derive (and in some cases 
+extend) many of the known sparsification results.
 % Much has been learned about herding since it was first proposed.

### 2014-02-19 arXiv/introduction.tex

--- 2014-02-19 arXiv/introduction.tex
@@
 
-The purpose of this paper is to serve as a tutorial on sparsification. While 
-most of the results presented are known, they are often distributed across the 
-theory, machine learning, and optimization literatures, and we hope to bring 
-them all into one place, as well as to provide a unifying perspective.
+The purpose of this paper is to serve as a tutorial on sparsification as it arises 
+in machine learning, computer science theory, and optimization. While most of the 
+results presented are known, they are often distributed across the these 
+literatures. Our objective is to bring them all into one place, 
+as well as to provide a unifying perspective.
 

### 2014-02-26 arXiv/introduction.tex
done with intro

--- 2014-02-19 arXiv/introduction.tex
@@
 \label{ex:graph}
-TODO: graph sparsification example
+Given an undirected graph $G$ with vertex set $\{1,\ldots,n\}$, let $E$ 
+denote its edge set, and let $m = |E|$. For a given 
+edge $e = (i,j)$, define $x_e \in \bR^n$ to be the vector with 
+$1$ in the $i$th coordinate, $-1$ in the $j$th coordinate, and 
+$0$s everywhere else. Let $X_e \eqdef mx_ex_e^{\top}$. 
+Define $S$ to be $\{X_e \mid e \in E\}$.
+
+The \emph{Laplacian} $L$ of $G$ is defined to be the matrix $L$ 
+with $L_{ij} = -1$ if $(i,j)$ is an edge, and $L_{ii}$ equal to 
+the degree of vertex $i$. It can be written as $L = \sum_{e \in E} x_ex_e^{\top}$. 
+In particular, $L \in \Conv(S)$.
+
+A $(S,k)$-sparse point can be written as a convex combination 
+\[ \sum_{r=1}^k w_rX_{e_r} = \sum_{r=1}^k (m\cdot w_r)x_{e_r}x_{e_r}^{\top} \]
+for some $k$ edges $e_1,\ldots,e_k$. Any such matrix is the Laplacian 
+of a \emph{weighted} graph (with edge weight $m\cdot w_r$ on edge $e_r$) 
+with $k$ edges and total edge weight $m$.
 \end{example}
 
-%% Many algorithms, such as SMC \citep{smc} and MCMC \citep{mcmc}, approximate a 
-%% distribution with samples. Most such algorithms require access to a full posterior and 
-%% may be prone to becoming stuck in local modes. In contrast, the herding algorithm introduced in 
-%% \citet{Welling:2009a} generates samples whose moments are guaranteed to converge 
-%% to any given collection of feasible moments. Moreover, \emph{only} the moments 
-%% are required; a full posterior is not needed.
-%% 
-%% Since its introduction, herding has been shown to have many desirable properties.
-%% For instance, \citet{Chen:2010a} showed that the moments converge at a rate of $O(1/T^{2})$ for 
-%% $T$ samples, which is even faster than the $O(1/T)$ rate that would be obtained 
-%% by sampling independently from the exact posterior. Furthermore, \citet{Bach:2012a} show that herding is 
-%% equivalent to conditional gradient, and \citet{Huszar:2012} show that 
-%% herding is equivalent to Bayesian quadrature. Herding was originally conceived of as a method 
-%% for inference in Markov random fields, and this line of work has been further explored in 
-%% \citet{Welling:2009a}, \citet{Gelfand:2010}, and \citet{Bornn:2013}.
-%% 
-%% In order to define herding, let $\sX$ be an observation space, $\phi : \sX \to \sF$ a 
-%% feature map to an inner product feature space, and $\bar\phi = \bE_{x \sim p}[\phi(x)]$ the empirical 
-%% moments of some distribution. Then the herding algorithm is the following simple procedure to 
-%% generate pseudosamples from $\sX$:
-%% \[
-%% x_{t} &\in \argmin_{x \in \sX} \ip{\phi(x)}{y_{t-1}} \\
-%% y_{t} &= y_{t-1}  + \bar \phi - \phi(x_{t}).
-%% \]
-%% In this paper we generalize the herding algorithm to minimize $\psi(x) + R^{*}(x)$ via the following set of updates, which 
-%% require only the convexity of $\psi$ and $R$:
-%% \[
-%% x_t &\in \argmin_{x} \psi(x) + \ip{x}{y_{t-1}} \\
-%% y_{t} &\in \argmax_{y} \ip{\bar{x}_t}{y} - R(y),
-%% \]
-%% where $\bar{x}_t$ is a certain weighted average of $x_1,\ldots,x_t$. Our 
-%% algorithm is described in more detail in Section~\ref{sec:algorithm}. This 
-%% generalization has several advantages:
-%% \begin{itemize}
-%% \item It allows us to cast herding in the framework of mirror descent \cite{Beck:2003}, 
-%%       which lets us better understand its convergence properties.
-%% \item It allows us to obtain generalizations of herding for infinite 
-%%       dimensional spaces that apply in cases where standard herding does not.
-%% \item It makes the known connections between herding, subgradient, and 
-%%       conditional gradient methods more clear. 
-%% \end{itemize}
-%% In particular, while it was known that herding was a form of conditional 
-%% gradient descent \cite{Bach:2012a} and hence (via the relationship in \citet{Bach:2012b}) also 
-%% a form of subgradient descent, our framework exposes all three algorithms 
-%% as special cases. A full table outlining the relationships for various choices of 
-%% $\psi$ and $R$ is given in Table~\ref{tab:connections}.
-%% 
-%% In summary, we introduce a novel gradient descent algorithm to investigate herding and its generalizations, and make the following contributions:
-%% \begin{itemize}
-%% \item In Section \ref{sec:algorithm} we introduce a pair of convex optimization algorithms, which we call primal and dual boosted mirror descent (\primal and \dual).
-We show how \primal and \dual relate to conditional gradient (\cgd) and mirror descent (\md).
-%% \item We then use \dual in Section \ref{sec:herding} to define two classes of generalized herding algorithms.
-Connections with the ``learning'' interpretation of herding are made by showing how to obtain herding as the limit of an entropically normalized case of \dual.
-%% \item Section \ref{sec:proofs} presents convergence rates for \primal and \dual.
-As special cases, we recover a $O(\log T/ T)$ convergence rate for classical herding and a $O(1/T)$ rate for weighted herding.
-Both of these results apply to the infinite dimensional case.
-%% \item In Section \ref{sec:lower-bounds} we prove a matching lower bound to the upper bound given in Section \ref{sec:proofs}, showing that any ``sparse'' herding-like method cannot converge faster than $O(1/T)$ in the the infinite dimensional case.% and that herding converges at the slower $O(\log T / T)$ rate.
-%% %\item Finally, in section \ref{sec:chen-proofs} we generalize the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \bmd. 
-%% \end{itemize}
+With these examples in mind, we are now ready to state our main result:
+\begin{theorem}
+\label{thm:main}
+Given vector spaces $V$ and $W$, let $f : V \times W \to \bR$ be a function 
+such that 
+\begin{enumerate}
+\item $f(\cdot,w)$ is linear for each fixed $w$
+\item $f(v, \cdot)$ is strongly convex with respect to a norm $\|\cdot\|$ 
+      for each fixed $v$.
+\end{enumerate}
+Then, for any compact set $S \subset V$, and any point $x \in \Conv(S)$, 
+there exists a $(S,k)$-sparse point $y$ such that 
+\[ \max_{w \in W} f(y,w) \leq \max_{w \in W} f(x,w) + \frac{2D^2}{k+1}, \]
+where $D = \sup_{s, s' \in S} \|s-s'\|_*$ and $\|\cdot\|_*$ is the dual norm 
+of $\|\cdot\|$.
+\end{theorem}
+The proof of Theorem~\ref{thm:main} is given in Section~\ref{todo}. For now, 
+to give a flavor for the relevance of Theorem~\ref{thm:main}, we show how 
+to obtain a sparsification result for Example~\ref{ex:marginal-polytope}.
+\begin{proposition}
+\label{prop:pseudosamples}
+Let $\phi : \sX \to \bR^d$ and let $\sM = \Conv(\{\phi(x) \mid x \in \sX\})$ 
+be the marginal polytope as defined in Example~\ref{ex:marginal-polytope}. 
+Let $D$ be the diameter of $\sM$ in the $l^2$ norm. Then, for any 
+$\mu \in \sM$, there exists points $x_1,\ldots,x_k$ and weights 
+$w_1,\ldots,w_k$ such that
+\[ \left\|\mu - \sum_{i=1}^k w_i \phi(x_i)\right\|_2 \leq \sqrt{\frac{4D^2}{k+1}} \]
+\end{proposition}
+\begin{proof}
+Let $V = W = \bR^d$, and apply Theorem~\ref{thm:main} to the function 
+$f(v,w) = (v-\mu)^{\top}w - \frac{1}{2}\|w\|_2^2$. The function $f$ is 
+linear in $v$ for each $w$, and is strongly convex in $w$ with respect to the 
+$l^2$ norm for each $v$. Also note that 
+$\sup_{w \in W} f(v,w) = \frac{1}{2}\|v-\mu\|_2^2$. Since the dual of the 
+$l^2$ norm is the $l^2$ norm, Theorem~\ref{thm:main} says that there exists 
+a $(S,k)$-sparse point $y$ such that
+\[ \frac{1}{2}\|\mu - y\|_2^2 \leq \frac{1}{2}\|\mu - \mu\|_2^2 + \frac{2D^2}{k+1}. \]
+Since $\|\mu-\mu\|_2 = 0$ and any $(S,k)$-sparse point $y$ can be written 
+as $\sum_{i=1}^k w_i \phi(x_i)$, we have the desired result.
+\end{proof}
+Proposition~\ref{prop:pseudosamples} says that, given any distribution 
+$\pi$ over $\sX$, we can find $O(1/\epsilon^2)$ pseudosamples that approximate 
+the moments of $\pi$ with accuracy $\epsilon$ in the $l^2$ norm. We note 
+that in some instances this can be improved to $O(1/\epsilon)$, see 
+for instance \cite{chen}.
+
+We will establish many more such results in the coming sections.
 

### 2014-02-28 arXiv/introduction.tex
edited intro and section 2; added lower bound result to section 2

@@
 in machine learning, computer science theory, and optimization. While most of the 
-results presented are known, they are often distributed across the these 
+results presented here are known, they are often distributed across these 
 literatures. Our objective is to bring them all into one place, 
@@
 
-By a ``sparsification result'', we mean, loosely, a method for approximating an 
+By a ``sparsification result'', we mean, loosely speaking, a method for approximating an 
 object (such as a matrix or a probability distribution) by a sparse linear 
@@
 sparsification} consists of approximating the adjacency matrix of a graph by a 
-weighted matrix that contains only a small number of edges \cite{?}. 
+weighted matrix that contains only a small number of edges \citep{?}. 
 \emph{Low-rank matrix approximation} consists of approximating a matrix by a 
-linear combination of a small number of rank-$1$ matrices \cite{?}. 
-\emph{Pseudo-sampling} (also known as \emph{herding} \cite{?}) consists of 
+linear combination of a small number of rank-$1$ matrices \citep{?}. 
+\emph{Pseudo-sampling} (also known as \emph{herding} \citep{?}) consists of 
 approximating the moments of a probability distribution using a weighted 
@@
 Nash equilibrium of a game by a weighted mixture of a small number of pure 
-strategies \cite{?}; finding low-complexity circuits that are indistinguishable 
-from a high-complexity circuit relative to a given oracle \cite{?}; and 
-approximating the maximum flow by a small number of resistive flows \cite{?}.
+strategies \citep{?}; finding low-complexity circuits that are indistinguishable 
+from a high-complexity circuit relative to a given oracle \citep{?}; and 
+approximating the maximum flow by a small number of resistive flows \citep{?}.
 
-In this paper, we will define a generalized notion of sparsity that captures 
-all of the above examples, and provide general sparsification algorithms that 
+\NA{Sparsification of 
+mathematical objects such as graphs, circuits, and probability measures is important
+because sparse approximations lead to objects that are much ``simpler'' than the 
+originals while remaining ``close'' to them. These simpler sparse approximations can
+then be manipulated more efficiently than the originals, typically because
+each atom can be operated on efficiently, which leads to, for example, 
+more efficient algorithms.}\fTBD{JHH: added some motivation.}
+
+In this paper we will define a generalized notion of sparsity that captures 
+all of the above examples and provide general sparsification algorithms that 
 can then be applied to each of these settings as special cases. However, in some 
@@
 In writing this tutorial we are indebted to several authors who have already 
-published unifying results between different notions of sparsity.
+published unifying results between different notions of sparsity \citep{?}.
 
@@
 Throughout, we will let $S$ denote a 
-compact subset of $\bR^n$ and let $\Conv(S)$ denote the convex hull of $S$ 
+compact subset of $\reals^n$ and let $\Conv(S)$ denote the convex hull of $S$ 
 (which will always be a closed set due to compactness of $S$). Two particularly 
@@
 \begin{definition}
-Given a set $S \subseteq \bR^n$, a \emph{$(S, k)$-sparse} point is a point 
+Given a set $S \subseteq \reals^n$, an \emph{$(S, k)$-sparse} point is a point 
 $x \in \Conv(S)$ that is a convex combination of at most $k$ extreme 
@@
 \label{ex:marginal-polytope}
-Given a space $\sX$ and a function $\phi : \sX \to \bR^d$, let 
+Given a space $\sX$ and a function $\phi : \sX \to \reals^d$, let 
 $S = \{\phi(x) \mid x \in \sX\}$. Then $\Conv(S)$ is called the 
@@
 $\phi(x)$ with respect to an arbitrary probability distribution 
-on $\sX$. A $(S,k)$-sparse point in $\Conv(S)$ is a point that 
+on $\sX$. An $(S,k)$-sparse point in $\Conv(S)$ is a point that 
 can be written as a weighted average $\sum_{i=1}^k w_i \phi(x_i)$. 
@@
 denote its edge set, and let $m = |E|$. For a given 
-edge $e = (i,j)$, define $x_e \in \bR^n$ to be the vector with 
+edge $e = (i,j)$, define $x_e \in \reals^n$ to be the vector with 
 $1$ in the $i$th coordinate, $-1$ in the $j$th coordinate, and 
-$0$s everywhere else. Let $X_e \eqdef mx_ex_e^{\top}$. 
-Define $S$ to be $\{X_e \mid e \in E\}$.
-
-The \emph{Laplacian} $L$ of $G$ is defined to be the matrix $L$ 
+zeros everywhere else. Let $X_e \eqdef mx_{e} x_{e}^{\top}$ and 
+define $S$ to be $\{X_e \mid e \in E\}$.
+The \emph{Laplacian} $L$ of graph $G$ is defined to be the matrix $L$ 
 with $L_{ij} = -1$ if $(i,j)$ is an edge, and $L_{ii}$ equal to 
-the degree of vertex $i$. It can be written as $L = \sum_{e \in E} x_ex_e^{\top}$. 
-In particular, $L \in \Conv(S)$.
-
-A $(S,k)$-sparse point can be written as a convex combination 
+the degree of vertex $i$. Equivalently, it can be written as 
+$L = \sum_{e \in E} x_{e} x_{e}^{\top}$ and note in particular $L \in \Conv(S)$.
+An $(S,k)$-sparse point can be written as a convex combination 
 \[ \sum_{r=1}^k w_rX_{e_r} = \sum_{r=1}^k (m\cdot w_r)x_{e_r}x_{e_r}^{\top} \]
@@
 \label{thm:main}
-Given vector spaces $V$ and $W$, let $f : V \times W \to \bR$ be a function 
+Given vector spaces $V$ and $W$\fTBD{JHH: hmm...you've switched from $\reals^{d}$ to vector spaces}, let $f : V \times W \to \reals$ be a function 
 such that 
@@
 \item $f(\cdot,w)$ is linear for each fixed $w$
-\item $f(v, \cdot)$ is strongly convex with respect to a norm $\|\cdot\|$ 
+\item $f(v, \cdot)$ is strongly convex\fTBD{JHH: we haven't defined strongly convex yet} with respect to a norm $\|\cdot\|$ 
       for each fixed $v$.
@@
 Then, for any compact set $S \subset V$, and any point $x \in \Conv(S)$, 
-there exists a $(S,k)$-sparse point $y$ such that 
+there exists an $(S,k)$-sparse point\fTBD{JHH: note that it should be {\bf an} $(S,k)$-sparse point} $y$ such that 
 \[ \max_{w \in W} f(y,w) \leq \max_{w \in W} f(x,w) + \frac{2D^2}{k+1}, \]
@@
 \end{theorem}
-The proof of Theorem~\ref{thm:main} is given in Section~\ref{todo}. For now, 
+This result is tight in the following sense:\fTBD{JHH: added complementary lower bound result}
+\begin{theorem}
+\label{thm:lower-bound}
+There exists a choice of $f$ satisfying the conditions of Theorem~\ref{thm:main}, 
+a compact set $S \subset V$, and a point $x \in \Conv(S)$ such that for any 
+$(S,k)$-sparse point $y$, 
+\[ \max_{w \in W} f(y,w) \ge \max_{w \in W} f(x,w) + \Theta(1/k). \]
+\end{theorem}
+The proofs of Theorems~\ref{thm:main} and \ref{thm:lower-bound} are given in Section~\ref{TODO}. For now, 
 to give a flavor for the relevance of Theorem~\ref{thm:main}, we show how 
@@
 \label{prop:pseudosamples}
-Let $\phi : \sX \to \bR^d$ and let $\sM = \Conv(\{\phi(x) \mid x \in \sX\})$ 
+Let $\phi : \sX \to \reals^d$ and let $\sM = \Conv(\{\phi(x) \mid x \in \sX\})$ 
 be the marginal polytope as defined in Example~\ref{ex:marginal-polytope}. 
@@
 \begin{proof}
-Let $V = W = \bR^d$, and apply Theorem~\ref{thm:main} to the function 
+Let $V = W = \reals^d$, and apply Theorem~\ref{thm:main} to the function 
 $f(v,w) = (v-\mu)^{\top}w - \frac{1}{2}\|w\|_2^2$. The function $f$ is 
@@
 \end{proof}
+An analogous result can be obtained by instead choosing $\phi : \mcX \to \mcH$, where $(\mcH, \|\cdot\|_{\mcH})$ is a reproducing kernel Hilbert space and using $\|\cdot\|_{\mcH}$ instead of $\|\cdot\|_{2}$.
 Proposition~\ref{prop:pseudosamples} says that, given any distribution 
 $\pi$ over $\sX$, we can find $O(1/\epsilon^2)$ pseudosamples that approximate 
-the moments of $\pi$ with accuracy $\epsilon$ in the $l^2$ norm. We note 
-that in some instances this can be improved to $O(1/\epsilon)$, see 
-for instance \cite{chen}.
+the moments of $\pi$ with accuracy $\epsilon$ in the $l^2$ (or Hilbert) norm.\fTBD{JHH: added Hilbert space discussion} As
+we shall see in Section \ref{sec:algorithm}, in some instances this 
+can be improved to $O(1/\epsilon)$.\fTBD{JHH: will give Chen extension, so no need
+to defer to Chen paper here}
+%We note that in some instances this can be improved to $O(1/\epsilon)$, see 
+%for instance \cite{chen}.
 
@@
 
+\NA{Proposed sections:
+\begin{enumerate}
+\item Intro
+\item Sparsity result and lower bound
+\item BMD 
+\item Proofs of sparsity result and lower bound
+\item Application: Herding, including Chen extension
+\item Application: Graph sparsification, low-rank matrix approximation, max flow
+\item Application: Circuits and boosting
+\item Application: Solving games 
+\item Concluding thoughts 
+\end{enumerate}
+\qquad(A) Appendix: BMD proofs}
+
 \end{document}

### 2014-02-28 arXiv/paper.tex
edited intro and section 2; added lower bound result to section 2

@@
 \usepackage{chngpage}
-\usepackage{fullpage}
+%\usepackage{fullpage}
 \input{latex-defs.tex}
@@
 
-\thanks{The authors contributed equally to this work.}
+\thanks{Both authors contributed equally to this work.}
 
@@
 focused on the idea of sparse or low-rank ``approximations'' to some object: 
-for instance, graph sparsification \cite{?}, low-rank SDP approximation \cite{?}, 
-and pseudosampling \cite{?}. In this paper we show that, whenever presented with 
+for instance, graph sparsification \citep{deCarliSilva:2011}, low-rank SDP approximation \citep{Arora:2007}, 
+and pseudosampling \citep{Welling:2009a,Chen:2010a}. In this paper we show that, when presented with 
 a strongly convex optimization problem whose dual problem is linear, we can obtain 
 sparse approximate solutions, and use this to re-derive (and in some cases 
-extend) many of the known sparsification results.
+extend) many of the known sparsification results. Our main technical tool
+is a generalization of the Frank-Wolfe (a.k.a.\ conditional gradient 
+descent) algorithm. 
 % Much has been learned about herding since it was first proposed.

### 2014-03-06 arXiv/introduction.tex
changed proposed sections
added some of section 3, changed theorem 2.6

@@
 \label{thm:main}
-Given vector spaces $V$ and $W$\fTBD{JHH: hmm...you've switched from $\reals^{d}$ to vector spaces}, let $f : V \times W \to \reals$ be a function 
+Given vector spaces $\V$ and $\W$\fTBD{JHH: hmm...you've switched from $\reals^{d}$ to vector spaces}, let $h : \W \to \bR$ be a convex function and let $f : \V \times \W \to \reals$ be a function 
 such that
@@
 \item $f(\cdot,w)$ is linear for each fixed $w$
-\item $f(v, \cdot)$ is strongly convex\fTBD{JHH: we haven't defined strongly convex yet} with respect to a norm $\|\cdot\|$ 
-      for each fixed $v$.
+\item $f(v, \cdot) - h(\cdot)$ is convex for each fixed $v$.
 \end{enumerate}
-Then, for any compact set $S \subset V$, and any point $x \in \Conv(S)$, 
+Suppose further that $h$ is strongly convex with respect to a norm $\|\cdot\|$. 
+\fTBD{JHH: we haven't defined strongly convex yet}
+Then, for any compact set $S \subset \V$, and any point $x \in \Conv(S)$, 
 there exists an $(S,k)$-sparse point\fTBD{JHH: note that it should be {\bf an} $(S,k)$-sparse point} $y$ such that 
-\[ \max_{w \in W} f(y,w) \leq \max_{w \in W} f(x,w) + \frac{2D^2}{k+1}, \]
+\[ \max_{w \in \W} f(y,w) \leq \max_{w \in \W} f(x,w) + \frac{2D^2}{k+1}, \]
 where $D = \sup_{s, s' \in S} \|s-s'\|_*$ and $\|\cdot\|_*$ is the dual norm 
@@
 There exists a choice of $f$ satisfying the conditions of Theorem~\ref{thm:main}, 
-a compact set $S \subset V$, and a point $x \in \Conv(S)$ such that for any 
+a compact set $S \subset \V$, and a point $x \in \Conv(S)$ such that for any 
 $(S,k)$-sparse point $y$, 
-\[ \max_{w \in W} f(y,w) \ge \max_{w \in W} f(x,w) + \Theta(1/k). \]
+\[ \max_{w \in \W} f(y,w) \ge \max_{w \in \W} f(x,w) + \Theta(1/k). \]
 \end{theorem}
@@
 \begin{proof}
-Let $V = W = \reals^d$, and apply Theorem~\ref{thm:main} to the function 
+Let $\V = \W = \reals^d$, and apply Theorem~\ref{thm:main} to the function 
 $f(v,w) = (v-\mu)^{\top}w - \frac{1}{2}\|w\|_2^2$. The function $f$ is 
@@
 \item Sparsity result and lower bound
-\item BMD 
-\item Proofs of sparsity result and lower bound
+\item BMD and proofs of sparsity result and lower bound
+\item Application: Low-rank matrix approximation, approximation of SDP matrix sums 
+\item Application: Graph sparsification, max flow
+\item Application: Solving games 
 \item Application: Herding, including Chen extension
-\item Application: Graph sparsification, low-rank matrix approximation, max flow
 \item Application: Circuits and boosting
-\item Application: Solving games 
 \item Concluding thoughts 

### 2014-03-07 arXiv/introduction.tex
minor changes

--- 2014-02-28 arXiv/introduction.tex
@@
 then be manipulated more efficiently than the originals, typically because
-each atom can be operated on efficiently, which leads to, for example, 
+each atom\fTBD{JHH: ``atom'' not defined yet} can be operated on efficiently, which leads to, for example, 
 more efficient algorithms.}\fTBD{JHH: added some motivation.}
@@
 \end{theorem}
-This result is tight in the following sense:\fTBD{JHH: added complementary lower bound result}
+This result is tight in the following sense:\fTBD{JS: can we generalize this to arbitrary norms? or some norms other than $\ell_{2}$?}\fTBD{JHH: rewrite to explicitly give $f$ and $S$}
 \begin{theorem}
@@
 \begin{enumerate}
-\item Intro
-\item Sparsity result and lower bound
-\item BMD 
-\item Proofs of sparsity result and lower bound
-\item Application: Herding, including Chen extension
-\item Application: Graph sparsification, low-rank matrix approximation, max flow
-\item Application: Circuits and boosting
+\item (B) Intro
+\item (B) Sparsity result and lower bound
+\item (JS) BMD and proofs of sparsity result and lower bound
+\item (JH) Application: Low-rank matrix approximation, approximation of SDP matrix sums 
+\item (JH) Application: Graph sparsification, max flow
 \item Application: Solving games 
+\item (JH) Application: Herding, including Chen extension
+\item (JS) Application: Circuits and boosting
 \item Concluding thoughts 

### 2014-03-07 arXiv/introduction.tex

--- 2014-03-06 arXiv/introduction.tex
@@
 \begin{enumerate}
-\item Intro
-\item Sparsity result and lower bound
-\item BMD and proofs of sparsity result and lower bound
-\item Application: Low-rank matrix approximation, approximation of SDP matrix sums 
-\item Application: Graph sparsification, max flow
+\item (B) Intro
+\item (B) Sparsity result and lower bound
+\item (JS) BMD and proofs of sparsity result and lower bound
+\item (JH) Application: Low-rank matrix approximation, approximation of SDP matrix sums 
+\item (JH) Application: Graph sparsification, max flow
 \item Application: Solving games 
-\item Application: Herding, including Chen extension
-\item Application: Circuits and boosting
+\item (JH) Application: Herding, including Chen extension
+\item (JS) Application: Circuits and boosting
 \item Concluding thoughts 

### 2014-03-07 arXiv/sparsity-result.tex

\documentclass[paper.tex]{subfiles}
\begin{document}

\section{Composite Mirror Descent and Sparsity}
\label{sec:algorithm}

Our sparsity result is algorithmic in nature, and relies on an algorithm called 
\emph{composite mirror descent} \cite{comid}. In composite mirror descent, we 
are given a sequence of functions $f_t(w) = g_t(w) + h_t(w)$, and the goal is to 
minimize
\[ \sum_{t=1}^T \alpha_t f_t(w), \]
for some known non-negative weights $\alpha_t$. We assume throughout that $f_t$, 
$g_t$, and $h_t$ are all convex. The composite mirror descent algorithm also 
depends on an auxiliary convex function $\psi(w)$:
\begin{algorithm}
\caption{The composite mirror descent algorithm.}
\label{alg:comid}
\begin{algorithmic}
\STATE {\bfseries Composite Mirror Descent}
\STATE Initialize $\theta_0 \gets 0$, $\psi_0 \gets \psi$
\FOR{$t = 1$ {\bfseries to} $T$}
  \STATE $\psi_t \gets \psi_{t-1} + \alpha_t h_t$
  \STATE $w_t \gets \argmin_{w \in \W} \psi_{t}(w) + \theta_{t-1}^{\top} w$
  \STATE $\theta_t \gets \theta_t + \alpha_t \nabla g_t(w_t)$
\ENDFOR
\STATE Output $w^* = \frac{\sum_{t=1}^T \alpha_t w_t}{\sum_{t=1}^T \alpha_t}$
\end{algorithmic}
\end{algorithm}

The implicit assumption of this algorithm is that 
$\argmin_{w \in \W} \psi_t(w) + \theta_{t-1}^{\top}w$ can be computed efficiently. 
Among other things, this will typically require that $h_t$ has some known parametric 
form (as opposed to just being provided as a black box).

% give example
As an example, suppose that $\psi(w) = h_t(w) = \frac{1}{2}\|w\|_2^2$. Then 
$\psi_t(w) = \frac{t+1}{2}\|w\|_2^2$, and 
$\argmin_{w \in \W} \psi_t(w) + \theta_{t-1}^{\top} w = -\frac{1}{t+1}\theta_{t-1}$.
There are also many other examples of functions for which the $\argmin$ can be 
computed efficiently; this is closely related to the theory of Fenchel duality, 
which we present in Section~\ref{sec:fenchel} together with additional examples.

% define strong convexity
The convergence rate of Algorithm~\ref{alg:comid} depends on the degree of 
convexity of the functions $\psi$ and $h_t$. We can measure this convexity 
using any norm $\|\cdot\|$ on $\W$.
\begin{definition}
A differentiable function $f$ is said to be \emph{$\rho$-strongly convex} with 
respect to a norm $\|\cdot\|$ if
\[ f(w) - f(w') - \nabla f(w')^{\top}(w - w') \geq \frac{\rho}{2}\|w-w'\|^2 \]
for all $w, w' \in \W$.
\end{definition}
When $\rho = 1$ we sometimes just say that $f$ is strongly convex with respect to $\|\cdot\|$.
If $\|\cdot\|$ is the $l^2$ norm and $f$ is twice-differentiable, then $f$ is $\rho$-strongly 
convex if and only if $\nabla^2 f(w) \succeq \rho I$ for all $w \in \W$. In particular, 
$f(w) = \frac{1}{2}\|w - w_0\|_2^2$ is $1$-strongly convex with respsect to $\|\cdot\|_2$ for 
any $w_0$, although this is not true for general norms (for instance, 
$\frac{1}{2}\|w-w_0\|_1^2$ is not strongly convex with respect to $\|\cdot\|_1$ for any value 
of $\rho$). See Section~\ref{sec:fenchel} for more details.

One additional notion we need is that of a dual norm.
\begin{definition}
Given a norm $\|\cdot\|$, the \emph{dual norm} $\|\cdot\|_*$ is defined as 
\[ \|\theta\|_* \eqdef \sup_{\|w\| \leq 1} \theta^{\top}w. \]
\end{definition}
In other words, the dual norm is exactly constructed that H\"{o}lder's 
inequality $\theta^{\top}w \leq \|\theta\|_* \|w\|$ holds. In particular, 
the dual norm of $\|\cdot\|_2$ is $\|\cdot\|_2$, and the dual norm of 
$\|\cdot\|_p$ is $\|\cdot\|_{\frac{p}{p-1}}$.

% provide convergence theorem
We are now ready to state the convergence rate of Algorithm~\ref{alg:comid}.
\begin{theorem}
\label{thm:comid}
Let $\|\cdot\|$ be any norm on $\W$. Suppose that $\psi$ is $\rho_0$-strongly convex 
with respect to $\|\cdot\|$, and that $h_t$ is $\rho_t$-strongly convex with respect 
to $\|\cdot\|$. Suppose that $\|\nabla g_t\|_* \leq r$ for all $t$. Let 
$A_t = \sum_{s=1}^t \alpha_s$ and $R_t = \rho_0 + \sum_{s=1}^t \alpha_s \rho_s$. 
Then, defining $f(w) \eqdef \frac{1}{A_T} \sum_{t=1}^{T} \alpha_t f_t(w)$, we have
\[ f(w^*) \leq \inf_{w} f(w) + \frac{r^2}{2A_T} \sum_{t=1}^T \frac{\alpha_t^2}{R_t}. \]
\end{theorem}
The proof is given in Appendix~\ref{apx:comid}.

To unpack Theorem~\ref{thm:comid}, let us consider several special cases.
\begin{corollary}
Suppose that $\psi \equiv 0$ and $h_t$ is $\rho$-strongly convex with respect to 
$\|\cdot\|$ for each $t$ for some fixed $\rho$. Then, for $\alpha_t = 1$, we have
\[ f(w^*) \leq \inf_w f(w) + \frac{r^2 \log(T+1)}{2\rho T}, \]
and for $\alpha_t = t$ we have
\[ f(w^*) \leq \inf_w f(w) + \frac{2r^2}{\rho (T+1)}. \]
\end{corollary}
\begin{proof}
For $\alpha_t = 1$, we have $A_T = T$ and $R_t = \rho t$. Then Theorem~\ref{thm:comid} 
yields a convergence rate of
\begin{align*}
f(w^*) &\leq \inf_w f(w) + \frac{r^2}{2T} \sum_{t=1}^T \frac{1}{\rho t} \\
 &\leq \inf_w f(w) + \frac{r^2 \log(T+1)}{2\rho T}.
\end{align*}
For $\alpha_t = t$, we have $A_T = \frac{T(T+1)}{2}$ and $R_t = \rho \frac{t(t+1)}{2}$. 
Then Theorem~\ref{thm:comid} yields a convergence rate of
\begin{align*}
f(w^*) &\leq \inf_w f(w) + \frac{r^2}{T(T+1)} \sum_{t=1}^T \frac{2t^2}{\rho t(t+1)} \\
 &\leq \inf_w f(w) + \frac{r^2}{T(T+1)} \sum_{t=1}^T \frac{2}{\rho} \\
 &= \inf_w f(w) + \frac{2r^2}{\rho (T+1)}.
\end{align*}
\end{proof}

\begin{corollary}
Suppose that $\psi$ is $\rho$-strongly convex with respect to $\|\cdot\|$, but make 
no assumptions on $h_t$ other than convexity. Then, for $\alpha_t = 1$, we have
\[ f(w^*) \leq \inf_w f(w) + \frac{r^2}{2\rho}. \]
\end{corollary}
\begin{proof}
We have $A_T = T$ and $R_t = \rho$. Then Theorem~\ref{thm:comid} yields
\begin{align*}
f(w^*) &\leq \inf_w f(w) + \frac{r^2}{2T} \sum_{t=1}^T \frac{1}{\rho} \\
 &= \inf_w f(w) + \frac{r^2}{2\rho}.
\end{align*}
\end{proof}

% give boosted version (cite Bach, Shai)
\subsection{Composite Mirror Descent and Sparsity}
\label{sec:boosted}
What relationship does Composite Mirror Descent have to sparsification? The answer lies
in a technique called \emph{boosting} which will allow us to cleverly pick the functions 
$f_t(w)$. Boosting is an ubiquitous idea spanning multiple literatures; % TODO provide citations
in our case, we suppose that we have fixed functions $h : \W \to \bR$ and 
$f : \V \times \W \to \bR$ such that:
\begin{enumerate}
\item $f(\cdot, w)$ is linear in $v$ for each $w \in \W$.
\item $f(v, \cdot) - h(\cdot)$ is convex in $w$ for each $v \in \V$.
\end{enumerate}
Let $v_t = \argmax_{v \in V} f(v, w_t)$. Then we will pick 
$h_t(w) = h(w)$ and $g_t(w) = f(v_t, w) - h(w)$. Note that, 
even though $g_t$ depends on $v_t$ and hence $w_t$, we can still 
run Algorithm~\ref{alg:comid}, since we do not need access to $g_t$ 
until \emph{after} selecting $w_t$.

This leads to the following primal-dual algorithm:
\begin{algorithm}
\caption{Primal-dual Composite Mirror Descent}
\label{alg:boosted-comid}
\begin{algorithmic}
\STATE Initialize $\theta_0 \gets 0$, $\psi_0 \gets \psi$
\FOR{$t = 1$ {\bfseries to} $T$}
  \STATE $\psi_t \gets \psi_{t-1} + \alpha_t h$
  \STATE $w_t \gets \argmin_{w \in \W} \psi_t(w) + \theta_{t-1}^{\top} w$
  \STATE $v_t \gets \argmax_{v \in \V} f(v, w_t)$
  \STATE $\theta_t \gets \theta_t + \alpha_t \left(\frac{\partial f(v_t, w_t)}{\partial w_t} - \frac{\partial h(w_t)}{\partial w_t}\right)$
\ENDFOR
\STATE Output $(v^*, w^*) = \left(\frac{\sum_{t=1}^T \alpha_t v_t}{\sum_{t=1}^T \alpha_t}, \frac{\sum_{t=1}^T \alpha_t w_t}{\sum_{t=1}^T \alpha_t}\right)$
\end{algorithmic}
\end{algorithm}

% give convergence theorem for boosted version
Algorithm~\ref{alg:boosted-comid} enjoys the following primal-dual convergence rate:
\begin{theorem}
TODO
\end{theorem}

\subsection{Fenchel duality}
\label{sec:fenchel}
% Fenchel conjugate properties
% -definition
% -convexity-smoothness duality
% -examples: L2, entropy, p-norm, matrix entropy, log-det
%  -compute Fenchel conjugate and derivative
% -cite Kakade paper, maybe others?

\subsection{Relations to Other Algorithms}
\label{sec:derived}
% derive gradient descent as a special case
% also Frank-Wolfe

\end{document}

### 2014-03-10 arXiv/sparsity-result.tex
added a bit more exposition

@@
 \begin{algorithm}
-\caption{The composite mirror descent algorithm.}
+\caption{Composite Mirror Descent}
 \label{alg:comid}
 \begin{algorithmic}
-\STATE {\bfseries Composite Mirror Descent}
 \STATE Initialize $\theta_0 \gets 0$, $\psi_0 \gets \psi$
@@
 \label{thm:comid}
-Let $\|\cdot\|$ be any norm on $\W$. Suppose that $\psi$ is $\rho_0$-strongly convex 
+Let $\|\cdot\|$ be any norm on $\W$. Suppose that $\psi$ is non-negative and $\rho_0$-strongly convex 
 with respect to $\|\cdot\|$, and that $h_t$ is $\rho_t$-strongly convex with respect 
@@
 Then, defining $f(w) \eqdef \frac{1}{A_T} \sum_{t=1}^{T} \alpha_t f_t(w)$, we have
-\[ f(w^*) \leq \inf_{w} f(w) + \frac{r^2}{2A_T} \sum_{t=1}^T \frac{\alpha_t^2}{R_t}. \]
+\[ f(w^*) \leq \inf_{w} \left\{ f(w) + \frac{1}{A_T} \psi(w)\right\} + \frac{r^2}{2A_T} \sum_{t=1}^T \frac{\alpha_t^2}{R_t}. \]
 \end{theorem}
@@
 \begin{corollary}
-Suppose that $\psi$ is $\rho$-strongly convex with respect to $\|\cdot\|$, but make 
-no assumptions on $h_t$ other than convexity. Then, for $\alpha_t = 1$, we have
-\[ f(w^*) \leq \inf_w f(w) + \frac{r^2}{2\rho}. \]
+Suppose that $\psi$ is $\rho_0$-strongly convex with respect to $\|\cdot\|$, but make 
+no assumptions on $h_t$ other than convexity. Then, for $\alpha_t = 1/\sqrt{T}$, we have
+\[ f(w^*) \leq \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{r^2}{2\rho_0\sqrt{T}}. \]
 \end{corollary}
 \begin{proof}
-We have $A_T = T$ and $R_t = \rho$. Then Theorem~\ref{thm:comid} yields
+We have $A_T = \sqrt{T}$ and $R_t = \rho_0$. Then Theorem~\ref{thm:comid} yields
 \begin{align*}
-f(w^*) &\leq \inf_w f(w) + \frac{r^2}{2T} \sum_{t=1}^T \frac{1}{\rho} \\
- &= \inf_w f(w) + \frac{r^2}{2\rho}.
+f(w^*) &\leq \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{r^2}{2\sqrt{T}} \sum_{t=1}^T \frac{1}{\rho_0 T} \\
+ &= \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{r^2}{2\rho_0\sqrt{T}}.
 \end{align*}
 \end{proof}
+So, if the $h_t$ are strongly convex, we can obtain an $O\left(\frac{1}{T}\right)$ 
+convergence rate. If the $h_t$ are only (weakly) convex, we can still obtain an 
+$O\left(\frac{1}{\sqrt{T}}\right)$ convergence rate, by ``deforming'' $f$ by a 
+strongly convex function $\psi$.
 
@@
 \end{enumerate}
-Let $v_t = \argmax_{v \in V} f(v, w_t)$. Then we will pick 
+We will also assume that there is a compact set $S \subset \V$ with respect 
+to which we are trying to obtain sparsity. Let 
+$v_t = \argmax_{v \in S} f(v, w_t)$. Then we will pick 
 $h_t(w) = h(w)$ and $g_t(w) = f(v_t, w) - h(w)$. Note that, 
@@
   \STATE $w_t \gets \argmin_{w \in \W} \psi_t(w) + \theta_{t-1}^{\top} w$
-  \STATE $v_t \gets \argmax_{v \in \V} f(v, w_t)$
+  \STATE $v_t \gets \argmax_{v \in S} f(v, w_t)$
   \STATE $\theta_t \gets \theta_t + \alpha_t \left(\frac{\partial f(v_t, w_t)}{\partial w_t} - \frac{\partial h(w_t)}{\partial w_t}\right)$
@@
 \begin{theorem}
-TODO
+Define the \emph{duality gap} $\Delta_T$ as
+\[ \Delta_T \eqdef \sup_{v \in \Conv(S)} f(v, w_T) - \inf_{w \in \W} f(v_T, w). \]
+Suppose that $\psi$ is $\rho_0$-strongly convex with 
+respect to a norm $\|\cdot\|$, and that $h$ is $\rho$-strongly 
+convex with respect to the same norm. Suppose that 
+$\|\nabla g_t\|_* \leq r$ for all $t$. Then, defining 
+$A_t \eqdef \sum_{s=1}^T \alpha_s$ as before, we have
+\[ \Delta_T \leq \frac{1}{A_T} \psi(w_T) + \frac{r^2}{2A_T} \sum_{t=1}^T \frac{\alpha_t^2}{\rho_0 + \rho A_t}. \]
+In addition, $v_T$ is $(S, T)$-sparse.
 \end{theorem}
+%\begin{remark}
+%If $f$ is a continuously differentiable function, the compactness of $S$ 
+%guarantees that $\sup_{v \in S} \|\nabla g(v, w)\|_*$ is finite for each 
+%$w$. 
+%\end{remark}
+\begin{corollary}
+Suppose that $\psi \equiv 0$ and $\rho > 0$, and set $\alpha_t = t$. Then 
+\[ \Delta_T \leq \frac{2r^2}{\rho (T+1)}. \]
+\end{corollary}
+
+\begin{corollary}
+Suppose that $\rho_0 > 0$ and set $\alpha_t = 1/\sqrt{T}$. Then
+\[ \Delta_T \leq \frac{1}{\sqrt{T}} \left[ \psi(w_T) + \frac{r^2}{2\rho_0} \right]. \]
+\end{corollary}
 

### 2014-03-27 arXiv/sparsity-result.tex
fixed some small errors

@@
 convex if and only if $\nabla^2 f(w) \succeq \rho I$ for all $w \in \W$. In particular, 
-$f(w) = \frac{1}{2}\|w - w_0\|_2^2$ is $1$-strongly convex with respsect to $\|\cdot\|_2$ for 
+$f(w) = \frac{1}{2}\|w - w_0\|_2^2$ is $1$-strongly convex with respect to $\|\cdot\|_2$ for 
 any $w_0$, although this is not true for general norms (for instance, 
@@
 Suppose that $\psi$ is $\rho_0$-strongly convex with respect to $\|\cdot\|$, but make 
-no assumptions on $h_t$ other than convexity. Then, for $\alpha_t = 1/\sqrt{T}$, we have
-\[ f(w^*) \leq \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{r^2}{2\rho_0\sqrt{T}}. \]
+no assumptions on $h_t$ other than convexity. Then, for $\alpha_t = 2/\sqrt{T}$,
+\fTBD{JHH: corrected choice of $\alpha_{t}$ and statement of theorem; also added in a necessary log factor} we have
+\[ f(w^*) \leq \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{2r^2 \log(T+1)}{\rho_0\sqrt{T}}. \]
 \end{corollary}
 \begin{proof}
-We have $A_T = \sqrt{T}$ and $R_t = \rho_0$. Then Theorem~\ref{thm:comid} yields
+We have $A_T \ge \sqrt{T}$ and $R_t = \rho_0$ so Theorem~\ref{thm:comid} yields
 \begin{align*}
-f(w^*) &\leq \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{r^2}{2\sqrt{T}} \sum_{t=1}^T \frac{1}{\rho_0 T} \\
- &= \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{r^2}{2\rho_0\sqrt{T}}.
+f(w^*) 
+&\leq \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{r^2}{2\sqrt{T}} \sum_{t=1}^T \frac{4}{\rho_0 t} \\
+&\leq \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{2r^2\log(T+1)}{\rho_0\sqrt{T}}.
 \end{align*}
@@
 convergence rate. If the $h_t$ are only (weakly) convex, we can still obtain an 
-$O\left(\frac{1}{\sqrt{T}}\right)$ convergence rate, by ``deforming'' $f$ by a 
+$O\left(\frac{\log(T+1)}{\sqrt{T}}\right)$ convergence rate, by ``deforming'' $f$ by a 
 strongly convex function $\psi$.
@@
   \STATE $v_t \gets \argmax_{v \in S} f(v, w_t)$
-  \STATE $\theta_t \gets \theta_t + \alpha_t \left(\frac{\partial f(v_t, w_t)}{\partial w_t} - \frac{\partial h(w_t)}{\partial w_t}\right)$
+  \STATE $\theta_t \gets \theta_{t-1} + \alpha_t \left(\frac{\partial f(v_t, w_t)}{\partial w_t} - \frac{\partial h(w_t)}{\partial w_t}\right)$
 \ENDFOR
@@
 \begin{corollary}
-Suppose that $\rho_0 > 0$ and set $\alpha_t = 1/\sqrt{T}$. Then
-\[ \Delta_T \leq \frac{1}{\sqrt{T}} \left[ \psi(w_T) + \frac{r^2}{2\rho_0} \right]. \]
+Suppose that $\rho_0 > 0$ and set $\alpha_t = 2/\sqrt{T}$. Then
+\[ \Delta_T \leq \frac{1}{\sqrt{T}} \left[ \psi(w_T) + \frac{2r^2\log(T+1)}{\rho_0} \right]. \]
 \end{corollary}

### 2014-03-27 arXiv/introduction.tex
fixed some small errors
rewrote theorem 2.7 (lower bound)

@@
 \label{thm:main}
-Given vector spaces $\V$ and $\W$\fTBD{JHH: hmm...you've switched from $\reals^{d}$ to vector spaces}, let $h : \W \to \bR$ be a convex function and let $f : \V \times \W \to \reals$ be a function 
+Given vector spaces $\V$ and $\W$\fTBD{JHH: hmm...you've switched from $\reals^{d}$ to vector spaces}, let $h : \W \to \reals$ be a convex function and let $f : \V \times \W \to \reals$ be a function 
 such that
@@
 \end{theorem}
-This result is tight in the following sense:\fTBD{JS: can we generalize this to arbitrary norms? or some norms other than $\ell_{2}$?}\fTBD{JHH: rewrite to explicitly give $f$ and $S$}
+This result is tight in the following sense:\fTBD{JS: can we generalize this to arbitrary norms? or some norms other than $\ell_{2}$?}
 \begin{theorem}
 \label{thm:lower-bound}
-There exists a choice of $f$ satisfying the conditions of Theorem~\ref{thm:main}, 
-a compact set $S \subset \V$, and a point $x \in \Conv(S)$ such that for any 
+Choose $S = \Delta_{n}$ as in Example \ref{ex:simplex} with $\V = \W = \Conv(S)$.
+Let $f(v,w) = \ip{v}{w} - \frac{1}{2}\|w\|^{2}$ (which satisfies the conditions of Theorem~\ref{thm:main}).
+Then with $x = 0$, for any
 $(S,k)$-sparse point $y$, 
-\[ \max_{w \in \W} f(y,w) \ge \max_{w \in \W} f(x,w) + \Theta(1/k). \]
+\[ \max_{w \in \W} f(y,w) \ge \max_{w \in \W} f(x,w) + \Theta(1/k), \]
+i.e.,
+\[ \frac{1}{2}\|y\|_{2}^{2} \ge \Theta(1/k). \]
 \end{theorem}
@@
 \item (JS) BMD and proofs of sparsity result and lower bound
-\item (JH) Application: Low-rank matrix approximation, approximation of SDP matrix sums 
+\item (JH) Application: Low-rank matrix approximation, approximation of PSD matrix sums 
 \item (JH) Application: Graph sparsification, max flow

### 2014-04-01 arXiv/sparsity-result.tex
added some defs; added a couple comments to sparsity result section

@@
 % give convergence theorem for boosted version
-Algorithm~\ref{alg:boosted-comid} enjoys the following primal-dual convergence rate:
+Algorithm~\ref{alg:boosted-comid} enjoys the following primal-dual convergence rate:\fTBD{JHH: did you mean do use $v_{T}$ and $w_{T}$ here, or should those be $v^{*}$ and $w^{*}$?}
 \begin{theorem}
@@
 
-\begin{corollary}
+\begin{corollary} \label{cor:duality-cor-2} % XXX: should come up with a better label for this
 Suppose that $\rho_0 > 0$ and set $\alpha_t = 2/\sqrt{T}$. Then
